### 2016-02-05 submission/manuscript.Rmd
Update manuscript to RMarkdown + add citations.

---
title: "Analyzing DECREASE trials for extent of data fabrication"
author: "Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
bibliography: ../bibliography/library.bib
csl: ../bibliography/ama.csl
---

The set of DECREASE trials investigated the effectiveness of beta-blockers to prevent perioperative mortality in non-cardiac surgery. Perioperative stands for the period of the surgical procedure, typically including admission, anaesthesia, surgery, and recovery. Upon being discovered as invalid [@commissie2011;@commissie2012;@commissie2013], the DECREASE studies were excluded from a 2014 meta-analysis [@bouri2014]. This corrected analysis found that conclusions on the effectiveness of beta-blockers in perioperative mortality were reversed and actually *increased* perioperative mortality instead of decreasing it.

The committee reports about the DECREASE studies indicate that data fabrication was likely, but the extent of the data fabrication was not clearly indicated or deemed estimable. In this project, we aim to investigate the extent of data fabrication in the DECREASE studies by statistical analysis to fill this gap. We expect to confirm the conclusions of the reports that data fabrication is likely.

To investigate the evidence for data fabrication in the DECREASE studies, we take three steps. First, we try to replicate the findings in the 2014 meta-analysis [@bouri2014], which should be possible with the trial-event data provided in that paper. Second, we evaluate the veracity of the DECREASE studies (i.e., the probability that these study findings occurred assuming no data fabrication occurred). Third, we reverse this assumption and assume that data fabrication did occur and estimate how many data points would need to be fabricated to reproduce the results of the DECREASE studies. We expand on the methods for these three steps below.

# Step 1: replicating Bouri et al. (2014)

To properly analyze the data from clinical trials that investigate the perioperative effect of beta-blockers on non-cardiac surgery mortality in high risk heart-patients, we first try to replicate Bouri et al. to ensure that (1) their results are reproducible and (2) we are using the correct estimates in the following steps.

In order to replicate the results of the meta-analysis, we extract the trial-event data available from figures 2 and 3 in the original paper. More specifically, we extract the raw trial-event data for the 2 (control vs experimental) by 2 (event or no event) design, which we use to recompute the natural logarithm of the risk ratio and its standard error. These data are depicted in Table 1.

*Table 1*. An example of clinical trial summary data used throughout this research.

|          |   Control    | Experimental |
|----------|--------------|--------------|
| No event | a            | b            |
| Event    | c            | d            |

Using the notation of Table 1, the log of the risk ratio is computed as

$ln(RR)=ln(\frac{d/(b+d)}{c/(a+c)})$

The standard error of this statistic is computed as

$se_{ln(RR)}=\sqrt{\frac{1}{a}+\frac{1}{b}+\frac{1}{c}+\frac{1}{d}}$

We use the `R` package `metafor` [@viechtbauer2010] to meta-analyze the risk ratios. First, we analyze all the non-DECREASE studies to verify the estimates in Figure 2 of Bouri and analyze all DECREASE studies to verify the estimates in Figure 3 of Bouri. Second, we meta-analyze all studies combined, including a dummy-predictor for the DECREASE and non-DECREASE studies to replicate results presented in Figure 4 of Bouri. These models will be run with a random-effects model, using the `REML` estimation method, which has been indicated to be the most optimal estimation method for heterogeneity in a random effects model [@viechtbauer2005]. Note that results need to be exponentiated because the meta-analysis is conducted on log transformed risk ratios.

## Checklist

1. Create datafile with the trial-event data from Bouri et al. (2014)
2. Code calculation of ln(RR) and standard error calculation in R
3. Implement the use of `metafor` with the `rma()` function, using random effects estimation method `REML`
4. Run DECREASE studies
5. Run non-DECREASE studies
6. Run meta-regression with dummy predictor comparing DECREASE and non-DECREASE studies

# Step 2: evaluating the veracity of DECREASE studies

We estimate the veracity of the DECREASE studies based on the estimates from the non-DECREASE studies, where we assume that the non-DECREASE studies provide the best known representation of the effect of beta-blockers on perioperative mortality. Veracity indicates the plausibility of the data in a random sampling context [@peters2015]. We assume to estimate the true population distribution of effects, not perturbed by statistical significance publication bias, because a substantial number of nonsignificant effects are included in the dataset.

The estimates of the non-DECREASE studies are obtained from step 1, which include the estimated risk ratio, and the 95% credibility interval as provided by the package `metafor`. Based on this, we assume a normal distribution of population effects with the estimated effect of step 1 as the mean of the distribution. The estimate in step 1 is accompanied by a 95% credibility interval (lower bound and upper bound), which denotes the bounds of the normal distribution that covers 95% of the density. The standard deviation is calculated as the distance from the mean to either bound, divided by 1.96.

With the estimated population effect distribution of true (that is pooled average based on non-DECREASE trials) effects of beta-blockers, the two-sided probability of the DECREASE trials’ findings can be estimated. These serve as the probability that the DECREASE trials’ findings occurred naturally, i.e., without data fabrication. The combined probability of the DECREASE trials is also calculated, by taking the product of the probability of the two independent studies.

## Checklist

1. Note the mean estimated ln(RR), credible interval upper- and lower bound from step 1
2. Calculate standard deviation of the ln(RR)
3. Compute two-sided probability for both DECREASE studies individually (for ln(RR))
4. Compute joint probability of DECREASE studies by taking the product of the two-sided probabilities for the individual studies.

# Step 3: estimating the amount of fabricated data

We estimate the number of data points (i.e., event or no event) fabricated in each DECREASE study individually (I and IV), if we were to assume data fabrication occurred. This step is carried out regardless of the results of step 2.

We first estimate the event probability in each condition (beta-blocker and control), for both the DECREASE- and non-DECREASE studies separately. Step 1 only provides an estimate of the overall risk ratio, but not of the odds that an event occurs conditional on exposure (beta-blocker or control) and stratum (DECREASE- versus non-DECREASE). To this end, we run four meta-analyses to estimate the ln(odds) of mortality per stratum. We apply the same methods as in step 1 to meta-analyze the individual effects (i.e., metafor package, REML random effects estimate, credible intervals). This results in eight meta-analyses to estimate the log odds.

The log odds for each condition, which is used in the meta-analysis, is calculated as

$ln(odds)=ln(\frac{p}{1-p})$

where *p* is the proportion of patients experiencing an event in that condition. The log odds are accompanied by the standard error

$se_{ln(odds)}=\frac{1}{np(1-p)}$

where *n* is the total amount of data points in that condition. This results in four estimates:

1. ln(odds) + credible interval in non-DECREASE/genuine studies, control group
2. ln(odds) + credible interval in non-DECREASE/genuine studies, beta-blockers group
3. ln(odds) + credible interval in DECREASE/faked studies, control group
4. ln(odds) + credible interval in DECREASE/faked studies, beta-blockers group

In order to test the number of data points fabricated in the DECREASE studies, we apply the inversion method to the data. The inversion method operates in 9 steps outlined below.

1. Take the entire sample from one DECREASE study, with the prespecified group sizes from that DECREASE study (e.g., 49 controls and 51 patients on beta-blockers). 
2. Randomly assign X={0, 1, 2, … , N} of the data points to be faked. 
3. Draw an ln(odds) from the corresponding population effect 1-4 outlined above. 
4. Sample an event with the probability calculated from the ln(odds) from the previous step and note the data point 0 (no event) or 1 (event) from either the genuine effect distribution (estimates 1 + 2), or faked effect distribution (estimates 3 + 4). 
5. When all data points have been generated, estimate the ln(RR) in this simulated sample with metafor and REML estimation.
6. Note whether this ln(RR) is more likely under the non-DECREASE effect distribution (coded 0) or under the DECREASE effect distribution (coded 1), based on two-sided p-values (larger p-value is more probable)
7. Repeat k times (k = number of iterations)
8. Calculate proportions of iterations where estimate was more likely under fabrication than under genuine data.
9. Note for how many data points from the fabricated (DECREASE) distribution the proportion is closest to .025 and .975 to provide a 95% confidence interval of data points fabricated

This procedure is carried out for both DECREASE studies included in Bouri et al. (2014)

### Checklist

1. Calculate the ln(odds) per condition for both non-DECREASE and DECREASE studies.
2. Calculate the standard error of the ln(odds) per effect
3. Compute the four meta-analyses to estimate the ln(odds) and credible intervals for each of the 2 (beta-blocker; control) by 2 (non-DECREASE; DECREASE) design.
4. Create function to apply inversion method
5. Apply inversion method to both DECREASE studies separately.

# References

### 2016-03-09 submission/manuscript.Rmd
Update markdown file with ln(odds) computation

@@
 
-where *n* is the total amount of data points in that condition. This results in four estimates:
+where *n* is the total amount of data points in that condition. In the case that no events occurred in the condition and *p* = 0, *p* is calculated as for example
+
+$p=\frac{c+.5}{a+c+.5}$
+
+where a and c are indicated in Table 1. This adjustment prevents computational errors and introduces minimal bias [@agresti2002].
+
+These meta-analyses result in four estimates:
 
@@
 Sample an event with the probability calculated from the ln(odds) from the previous step and note the data point 0 (no event) or 1 (event) from either the genuine effect distribution (estimates 1 + 2), or faked effect distribution (estimates 3 + 4).
-5. When all data points have been generated, estimate the ln(RR) in this simulated sample with metafor and REML estimation.
+5. When all data points have been generated, estimate the ln(RR) in this simulated sample with `metafor` and REML estimation.
 6.

### 2016-03-09 submission/manuscript.Rmd
Update RMarkdown file

--- 2016-02-05 submission/manuscript.Rmd
@@
 output: word_document
+csl: ../bibliography/ama.csl
 bibliography: ../bibliography/library.bib
-csl: ../bibliography/ama.csl
 ---
 
-The set of DECREASE trials investigated the effectiveness of beta-blockers to prevent perioperative mortality in non-cardiac surgery.
-Perioperative stands for the period of the surgical procedure, typically including admission, anaesthesia, surgery, and recovery.
-Upon being discovered as invalid [@commissie2011;@commissie2012;@commissie2013], the DECREASE studies were excluded from a 2014 meta-analysis [@bouri2014].
-This corrected analysis found that conclusions on the effectiveness of beta-blockers in perioperative mortality were reversed and actually *increased* perioperative mortality instead of decreasing it.
+The effectiveness of beta-blockers in preventing perioperative mortatility in non-cardiac surgery has been subject of discussion due to findings of misconduct.
+[@commissie2011;@commissie2012;@commissie2013;@bouri2014].
+Perioperative mortality is the deathrate of patients during the period of the surgical procedure, typically including admission, anaesthesia, surgery, and recovery.
+The trials that were found to be harmed were the DECREASE trials[@commissie2011;@commissie2012;@commissie2013].
+Upon being discovered as invalid, the DECREASE studies were excluded from a 2014 meta-analysis [@bouri2014].
+The corrected analysis found that conclusions on the effectiveness of beta-blockers in perioperative mortality were reversed --- beta-blockers seemed to actually *increase* perioperative mortality instead of decrease it.
 
-The committee reports about the DECREASE studies indicate that data fabrication was likely, but the extent of the data fabrication was not clearly indicated or deemed estimable.
-In this project, we aim to investigate the extent of data fabrication in the DECREASE studies by statistical analysis to fill this gap.
+The committee reports about the DECREASE studies indicate that data fabrication was likely[@commissie2011;@commissie2012;@commissie2013], but the extent of the data fabrication was not clearly indicated or deemed estimable.
+In this paper, we aim to estimate the extent of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999].
 We expect to confirm the conclusions of the reports that data fabrication is likely.
 
-To investigate the evidence for data fabrication in the DECREASE studies, we take three steps.
-First, we try to replicate the findings in the 2014 meta-analysis [@bouri2014], which should be possible with the trial-event data provided in that paper.
+To investigate the evidence for data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999], we take three steps.
+First, we try to replicate the findings from the 2014 meta-analysis [@bouri2014], which should be possible with the trial-event data provided in that paper.
 Second, we evaluate the veracity of the DECREASE studies (i.e., the probability that these study findings occurred assuming no data fabrication occurred).
-Third, we reverse this assumption and assume that data fabrication did occur and estimate how many data points would need to be fabricated to reproduce the results of the DECREASE studies.
+Third, we reverse this assumption and assume that data fabrication did occur and estimate how many data points would have to be fabricated to reproduce the results of the DECREASE studies.
 We expand on the methods for these three steps below.
@@
 
-To properly analyze the data from clinical trials that investigate the perioperative effect of beta-blockers on non-cardiac surgery mortality in high risk heart-patients, we first try to replicate Bouri et al.
-to ensure that (1) their results are reproducible and (2) we are using the correct estimates in the following steps.
-
+We first replicate parts of Bouri et al.
+This helps ensure that (1) their results are reproducible and (2) we are using the correct estimates in the following steps.
 In order to replicate the results of the meta-analysis, we extract the trial-event data available from figures 2 and 3 in the original paper.
 More specifically, we extract the raw trial-event data for the 2 (control vs experimental) by 2 (event or no event) design, which we use to recompute the natural logarithm of the risk ratio and its standard error.
-These data are depicted in Table 1.
+
+Using the notation of Table 1, the log of the risk ratio per study is computed as
+
+$ln(RR)=ln(\frac{d/(b+d)}{c/(a+c)})$
+
+The standard error of this statistic is computed as
+
+$se_{ln(RR)}=\sqrt{\frac{1}{a}+\frac{1}{b}+\frac{1}{c}+\frac{1}{d}}$
+
+In case there is a zero-count in either *d* or *c* from Table 1, a half (i.e., .5) is added to each cell to ensure that computations are possible (i.e., *a*, *b*, *c*, and *d*).
+This is automatically done by `Review Manager`, the software used by Bouri et al., and is the default way of dealing with zero-counts in analyzing contingency tables [@agresti2002].
 
@@
 
-Using the notation of Table 1, the log of the risk ratio is computed as
-
-$ln(RR)=ln(\frac{d/(b+d)}{c/(a+c)})$
-
-The standard error of this statistic is computed as
-
-$se_{ln(RR)}=\sqrt{\frac{1}{a}+\frac{1}{b}+\frac{1}{c}+\frac{1}{d}}$
-
 We use the `R` package `metafor` [@viechtbauer2010] to meta-analyze the risk ratios.
@@
 Second, we meta-analyze all studies combined, including a dummy-predictor for the DECREASE and non-DECREASE studies to replicate results presented in Figure 4 of Bouri.
-These models will be run with a random-effects model, using the `REML` estimation method, which has been indicated to be the most optimal estimation method for heterogeneity in a random effects model [@viechtbauer2005].
-Note that results need to be exponentiated because the meta-analysis is conducted on log transformed risk ratios.
+These estimates are made with random-effects model using the `REML` estimation method.
+`REML` estimation has been indicated to be the most optimal estimation method for heterogeneity in a random effects model [@viechtbauer2005].
+Prior to comparison of the estimates, results need to be exponentiated because the meta-analysis is conducted on log transformed risk ratios.
 
@@
 
-We estimate the veracity of the DECREASE studies based on the estimates from the non-DECREASE studies, where we assume that the non-DECREASE studies provide the best known representation of the effect of beta-blockers on perioperative mortality.
-Veracity indicates the plausibility of the data in a random sampling context [@peters2015].
-We assume to estimate the true population distribution of effects, not perturbed by statistical significance publication bias, because a substantial number of nonsignificant effects are included in the dataset.
+We estimate the probability of the results in the DECREASE studies based on the estimates from the non-DECREASE studies.
+In order to do this, we assume that the non-DECREASE studies provide the best known representation of the true effect of beta-blockers on perioperative mortality.
+This is also known as estimating the veracity [@peters2015], which indicates the plausibility of the data in a random sampling context.
+We assume to estimate the true population distribution of effects, not perturbed by publication bias due to statistical significance, because a substantial number of nonsignificant effects are included in the dataset.
 
+Based on the estimated mean effect and credible interval in the non-DECREASE studies, we compute the probability of the DECREASE trials.
 The estimates of the non-DECREASE studies are obtained from step 1, which include the estimated risk ratio, and the 95% credibility interval as provided by the package `metafor`.
-Based on this, we assume a normal distribution of population effects with the estimated effect of step 1 as the mean of the distribution.
-The estimate in step 1 is accompanied by a 95% credibility interval (lower bound and upper bound), which denotes the bounds of the normal distribution that covers 95% of the density.
+Based on this, we assume a normal distribution of population effects with the estimated effect as the mean of the distribution.
+The 95% credibility interval  denotes the bounds of the normal distribution that covers 95% of the density.
 The standard deviation is calculated as the distance from the mean to either bound, divided by 1.96.
 
-With the estimated population effect distribution of true (that is pooled average based on non-DECREASE trials) effects of beta-blockers, the two-sided probability of the DECREASE trials’ findings can be estimated.
-These serve as the probability that the DECREASE trials’ findings occurred naturally, i.e., without data fabrication.
+The two-sided probability of the DECREASE trials’ findings serves as the probability that the DECREASE trials’ findings occurred naturally.
+In other words, this is the estimated probability of the DECREASE trials without data fabrication, assuming that the non-DECREASE trials are genuine.
 The combined probability of the DECREASE trials is also calculated, by taking the product of the probability of the two independent studies.
@@
 
-We estimate the number of data points (i.e., event or no event) fabricated in each DECREASE study individually (I and IV), if we were to assume data fabrication occurred.
+We estimate the number of data points fabricated in each DECREASE study individually (I and IV), if we were to assume data fabrication occurred.
 This step is carried out regardless of the results of step 2.
+This provides an indication of the severity of the data fabrication as concluded by the misconduct reports [@commissie2011;@commissie2012;@commissie2013;@bouri2014].
 
-We first estimate the event probability in each condition (beta-blocker and control), for both the DECREASE- and non-DECREASE studies separately.
-Step 1 only provides an estimate of the overall risk ratio, but not of the odds that an event occurs conditional on exposure (beta-blocker or control) and stratum (DECREASE- versus non-DECREASE).
+We first estimate the mortality probability in each condition (beta-blocker and control), for both the DECREASE- and non-DECREASE studies separately.
 To this end, we run four meta-analyses to estimate the ln(odds) of mortality per stratum.
@@
 This results in eight meta-analyses to estimate the log odds.
+These separate analyses are necessary, because step 1 only provides an estimate of the overall risk ratio, but not of the odds that an event occurs conditional on exposure (beta-blocker or control) and study set (DECREASE- versus non-DECREASE).
 
-The log odds for each condition, which is used in the meta-analysis, is calculated as
+The log odds per study are calculated for each condition. For example, for the control condition, based on Table 1, it is calculated as
 
-$ln(odds)=ln(\frac{p}{1-p})$
+$ln(odds)=ln(\frac{c/(a+c)}{1-(c/(a+c))})$
 
-where *p* is the proportion of patients experiencing an event in that condition. The log odds are accompanied by the standard error
+accompanied by the standard error
 
-$se_{ln(odds)}=\frac{1}{np(1-p)}$
+$se_{ln(odds)}=\frac{1}{(a+c)(c/(a+c))(1-(c/(a+c)))}$
 
-where *n* is the total amount of data points in that condition. This results in four estimates:
+For the beta-blocker condition *c* is replaced by *d*, and *a* is replaced by *b*.
+If there are no events in that condition (i.e., zero-count), a half is added to each cell (see also Step 1).[@agresti2002] Subsequently, meta-analyses are conducted per condition (beta-blocker and control), for both the DECREASE- and non-DECREASE studies separately.
 
-1. ln(odds) + credible interval in non-DECREASE/genuine studies, control group
-2. ln(odds) + credible interval in non-DECREASE/genuine studies, beta-blockers group
-3. ln(odds) + credible interval in DECREASE/faked studies, control group
-4. ln(odds) + credible interval in DECREASE/faked studies, beta-blockers group
+We apply the inversion method to estimate the number of data points fabricated in the DECREASE studies.
+To this end, we assume that the DECREASE data are in fact fabricated, opposed to Step 2 where we assumed they were genuine.
+The inversion method iteratively simulates data where *X* data points in that study are fabricated, where *X* varies from 0 through *N*.
+Each simulated sample is based on the effect distributions estimated earlier.
+This procedure is carried out for both DECREASE I and DECREASE IV.
+The analysis script will be made publicly available, and we include the step-by-step process here.
 
-In order to test the number of data points fabricated in the DECREASE studies, we apply the inversion method to the data. The inversion method operates in 9 steps outlined below.
-
-1. Take the entire sample from one DECREASE study, with the prespecified group sizes from that DECREASE study (e.g., 49 controls and 51 patients on beta-blockers). 
-2. Randomly assign X={0, 1, 2, … , N} of the data points to be faked. 
-3. Draw an ln(odds) from the corresponding population effect 1-4 outlined above. 
-4.
-Sample an event with the probability calculated from the ln(odds) from the previous step and note the data point 0 (no event) or 1 (event) from either the genuine effect distribution (estimates 1 + 2), or faked effect distribution (estimates 3 + 4).
-5. When all data points have been generated, estimate the ln(RR) in this simulated sample with metafor and REML estimation.
+1. Take the entire sample from one DECREASE study, with the prespecified group sizes from that DECREASE study (e.g., 49 controls and 51 patients on beta-blockers, N=100). 
+2. Randomly assign X={0, 1, 2, … , N} of the data points to be fabricated. 
+3. Draw an ln(odds) from the corresponding estimated effect distribution (i.e., fabricated/genuine control/beta-blocker condition). 
+4. Sample an event with the probability calculated from the ln(odds) from the previous step and note the data point 0 (no event) or 1 (event). 
+5. When all data points have been generated, estimate the ln(RR) in this simulated sample with `metafor` and `REML` estimation.
 6.
-Note whether this ln(RR) is more likely under the non-DECREASE effect distribution (coded 0) or under the DECREASE effect distribution (coded 1), based on two-sided p-values (larger p-value is more probable)
-7. Repeat k times (k = number of iterations)
+Note whether this ln(RR) is more likely under the non-DECREASE effect distribution (coded 0) or under the DECREASE effect distribution (coded 1), based on two-sided p-values (larger p-value is more probable) from the estimated effect distributions
+7. Repeat k times (k = number of iterations; these have yet to be determined but will be at least 1,000)
 8. Calculate proportions of iterations where estimate was more likely under fabrication than under genuine data.
-9. Note for how many data points from the fabricated (DECREASE) distribution the proportion is closest to .025 and .975 to provide a 95% confidence interval of data points fabricated
-
-This procedure is carried out for both DECREASE studies included in Bouri et al. (2014)
+9. Note for how many data points from the fabricated (DECREASE) distribution the proportion is closest to .025 and .975 to provide a 95% confidence interval of data points fabricated.
 
@@
 1. Calculate the ln(odds) per condition for both non-DECREASE and DECREASE studies.
-2. Calculate the standard error of the ln(odds) per effect
+2. Calculate the standard error of the ln(odds) per effect.
 3. Compute the four meta-analyses to estimate the ln(odds) and credible intervals for each of the 2 (beta-blocker; control) by 2 (non-DECREASE; DECREASE) design.
-4. Create function to apply inversion method
+4. Create function to apply inversion method.
 5. Apply inversion method to both DECREASE studies separately.

### 2016-03-11 submission/manuscript.Rmd
Add computation of ln(RR)

@@
 
-# Step 1: replicating Bouri et al. (2014)
+# Step 1: reproducing Bouri et al. (2014)
+
+```{r, echo = FALSE}
+suppressPackageStartupMessages(if(!require(metafor)){install.package('metafor')})
+suppressPackageStartupMessages(library(metafor))
+
+bouri_data <- read.csv('../data/bouri_data.csv')
+
+bouri_data$a <- bouri_data$control_n - bouri_data$control_events
+bouri_data$b <- bouri_data$beta_n - bouri_data$beta_events
+bouri_data$c <- bouri_data$control_events
+bouri_data$d <- bouri_data$beta_event
+
+# Add .5 to studies with zero-counts
+bouri_data[c(4, 7, 9), 7:10] <-  bouri_data[c(4, 7, 9), 7:10] + .5
+
+bouri_data$ln_rr <- log((bouri_data$d/(bouri_data$b+bouri_data$d))/(bouri_data$c/(bouri_data$a+bouri_data$c)))
+bouri_data$se_ln_rr <- sqrt((1/bouri_data$a)+(1/bouri_data$b)+(1/bouri_data$c)+(1/bouri_data$d))
+```
 

### 2016-03-11 submission/manuscript.Rmd
Add initial meta-analysis models for step 1

@@
 bouri_data$se_ln_rr <- sqrt((1/bouri_data$a)+(1/bouri_data$b)+(1/bouri_data$c)+(1/bouri_data$d))
+
+mod_clean <- metafor::rma(yi = ln_rr,
+             sei = se_ln_rr,
+             method = "REML",
+             data = bouri_data[bouri_data$decrease == 0, ],
+             weighted = TRUE)
+mod_dirty <- metafor::rma(yi = ln_rr,
+             sei = se_ln_rr,
+             method = "REML",
+             data = bouri_data[bouri_data$decrease == 1, ],
+             weighted = TRUE)
 ```

### 2016-03-17 submission/manuscript.Rmd
Implement + draft step 2

@@
 Perioperative mortality is the deathrate of patients during the period of the surgical procedure, typically including admission, anaesthesia, surgery, and recovery.
-The trials that were found to be harmed were the DECREASE trials[@commissie2011;@commissie2012;@commissie2013].
-Upon being discovered as invalid, the DECREASE studies were excluded from a 2014 meta-analysis [@bouri2014].
-The corrected analysis found that conclusions on the effectiveness of beta-blockers in perioperative mortality were reversed --- beta-blockers seemed to actually *increase* perioperative mortality instead of decrease it.
 
-The committee reports about the DECREASE studies indicate that data fabrication was likely[@commissie2011;@commissie2012;@commissie2013], but the extent of the data fabrication was not clearly indicated or deemed estimable.
+The trials that were found to be suspect were the DECREASE trials [@commissie2011;@commissie2012;@commissie2013].
+Upon being discovered as (potentially) fabricated, the DECREASE studies were excluded from a 2014 meta-analysis [@bouri2014].
+The corrected analyses found that conclusions on the effectiveness of beta-blockers in perioperative mortality were reversed --- beta-blockers seemed to actually *increase* perioperative mortality instead of decrease it.
+
+The committee reports about the DECREASE studies indicate that data fabrication was likely [@commissie2011;@commissie2012;@commissie2013], but the extent of the data fabrication was not clearly indicated or deemed estimable.
 In this paper, we aim to estimate the extent of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999].
@@
 To investigate the evidence for data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999], we take three steps.
-First, we try to replicate the findings from the 2014 meta-analysis [@bouri2014], which should be possible with the trial-event data provided in that paper.
+First, we try to replicate the findings from the 2014 meta-analysis [@bouri2014], which should be possible with the trial-event data provided in this meta-analysis.
 Second, we evaluate the veracity of the DECREASE studies (i.e., the probability that these study findings occurred assuming no data fabrication occurred).
@@
 # Step 1: reproducing Bouri et al. (2014)
+
+In order to ensure we are using similar analytic procedures as in the 2014 meta-analysis [@bouri2014], we initially reproduce the estimates.
+This ensured that (1) their results are reproducible and (2) we are using the correct estimates in the following steps.
+In order to replicate the results of the meta-analysis, we extract the trial-event data available from figures 2 and 3 in the original paper[@bouri2014].
+More specifically, we extract the raw trial-event data for the 2 (control vs experimental) by 2 (event or no event) design, which we use to recompute the natural logarithm of the risk ratio and its standard error.
+The extracted trial-event data is available at osf.io/aykeh and our analysis plan was preregistered at osf.io/vnmzc.
+
+## Method
 
@@
 
-We first replicate parts of Bouri et al.
-This helps ensure that (1) their results are reproducible and (2) we are using the correct estimates in the following steps.
-In order to replicate the results of the meta-analysis, we extract the trial-event data available from figures 2 and 3 in the original paper.
-More specifically, we extract the raw trial-event data for the 2 (control vs experimental) by 2 (event or no event) design, which we use to recompute the natural logarithm of the risk ratio and its standard error.
+We computed the risk ratio (RR) for each study and meta-analyzed the effects.
+We used the `R` package `metafor` [@viechtbauer2010] to meta-analyze the RRs.
+We estimated a weighted random-effects model using the restricted maximum-likelihood estimator (i.e., `REML`)[@viechtbauer2005] to estimate the variance of effects.
+We used the default weighting procedure in the `metafor` package, because the weights provided in the original paper were not accounted for.
+The 2014 meta-analysis [@bouri2014] did not specify the effect variance estimate used, so minor discrepancies between our estimates and the original estimates might originate from these differences in the estimation procedure.
 
-Using the notation of Table 1, the log of the risk ratio per study is computed as
+## Results
 
-$ln(RR)=ln(\frac{d/(b+d)}{c/(a+c)})$
+First, we analyzed the RRs to verify the estimates in Figure 2 and found that we were able to reproduce the estimates for the different sets of studies.
+The original Figure 2 [@bouri2014] differentiated between the estimates from the non-DECREASE (i.e., secure) trials and the DECREASE (i.e., non-secure trials).
+We confirmed the estimates of the effect size and the variance for both the non-DECREASE- and the DECREASE trials, save for some minor discrepancies due to the estimation method.
+Table `r tab <- 1; tab` depicts the original and reproduced values for both sets of studies.
 
-The standard error of this statistic is computed as
+*Table `r tab`*. The original- and reproduced meta-analytic results based on the data provided in the 2014 meta-analysis.[@bouri2014]
 
-$se_{ln(RR)}=\sqrt{\frac{1}{a}+\frac{1}{b}+\frac{1}{c}+\frac{1}{d}}$
+|              |            | Risk ratio   | $\tau^2$      | Confidence interval |
+|--------------|------------|---------------------|-------------------|---|
+| Non-DECREASE | Original   | 1.27                   | 0.00                 | [1.01; 1.60] |
+|              | Reproduced | `r round(exp(mod_clean$b), 2)`  | `r round(mod_clean$tau2, 2)`                 | `r sprintf('[%s; %s]', round(exp(mod_clean$ci.lb), 2), round(exp(mod_clean$ci.ub), 2))` |
+| DECREASE     | Original   | 0.42                   | 0.29                 | [0.15; 1.23] |
+|              | Reproduced | `r round(exp(mod_dirty$b), 2)` | `r round(mod_dirty$tau2, 2)` | `r sprintf('[%s; %s]', round(exp(mod_dirty$ci.lb), 2), round(exp(mod_dirty$ci.ub), 2))` |
 
-In case there is a zero-count in either *d* or *c* from Table 1, a half (i.e., .5) is added to each cell to ensure that computations are possible (i.e., *a*, *b*, *c*, and *d*).
-This is automatically done by `Review Manager`, the software used by Bouri et al., and is the default way of dealing with zero-counts in analyzing contingency tables [@agresti2002].
+```{r, echo = FALSE}
+mod_mods <- metafor::rma(yi = ln_rr,
+             sei = se_ln_rr,
+             mod =~ as.factor(decrease), # 1 are the decrease studies
+             method = "REML",
+             data = bouri_data,
+             weighted = TRUE)
+```
 
-*Table 1*. An example of clinical trial summary data used throughout this research.
+Second, we meta-analyzed all studies combined, including a dummy-predictor for the DECREASE and non-DECREASE studies to replicate results presented in Figure 4 of the 2014 meta-analysis.[@bouri2014] Oddly enough our reproduced results deviated more from the asssumption that the subgroups are in fact from the same population; whereas the original test for subgroup differences was $\chi^2(1)=3.91,p=.05$, our reproduced analyses indicated that $\chi^2(1)=`r round(mod_mods$QM, 2)`,p=`r round(mod_mods$QMp, 3)`$.
+Additionally, the original analyses showed substantial heterogeneity ($I^2=74.4$%), but our reproduced analyses showed no such thing ($I^2=`r round(mod_mods$I2, 3)`$%).
+Different estimates of the heterogeneity (e.g., DerSimonian-Laird instead of `REML`) did not change this difference.
 
-|          |   Control    | Experimental |
-|----------|--------------|--------------|
-| No event | a            | b            |
-| Event    | c            | d            |
-
-We use the `R` package `metafor` [@viechtbauer2010] to meta-analyze the risk ratios.
-First, we analyze all the non-DECREASE studies to verify the estimates in Figure 2 of Bouri and analyze all DECREASE studies to verify the estimates in Figure 3 of Bouri.
-Second, we meta-analyze all studies combined, including a dummy-predictor for the DECREASE and non-DECREASE studies to replicate results presented in Figure 4 of Bouri.
-These estimates are made with random-effects model using the `REML` estimation method.
-`REML` estimation has been indicated to be the most optimal estimation method for heterogeneity in a random effects model [@viechtbauer2005].
-Prior to comparison of the estimates, results need to be exponentiated because the meta-analysis is conducted on log transformed risk ratios.
-
-## Checklist
-
-1. Create datafile with the trial-event data from Bouri et al. (2014)
-2. Code calculation of ln(RR) and standard error calculation in R
-3. Implement the use of `metafor` with the `rma()` function, using random effects estimation method `REML`
-4. Run DECREASE studies
-5. Run non-DECREASE studies
-6. Run meta-regression with dummy predictor comparing DECREASE and non-DECREASE studies
+<!--- Emailed bouri about this on 2016 03 16--->
 
@@
 
-We estimate the probability of the results in the DECREASE studies based on the estimates from the non-DECREASE studies.
-In order to do this, we assume that the non-DECREASE studies provide the best known representation of the true effect of beta-blockers on perioperative mortality.
-This is also known as estimating the veracity [@peters2015], which indicates the plausibility of the data in a random sampling context.
-We assume to estimate the true population distribution of effects, not perturbed by publication bias due to statistical significance, because a substantial number of nonsignificant effects are included in the dataset.
+We estimated the probability of the results in the DECREASE studies based on the estimates from the non-DECREASE studies.
+In order to do this, we assumed that the non-DECREASE studies provide the best known representation of the true effect of beta-blockers on perioperative mortality.
+This probability is also known as the veracity of the data [@peters2015], which indicates the probability of the observed data in a random sampling context.
+We assumed to estimate the true population distribution of effects, not perturbed by publication bias due to statistical significance, because a substantial number of nonsignificant effects are included in the dataset.
+To estimate the true population effect we only used the non-DECREASE studies.
 
-Based on the estimated mean effect and credible interval in the non-DECREASE studies, we compute the probability of the DECREASE trials.
-The estimates of the non-DECREASE studies are obtained from step 1, which include the estimated risk ratio, and the 95% credibility interval as provided by the package `metafor`.
-Based on this, we assume a normal distribution of population effects with the estimated effect as the mean of the distribution.
-The 95% credibility interval  denotes the bounds of the normal distribution that covers 95% of the density.
-The standard deviation is calculated as the distance from the mean to either bound, divided by 1.96.
+## Method
 
-The two-sided probability of the DECREASE trials’ findings serves as the probability that the DECREASE trials’ findings occurred naturally.
-In other words, this is the estimated probability of the DECREASE trials without data fabrication, assuming that the non-DECREASE trials are genuine.
-The combined probability of the DECREASE trials is also calculated, by taking the product of the probability of the two independent studies.
+```{r, echo = FALSE}
+mod_clean_pred <- predict(mod_clean)
+```
 
-## Checklist
+Based on the estimated mean log RR and credible interval in the non-DECREASE studies, we computed the probability of log RR in the DECREASE trials.
+The estimates of the non-DECREASE studies were obtained from step 1, which include the estimated log RR (i.e., `r round(mod_clean$b, 2)`), and the 95% credibility interval as provided by the package `metafor` (i.e., `r sprintf('[%s; %s]', round(mod_clean_pred$cr.lb, 3), round(mod_clean_pred$cr.ub, 3))`.
+We assumed a normal distribution of population effects with the estimated effect as the mean of the distribution.
+The 95% credibility interval denotes the bounds of the normal distribution that covers 95% of the density, where the standard deviation is calculated as the distance from the mean to either bound, divided by 1.96.
+This allows for an approximation of the population effect distribution, as depicted in Figure `r fig <- 1; fig`.
 
-1. Note the mean estimated ln(RR), credible interval upper- and lower bound from step 1
-2. Calculate standard deviation of the ln(RR)
-3. Compute two-sided probability for both DECREASE studies individually (for ln(RR))
-4. Compute joint probability of DECREASE studies by taking the product of the two-sided probabilities for the individual studies.
+*Figure `r fig`*. Density plot of the estimated true effect distribution, with the position of the DECREASE studies highlighted.
+```{r, echo = FALSE}
+temp <- seq(-3, 3, .001)
+
+# pdf('../figures/fig1', width = 8, height = 6)
+# use http://pdf2png.com/ when preparing manuscript
+# for high quality plots
+par(mar = c(.5,0,0,0))
+plot(x = temp,
+     dnorm(x = temp,
+      mod_clean$b,
+      (mod_clean_pred$pred - mod_clean_pred$cr.lb)/1.96),
+     type = 'l',
+     bty = 'n',
+     xaxt = 'n',
+     yaxt = 'n',
+     xlab = '',
+     ylab = '')
+
+points(x = bouri_data$ln_rr[bouri_data$decrease == 1][1], y = 0)
+text(x = bouri_data$ln_rr[bouri_data$decrease == 1][1],
+     y = .15,
+     labels = 'DECREASE-I',
+     cex = .8)
+
+points(x = bouri_data$ln_rr[bouri_data$decrease == 1][2], y = 0)
+text(x = bouri_data$ln_rr[bouri_data$decrease == 1][2],
+     y = -.1,
+     labels = 'DECREASE-IV',
+     cex = .8)
+
+# dev.off()
+```
+
+Based on the estimated effect distribution from the non-DECREASE trials, we calculated the probability of each DECREASE trial or a more extreme result.
+In other words, we computed the *p*-value for the hypothesis that the DECREASE studies occurred based on the information available from the other trials.
+
+## Results
+
+```{r, echo = FALSE}
+prob <- pnorm(q = bouri_data$ln_rr[bouri_data$decrease == 1],
+      mod_clean_pred$pred,
+      (mod_clean_pred$pred - mod_clean_pred$cr.lb)/1.96) * 2
+```
+
+As Figure `r fig` already indicates, the DECREASE trials are highly unlikely under the estimated effect distribution based on the non-DECREASE trials.
+More specifically, DECREASE-I (or more extreme) has a probability of `r prob[1]` (much less than 1 in a quintillion) and DECREASE-IV `r prob[2]` (3 in a billion), which indicates they are unlikely to come from the same population effect distribution.
+Observing two of such extremely unlikely studies in this population effect distribution is then almost impossible, `r prob[1]*prob[2]`.
+This would therefore indicate that the DECREASE trials are severely different from the non-DECREASE studies.
 
@@
 
+# Discussion
+
 # References

### 2016-03-30 functions/inversion_method_certain.R
Program inversion method, saving objects for efficiency

@@ -1,0 +1,183 @@
+xs <- 0:(bouri_data$control_n[bouri_data$decrease == 1][1] + 
+           bouri_data$beta_n[bouri_data$decrease == 1][1])
+
+result_almost_i_certain <- matrix(nrow = iterations, ncol = length(xs))
+
+# Loop for DECREASE-I
+j <- 1
+
+for (x in xs){
+  for (iter in 1:iterations){
+    # Setup data handling
+    working_df <- data.frame(condition = c(rep('control', bouri_data$control_n[bouri_data$decrease == 1][1]),
+                                           rep('beta', bouri_data$beta_n[bouri_data$decrease == 1][1])), 
+                             fabricated = rep(NA, max(xs)),
+                             event = rep(NA, max(xs)))
+    working_df$fabricated <- sample(x = c(rep(0, max(xs) - x), rep(1, x)), replace = FALSE)
+    
+    # Sample the proportion of events from population effects
+    # min() ensures that probabilities are truncated to be at most 1
+    cont_prop <- min(exp(rnorm(1,
+                               cont_pred,
+                               cont_var)), 1)
+    beta_prop <- min(exp(rnorm(1,
+                               beta_pred,
+                               beta_var)), 1)
+    
+    cont_decrease_prop <- exp(cont_decrease_pred)
+    beta_decrease_prop <- exp(beta_decrease_pred)
+    
+    # Simulate the different data
+    # Control non-fabricated
+    working_df$event[working_df$condition == 'control' &
+                       working_df$fabricated == 0] <- rbinom(n = sum(working_df$condition == 'control' & working_df$fabricated == 0),
+                                                             size = 1,
+                                                             cont_prop)
+    
+    # Beta non-fabricated
+    working_df$event[working_df$condition == 'beta' &
+                       working_df$fabricated == 0] <- rbinom(n = sum(working_df$condition == 'beta' & working_df$fabricated == 0),
+                                                             size = 1,
+                                                             beta_prop)
+    
+    # Control fabricated
+    working_df$event[working_df$condition == 'control' &
+                       working_df$fabricated == 1] <- rbinom(n = sum(working_df$condition == 'control' & working_df$fabricated == 1),
+                                                             size = 1,
[…]

### 2016-03-30 functions/inversion_method_uncertain.R

@@ -1,0 +1,191 @@
+xs <- 0:(bouri_data$control_n[bouri_data$decrease == 1][1] + 
+           bouri_data$beta_n[bouri_data$decrease == 1][1])
+
+result_almost_i_uncertain <- matrix(nrow = iterations, ncol = length(xs))
+
+# Loop for DECREASE-I
+j <- 1
+
+for (x in xs){
+  for (iter in 1:iterations){
+    # Setup data handling
+    working_df <- data.frame(condition = c(rep('control', bouri_data$control_n[bouri_data$decrease == 1][1]),
+                                           rep('beta', bouri_data$beta_n[bouri_data$decrease == 1][1])), 
+                             fabricated = rep(NA, max(xs)),
+                             event = rep(NA, max(xs)))
+    working_df$fabricated <- sample(x = c(rep(0, max(xs) - x), rep(1, x)), replace = FALSE)
+    
+    # Sample the proportion of events from population effects
+    # min() ensures that probabilities are truncated to be at most 1
+    cont_prop <- min(exp(rnorm(1,
+                               cont_pred,
+                               cont_var)), 1)
+    beta_prop <- min(exp(rnorm(1,
+                               beta_pred,
+                               beta_var)), 1)
+    
+    cont_decrease_prop <- min(exp(rnorm(1,
+                                        cont_decrease_pred,
+                                        cont_decrease_var)), 1)
+    beta_decrease_prop <- min(exp(rnorm(1,
+                                        beta_decrease_pred,
+                                        beta_decrease_var)), 1)
+    
+    # Simulate the different data
+    # Control non-fabricated
+    working_df$event[working_df$condition == 'control' &
+                       working_df$fabricated == 0] <- rbinom(n = sum(working_df$condition == 'control' & working_df$fabricated == 0),
+                                                             size = 1,
+                                                             cont_prop)
+    
+    # Beta non-fabricated
+    working_df$event[working_df$condition == 'beta' &
+                       working_df$fabricated == 0] <- rbinom(n = sum(working_df$condition == 'beta' & working_df$fabricated == 0),
+                                                             size = 1,
+                                                             beta_prop)
+    
+    # Control fabricated
+    working_df$event[working_df$condition == 'control' &
[…]

### 2016-03-30 submission/manuscript.Rmd
Update manuscript

@@
 suppressPackageStartupMessages(library(metafor))
+suppressPackageStartupMessages(if(!require(latex2exp)){install.package('latex2exp')})
+suppressPackageStartupMessages(library(latex2exp))
 
@@
 We used the default weighting procedure in the `metafor` package, because the weights provided in the original paper were not accounted for.
+In the case that there was a zero-count for a cell, .5 was added to each cell of the trial [@agresti2002].
 The 2014 meta-analysis [@bouri2014] did not specify the effect variance estimate used, so minor discrepancies between our estimates and the original estimates might originate from these differences in the estimation procedure.
@@
 mod_clean_pred <- predict(mod_clean)
+mod_dirty_pred <- predict(mod_dirty)
 ```
@@
 
-*Figure `r fig`*. Density plot of the estimated true effect distribution, with the position of the DECREASE studies highlighted.
 ```{r, echo = FALSE}
@@
 
+*Figure `r fig`*. Density plot of the estimated true effect distribution, with the position of the DECREASE studies highlighted.
+
 Based on the estimated effect distribution from the non-DECREASE trials, we calculated the probability of each DECREASE trial or a more extreme result.
@@
 As Figure `r fig` already indicates, the DECREASE trials are highly unlikely under the estimated effect distribution based on the non-DECREASE trials.
-More specifically, DECREASE-I (or more extreme) has a probability of `r prob[1]` (much less than 1 in a quintillion) and DECREASE-IV `r prob[2]` (3 in a billion), which indicates they are unlikely to come from the same population effect distribution.
-Observing two of such extremely unlikely studies in this population effect distribution is then almost impossible, `r prob[1]*prob[2]`.
+More specifically, DECREASE-I (or more extreme) has a probability of `r prob[1]` (less than 1 in a quintillion) and DECREASE-IV `r prob[2]` (3 in a billion), which indicates they are unlikely to come from the same population effect distribution.
+Observing two of such extremely unlikely studies in this population effect distribution is highly improbable, `r prob[1]*prob[2]`.
 This would therefore indicate that the DECREASE trials are severely different from the non-DECREASE studies.
+
+We do note that the variance in the non-DECREASE studies was estimated at 0, indicating that there would supposedly be a fixed-effect of beta-blockers on perioperative mortality despite differences in type of beta-blockers administered and dosage/duration of the beta-blockers.
+This homogeneity is not excessive, considering that the *Q*-statistic is sufficiently large to begin with (i.e., `r round(mod_clean$QE, .2)`).[@ioannidis2006] Nonetheless, the original non-DECREASE trials show substantive variation in the treatment, for example the usage of different beta-blockers (i.e., atenolol, propanolol, bisoprolol, metoprolol), which have been argued to affect effectiveness of the interventions REF.
+As such, the lack of heterogeneity can be genuine and indicative of no difference between these beta-blockers and their dosage/duration.
+Nonetheless, it could also be due to uncertainty in the estimate of heterogeneity due to the small set of studies included.
+
+Considering that the variance (i.e., heterogeneity) estimate is 0 and uncertain due to the small amount of studies included, sensitivity analyses of the results are in place.
+The probability of observing both DECREASE trials stays approximately below 1 out of 1000 until the variance estimate is .25, and the probability stays approximately below 1 out of 100 until the variance estimate is .43 (see Figure `r fig <- fig + 1; fig`).
+Further research would be worthwhile to better estimate the variance in the effects of beta-blockers on perioperative mortality.
+
+```{r, echo = FALSE}
+tau2 <- seq(0, 2, .01)
+
+prob_sens_t <- NULL
+i <- 1
+
+for (t2 in tau2){
+  mod_clean_sens <- metafor::rma(yi = ln_rr,
+                                 sei = se_ln_rr, tau2 = t2,
+                                 method = "REML",
+                                 data = bouri_data[bouri_data$decrease == 0, ],
+                                 weighted = TRUE)
+  mod_clean_pred_sens <- predict(mod_clean_sens)
+  prob_sens <- pnorm(q = bouri_data$ln_rr[bouri_data$decrease == 1],
+                     mod_clean_pred_sens$pred,
+                     (mod_clean_pred_sens$pred - mod_clean_pred_sens$cr.lb)/1.96) * 2
+  prob_sens_t[i] <- prob_sens[1] * prob_sens[2]
+
+  i <- i + 1
+}
+
+par(mar = c(4, 4, 0, 0))
+plot(tau2, prob_sens_t, type = 'l', xlab = TeX('$\\tau^2$'),
+     ylab = 'p-value', bty = 'n')
+points(x = tau2[c(26, 44)], y = prob_sens_t[c(26, 44)])
+text(x = tau2[c(26, 44)],
+     y = prob_sens_t[c(26, 44)] + .02,
+     labels = c(TeX('$\\tau^2=.25$'), TeX('$\\tau^2=.43$')),
+     cex = .8)
+```
+
+*Figure `r fig`*.
+Sensitivity analyses for the *p*-value that indicates the probability of observing the DECREASE studies, or more extreme, based on the estimated population effect and the accompanying variance estimate.
 
@@
 
-We estimate the number of data points fabricated in each DECREASE study individually (I and IV), if we were to assume data fabrication occurred.
-This step is carried out regardless of the results of step 2.
-This provides an indication of the severity of the data fabrication as concluded by the misconduct reports [@commissie2011;@commissie2012;@commissie2013;@bouri2014].
+We estimated the number of data points that would have to be fabricated to arrive at the estimates from the DECREASE studies, given that the non-DECREASE studies represent the population effect.
+This assumes, in contrast to Step 2, that the DECREASE studies were in fact fabricated.
+The estimates from Step 3 provide an indication to the extent of the potential data fabrication in the DECREASE studies.[@commissie2011;@commissie2012;@commissie2013;@bouri2014]
 
-We first estimate the mortality probability in each condition (beta-blocker and control), for both the DECREASE- and non-DECREASE studies separately.
-To this end, we run four meta-analyses to estimate the ln(odds) of mortality per stratum.
-We apply the same methods as in step 1 to meta-analyze the individual effects (i.e., metafor package, REML random effects estimate, credible intervals).
-This results in eight meta-analyses to estimate the log odds.
-These separate analyses are necessary, because step 1 only provides an estimate of the overall risk ratio, but not of the odds that an event occurs conditional on exposure (beta-blocker or control) and study set (DECREASE- versus non-DECREASE).
+## Method
 
-The log odds per study are calculated for each condition. For example, for the control condition, based on Table 1, it is calculated as
+In order to estimate the number of fabricated data points, we meta-analyzed the log odds per condition for both the DECREASE- and non-DECREASE studies separately.
+For each of these, we ran a meta-analysis applying the same methods as in Step 1.
+These separate analyses were necessary, because Step 1 only provided estimates of the overall risk ratio, but not of the odds conditional on exposure (beta-blocker or control) and study set (DECREASE- versus non-DECREASE).
 
-$ln(odds)=ln(\frac{c/(a+c)}{1-(c/(a+c))})$
+We applied the inversion method to estimate the number of data points fabricated in the DECREASE-I and DECREASE-IV studies.
+To this end, we assumed that the DECREASE data are in fact fabricated, opposed to Step 2 where we assumed they were genuine.
+The inversion method iteratively simulates data with *X* fabricated data points; *X* varies from 0 through *N*.
+For each *X*, 10,000 iterations were conducted.
+Data fabrication occurred using a binomial distribution, with the population effect from the DECREASE studies serving as the probability of an event occurring.
+The population effect was either drawn from the estimated population distribution, which introduced uncertainty (i.e., uncertain estimate), or was equated to the point estimate (i.e., certain estimate; not preregistered).
+The remaining genuine data, *N - X*, were simulated in a similar fashion, where we only used the effect distribution from the non-DECREASE studies.
+Using the (partially) fabricated dataset, we estimated the risk ratio from the data and noted whether the two-sided *p*-value was larger under the non-DECREASE effect distribution (coded 0) or the DECREASE effect distribution.
 
-accompanied by the standard error
+Based on the results of the inversion method, we obtain a confidence interval of data points fabricated in the DECREASE studies.
+After running all iterations for one *X*, the proportion of trials that were more likely under the DECREASE population effect than under the non-DECREASE population effect were computed.
+This resulted in a proportion for each *X* from 0 to *N*.
+Those proportions closest to .025 and .975 denoted the 95% confidence interval of data points fabricated.
 
-$se_{ln(odds)}=\frac{1}{(a+c)(c/(a+c))(1-(c/(a+c)))}$
+## Results
 
-For the beta-blocker condition *c* is replaced by *d*, and *a* is replaced by *b*.
-If there are no events in that condition (i.e., zero-count), a half is added to each cell (see also Step 1).[@agresti2002] Subsequently, meta-analyses are conducted per condition (beta-blocker and control), for both the DECREASE- and non-DECREASE studies separately.
+```{r, echo = FALSE}
+# Log odds for control conditions
+numerator <- bouri_data$c/(bouri_data$a + bouri_data$c)
+denominator <- 1 - (bouri_data$c / (bouri_data$a + bouri_data$c))
+bouri_data$ln_odds_cont <- log(numerator / denominator)
 
-We apply the inversion method to estimate the number of data points fabricated in the DECREASE studies.
-To this end, we assume that the DECREASE data are in fact fabricated, opposed to Step 2 where we assumed they were genuine.
-The inversion method iteratively simulates data where *X* data points in that study are fabricated, where *X* varies from 0 through *N*.
-Each simulated sample is based on the effect distributions estimated earlier.
-This procedure is carried out for both DECREASE I and DECREASE IV.
-The analysis script will be made publicly available, and we include the step-by-step process here.
+bouri_data$se_ln_odds_cont <- 1 / ((bouri_data$a + bouri_data$c) * 
+                                     (bouri_data$c / (bouri_data$a + bouri_data$c)) *
+                                     (1 - (bouri_data$c / (bouri_data$a + bouri_data$c))))
 
-1. Take the entire sample from one DECREASE study, with the prespecified group sizes from that DECREASE study (e.g., 49 controls and 51 patients on beta-blockers, N=100). 
-2. Randomly assign X={0, 1, 2, … , N} of the data points to be fabricated. 
-3. Draw an ln(odds) from the corresponding estimated effect distribution (i.e., fabricated/genuine control/beta-blocker condition). 
-4. Sample an event with the probability calculated from the ln(odds) from the previous step and note the data point 0 (no event) or 1 (event). 
-5. When all data points have been generated, estimate the ln(RR) in this simulated sample with `metafor` and `REML` estimation.
-6.
-Note whether this ln(RR) is more likely under the non-DECREASE effect distribution (coded 0) or under the DECREASE effect distribution (coded 1), based on two-sided p-values (larger p-value is more probable) from the estimated effect distributions
-7. Repeat k times (k = number of iterations; these have yet to be determined but will be at least 1,000)
-8. Calculate proportions of iterations where estimate was more likely under fabrication than under genuine data.
-9. Note for how many data points from the fabricated (DECREASE) distribution the proportion is closest to .025 and .975 to provide a 95% confidence interval of data points fabricated.
+# Log odds for beta-blocker conditions
+numerator <- bouri_data$d/(bouri_data$b + bouri_data$d)
+denominator <- 1 - (bouri_data$d / (bouri_data$b + bouri_data$d))
+bouri_data$ln_odds_beta <- log(numerator / denominator)
 
-### Checklist
+bouri_data$se_ln_odds_beta <- 1 / ((bouri_data$b + bouri_data$d) * 
+                                     (bouri_data$d / (bouri_data$b + bouri_data$d)) *
+                                     (1 - (bouri_data$d / (bouri_data$b + bouri_data$d))))
 
-1. Calculate the ln(odds) per condition for both non-DECREASE and DECREASE studies.
-2. Calculate the standard error of the ln(odds) per effect.
-3. Compute the four meta-analyses to estimate the ln(odds) and credible intervals for each of the 2 (beta-blocker; control) by 2 (non-DECREASE; DECREASE) design.
-4. Create function to apply inversion method.
-5. Apply inversion method to both DECREASE studies separately.
+mod_cont <- rma(yi = ln_odds_cont,
+                sei = se_ln_odds_cont,
+                data = bouri_data[bouri_data$decrease == 0, ],
+                method = 'REML',
+                weighted = TRUE)
+mod_beta <- rma(yi = ln_odds_beta,
+                sei = se_ln_odds_beta,
+                data = bouri_data[bouri_data$decrease == 0, ],
+                method = 'REML',
+                weighted = TRUE)
+
+mod_cont_decrease <- rma(yi = ln_odds_cont,
+                         sei = se_ln_odds_cont,
+                         data = bouri_data[bouri_data$decrease == 1, ],
+                         method = 'REML',
+                         weighted = TRUE)
+mod_beta_decrease <- rma(yi = ln_odds_beta,
+                         sei = se_ln_odds_beta,
+                         data = bouri_data[bouri_data$decrease == 1, ],
+                         method = 'REML',
+                         weighted = TRUE)
+
+cont_pred <- predict(mod_cont)$pred
+cont_cr_lb <- predict(mod_cont)$cr.lb
+cont_cr_ub <- predict(mod_cont)$cr.ub
+cont_var <- (cont_pred - cont_cr_lb) / 1.96
+
+beta_pred <- predict(mod_beta)$pred
+beta_cr_lb <- predict(mod_beta)$cr.lb
+beta_cr_ub <- predict(mod_beta)$cr.ub
+beta_var <- (beta_pred - beta_cr_lb) / 1.96
+
+cont_decrease_pred <- predict(mod_cont_decrease)$pred
+cont_decrease_cr_lb <- predict(mod_cont_decrease)$cr.lb
+cont_decrease_cr_ub <- predict(mod_cont_decrease)$cr.ub
+cont_decrease_var <- (cont_decrease_pred - cont_decrease_cr_lb) / 1.96
+
+beta_decrease_pred <- predict(mod_beta_decrease)$pred
+beta_decrease_cr_lb <- predict(mod_beta_decrease)$cr.lb
+beta_decrease_cr_ub <- predict(mod_beta_decrease)$cr.ub
+beta_decrease_var <- (beta_decrease_pred - beta_decrease_cr_lb) / 1.96
+
+iterations <- 1000
+
+# Uncomment if you want to rerun the inversion method
+# > t1
+# [1] "2016-03-29 13:34:31 CEST"
+# > t2
+# [1] "2016-03-29 15:55:27 CEST"
+# t1 <- Sys.time()
+# source('../functions/inversion_method_uncertain.R')
+# source('../functions/inversion_method_certain.R')
+# t2 <- Sys.time()
+
+load('../data/result_almost_i_uncertain')
+load('../data/result_almost_i_certain')
+
+load('../data/result_almost_iv_uncertain')
+load('../data/result_almost_iv_certain')
+
+res_i_uncertain <- apply(result_almost_i_uncertain, 2, sum) / iterations
+res_i_certain <- apply(result_almost_i_certain, 2, sum) / iterations
+
+par(mar = c(2.5, 5, 0.5, 0.5),
+    mfrow = c(2, 1))
+plot(res_i_uncertain,
+     xlab = '',
+     ylab = 'Proportion',
+     ylim = 0:1,
+     type = 'l',
+     lty = 1)
+lines(res_i_certain, pch = 2, type = 'l', lty = 2)
+abline(h = c(.025, .975), lty = 3)
+
+clip(x1 = 0, x2 = length(res_i_uncertain), y1 = 0.025, y2 = .975)
+abline(v = which(abs(res_i_uncertain - .975) == min(abs(res_i_uncertain - .975))), lty = 1)
+abline(v = which(abs(res_i_certain - .975) == min(abs(res_i_certain - .975))), lty = 2)
+
+res_iv_uncertain <- apply(result_almost_iv_uncertain, 2, sum) / iterations
+res_iv_certain <- apply(result_almost_iv_certain, 2, sum) / iterations
+
+par(mar = c(4, 5, 0.5, 0.5))
+plot(res_iv_uncertain,
+     xlab = 'Data-points fabricated',
+     ylab = 'Proportion',
+     ylim = 0:1,
+     type = 'l',
+     lty = 1)
+lines(res_iv_certain, pch = 2, type = 'l',
+      lty = 2)
+abline(h = c(.025, .975), lty = 3)
+
+clip(x1 = 0, x2 = length(res_iv_uncertain), y1 = 0.025, y2 = .975)
+abline(v = which(abs(res_iv_uncertain - .975) == min(abs(res_iv_uncertain - .975))), lty = 1)
+abline(v = which(abs(res_iv_certain - .975) == min(abs(res_iv_certain - .975))), lty = 2)
+
+```
 

### 2016-04-15 submission/manuscript.Rmd
Interim commit

@@
 
+# load('data/result_almost_i_uncertain')
+# load('data/result_almost_i_certain')
+# 
+# load('data/result_almost_iv_uncertain')
+# load('data/result_almost_iv_certain')
+
 res_i_uncertain <- apply(result_almost_i_uncertain, 2, sum) / iterations
@@
 abline(h = c(.025, .975), lty = 3)
+
+legend(x = 0, y = .5, bty = 'n', lty = c(1, 2), legend = c("Distribution estimate", "Point estimate"))
 
@@
 
+legend(x = 0, y = .55, bty = 'n', lty = c(1, 2), legend = c("Distribution estimate", "Point estimate"))
+
 clip(x1 = 0, x2 = length(res_iv_uncertain), y1 = 0.025, y2 = .975)
@@
 abline(v = which(abs(res_iv_certain - .975) == min(abs(res_iv_certain - .975))), lty = 2)
+```
 
-```
+| -           | Estimate              | Lowerbound  | Upperbound | N    |
+|-------------|-----------------------|-------------|------------|------|
+| DECREASE-I  | Distribution estimate | 0           | 79;85;90   | 113  |
+| -           | Point estimate        | 0           | 108;110    | 113  |
+| DECREASE-IV | Distribution estimate | 0           | 1054       | 1067 |
+| -           | Point estimate        | 0           | 718        | 1067 |
 

### 2016-04-26 submission/manuscript.Rmd
Update manuscript

@@
 author: "Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: word_document
+date: '`r format(Sys.time(), "%d %B, %Y")`'
+output: pdf_document
 csl: ../bibliography/ama.csl
@@
 Upon being discovered as (potentially) fabricated, the DECREASE studies were excluded from a 2014 meta-analysis [@bouri2014].
-The corrected analyses found that conclusions on the effectiveness of beta-blockers in perioperative mortality were reversed --- beta-blockers seemed to actually *increase* perioperative mortality instead of decrease it.
+The meta-analyses restricted to non-DECREASE studies found that conclusions on the effectiveness of beta-blockers in perioperative mortality were reversed --- beta-blockers seemed to actually *increase* perioperative mortality instead of decreasing it.
 
-The committee reports about the DECREASE studies indicate that data fabrication was likely [@commissie2011;@commissie2012;@commissie2013], but the extent of the data fabrication was not clearly indicated or deemed estimable.
+The committee reports about the DECREASE studies indicate that data fabrication was likely [@commissie2011;@commissie2012;@commissie2013], but the extent of the data fabrication remained unclear.
 In this paper, we aim to estimate the extent of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999].
-We expect to confirm the conclusions of the reports that data fabrication is likely.
 
-To investigate the evidence for data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999], we take three steps.
-First, we try to replicate the findings from the 2014 meta-analysis [@bouri2014], which should be possible with the trial-event data provided in this meta-analysis.
-Second, we evaluate the veracity of the DECREASE studies (i.e., the probability that these study findings occurred assuming no data fabrication occurred).
-Third, we reverse this assumption and assume that data fabrication did occur and estimate how many data points would have to be fabricated to reproduce the results of the DECREASE studies.
-We expand on the methods for these three steps below.
+To investigate the evidence for data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999], we took three steps.
+First, we tried to replicate the findings from the 2014 meta-analysis [@bouri2014].
+Second, we evaluated the veracity of the DECREASE studies (i.e., the probability that these study findings occurred assuming no data fabrication).
+Third, we reversed this assumption and assumed that data fabrication did occur, estimating how many data points would have to be fabricated to reproduce the results of the DECREASE studies.
 
@@
 
-In order to ensure we are using similar analytic procedures as in the 2014 meta-analysis [@bouri2014], we initially reproduce the estimates.
+## Methods
+
+To ensure that we were using similar analytic procedures as in the 2014 meta-analysis [@bouri2014], we initially reproduce Bouri et al.'s estimates.
 This ensured that (1) their results are reproducible and (2) we are using the correct estimates in the following steps.
-In order to replicate the results of the meta-analysis, we extract the trial-event data available from figures 2 and 3 in the original paper[@bouri2014].
-More specifically, we extract the raw trial-event data for the 2 (control vs experimental) by 2 (event or no event) design, which we use to recompute the natural logarithm of the risk ratio and its standard error.
-The extracted trial-event data is available at osf.io/aykeh and our analysis plan was preregistered at osf.io/vnmzc.
-
-## Method
+Using figures 2 and 3 from the original paper[@bouri2014], we extracted the raw event data for the 2 (control vs experimental) by 2 (event or no event) design, which we used to recompute the natural logarithm of the risk ratio and its standard error.
+The extracted event data is available at [osf.io/aykeh](osf.io/aykeh) and our analysis plan was preregistered at [osf.io/vnmzc](osf.io/vnmzc).
 
@@
 bouri_data <- read.csv('../data/bouri_data.csv')
+bouri_data <- read.csv('data/bouri_data.csv')
 
@@
 
-We computed the risk ratio (RR) for each study and meta-analyzed the effects.
-We used the `R` package `metafor` [@viechtbauer2010] to meta-analyze the RRs.
+We computed the log risk ratio (i.e., log RR) for each study and pooled these using the `R` package `metafor` [@viechtbauer2010].
 We estimated a weighted random-effects model using the restricted maximum-likelihood estimator (i.e., `REML`)[@viechtbauer2005] to estimate the variance of effects.
-We used the default weighting procedure in the `metafor` package, because the weights provided in the original paper were not accounted for.
-In the case that there was a zero-count for a cell, .5 was added to each cell of the trial [@agresti2002].
-The 2014 meta-analysis [@bouri2014] did not specify the effect variance estimate used, so minor discrepancies between our estimates and the original estimates might originate from these differences in the estimation procedure.
+We used the default weighting procedure in the `metafor` package.
+When there was a zero-count for a cell, .5 was added to each cell of that trial [@agresti2002].
+The 2014 meta-analysis [@bouri2014] did not specify the effect variance estimate used.
+Hence, minor discrepancies between our estimates and the original estimates might be due to differences in the estimation procedure.
 
@@
 First, we analyzed the RRs to verify the estimates in Figure 2 and found that we were able to reproduce the estimates for the different sets of studies.
-The original Figure 2 [@bouri2014] differentiated between the estimates from the non-DECREASE (i.e., secure) trials and the DECREASE (i.e., non-secure trials).
+The original Figure 2 [@bouri2014] differentiated between the estimates from the non-DECREASE trials and the DECREASE trials.
 We confirmed the estimates of the effect size and the variance for both the non-DECREASE- and the DECREASE trials, save for some minor discrepancies due to the estimation method.
@@
 
-Second, we meta-analyzed all studies combined, including a dummy-predictor for the DECREASE and non-DECREASE studies to replicate results presented in Figure 4 of the 2014 meta-analysis.[@bouri2014] Oddly enough our reproduced results deviated more from the asssumption that the subgroups are in fact from the same population; whereas the original test for subgroup differences was $\chi^2(1)=3.91,p=.05$, our reproduced analyses indicated that $\chi^2(1)=`r round(mod_mods$QM, 2)`,p=`r round(mod_mods$QMp, 3)`$.
-Additionally, the original analyses showed substantial heterogeneity ($I^2=74.4$%), but our reproduced analyses showed no such thing ($I^2=`r round(mod_mods$I2, 3)`$%).
-Different estimates of the heterogeneity (e.g., DerSimonian-Laird instead of `REML`) did not change this difference.
+Second, we meta-analyzed all studies combined, including a dummy predictor for the DECREASE and non-DECREASE studies to replicate results presented in Figure 4 of the 2014 meta-analysis.[@bouri2014] Surprisingly, our results showed more evidence against equal subgroups than the original meta-analysis [@bouri2014].
+Whereas the original test for subgroup differences was $\chi^2(1)=3.91,p=.05$, our analyses indicated that $\chi^2(1)=`r round(mod_mods$QM, 2)`,p=`r round(mod_mods$QMp, 3)`$.
+Additionally, the original analyses showed substantial heterogeneity ($I^2=74.4$\%) where we found an $I^2=`r round(mod_mods$I2, 3)`$\%.
+Different estimates of the heterogeneity (e.g., `DerSimonian-Laird` instead of `REML`) did not change this difference.
 
@@
 
-We estimated the probability of the results in the DECREASE studies based on the estimates from the non-DECREASE studies.
-In order to do this, we assumed that the non-DECREASE studies provide the best known representation of the true effect of beta-blockers on perioperative mortality.
-This probability is also known as the veracity of the data [@peters2015], which indicates the probability of the observed data in a random sampling context.
-We assumed to estimate the true population distribution of effects, not perturbed by publication bias due to statistical significance, because a substantial number of nonsignificant effects are included in the dataset.
-To estimate the true population effect we only used the non-DECREASE studies.
+We estimated the probability of obtaining the results in the DECREASE studies based on the estimates from the non-DECREASE studies.
+To this end, we assumed that the non-DECREASE studies provide a valid representation of the true effect of beta-blockers on perioperative mortality.
+The estimated probability is also known as the veracity of the data [@peters2015], which indicates the probability of the observed data under a given true effect.
+We assumed that the non-DECREASE studies estimate the true population distribution of effects, not perturbed by publication bias due to statistical (non)significance.
+Publication bias was assumed to not be a problem because a substantial number of nonsignificant effects are included in the dataset.
 
@@
 Based on the estimated mean log RR and credible interval in the non-DECREASE studies, we computed the probability of log RR in the DECREASE trials.
-The estimates of the non-DECREASE studies were obtained from step 1, which include the estimated log RR (i.e., `r round(mod_clean$b, 2)`), and the 95% credibility interval as provided by the package `metafor` (i.e., `r sprintf('[%s; %s]', round(mod_clean_pred$cr.lb, 3), round(mod_clean_pred$cr.ub, 3))`.
+The estimates of the non-DECREASE studies were obtained from step 1, which include the estimated log RR (i.e., `r round(mod_clean$b, 2)`), and the 95\% credibility interval as provided by the package `metafor` (i.e., `r sprintf('[%s; %s]', round(mod_clean_pred$cr.lb, 3), round(mod_clean_pred$cr.ub, 3))`.
 We assumed a normal distribution of population effects with the estimated effect as the mean of the distribution.
-The 95% credibility interval denotes the bounds of the normal distribution that covers 95% of the density, where the standard deviation is calculated as the distance from the mean to either bound, divided by 1.96.
+The 95\% credibility interval denotes the bounds of the normal distribution that covers 95\% of the density, where the standard deviation is calculated as the distance from the mean to either bound, divided by 1.96.
 This allows for an approximation of the population effect distribution, as depicted in Figure `r fig <- 1; fig`.
@@
 
-*Figure `r fig`*. Density plot of the estimated true effect distribution, with the position of the DECREASE studies highlighted.
+*Figure `r fig`*. Density plot of the estimated true effect distribution based on the non-DECREASE studies, with the position of the DECREASE studies highlighted.
 
-Based on the estimated effect distribution from the non-DECREASE trials, we calculated the probability of each DECREASE trial or a more extreme result.
+Based on the estimated effect distribution from the non-DECREASE trials, we calculated the probability of each DECREASE trial result, or a more extreme result.
 In other words, we computed the *p*-value for the hypothesis that the DECREASE studies occurred based on the information available from the other trials.
@@
 
-As Figure `r fig` already indicates, the DECREASE trials are highly unlikely under the estimated effect distribution based on the non-DECREASE trials.
-More specifically, DECREASE-I (or more extreme) has a probability of `r prob[1]` (less than 1 in a quintillion) and DECREASE-IV `r prob[2]` (3 in a billion), which indicates they are unlikely to come from the same population effect distribution.
-Observing two of such extremely unlikely studies in this population effect distribution is highly improbable, `r prob[1]*prob[2]`.
+Figure `r fig` indicates that the DECREASE trials are highly unlikely under the estimated effect distribution based on the non-DECREASE trials.
+More specifically, DECREASE-I (or more extreme) has a probability of $`r prob[1]`$ (less than 1 in a quintillion) and DECREASE-IV $`r prob[2]`$ (3 in a billion), which indicates they are unlikely to come from the same population effect distribution.
+Observing two of such extremely unlikely results in this population effect distribution is highly improbable, $`r prob[1]*prob[2]`$.
 This would therefore indicate that the DECREASE trials are severely different from the non-DECREASE studies.
 
-We do note that the variance in the non-DECREASE studies was estimated at 0, indicating that there would supposedly be a fixed-effect of beta-blockers on perioperative mortality despite differences in type of beta-blockers administered and dosage/duration of the beta-blockers.
-This homogeneity is not excessive, considering that the *Q*-statistic is sufficiently large to begin with (i.e., `r round(mod_clean$QE, .2)`).[@ioannidis2006] Nonetheless, the original non-DECREASE trials show substantive variation in the treatment, for example the usage of different beta-blockers (i.e., atenolol, propanolol, bisoprolol, metoprolol), which have been argued to affect effectiveness of the interventions REF.
-As such, the lack of heterogeneity can be genuine and indicative of no difference between these beta-blockers and their dosage/duration.
-Nonetheless, it could also be due to uncertainty in the estimate of heterogeneity due to the small set of studies included.
+We note that the variance in the non-DECREASE studies was estimated at 0, indicating that there would supposedly be a fixed-effect of beta-blockers on perioperative mortality despite differences in type of beta-blockers administered and dosage/duration of the beta-blockers.
+This homogeneity is not excessive, considering that the *Q*-statistic is sufficiently large to begin with (i.e., `r round(mod_clean$QE, .2)`).[@ioannidis2006] Nonetheless, the original non-DECREASE trials show substantive variation in the treatment, for example the usage of different beta-blockers (i.e., atenolol, propanolol, bisoprolol, metoprolol) in different (standardized) dosages, which have been argued to affect effectiveness of the interventions.
+As such, the lack of heterogeneity can be genuine and indicative of no difference between these different beta-blockers and their dosage/duration.
+Nonetheless, it could also be due to uncertainty in the estimate of heterogeneity $\tau^2$ due to the small set of studies included.
 
-Considering that the variance (i.e., heterogeneity) estimate is 0 and uncertain due to the small amount of studies included, sensitivity analyses of the results are in place.
+Considering that the variance (i.e., heterogeneity) estimate is 0 and uncertain due to the small number of studies included, sensitivity analyses of the results are in place.
 The probability of observing both DECREASE trials stays approximately below 1 out of 1000 until the variance estimate is .25, and the probability stays approximately below 1 out of 100 until the variance estimate is .43 (see Figure `r fig <- fig + 1; fig`).
@@
 *Figure `r fig`*.
-Sensitivity analyses for the *p*-value that indicates the probability of observing the DECREASE studies, or more extreme, based on the estimated population effect and the accompanying variance estimate.
+Sensitivity analyses for the *p*-value that indicates the probability of observing the results from the DECREASE studies, or more extreme results, based on the estimated true effect (non-DECREASE trials) and the accompanying variance estimate.
 
@@
 
-We estimated the number of data points that would have to be fabricated to arrive at the estimates from the DECREASE studies, given that the non-DECREASE studies represent the population effect.
-This assumes, in contrast to Step 2, that the DECREASE studies were in fact fabricated.
-The estimates from Step 3 provide an indication to the extent of the potential data fabrication in the DECREASE studies.[@commissie2011;@commissie2012;@commissie2013;@bouri2014]
+We estimated the number of data points that would need to be fabricated to arrive at the estimates from the DECREASE studies, given that the non-DECREASE studies represent the true effect.
+This assumes, in contrast to Step 2, that the DECREASE studies in fact contain fabrication data.
+The estimates from Step 3 provide an indication of the extent of the potential data fabrication in the DECREASE studies under this assumption.[@commissie2011;@commissie2012;@commissie2013;@bouri2014]
 
@@
 
-In order to estimate the number of fabricated data points, we meta-analyzed the log odds per condition for both the DECREASE- and non-DECREASE studies separately.
+In order to estimate the number of fabricated data points, we meta-analyzed the log odds of mortality per condition for both the DECREASE- and non-DECREASE studies separately.
 For each of these, we ran a meta-analysis applying the same methods as in Step 1.
@@
 
-We applied the inversion method to estimate the number of data points fabricated in the DECREASE-I and DECREASE-IV studies.
-To this end, we assumed that the DECREASE data are in fact fabricated, opposed to Step 2 where we assumed they were genuine.
-The inversion method iteratively simulates data with *X* fabricated data points; *X* varies from 0 through *N*.
+We applied the inversion method to estimate the number of data points fabricated in the DECREASE-I and DECREASE-IV studies.[@casella2002] The inversion method iteratively simulates data with *X* fabricated data points; *X* varying from 0 through *N*.
 For each *X*, 10,000 iterations were conducted.
@@
 Using the (partially) fabricated dataset, we estimated the risk ratio from the data and noted whether the two-sided *p*-value was larger under the non-DECREASE effect distribution (coded 0) or the DECREASE effect distribution.
+<!--- Marcel's Likelihood approach should be incorporated still; DEVIATION FROM PREREGISTRATION --->
 
@@
 This resulted in a proportion for each *X* from 0 to *N*.
-Those proportions closest to .025 and .975 denoted the 95% confidence interval of data points fabricated.
+Those proportions closest to .025 and .975 denoted the 95\% confidence interval of data points fabricated.
 
@@
 
-| -           | Estimate              | Lowerbound  | Upperbound | N    |
+|            | Estimate              | Lowerbound  | Upperbound | N    |
 |-------------|-----------------------|-------------|------------|------|
 | DECREASE-I  | Distribution estimate | 0           | 79;85;90   | 113  |
-| -           | Point estimate        | 0           | 108;110    | 113  |
+|            | Point estimate        | 0           | 108;110    | 113  |
 | DECREASE-IV | Distribution estimate | 0           | 1054       | 1067 |
-| -           | Point estimate        | 0           | 718        | 1067 |
+|            | Point estimate        | 0           | 718        | 1067 |
 
@@
 
+<!---
+GtR
+1. Is natuurlijk wle gevaarlijk als we in gedachten nemen dat DECREASE op een heel
+andere benadering van perioperatieve blockade is gestoeld.
+Ik begrijp dat een dergeliijk bezwaar het fundament van onze exercitie aantast,
+maar we moeten daar goed over nadenken.
+In discussie kunnen we zeggen:
+Decrease approach is niet wezenlijk anders dan non-decrease enonze analyse geldt
+Decrease appraoch is wezenijk anders qua aanpak en effect en daarom kunnen
+verschillen te wijten zijin aan fraude maar ook aan beter effect en die twee
+kunnen wij niet scheiden. 
+Nieuwe decrease aanpak trials zijj nodig (zie ook paper uit NTvG van Marleen)
+2. In Discussion we must debate the statistical aspects and the content aspects where DECREASE have a very different approach to the prevention of periop mortality than non-DECREASE.
+This makes our own work much more difficult to interpret
+--->
+
 # References

### 2016-04-26 functions/inversion_method_certain.R
Alter inversion method to NON-PREREGISTERED outcome: change to Likelihood comparison (more precise)

@@ -1,11 +1,9 @@
 xs <- 0:(bouri_data$control_n[bouri_data$decrease == 1][1] + 
            bouri_data$beta_n[bouri_data$decrease == 1][1])
 
-result_almost_i_certain <- matrix(nrow = iterations, ncol = length(xs))
+result_almost_i_certain <- rep(0, length(xs))
 
 # Loop for DECREASE-I
-j <- 1
-
 for (x in xs){
   for (iter in 1:iterations){
     # Setup data handling
@@ -52,51 +50,34 @@ for (x in xs){
                                                              size = 1,
                                                              beta_decrease_prop)
     
-    # Estimate the ln(RR) for this sample
-    a <- sum(working_df$condition == 'control' & working_df$event == 0)
-    b <- sum(working_df$condition == 'beta' & working_df$event == 0)
-    c <- sum(working_df$condition == 'control' & working_df$event == 1)
-    d <- sum(working_df$condition == 'beta' & working_df$event == 1)
+    # Calculate the cell counts
+    n_exp_mort <- sum(working_df$event[working_df$condition == 'beta'])
+    n_exp_liv <- sum(working_df$event[working_df$condition == 'beta'] == 0)
+    n_cont_mort <- sum(working_df$event[working_df$condition == 'control'])
+    n_cont_liv <- sum(working_df$event[working_df$condition == 'control'] == 0)
     
-    # Add .5 to studies with zero-count
-    if (0 %in% c(a, b, c, d)){
-      a <- a + .5
-      b <- b + .5
-      c <- c + .5
-      d <- d + .5
-    }
     
-    ln_rr <- log((d/(b+d))/(c/(a+c)))
+    # Likelihood under DECREASE
+    l_fab <- beta_decrease_prop^n_exp_mort *
+      (1 - beta_decrease_prop)^n_exp_liv *
+      cont_decrease_prop^n_cont_mort * (1 - cont_decrease_prop)^n_cont_liv
+    l_tru <- beta_prop^n_exp_mort *
+      (1 - beta_prop)^n_exp_liv *
+      cont_prop^n_cont_mort * (1 - cont_prop)^n_cont_liv
     
-    # Determine whether more likely under DECREASE or non-DECREASE
-    # TRUE indicates more likely under DECREASE
-    # the computation just circumvents having to adjust left- or 
[…]

### 2016-04-26 functions/inversion_method_uncertain.R

@@ -1,11 +1,9 @@
 xs <- 0:(bouri_data$control_n[bouri_data$decrease == 1][1] + 
            bouri_data$beta_n[bouri_data$decrease == 1][1])
 
-result_almost_i_uncertain <- matrix(nrow = iterations, ncol = length(xs))
+result_almost_i_uncertain <- rep(0, length(xs))
 
 # Loop for DECREASE-I
-j <- 1
-
 for (x in xs){
   for (iter in 1:iterations){
     # Setup data handling
@@ -56,51 +54,34 @@ for (x in xs){
                                                              size = 1,
                                                              beta_decrease_prop)
     
-    # Estimate the ln(RR) for this sample
-    a <- sum(working_df$condition == 'control' & working_df$event == 0)
-    b <- sum(working_df$condition == 'beta' & working_df$event == 0)
-    c <- sum(working_df$condition == 'control' & working_df$event == 1)
-    d <- sum(working_df$condition == 'beta' & working_df$event == 1)
+    # Calculate the cell counts
+    n_exp_mort <- sum(working_df$event[working_df$condition == 'beta'])
+    n_exp_liv <- sum(working_df$event[working_df$condition == 'beta'] == 0)
+    n_cont_mort <- sum(working_df$event[working_df$condition == 'control'])
+    n_cont_liv <- sum(working_df$event[working_df$condition == 'control'] == 0)
     
-    # Add .5 to studies with zero-count
-    if (0 %in% c(a, b, c, d)){
-      a <- a + .5
-      b <- b + .5
-      c <- c + .5
-      d <- d + .5
-    }
     
-    ln_rr <- log((d/(b+d))/(c/(a+c)))
+    # Likelihood under DECREASE
+    l_fab <- beta_decrease_prop^n_exp_mort *
+      (1 - beta_decrease_prop)^n_exp_liv *
+      cont_decrease_prop^n_cont_mort * (1 - cont_decrease_prop)^n_cont_liv
+    l_tru <- beta_prop^n_exp_mort *
+      (1 - beta_prop)^n_exp_liv *
+      cont_prop^n_cont_mort * (1 - cont_prop)^n_cont_liv
     
-    # Determine whether more likely under DECREASE or non-DECREASE
-    # TRUE indicates more likely under DECREASE
-    # the computation just circumvents having to adjust left- or 
[…]

### 2016-04-29 submission/manuscript.Rmd
Minor manuscript adjustment
Rerun with Likelihood as endpoint for step 3

@@
 bouri_data <- read.csv('../data/bouri_data.csv')
-bouri_data <- read.csv('data/bouri_data.csv')
+# bouri_data <- read.csv('data/bouri_data.csv')
 
@@
 
-iterations <- 1000
+iterations <- 10000
 
@@
 # source('../functions/inversion_method_uncertain.R')
+# # source('functions/inversion_method_uncertain.R')
+# t2 <- Sys.time()
 # source('../functions/inversion_method_certain.R')
-# t2 <- Sys.time()
+# # source('functions/inversion_method_certain.R')
+# t3 <- Sys.time()
 
@@
 load('../data/result_almost_iv_certain')
-
+# 
 # load('data/result_almost_i_uncertain')

### 2016-04-29 submission/manuscript.Rmd
Add confidence interval plots

@@
 load('../data/result_almost_iv_certain')
-# 
+
 # load('data/result_almost_i_uncertain')
@@
 
-res_i_uncertain <- apply(result_almost_i_uncertain, 2, sum) / iterations
-res_i_certain <- apply(result_almost_i_certain, 2, sum) / iterations
+res_i_uncertain <- result_almost_i_uncertain / iterations
+res_i_certain <- result_almost_i_certain / iterations
+
+res_iv_uncertain <- result_almost_iv_uncertain / iterations
+res_iv_certain <- result_almost_iv_certain / iterations
+
+res_i_uncertain_lb <- NULL
+res_i_uncertain_ub <- NULL
+res_i_certain_lb <- NULL
+res_i_certain_ub <- NULL
+
+res_iv_uncertain_lb <- NULL
+res_iv_uncertain_ub <- NULL
+res_iv_certain_lb <- NULL
+res_iv_certain_ub <- NULL
+
+j <- 1
+for (i in seq(0, 1, .01))
+{
+  lb <- .5 - (i / 2)
+  ub <- .5 + (i / 2)
+
+  res_i_uncertain_lb[j] <- which(abs(res_i_uncertain - lb) == min(abs(res_i_uncertain - lb)))[1]
+  res_i_uncertain_ub[j] <- which(abs(res_i_uncertain - ub) == min(abs(res_i_uncertain - ub)))[1]
+
+  res_i_certain_lb[j] <- which(abs(res_i_certain - lb) == min(abs(res_i_certain - lb)))[1]
+  res_i_certain_ub[j] <- which(abs(res_i_certain - ub) == min(abs(res_i_certain - ub)))[1]
+
+  res_iv_uncertain_lb[j] <- which(abs(res_iv_uncertain - lb) == min(abs(res_iv_uncertain - lb)))[1]
+  res_iv_uncertain_ub[j] <- which(abs(res_iv_uncertain - ub) == min(abs(res_iv_uncertain - ub)))[1]
+
+  res_iv_certain_lb[j] <- which(abs(res_iv_certain - lb) == min(abs(res_iv_certain - lb)))[1]
+  res_iv_certain_ub[j] <- which(abs(res_iv_certain - ub) == min(abs(res_iv_certain - ub)))[1]
+
+  j <- j + 1
+}
+
+# Manually correct some errors in the bound seeking from the loop above
+# errors arise from Monte Carlo error:
+# .022 .021 .023 <-- selects .021, but should select 0
+# considering that running >10000 takes ages and does not guarantee
+# it will fix the problem, I do it like this
+res_i_uncertain_lb <- ifelse(res_i_uncertain_lb == 4, 0, res_i_uncertain_lb)
+res_i_certain_lb <- ifelse(res_i_certain_lb == 1, 0, res_i_certain_lb)
+res_iv_uncertain_lb <- ifelse(res_iv_uncertain_lb == 8, 0, res_iv_uncertain_lb)
+res_iv_certain_lb[98:101] <- 0
+res_iv_certain_ub[98:101] <- 1067
 
 par(mar = c(2.5, 5, 0.5, 0.5),
-    mfrow = c(2, 1))
+    mfrow = c(2, 2))
 plot(res_i_uncertain,
@@
 
-legend(x = 0, y = .5, bty = 'n', lty = c(1, 2), legend = c("Distribution estimate", "Point estimate"))
+legend(x = 0, y = .9, bty = 'n', lty = c(1, 2), legend = c("Distribution estimate", "Point estimate"))
 
-clip(x1 = 0, x2 = length(res_i_uncertain), y1 = 0.025, y2 = .975)
-abline(v = which(abs(res_i_uncertain - .975) == min(abs(res_i_uncertain - .975))), lty = 1)
-abline(v = which(abs(res_i_certain - .975) == min(abs(res_i_certain - .975))), lty = 2)
+plot(x = NA, y = NA,
+     xlim = c(0,1), ylim = c(0, length(res_i_certain)),
+     xlab = "Confidence in %", ylab = "Data points fabricated")
 
-res_iv_uncertain <- apply(result_almost_iv_uncertain, 2, sum) / iterations
-res_iv_certain <- apply(result_almost_iv_certain, 2, sum) / iterations
+lines(x = seq(0, 1, .01), y = res_i_uncertain_lb)
+lines(x = seq(0, 1, .01), y = res_i_uncertain_ub)
+lines(x = seq(0, 1, .01), y = res_i_certain_lb, lty = 2, pch = 2, type = 'l')
+lines(x = seq(0, 1, .01), y = res_i_certain_ub, lty = 2, pch = 2, type = 'l')
+abline(v = .95, lty = 3)
+
+# clip(x1 = 0, x2 = length(res_i_uncertain), y1 = 0.025, y2 = .975)
+# abline(v = which(abs(res_i_uncertain - .975) == min(abs(res_i_uncertain - .975))), lty = 1)
+# abline(v = which(abs(res_i_certain - .975) == min(abs(res_i_certain - .975))), lty = 2)
 
@@
 
-legend(x = 0, y = .55, bty = 'n', lty = c(1, 2), legend = c("Distribution estimate", "Point estimate"))
+plot(x = NA, y = NA,
+     xlim = c(0,1), ylim = c(0, length(res_iv_certain)),
+     xlab = "Confidence in %", ylab = "Data points fabricated")
 
-clip(x1 = 0, x2 = length(res_iv_uncertain), y1 = 0.025, y2 = .975)
-abline(v = which(abs(res_iv_uncertain - .975) == min(abs(res_iv_uncertain - .975))), lty = 1)
-abline(v = which(abs(res_iv_certain - .975) == min(abs(res_iv_certain - .975))), lty = 2)
+lines(x = seq(0, 1, .01), y = res_iv_uncertain_lb)
+lines(x = seq(0, 1, .01), y = res_iv_uncertain_ub)
+lines(x = seq(0, 1, .01), y = res_iv_certain_lb, lty = 2, pch = 2, type = 'l')
+lines(x = seq(0, 1, .01), y = res_iv_certain_ub, lty = 2, pch = 2, type = 'l')
+abline(v = .95, lty = 3)
 ```
-
-|            | Estimate              | Lowerbound  | Upperbound | N    |
-|-------------|-----------------------|-------------|------------|------|
-| DECREASE-I  | Distribution estimate | 0           | 79;85;90   | 113  |
-|            | Point estimate        | 0           | 108;110    | 113  |
-| DECREASE-IV | Distribution estimate | 0           | 1054       | 1067 |
-|            | Point estimate        | 0           | 718        | 1067 |
+*Figure `r fig <- fig + 1; fig`*.
 

### 2016-07-28 submission/manuscript.Rmd
Minor correction in code.

@@
 ```{r, echo = FALSE}
-suppressPackageStartupMessages(if(!require(metafor)){install.package('metafor')})
+suppressPackageStartupMessages(if(!require(metafor)){install.packages('metafor')})
 suppressPackageStartupMessages(library(metafor))
-suppressPackageStartupMessages(if(!require(latex2exp)){install.package('latex2exp')})
+suppressPackageStartupMessages(if(!require(latex2exp)){install.packages('latex2exp')})
 suppressPackageStartupMessages(library(latex2exp))

### 2017-01-26 submission/manuscript.Rmd
Revise up to step 3 methods

--- 2016-04-29 submission/manuscript.Rmd
@@
 date: '`r format(Sys.time(), "%d %B, %Y")`'
-output: pdf_document
+output:
+  pdf_document:
+    keep_tex: yes
+  word_document: default
 csl: ../bibliography/ama.csl
@@
 
-The effectiveness of beta-blockers in preventing perioperative mortatility in non-cardiac surgery has been subject of discussion due to findings of misconduct.
-[@commissie2011;@commissie2012;@commissie2013;@bouri2014].
-Perioperative mortality is the deathrate of patients during the period of the surgical procedure, typically including admission, anaesthesia, surgery, and recovery.
+<!-- Introduction
+What should be in here?
+1. Effectiveness of beta-blockers (decrease probability))
+2. Effects of findings of misconduct
+3. Why this study. -->
+The effect of beta-blockers on perioperative mortality in non-cardiac surgery for patients with coronary artery disease (CAD) has been subject of discussion [@Coleg5210] due to findings of research  misconduct [@commissie2011;@commissie2012;@commissie2013] that were central to claims of their effectiveness [@Devereaux313;@Angeli2010;@bouri2014].
+Perioperative mortality is the deathrate of patients during the period of the surgical procedure, which typically includes admission, anaesthesia, surgery, and recovery.
+This related finding of research misconduct altered meta-analytic results of perioperative mortality in non-cardiac surgery for CAD patients substantively[@bouri2014], where meta-analyses that include these studies conclude that beta-blockers decrease perioperative mortality [@Devereaux313;@Angeli2010;@bouri2014] but a meta-analysis that excludes these studies concludes that beta-blockers increase perioperative mortality[@bouri2014].
 
-The trials that were found to be suspect were the DECREASE trials [@commissie2011;@commissie2012;@commissie2013].
-Upon being discovered as (potentially) fabricated, the DECREASE studies were excluded from a 2014 meta-analysis [@bouri2014].
-The meta-analyses restricted to non-DECREASE studies found that conclusions on the effectiveness of beta-blockers in perioperative mortality were reversed --- beta-blockers seemed to actually *increase* perioperative mortality instead of decreasing it.
+The trials deemed to be the result of research misconduct were the Dutch DECREASE trials[@commissie2011;@commissie2012;@commissie2013].
+The committees that investigated the integrity of the DECREASE studies reported that data fabrication was likely [@commissie2011;@commissie2012;@commissie2013], but that the extent of the data fabrication remained unclear.
+Given that the inclusion or exclusion of these trials in meta-analyses is crucial in determining whether beta-blockers are indeed effective during the perioperative period and the latest guidelines still recommend the usage of beta-blockers [@esc2014], we aim to estimate the extent of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999] in order to further this debate.
 
-The committee reports about the DECREASE studies indicate that data fabrication was likely [@commissie2011;@commissie2012;@commissie2013], but the extent of the data fabrication remained unclear.
-In this paper, we aim to estimate the extent of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999].
+The reports on the DECREASE trials primarily focused on the provenance of the raw data and patient files (e.g., informed consent, whether data corresponded to patient files), but neglected the extent to which the DECREASE studies deviated from similar trials that tested the effectiveness of beta-blockers on perioperative mortality.
+Comparing how (dis)similar trials are from comparable studies is not a new idea [@Buyse1999-jq] and has proven effective in detecting data fabrication [@Knepper2016-la].
+This is a different approach from the forensic statistical methods previously applied [@commissie2013], where results within one study are evaluated as opposed to other similar studies.
+This explains the discrepancy between the feasibility of evaluating the DECREASE trials by the statistical expert of the committee report and ours.
 
-To investigate the evidence for data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999], we took three steps.
-First, we tried to replicate the findings from the 2014 meta-analysis [@bouri2014].
-Second, we evaluated the veracity of the DECREASE studies (i.e., the probability that these study findings occurred assuming no data fabrication).
-Third, we reversed this assumption and assumed that data fabrication did occur, estimating how many data points would have to be fabricated to reproduce the results of the DECREASE studies.
+<!-- Hier evt nog wat vertellen over FOI? -->
+<!-- Finally, we share our unsuccesful experiences in trying to further investigate the issue by requesting the DECREASE-VI raw data and call for more transparency. -->
+To statistically investigate the evidence of data fabrication in the results of the DECREASE studies [@dunkelgrun2009;@poldermans1999], we took three steps.
+First, we replicate the findings from a 2014 meta-analysis [@bouri2014] that contains sufficient information to estimate the deviation of the DECREASE-I and DECREASE-IV studies from the remaining studies.
+Second, we evaluate the veracity of the DECREASE-I and DECREASE-IV studies (i.e., the probability that these study findings occurred assuming no data fabrication occurred).
+Third, we reverse this assumption and assume that data fabrication did occur and estimate how many data points would have to be fabricated to reproduce the results of the DECREASE-I and DECREASE-IV studies if the remaining studies are regarded as estimating the true effect of beta-blockers on perioperative mortality.
 
-# Step 1: reproducing Bouri et al. (2014)
+# Step 1: reproducing meta-analysis of Bouri et al. (2014)
 
@@
 
-To ensure that we were using similar analytic procedures as in the 2014 meta-analysis [@bouri2014], we initially reproduce Bouri et al.'s estimates.
-This ensured that (1) their results are reproducible and (2) we are using the correct estimates in the following steps.
-Using figures 2 and 3 from the original paper[@bouri2014], we extracted the raw event data for the 2 (control vs experimental) by 2 (event or no event) design, which we used to recompute the natural logarithm of the risk ratio and its standard error.
-The extracted event data is available at [osf.io/aykeh](osf.io/aykeh) and our analysis plan was preregistered at [osf.io/vnmzc](osf.io/vnmzc).
+To ensure that we used similar analysis procedures as in the 2014 meta-analysis [@bouri2014], we initially reproduced Bouri et al.'s estimates.
+This ensured that (1) their results are reproducible and (2) we are using the correct estimates in subsequent steps of our analyses.
+Using figures 2 and 3 from the original paper [@bouri2014], we extracted the raw event data for the 2 (control vs experimental) by 2 (event or no event) design, which we used to recompute the natural logarithm of the risk ratio and its standard error.
+The extracted event data is available at [osf.io/aykeh](https://osf.io/aykeh) and our analysis plan was preregistered at [osf.io/vnmzc](https://osf.io/vnmzc).
 
 ```{r, echo = FALSE}
-suppressPackageStartupMessages(if(!require(metafor)){install.package('metafor')})
+suppressPackageStartupMessages(if(!require(metafor)){install.packages('metafor', repos = "http://cran.us.r-project.org")})
 suppressPackageStartupMessages(library(metafor))
-suppressPackageStartupMessages(if(!require(latex2exp)){install.package('latex2exp')})
+suppressPackageStartupMessages(if(!require(latex2exp)){install.packages('latex2exp', repos = "http://cran.us.r-project.org")})
 suppressPackageStartupMessages(library(latex2exp))
+suppressPackageStartupMessages(library(knitr))
 
-bouri_data <- read.csv('../data/bouri_data.csv')
-# bouri_data <- read.csv('data/bouri_data.csv')
+x12 <- '../'
+# x12 <- ''
+bouri_data <- read.csv(sprintf('%sdata/bouri_data.csv', x12))
 
@@
 We computed the log risk ratio (i.e., log RR) for each study and pooled these using the `R` package `metafor` [@viechtbauer2010].
-We estimated a weighted random-effects model using the restricted maximum-likelihood estimator (i.e., `REML`)[@viechtbauer2005] to estimate the variance of effects.
+We estimated a weighted random-effects model using the restricted maximum-likelihood estimator (i.e., `REML`) [@viechtbauer2005] to estimate the variance of effects.
 We used the default weighting procedure in the `metafor` package.
-When there was a zero-count for a cell, .5 was added to each cell of that trial [@agresti2002].
-The 2014 meta-analysis [@bouri2014] did not specify the effect variance estimate used.
-Hence, minor discrepancies between our estimates and the original estimates might be due to differences in the estimation procedure.
+When there was a zero-count for a cell, .5 was added to each cell of that trial, as is common in meta-analyses on risk- and odds ratios in order to deal with computational artefacts [@agresti2002].
+The 2014 meta-analysis [@bouri2014] did not specify the effect variance estimate used; hence, minor discrepancies between our estimates and the original estimates could be due to differences in the estimation procedure.
 
@@
 
-First, we analyzed the RRs to verify the estimates in Figure 2 and found that we were able to reproduce the estimates for the different sets of studies.
-The original Figure 2 [@bouri2014] differentiated between the estimates from the non-DECREASE trials and the DECREASE trials.
-We confirmed the estimates of the effect size and the variance for both the non-DECREASE- and the DECREASE trials, save for some minor discrepancies due to the estimation method.
-Table `r tab <- 1; tab` depicts the original and reproduced values for both sets of studies.
+We analyzed the RRs to verify the original estimates (Figure 2 of the 2014 meta-analysis [@bouri2014]) and found that we were able to reproduce the estimates for the different sets of studies.
+Bouri et al.
+differentiated between the estimates from the non-DECREASE trials and the DECREASE trials.
+We confirmed the effect size estimates and the variance estimates for both the non-DECREASE- and the DECREASE trials, save for some minor discrepancies due to the estimation method.
+Table 1 depicts the original and reproduced values for both sets of studies.
 
-*Table `r tab`*. The original- and reproduced meta-analytic results based on the data provided in the 2014 meta-analysis.[@bouri2014]
+```{r echo=FALSE,results='asis'}
+table1 <- matrix('', ncol = 5, nrow = 5)
+table1[1,3] <- 'Risk ratio'
+table1[1,4] <- '$\\tau^2$'
+table1[1,5] <- 'Confidence interval'
+table1[2,1] <- 'Non-DECREASE'
+table1[4,1] <- 'DECREASE'
+table1[c(2,4),2] <- 'Original'
+table1[c(3,5),2] <- 'Reproduced'
+table1[2,3] <- 1.27
+table1[2,4] <- 0.00
+table1[2,5] <- '[1.01; 1.60]'
 
-|              |            | Risk ratio   | $\tau^2$      | Confidence interval |
-|--------------|------------|---------------------|-------------------|---|
-| Non-DECREASE | Original   | 1.27                   | 0.00                 | [1.01; 1.60] |
-|              | Reproduced | `r round(exp(mod_clean$b), 2)`  | `r round(mod_clean$tau2, 2)`                 | `r sprintf('[%s; %s]', round(exp(mod_clean$ci.lb), 2), round(exp(mod_clean$ci.ub), 2))` |
-| DECREASE     | Original   | 0.42                   | 0.29                 | [0.15; 1.23] |
-|              | Reproduced | `r round(exp(mod_dirty$b), 2)` | `r round(mod_dirty$tau2, 2)` | `r sprintf('[%s; %s]', round(exp(mod_dirty$ci.lb), 2), round(exp(mod_dirty$ci.ub), 2))` |
+table1[3,3] <- round(exp(mod_clean$b), 2)
+table1[3,4] <- round(mod_clean$tau2, 2)
+table1[3,5] <- sprintf('[%s; %s]', round(exp(mod_clean$ci.lb), 2), round(exp(mod_clean$ci.ub), 2))
+
+table1[4,3] <- 0.42
+table1[4,4] <- 0.29
+table1[4,5] <- '[0.15; 1.23]'
+
+table1[5,3] <- round(exp(mod_dirty$b), 2)
+table1[5,4] <- round(mod_dirty$tau2, 2)
+table1[5,5] <- sprintf('[%s; %s]', round(exp(mod_dirty$ci.lb), 2), round(exp(mod_dirty$ci.ub), 2))
+
+kable(table1, caption = 'The original- and reproduced meta-analytic results based on the data provided in the 2014 meta-analysis by Bouri et al.')
+```
 
@@
 
-Second, we meta-analyzed all studies combined, including a dummy predictor for the DECREASE and non-DECREASE studies to replicate results presented in Figure 4 of the 2014 meta-analysis.[@bouri2014] Surprisingly, our results showed more evidence against equal subgroups than the original meta-analysis [@bouri2014].
-Whereas the original test for subgroup differences was $\chi^2(1)=3.91,p=.05$, our analyses indicated that $\chi^2(1)=`r round(mod_mods$QM, 2)`,p=`r round(mod_mods$QMp, 3)`$.
+Second, we meta-analyzed all studies combined, including a dummy predictor for the DECREASE and non-DECREASE studies to replicate results presented in Figure 4 of the 2014 meta-analysis [@bouri2014].
+Surprisingly, our results showed more evidence for unequal subgroups than the original meta-analysis [@bouri2014]; whereas the original test for subgroup differences was $\chi^2(1)=3.91,p=.05$, our analyses indicated that $\chi^2(1)=`r round(mod_mods$QM, 2)`,p=`r round(mod_mods$QMp, 3)`$.
 Additionally, the original analyses showed substantial heterogeneity ($I^2=74.4$\%) where we found an $I^2=`r round(mod_mods$I2, 3)`$\%.
-Different estimates of the heterogeneity (e.g., `DerSimonian-Laird` instead of `REML`) did not change this difference.
+Different heterogeneity estimators (e.g., `DerSimonian-Laird` instead of `REML`) did not resolve this difference.
+We tried to clarify these discrepancies by e-mailing the original authors, but failed to receive a response.
 
-<!--- Emailed bouri about this on 2016 03 16--->
+<!--- Emailed bouri about this on 2016 03 16 -->
 
@@
 
-We estimated the probability of obtaining the results in the DECREASE studies based on the estimates from the non-DECREASE studies.
-To this end, we assumed that the non-DECREASE studies provide a valid representation of the true effect of beta-blockers on perioperative mortality.
+Based on the non-DECREASE estimates from Step 1, we estimated the probability of obtaining the results in the DECREASE studies.
+To this end, we assumed that the non-DECREASE studies provide a valid representation of the true effect of beta-blockers on perioperative mortality (similar to Bouri et al.
+[@bouri2014]; but see the discussion).
 The estimated probability is also known as the veracity of the data [@peters2015], which indicates the probability of the observed data under a given true effect.
 We assumed that the non-DECREASE studies estimate the true population distribution of effects, not perturbed by publication bias due to statistical (non)significance.
-Publication bias was assumed to not be a problem because a substantial number of nonsignificant effects are included in the dataset.
+Publication bias was assumed to not be a problem because a substantial number of nonsignificant effects are included in the dataset (9 of 11 nonsignificant).
 
@@
 
-Based on the estimated mean log RR and credible interval in the non-DECREASE studies, we computed the probability of log RR in the DECREASE trials.
-The estimates of the non-DECREASE studies were obtained from step 1, which include the estimated log RR (i.e., `r round(mod_clean$b, 2)`), and the 95\% credibility interval as provided by the package `metafor` (i.e., `r sprintf('[%s; %s]', round(mod_clean_pred$cr.lb, 3), round(mod_clean_pred$cr.ub, 3))`.
+Based on the estimated mean log RR and its credible interval in the non-DECREASE studies, we computed the probability of log RR in the DECREASE trials.
+The estimates of the non-DECREASE studies were obtained from Step 1, which include the estimated log RR (i.e., `r round(mod_clean$b, 2)`), and its 95\% credibility interval as provided by the package `metafor` (i.e., `r sprintf('[%s; %s]', round(mod_clean_pred$cr.lb, 3), round(mod_clean_pred$cr.ub, 3))`.
 We assumed a normal distribution of population effects with the estimated effect as the mean of the distribution.
@@
 
-# pdf('../figures/fig1', width = 8, height = 6)
-# use http://pdf2png.com/ when preparing manuscript
-# for high quality plots
+pdf(sprintf('%sfigures/fig1.pdf', x12), width = 8, height = 6)
 par(mar = c(.5,0,0,0))
@@
 
-# dev.off()
+invisible(dev.off())
 ```
 
-*Figure `r fig`*. Density plot of the estimated true effect distribution based on the non-DECREASE studies, with the position of the DECREASE studies highlighted.
+```{r figure 1, fig.cap = 'Density plot of the estimated true effect distribution based on the non-DECREASE studies only, with the position of the DECREASE studies highlighted.', out.width = '80%', fig.asp = .5, fig.align = 'center', echo = FALSE}
+knitr::include_graphics(sprintf("%sfigures/fig1.pdf", x12))
+```
 
@@
 Figure `r fig` indicates that the DECREASE trials are highly unlikely under the estimated effect distribution based on the non-DECREASE trials.
-More specifically, DECREASE-I (or more extreme) has a probability of $`r prob[1]`$ (less than 1 in a quintillion) and DECREASE-IV $`r prob[2]`$ (3 in a billion), which indicates they are unlikely to come from the same population effect distribution.
-Observing two of such extremely unlikely results in this population effect distribution is highly improbable, $`r prob[1]*prob[2]`$.
+More specifically, DECREASE-I (or more extreme) has a probability of $`r prob[1]`$ (less than 1 in a quintillion) and DECREASE-IV $`r prob[2]`$ (3 in a billion), which indicates they are unlikely to come from the same population effect distribution as the non-DECREASE trials.
+Observing two of such extremely unlikely results in this population effect distribution jointly is highly improbable, $`r prob[1]*prob[2]`$.
 This would therefore indicate that the DECREASE trials are severely different from the non-DECREASE studies.
 
-We note that the variance in the non-DECREASE studies was estimated at 0, indicating that there would supposedly be a fixed-effect of beta-blockers on perioperative mortality despite differences in type of beta-blockers administered and dosage/duration of the beta-blockers.
-This homogeneity is not excessive, considering that the *Q*-statistic is sufficiently large to begin with (i.e., `r round(mod_clean$QE, .2)`).[@ioannidis2006] Nonetheless, the original non-DECREASE trials show substantive variation in the treatment, for example the usage of different beta-blockers (i.e., atenolol, propanolol, bisoprolol, metoprolol) in different (standardized) dosages, which have been argued to affect effectiveness of the interventions.
-As such, the lack of heterogeneity can be genuine and indicative of no difference between these different beta-blockers and their dosage/duration.
-Nonetheless, it could also be due to uncertainty in the estimate of heterogeneity $\tau^2$ due to the small set of studies included.
-
-Considering that the variance (i.e., heterogeneity) estimate is 0 and uncertain due to the small number of studies included, sensitivity analyses of the results are in place.
+Results from Step 1 indicated that no variance (i.e., homogeneity; $\tau^2=0$) of the effects was observed; given the small number of studies included (i.e., `r mod_clean$k`) this estimate is uncertain, however.
+We conducted sensitivity analyses in order to see how dependent results are on the specific heterogeneity estimate.
 The probability of observing both DECREASE trials stays approximately below 1 out of 1000 until the variance estimate is .25, and the probability stays approximately below 1 out of 100 until the variance estimate is .43 (see Figure `r fig <- fig + 1; fig`).
-Further research would be worthwhile to better estimate the variance in the effects of beta-blockers on perioperative mortality.
+Further research would be worthwhile to better estimate the variance in the effects of beta-blockers on perioperative mortality and what moderators affect the effect size (e.g., type of beta-blocker).
 
 ```{r, echo = FALSE}
+pdf(sprintf('%sfigures/fig2.pdf', x12), width = 8, height = 6)
 tau2 <- seq(0, 2, .01)
@@
      cex = .8)
+invisible(dev.off())
 ```
-
-*Figure `r fig`*.
-Sensitivity analyses for the *p*-value that indicates the probability of observing the results from the DECREASE studies, or more extreme results, based on the estimated true effect (non-DECREASE trials) and the accompanying variance estimate.
+```{r figure 2, fig.cap = 'Sensitivity analyses for the p-value that indicates the probability of observing the results from the DECREASE studies, or more extreme results, based on the estimated true effect (non-DECREASE trials) and the accompanying variance estimate.', out.width = '80%', fig.asp = .5, fig.align = 'center', echo = FALSE}
+knitr::include_graphics(sprintf("%sfigures/fig2.pdf", x12))
+```
 
@@
 We estimated the number of data points that would need to be fabricated to arrive at the estimates from the DECREASE studies, given that the non-DECREASE studies represent the true effect.
-This assumes, in contrast to Step 2, that the DECREASE studies in fact contain fabrication data.
-The estimates from Step 3 provide an indication of the extent of the potential data fabrication in the DECREASE studies under this assumption.[@commissie2011;@commissie2012;@commissie2013;@bouri2014]
+This assumes, in contrast to Step 2, that the DECREASE studies in fact contain fabricated data.
+The estimates from Step 3 provide an indication of the extent of the potential data fabrication in the DECREASE studies under this assumption [@commissie2011;@commissie2012;@commissie2013;@bouri2014].
 
@@
 
-In order to estimate the number of fabricated data points, we meta-analyzed the log odds of mortality per condition for both the DECREASE- and non-DECREASE studies separately.
-For each of these, we ran a meta-analysis applying the same methods as in Step 1.
-These separate analyses were necessary, because Step 1 only provided estimates of the overall risk ratio, but not of the odds conditional on exposure (beta-blocker or control) and study set (DECREASE- versus non-DECREASE).
+In order to estimate the number of fabricated data points, we estimated the population effect distribution of mortality (in log odds) per condition in both the DECREASE- and non-DECREASE studies separately.
+For each of these, we ran a meta-analysis applying the same, verified methods used in Step 1.
+These separate analyses were necessary, considering that Step 1 only provided estimates of the overall risk ratio but not the odds conditional on exposure (beta-blocker or control) and study set (DECREASE- versus non-DECREASE).
+This resulted in four meta-analytic mortality estimates with an accompanying variance.
+Throughout the simulations, we use the point estimates to simulate genuine- and fabricated data, but supplement this by using the distribution estimates as sensitivity analyses.
 
-We applied the inversion method to estimate the number of data points fabricated in the DECREASE-I and DECREASE-IV studies.[@casella2002] The inversion method iteratively simulates data with *X* fabricated data points; *X* varying from 0 through *N*.
-For each *X*, 10,000 iterations were conducted.
-Data fabrication occurred using a binomial distribution, with the population effect from the DECREASE studies serving as the probability of an event occurring.
-The population effect was either drawn from the estimated population distribution, which introduced uncertainty (i.e., uncertain estimate), or was equated to the point estimate (i.e., certain estimate; not preregistered).
-The remaining genuine data, *N - X*, were simulated in a similar fashion, where we only used the effect distribution from the non-DECREASE studies.
-Using the (partially) fabricated dataset, we estimated the risk ratio from the data and noted whether the two-sided *p*-value was larger under the non-DECREASE effect distribution (coded 0) or the DECREASE effect distribution.
-<!--- Marcel's Likelihood approach should be incorporated still; DEVIATION FROM PREREGISTRATION --->
+We applied the inversion method to estimate the number of fabricated data points in the DECREASE-I and DECREASE-IV studies [@casella2002].
+With the inversion method, we iteratively hypothesized that *X* out of *N* data points were fabricated (i.e., $0, 1, ..., N$) in the DECREASE-I and DECREASE-IV studies separately.
+For each study, we simulated 10,000 datasets per hypothesis of *X* fabricated data points and *N-X* genuine data points.
+For each simulated dataset (exact simulation procedure in next paragraph), we determined the likelihood of the results with
+\begin{equation}
+\label{eq1}
+L(\theta|\pi_{E},\pi_{C})=\pi_{E}^{n_{11}}(1-\pi_{E})^{n_{12}} \times \pi_{C}^{n_{21}}(1 - \pi_{C})^{n_{22}}
+\end{equation}
+where $\pi_{E}$ indicates the mortality rate of the experimental condition as drawn from the meta-analytic effect distribution ($\pi_{C}$ indicates the same for the control condition).
+The likelihood was computed under both the fabricated effect estimates (i.e., $L_{fabricated}$) and the genuine data (i.e., $L_{genuine}$).
+Table 2 indicates which cell sizes the various $n_{XX}$ refer to within the (simulated) data.
+We compared the likelihoods to determine whether the simulated data were more likely to arise from the genuine- ($L_{genuine}$) or from the fabricated data ($L_{fabricated}$).
+Note that using the likelihoods for this is a minor deviation from the preregistration, where we initially planned on using $p$-value comparisons ([osf.io/vnmzc](https://osf.io/vnmzc)).
 
-Based on the results of the inversion method, we obtain a confidence interval of data points fabricated in the DECREASE studies.
-After running all iterations for one *X*, the proportion of trials that were more likely under the DECREASE population effect than under the non-DECREASE population effect were computed.
-This resulted in a proportion for each *X* from 0 to *N*.
-Those proportions closest to .025 and .975 denoted the 95\% confidence interval of data points fabricated.
+```{r echo=FALSE,results='asis'}
+table2 <- matrix('', ncol = 3, nrow = 3)
+table2[1,1] <- ''
+table2[1,2] <- 'Dead'
+table2[1,3] <- 'Alive'
+
+table2[2,1] <- 'Beta-blockers'
+table2[2,2] <- '$n_{11}$'
+table2[2,3] <- '$n_{12}$'
+
+table2[3,1] <- 'Control'
+table2[3,2] <- '$n_{21}$'
+table2[3,3] <- '$n_{22}$'
+
+kable(table2, caption = 'Outcome possibilities within a simulated 2 (beta-blocker v control) by 2 (dead v alive) clinical trial.')
+```
+
+For each hypothesis of *X* out of *N* fabricated datapoints, we computed the probability that the fabricated data are more likely than the genuine data ($p_F=P(L_{fabricated}>L_{genuine})$).
+Based on $p_F$, we computed the confidence interval for $X$ (i.e., $X_{LB};X_{UB}$).
+For a 95\% confidence interval, the lowerbound is equal to the $p_F$ closest to .025, whereas the upperbound is equal to the $p_F$ closest to .975.
+
+We computed $p_F$ for all possibilities of *X* out of *N* fabricated datapoints in 10,000 randomly generated datsets, which were generated in three steps. For each dataset we:
+
+1. Sampled (without replacement) *X* fictitious participants that would be the result of data fabrication (sampled across all conditions).
+
+2.
+Determined the population mortality rate for each condition (i.e., for each cell as in Table 3).
+The meta-analytic point estimate was used or a population effect was randomly drawn from the meta-analytic effect distribution.
+
+3. Simulated the number of deaths for the different conditions using a binomial distribution based on the mortality rate as determined in 2, resulting in the cell counts as in Table 3.
+
+Based on the meta-analytic effect from 2 and the cell sizes from 3, we computed the  likelihoods of $L_{fabricated}$ and $L_{genuine}$ using Equation \ref{eq1}.
+As mentioned before, $p_F$ was computed as the probability that the data were more likely under the estimates resulting from the (allegedly) fabricated data (i.e., the DECREASE trials) than under the estimates resulting from the genuine data (i.e., the non-DECREASE trials; $p_F=P(L_{fabricated}>L_{genuine})$).
 
@@
 
-par(mar = c(2.5, 5, 0.5, 0.5),
+# pdf(sprintf('%sfigures/fig3.pdf', x12), width = 8, height = 6)
+# 
+# par(mar = c(2.5, 5, 0.5, 0.5),
+#     mfrow = c(2, 2))
+# plot(res_i_uncertain,
+#      xlab = '',
+#      ylab = 'Proportion',
+#      ylim = 0:1,
+#      type = 'l',
+#      lty = 1)
+# lines(res_i_certain, pch = 2, type = 'l', lty = 2)
+# abline(h = c(.025, .975), lty = 3)
+# 
+# legend(x = 0, y = .9, bty = 'n', lty = c(1, 2), legend = c("Distribution estimate", "Point estimate"))
+# 
+# plot(x = NA, y = NA,
+#      xlim = c(0,1), ylim = c(0, length(res_i_certain)),
+#      xlab = "Confidence in %", ylab = "Data points fabricated")
+# 
+# lines(x = seq(0, 1, .01), y = res_i_uncertain_lb)
+# lines(x = seq(0, 1, .01), y = res_i_uncertain_ub)
+# lines(x = seq(0, 1, .01), y = res_i_certain_lb, lty = 2, pch = 2, type = 'l')
+# lines(x = seq(0, 1, .01), y = res_i_certain_ub, lty = 2, pch = 2, type = 'l')
+# abline(v = .95, lty = 3)
+# 
+# # clip(x1 = 0, x2 = length(res_i_uncertain), y1 = 0.025, y2 = .975)
+# # abline(v = which(abs(res_i_uncertain - .975) == min(abs(res_i_uncertain - .975))), lty = 1)
+# # abline(v = which(abs(res_i_certain - .975) == min(abs(res_i_certain - .975))), lty = 2)
+# 
+# par(mar = c(4, 5, 0.5, 0.5))
+# plot(res_iv_uncertain,
+#      xlab = 'Data-points fabricated',
+#      ylab = 'Proportion',
+#      ylim = 0:1,
+#      type = 'l',
+#      lty = 1)
+# lines(res_iv_certain, pch = 2, type = 'l',
+#       lty = 2)
+# abline(h = c(.025, .975), lty = 3)
+# 
+# plot(x = NA, y = NA,
+#      xlim = c(0,1), ylim = c(0, length(res_iv_certain)),
+#      xlab = "Confidence in %", ylab = "Data points fabricated")
+# 
+# lines(x = seq(0, 1, .01), y = res_iv_uncertain_lb)
+# lines(x = seq(0, 1, .01), y = res_iv_uncertain_ub)
+# lines(x = seq(0, 1, .01), y = res_iv_certain_lb, lty = 2, pch = 2, type = 'l')
+# lines(x = seq(0, 1, .01), y = res_iv_certain_ub, lty = 2, pch = 2, type = 'l')
+# abline(v = .95, lty = 3)
+# invisible(dev.off())
+
+pdf(sprintf('%sfigures/fig3.pdf', x12), width = 8, height = 6)
+
+par(mar = c(5, 5, 2.5, 0),
     mfrow = c(2, 2))
 plot(res_i_uncertain,
-     xlab = '',
+     xlab = 'Data points fabricated',
      ylab = 'Proportion',
+     main = 'DECREASE I',
      ylim = 0:1,
@@
 legend(x = 0, y = .9, bty = 'n', lty = c(1, 2), legend = c("Distribution estimate", "Point estimate"))
+
+par(mar = c(5, 0, 2.5, 0.5))
+
+plot(res_iv_uncertain,
+     xlab = 'Data-points fabricated',
+     ylab = '',
+     main = 'DECREASE IV',
+     ylim = 0:1,
+     yaxt = 'n',
+     type = 'l',
+     lty = 1)
+lines(res_iv_certain, pch = 2, type = 'l',
+      lty = 2)
+abline(h = c(.025, .975), lty = 3)
+
+par(mar = c(4, 5, 0.5, 0))
 
@@
 
-par(mar = c(4, 5, 0.5, 0.5))
-plot(res_iv_uncertain,
-     xlab = 'Data-points fabricated',
-     ylab = 'Proportion',
-     ylim = 0:1,
-     type = 'l',
-     lty = 1)
-lines(res_iv_certain, pch = 2, type = 'l',
-      lty = 2)
-abline(h = c(.025, .975), lty = 3)
+par(mar = c(4, 0, 0.5, 0.5))
 
@@
      xlim = c(0,1), ylim = c(0, length(res_iv_certain)),
-     xlab = "Confidence in %", ylab = "Data points fabricated")
+     xlab = "Confidence in %", ylab = "",
+     yaxt = 'n')
 
@@
 abline(v = .95, lty = 3)
+invisible(dev.off())
 ```
-*Figure `r fig <- fig + 1; fig`*.
+
+```{r figure 3, fig.cap = 'Test', out.width = '80%', fig.asp = .5, fig.align = 'center', echo = FALSE}
+knitr::include_graphics(sprintf("%sfigures/fig3.pdf", x12))
+```
+
+```{r}
+res_i_certain
+res_i_uncertain
+res_iv_certain
+res_iv_uncertain
+```
 
@@
 This makes our own work much more difficult to interpret
---->
+-->
+
+<!--Move to discussion?-->
+We note that the variance in the non-DECREASE studies was estimated at 0, indicating that there would supposedly be a fixed-effect of beta-blockers on perioperative mortality despite differences in type of beta-blockers administered and dosage/duration of the beta-blockers.
+This homogeneity is not excessive, considering that the *Q*-statistic is sufficiently large to begin with (i.e., `r round(mod_clean$QE, .2)`).[@ioannidis2006] Nonetheless, the original non-DECREASE trials show substantive variation in the treatment, for example the usage of different beta-blockers (i.e., atenolol, propanolol, bisoprolol, metoprolol) in different (standardized) dosages, which have been argued to affect effectiveness of the interventions[@klei2015].
+As such, the lack of heterogeneity can be genuine and indicative of no difference between these different beta-blockers and their dosage/duration.
+Nonetheless, it could also be due to uncertainty in the estimate of heterogeneity $\tau^2$ due to the small set of studies included.
+
+However, due to various types of beta-blockers, evidence might be moderated[@klei2015].
+
+Additionally, during the process of this research paper, the first author tried to ascertain a raw dataset of one of the DECREASE trials via a Freedom of Information request (FOI).
+The Erasmus MC refused to share these data
+
+The ESC/ESA guidelines from 2014[@] were published in October 2014, whereas the meta-analysis indicating a reversal of the effectiveness of beta-blockers was published online at the end of July 2013[@].
+The committee revising the ESC/ESA guidelines stated: "The respective writing committees independently performed their literature review and analysis, and then developed their recommendations."[@Kristensen_2014] Regardless, the revisions made "recommend continuation of beta-blocker therapy in the perioperative period in patients currently receiving this medication" and do not recommend the "initiation of beta-blockers in patients undergoing low-risk surgery"[@guarracino2015].
+Review pieces note no substantial clinical change with respect to beta-blockers.[@Port2016;@Kristensen2016]
+
+The revisions to the ESC/ESA guidelines do not discourage use of beta-blockers, despite the obvious lack of evidence for systematic evidence of the effectiveness of beta-blockers to reduce perioperative mortality in trials.
+If anything, the trials show that insufficient evidence has been collected.
+This becomes apparent when the estimated effect for all non-DECREASE trials is inspected; its *p*-value is .04, which is insufficient evidence and is actually more likely when there is no effect in of beta-blockers in reality[@van_Aert_2016]
+
+In order , the first author tried to ascertain the data of DECREASE VI, the only DECREASE study for which a dataset is still available[@Coleg5210] via a Freedom of Information (FOI) request.
+Within this paper, we inspected data anomalies based on summary results, which contain less information to be able to determine what isThe board of Erasmus Medical Centrum (EMC) was unwilling to comply with this request, stating that it was not a matter of the board.
+Regardless of whether it was a board related matter,
 

### 2017-01-26 submission/manuscript.tex
Revise up to step 3 methods

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Analyzing DECREASE trials for extent of data fabrication},
            pdfauthor={Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\ImportTok}[1]{{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
\newcommand{\BuiltInTok}[1]{{#1}}
\newcommand{\ExtensionTok}[1]{{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Analyzing DECREASE trials for extent of data fabrication}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{26 January, 2017}

\begin{document}
\maketitle

The effect of beta-blockers on perioperative mortality in non-cardiac
surgery for patients with coronary artery disease (CAD) has been subject
of discussion\textsuperscript{1} due to findings of research
misconduct\textsuperscript{2--4} that were central to claims of their
effectiveness\textsuperscript{5--7}. Perioperative mortality is the
deathrate of patients during the period of the surgical procedure, which
typically includes admission, anaesthesia, surgery, and recovery. This
related finding of research misconduct altered meta-analytic results of
perioperative mortality in non-cardiac surgery for CAD patients
substantively\textsuperscript{7}, where meta-analyses that include these
studies conclude that beta-blockers decrease perioperative
mortality\textsuperscript{5--7} but a meta-analysis that excludes these
studies concludes that beta-blockers increase perioperative
mortality\textsuperscript{7}.

The trials deemed to be the result of research misconduct were the Dutch
DECREASE trials\textsuperscript{2--4}. The committees that investigated
the integrity of the DECREASE studies reported that data fabrication was
likely\textsuperscript{2--4}, but that the extent of the data
fabrication remained unclear. Given that the inclusion or exclusion of
these trials in meta-analyses is crucial in determining whether
beta-blockers are indeed effective during the perioperative period and
the latest guidelines still recommend the usage of
beta-blockers\textsuperscript{8}, we aim to estimate the extent of data
fabrication in the DECREASE studies\textsuperscript{9,10} in order to
further this debate.

The reports on the DECREASE trials primarily focused on the provenance
of the raw data and patient files (e.g., informed consent, whether data
corresponded to patient files), but neglected the extent to which the
DECREASE studies deviated from similar trials that tested the
effectiveness of beta-blockers on perioperative mortality. Comparing how
(dis)similar trials are from comparable studies is not a new
idea\textsuperscript{11} and has proven effective in detecting data
fabrication\textsuperscript{12}. This is a different approach from the
forensic statistical methods previously applied\textsuperscript{4},
where results within one study are evaluated as opposed to other similar
studies. This explains the discrepancy between the feasibility of
evaluating the DECREASE trials by the statistical expert of the
committee report and ours.

To statistically investigate the evidence of data fabrication in the
results of the DECREASE studies\textsuperscript{9,10}, we took three
steps. First, we replicate the findings from a 2014
meta-analysis\textsuperscript{7} that contains sufficient information to
estimate the deviation of the DECREASE-I and DECREASE-IV studies from
the remaining studies. Second, we evaluate the veracity of the
DECREASE-I and DECREASE-IV studies (i.e., the probability that these
study findings occurred assuming no data fabrication occurred). Third,
we reverse this assumption and assume that data fabrication did occur
and estimate how many data points would have to be fabricated to
reproduce the results of the DECREASE-I and DECREASE-IV studies if the
remaining studies are regarded as estimating the true effect of
beta-blockers on perioperative mortality.

\section{Step 1: reproducing meta-analysis of Bouri et al.
(2014)}\label{step-1-reproducing-meta-analysis-of-bouri-et-al.-2014}

\subsection{Methods}\label{methods}

To ensure that we used similar analysis procedures as in the 2014
meta-analysis\textsuperscript{7}, we initially reproduced Bouri et al.'s
estimates. This ensured that (1) their results are reproducible and (2)
we are using the correct estimates in subsequent steps of our analyses.
Using figures 2 and 3 from the original paper\textsuperscript{7}, we
extracted the raw event data for the 2 (control vs experimental) by 2
(event or no event) design, which we used to recompute the natural
logarithm of the risk ratio and its standard error. The extracted event
data is available at \href{https://osf.io/aykeh}{osf.io/aykeh} and our
analysis plan was preregistered at
\href{https://osf.io/vnmzc}{osf.io/vnmzc}.

We computed the log risk ratio (i.e., log RR) for each study and pooled
these using the \texttt{R} package \texttt{metafor}\textsuperscript{13}.
We estimated a weighted random-effects model using the restricted
maximum-likelihood estimator (i.e., \texttt{REML})\textsuperscript{14}
to estimate the variance of effects. We used the default weighting
procedure in the \texttt{metafor} package. When there was a zero-count
for a cell, .5 was added to each cell of that trial, as is common in
meta-analyses on risk- and odds ratios in order to deal with
computational artefacts\textsuperscript{15}. The 2014
meta-analysis\textsuperscript{7} did not specify the effect variance
estimate used; hence, minor discrepancies between our estimates and the
original estimates could be due to differences in the estimation
procedure.

\subsection{Results}\label{results}

We analyzed the RRs to verify the original estimates (Figure 2 of the
2014 meta-analysis\textsuperscript{7}) and found that we were able to
reproduce the estimates for the different sets of studies. Bouri et al.
differentiated between the estimates from the non-DECREASE trials and
the DECREASE trials. We confirmed the effect size estimates and the
variance estimates for both the non-DECREASE- and the DECREASE trials,
save for some minor discrepancies due to the estimation method. Table 1
depicts the original and reproduced values for both sets of studies.

\begin{longtable}[]{@{}lllll@{}}
\caption{The original- and reproduced meta-analytic results based on the
data provided in the 2014 meta-analysis by Bouri et al.}\tabularnewline
\toprule
& & Risk ratio & \(\tau^2\) & Confidence interval\tabularnewline
Non-DECREASE & Original & 1.27 & 0 & {[}1.01; 1.60{]}\tabularnewline
& Reproduced & 1.28 & 0 & {[}1.01; 1.62{]}\tabularnewline
DECREASE & Original & 0.42 & 0.29 & {[}0.15; 1.23{]}\tabularnewline
& Reproduced & 0.44 & 0.24 & {[}0.15; 1.23{]}\tabularnewline
\bottomrule
\end{longtable}

Second, we meta-analyzed all studies combined, including a dummy
predictor for the DECREASE and non-DECREASE studies to replicate results
presented in Figure 4 of the 2014 meta-analysis\textsuperscript{7}.
Surprisingly, our results showed more evidence for unequal subgroups
than the original meta-analysis\textsuperscript{7}; whereas the original
test for subgroup differences was \(\chi^2(1)=3.91,p=.05\), our analyses
indicated that \(\chi^2(1)=6.12,p=0.013\). Additionally, the original
analyses showed substantial heterogeneity (\(I^2=74.4\)\%) where we
found an \(I^2=0\)\%. Different heterogeneity estimators (e.g.,
\texttt{DerSimonian-Laird} instead of \texttt{REML}) did not resolve
this difference. We tried to clarify these discrepancies by e-mailing
the original authors, but failed to receive a response.

\section{Step 2: evaluating the veracity of DECREASE
studies}\label{step-2-evaluating-the-veracity-of-decrease-studies}

Based on the non-DECREASE estimates from Step 1, we estimated the
probability of obtaining the results in the DECREASE studies. To this
end, we assumed that the non-DECREASE studies provide a valid
representation of the true effect of beta-blockers on perioperative
mortality (similar to Bouri et al.\textsuperscript{7}; but see the
discussion). The estimated probability is also known as the veracity of
the data\textsuperscript{16}, which indicates the probability of the
observed data under a given true effect. We assumed that the
non-DECREASE studies estimate the true population distribution of
effects, not perturbed by publication bias due to statistical
(non)significance. Publication bias was assumed to not be a problem
because a substantial number of nonsignificant effects are included in
the dataset (9 of 11 nonsignificant).

\subsection{Method}\label{method}

Based on the estimated mean log RR and its credible interval in the
non-DECREASE studies, we computed the probability of log RR in the
DECREASE trials. The estimates of the non-DECREASE studies were obtained
from Step 1, which include the estimated log RR (i.e., 0.25), and its
95\% credibility interval as provided by the package \texttt{metafor}
(i.e., {[}0.01; 0.484{]}. We assumed a normal distribution of population
effects with the estimated effect as the mean of the distribution. The
95\% credibility interval denotes the bounds of the normal distribution
that covers 95\% of the density, where the standard deviation is
calculated as the distance from the mean to either bound, divided by
1.96. This allows for an approximation of the population effect
distribution, as depicted in Figure 1.

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{../figures/fig1}

}

\caption{Density plot of the estimated true effect distribution based on the non-DECREASE studies only, with the position of the DECREASE studies highlighted.}\label{fig:figure 1}
\end{figure}

Based on the estimated effect distribution from the non-DECREASE trials,
we calculated the probability of each DECREASE trial result, or a more
extreme result. In other words, we computed the \emph{p}-value for the
hypothesis that the DECREASE studies occurred based on the information
available from the other trials.

\subsection{Results}\label{results-1}

Figure 1 indicates that the DECREASE trials are highly unlikely under
the estimated effect distribution based on the non-DECREASE trials. More
specifically, DECREASE-I (or more extreme) has a probability of
\(2.9610213\times 10^{-53}\) (less than 1 in a quintillion) and
DECREASE-IV \(3.0974376\times 10^{-9}\) (3 in a billion), which
indicates they are unlikely to come from the same population effect
distribution as the non-DECREASE trials. Observing two of such extremely
unlikely results in this population effect distribution jointly is
highly improbable, \(9.1715786\times 10^{-62}\). This would therefore
indicate that the DECREASE trials are severely different from the
non-DECREASE studies.

Results from Step 1 indicated that no variance (i.e., homogeneity;
\(\tau^2=0\)) of the effects was observed; given the small number of
studies included (i.e., 9) this estimate is uncertain, however. We
conducted sensitivity analyses in order to see how dependent results are
on the specific heterogeneity estimate. The probability of observing
both DECREASE trials stays approximately below 1 out of 1000 until the
variance estimate is .25, and the probability stays approximately below
1 out of 100 until the variance estimate is .43 (see Figure 2). Further
research would be worthwhile to better estimate the variance in the
effects of beta-blockers on perioperative mortality and what moderators
affect the effect size (e.g., type of beta-blocker).

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{../figures/fig2}

}

\caption{Sensitivity analyses for the p-value that indicates the probability of observing the results from the DECREASE studies, or more extreme results, based on the estimated true effect (non-DECREASE trials) and the accompanying variance estimate.}\label{fig:figure 2}
\end{figure}

\section{Step 3: estimating the amount of fabricated
data}\label{step-3-estimating-the-amount-of-fabricated-data}

We estimated the number of data points that would need to be fabricated
to arrive at the estimates from the DECREASE studies, given that the
non-DECREASE studies represent the true effect. This assumes, in
contrast to Step 2, that the DECREASE studies in fact contain fabricated
data. The estimates from Step 3 provide an indication of the extent of
the potential data fabrication in the DECREASE studies under this
assumption\textsuperscript{2--4,7}.

\subsection{Method}\label{method-1}

In order to estimate the number of fabricated data points, we estimated
the population effect distribution of mortality (in log odds) per
condition in both the DECREASE- and non-DECREASE studies separately. For
each of these, we ran a meta-analysis applying the same, verified
methods used in Step 1. These separate analyses were necessary,
considering that Step 1 only provided estimates of the overall risk
ratio but not the odds conditional on exposure (beta-blocker or control)
and study set (DECREASE- versus non-DECREASE). This resulted in four
meta-analytic mortality estimates with an accompanying variance.
Throughout the simulations, we use the point estimates to simulate
genuine- and fabricated data, but supplement this by using the
distribution estimates as sensitivity analyses.

We applied the inversion method to estimate the number of fabricated
data points in the DECREASE-I and DECREASE-IV
studies\textsuperscript{17}. With the inversion method, we iteratively
hypothesized that \emph{X} out of \emph{N} data points were fabricated
(i.e., \(0, 1, ..., N\)) in the DECREASE-I and DECREASE-IV studies
separately. For each study, we simulated 10,000 datasets per hypothesis
of \emph{X} fabricated data points and \emph{N-X} genuine data points.
For each simulated dataset (exact simulation procedure in next
paragraph), we determined the likelihood of the results with

\begin{equation}
\label{eq1}
L(\theta|\pi_{E},\pi_{C})=\pi_{E}^{n_{11}}(1-\pi_{E})^{n_{12}} \times \pi_{C}^{n_{21}}(1 - \pi_{C})^{n_{22}}
\end{equation}

where \(\pi_{E}\) indicates the mortality rate of the experimental
condition as drawn from the meta-analytic effect distribution
(\(\pi_{C}\) indicates the same for the control condition). The
likelihood was computed under both the fabricated effect estimates
(i.e., \(L_{fabricated}\)) and the genuine data (i.e., \(L_{genuine}\)).
Table 2 indicates which cell sizes the various \(n_{XX}\) refer to
within the (simulated) data. We compared the likelihoods to determine
whether the simulated data were more likely to arise from the genuine-
(\(L_{genuine}\)) or from the fabricated data (\(L_{fabricated}\)). Note
that using the likelihoods for this is a minor deviation from the
preregistration, where we initially planned on using \(p\)-value
comparisons (\href{https://osf.io/vnmzc}{osf.io/vnmzc}).

\begin{longtable}[]{@{}lll@{}}
\caption{Outcome possibilities within a simulated 2 (beta-blocker v
control) by 2 (dead v alive) clinical trial.}\tabularnewline
\toprule
& Dead & Alive\tabularnewline
Beta-blockers & \(n_{11}\) & \(n_{12}\)\tabularnewline
Control & \(n_{21}\) & \(n_{22}\)\tabularnewline
\bottomrule
\end{longtable}

For each hypothesis of \emph{X} out of \emph{N} fabricated datapoints,
we computed the probability that the fabricated data are more likely
than the genuine data (\(p_F=P(L_{fabricated}>L_{genuine})\)). Based on
\(p_F\), we computed the confidence interval for \(X\) (i.e.,
\(X_{LB};X_{UB}\)). For a 95\% confidence interval, the lowerbound is
equal to the \(p_F\) closest to .025, whereas the upperbound is equal to
the \(p_F\) closest to .975.

We computed \(p_F\) for all possibilities of \emph{X} out of \emph{N}
fabricated datapoints in 10,000 randomly generated datsets, which were
generated in three steps. For each dataset we:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Sampled (without replacement) \emph{X} fictitious participants that
  would be the result of data fabrication (sampled across all
  conditions).
\item
  Determined the population mortality rate for each condition (i.e., for
  each cell as in Table 3). The meta-analytic point estimate was used or
  a population effect was randomly drawn from the meta-analytic effect
  distribution.
\item
  Simulated the number of deaths for the different conditions using a
  binomial distribution based on the mortality rate as determined in 2,
  resulting in the cell counts as in Table 3.
\end{enumerate}

Based on the meta-analytic effect from 2 and the cell sizes from 3, we
computed the likelihoods of \(L_{fabricated}\) and \(L_{genuine}\) using
Equation \ref{eq1}. As mentioned before, \(p_F\) was computed as the
probability that the data were more likely under the estimates resulting
from the (allegedly) fabricated data (i.e., the DECREASE trials) than
under the estimates resulting from the genuine data (i.e., the
non-DECREASE trials; \(p_F=P(L_{fabricated}>L_{genuine})\)).

\subsection{Results}\label{results-2}

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{../figures/fig3}

}

\caption{Test}\label{fig:figure 3}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{res_i_certain}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   [1] 0.1507 0.1624 0.1700 0.1674 0.1792 0.1807 0.1868 0.1867 0.1879 0.1930
##  [11] 0.2012 0.2182 0.2083 0.2114 0.2257 0.2296 0.2361 0.2476 0.2377 0.2488
##  [21] 0.2545 0.2631 0.2742 0.2788 0.2823 0.2894 0.3002 0.2996 0.3032 0.3094
##  [31] 0.3180 0.3243 0.3305 0.3352 0.3380 0.3547 0.3575 0.3649 0.3740 0.3873
##  [41] 0.3852 0.3957 0.4099 0.4088 0.4113 0.4205 0.4314 0.4378 0.4430 0.4524
##  [51] 0.4643 0.4655 0.4651 0.4851 0.4983 0.4931 0.5006 0.5106 0.5148 0.5241
##  [61] 0.5338 0.5428 0.5528 0.5549 0.5558 0.5768 0.5752 0.5797 0.5912 0.6043
##  [71] 0.6060 0.6177 0.6243 0.6283 0.6346 0.6433 0.6515 0.6586 0.6667 0.6684
##  [81] 0.6787 0.6895 0.6947 0.6978 0.7021 0.7121 0.7150 0.7264 0.7287 0.7402
##  [91] 0.7401 0.7522 0.7506 0.7533 0.7591 0.7762 0.7832 0.7782 0.7821 0.7873
## [101] 0.7894 0.7925 0.8009 0.8095 0.8175 0.8135 0.8211 0.8251 0.8306 0.8363
## [111] 0.8349 0.8372 0.8453
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{res_i_uncertain}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   [1] 0.1224 0.1247 0.1263 0.1213 0.1224 0.1318 0.1288 0.1355 0.1370 0.1419
##  [11] 0.1502 0.1448 0.1557 0.1582 0.1619 0.1730 0.1801 0.1758 0.1796 0.1869
##  [21] 0.1938 0.1968 0.2086 0.2106 0.2239 0.2267 0.2350 0.2414 0.2446 0.2602
##  [31] 0.2619 0.2716 0.2697 0.2896 0.3045 0.3067 0.3302 0.3313 0.3332 0.3543
##  [41] 0.3552 0.3801 0.3755 0.3816 0.4068 0.4056 0.4106 0.4231 0.4297 0.4384
##  [51] 0.4573 0.4694 0.4742 0.4911 0.4864 0.5000 0.5125 0.5172 0.5247 0.5337
##  [61] 0.5451 0.5518 0.5554 0.5640 0.5666 0.5798 0.5924 0.5896 0.6015 0.6100
##  [71] 0.6154 0.6297 0.6251 0.6376 0.6446 0.6480 0.6616 0.6649 0.6725 0.6786
##  [81] 0.6887 0.6913 0.6999 0.7077 0.7116 0.7248 0.7153 0.7227 0.7302 0.7379
##  [91] 0.7432 0.7499 0.7534 0.7530 0.7567 0.7708 0.7712 0.7762 0.7723 0.7988
## [101] 0.7908 0.7958 0.7998 0.8060 0.8082 0.8051 0.8234 0.8162 0.8240 0.8323
## [111] 0.8445 0.8647 0.8862
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{res_iv_certain}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##    [1] 0.0223 0.0255 0.0234 0.0240 0.0263 0.0263 0.0226 0.0232 0.0220
##   [10] 0.0234 0.0251 0.0254 0.0247 0.0234 0.0276 0.0249 0.0268 0.0244
##   [19] 0.0244 0.0275 0.0264 0.0267 0.0267 0.0250 0.0279 0.0253 0.0254
##   [28] 0.0268 0.0267 0.0267 0.0267 0.0257 0.0264 0.0279 0.0264 0.0274
##   [37] 0.0285 0.0295 0.0269 0.0290 0.0301 0.0305 0.0288 0.0281 0.0271
##   [46] 0.0293 0.0280 0.0271 0.0266 0.0289 0.0283 0.0277 0.0299 0.0283
##   [55] 0.0296 0.0326 0.0344 0.0291 0.0321 0.0299 0.0287 0.0302 0.0289
##   [64] 0.0301 0.0297 0.0312 0.0315 0.0291 0.0308 0.0300 0.0298 0.0319
##   [73] 0.0283 0.0339 0.0311 0.0317 0.0313 0.0322 0.0324 0.0342 0.0302
##   [82] 0.0334 0.0332 0.0340 0.0334 0.0328 0.0349 0.0334 0.0348 0.0339
##   [91] 0.0339 0.0326 0.0350 0.0324 0.0353 0.0307 0.0339 0.0386 0.0379
##  [100] 0.0343 0.0347 0.0356 0.0370 0.0358 0.0343 0.0345 0.0356 0.0373
##  [109] 0.0399 0.0382 0.0376 0.0372 0.0376 0.0383 0.0374 0.0391 0.0349
##  [118] 0.0389 0.0377 0.0406 0.0443 0.0394 0.0402 0.0415 0.0406 0.0415
##  [127] 0.0412 0.0433 0.0430 0.0376 0.0399 0.0444 0.0418 0.0445 0.0464
##  [136] 0.0450 0.0461 0.0421 0.0447 0.0398 0.0418 0.0440 0.0461 0.0463
##  [145] 0.0500 0.0447 0.0438 0.0439 0.0492 0.0465 0.0464 0.0466 0.0487
##  [154] 0.0482 0.0467 0.0530 0.0501 0.0466 0.0510 0.0470 0.0540 0.0517
##  [163] 0.0542 0.0523 0.0549 0.0488 0.0509 0.0550 0.0550 0.0525 0.0541
##  [172] 0.0522 0.0579 0.0551 0.0548 0.0533 0.0590 0.0573 0.0557 0.0572
##  [181] 0.0548 0.0531 0.0598 0.0566 0.0613 0.0595 0.0599 0.0545 0.0627
##  [190] 0.0581 0.0625 0.0603 0.0643 0.0595 0.0650 0.0621 0.0657 0.0673
##  [199] 0.0648 0.0666 0.0684 0.0676 0.0654 0.0673 0.0679 0.0702 0.0687
##  [208] 0.0749 0.0713 0.0731 0.0696 0.0709 0.0702 0.0719 0.0764 0.0758
##  [217] 0.0733 0.0790 0.0725 0.0773 0.0770 0.0765 0.0778 0.0822 0.0789
##  [226] 0.0801 0.0814 0.0858 0.0820 0.0821 0.0854 0.0803 0.0824 0.0831
##  [235] 0.0864 0.0895 0.0845 0.0943 0.0883 0.0963 0.0930 0.0919 0.0871
##  [244] 0.0921 0.0940 0.0896 0.0967 0.0941 0.1000 0.0969 0.0980 0.0951
##  [253] 0.0988 0.1051 0.1031 0.1013 0.1059 0.0974 0.1055 0.1064 0.1074
##  [262] 0.1058 0.1068 0.1064 0.1117 0.1100 0.1069 0.1174 0.1078 0.1120
##  [271] 0.1107 0.1124 0.1180 0.1165 0.1190 0.1238 0.1187 0.1207 0.1219
##  [280] 0.1310 0.1247 0.1233 0.1288 0.1228 0.1319 0.1248 0.1337 0.1326
##  [289] 0.1339 0.1346 0.1329 0.1365 0.1366 0.1443 0.1395 0.1391 0.1377
##  [298] 0.1433 0.1424 0.1435 0.1466 0.1503 0.1519 0.1494 0.1459 0.1519
##  [307] 0.1578 0.1559 0.1500 0.1585 0.1550 0.1520 0.1651 0.1632 0.1662
##  [316] 0.1605 0.1623 0.1672 0.1663 0.1654 0.1730 0.1665 0.1721 0.1663
##  [325] 0.1789 0.1690 0.1754 0.1792 0.1782 0.1825 0.1827 0.1846 0.1820
##  [334] 0.1861 0.1884 0.1893 0.1936 0.1931 0.1956 0.1903 0.1942 0.1914
##  [343] 0.1901 0.2079 0.2054 0.2033 0.1976 0.2029 0.2099 0.2079 0.2084
##  [352] 0.2212 0.2063 0.2079 0.2232 0.2101 0.2222 0.2166 0.2184 0.2144
##  [361] 0.2220 0.2219 0.2351 0.2272 0.2271 0.2311 0.2337 0.2291 0.2382
##  [370] 0.2409 0.2362 0.2391 0.2327 0.2468 0.2372 0.2416 0.2520 0.2460
##  [379] 0.2548 0.2598 0.2457 0.2585 0.2620 0.2636 0.2608 0.2623 0.2661
##  [388] 0.2669 0.2655 0.2664 0.2662 0.2679 0.2702 0.2770 0.2752 0.2769
##  [397] 0.2836 0.2842 0.2828 0.2806 0.2860 0.2825 0.2788 0.2874 0.2987
##  [406] 0.2879 0.2975 0.2895 0.2920 0.3043 0.3040 0.3044 0.3065 0.3083
##  [415] 0.3066 0.3182 0.3098 0.3150 0.3167 0.3066 0.3137 0.3200 0.3153
##  [424] 0.3271 0.3237 0.3273 0.3252 0.3359 0.3381 0.3298 0.3320 0.3384
##  [433] 0.3328 0.3353 0.3417 0.3377 0.3521 0.3491 0.3432 0.3503 0.3540
##  [442] 0.3581 0.3569 0.3499 0.3632 0.3561 0.3652 0.3658 0.3703 0.3630
##  [451] 0.3637 0.3745 0.3743 0.3858 0.3751 0.3728 0.3801 0.3978 0.3775
##  [460] 0.3776 0.3882 0.3847 0.3893 0.3876 0.3899 0.3889 0.3906 0.3921
##  [469] 0.4023 0.4096 0.4099 0.4129 0.4029 0.4086 0.4096 0.4102 0.4108
##  [478] 0.4269 0.4119 0.4128 0.4277 0.4196 0.4194 0.4236 0.4307 0.4253
##  [487] 0.4279 0.4335 0.4413 0.4325 0.4449 0.4420 0.4371 0.4446 0.4488
##  [496] 0.4494 0.4533 0.4548 0.4574 0.4723 0.4585 0.4768 0.4649 0.4670
##  [505] 0.4617 0.4672 0.4727 0.4755 0.4753 0.4681 0.4862 0.4814 0.4835
##  [514] 0.4873 0.4832 0.4794 0.4749 0.4809 0.4973 0.4886 0.4986 0.4971
##  [523] 0.5020 0.4994 0.5077 0.5026 0.5016 0.5028 0.5094 0.5111 0.5157
##  [532] 0.5163 0.5252 0.5159 0.5145 0.5301 0.5237 0.5264 0.5364 0.5265
##  [541] 0.5337 0.5342 0.5420 0.5380 0.5473 0.5335 0.5496 0.5520 0.5463
##  [550] 0.5514 0.5471 0.5465 0.5572 0.5573 0.5569 0.5582 0.5618 0.5581
##  [559] 0.5711 0.5598 0.5705 0.5692 0.5758 0.5841 0.5741 0.5778 0.5785
##  [568] 0.5856 0.5898 0.5809 0.5956 0.5882 0.5916 0.6016 0.5906 0.5955
##  [577] 0.6025 0.6019 0.5995 0.5995 0.6165 0.6099 0.6058 0.6030 0.6195
##  [586] 0.6180 0.6142 0.6274 0.6210 0.6202 0.6308 0.6285 0.6253 0.6296
##  [595] 0.6348 0.6360 0.6446 0.6328 0.6408 0.6404 0.6390 0.6490 0.6529
##  [604] 0.6408 0.6524 0.6547 0.6440 0.6573 0.6582 0.6538 0.6579 0.6646
##  [613] 0.6699 0.6688 0.6707 0.6745 0.6714 0.6767 0.6739 0.6773 0.6863
##  [622] 0.6772 0.6827 0.6887 0.6903 0.6890 0.6941 0.7070 0.6965 0.6978
##  [631] 0.7017 0.7021 0.7107 0.6981 0.7082 0.7076 0.7137 0.7035 0.7197
##  [640] 0.7202 0.7168 0.7055 0.7230 0.7291 0.7289 0.7301 0.7281 0.7277
##  [649] 0.7348 0.7312 0.7336 0.7395 0.7424 0.7412 0.7415 0.7463 0.7491
##  [658] 0.7462 0.7457 0.7517 0.7573 0.7553 0.7580 0.7545 0.7525 0.7570
##  [667] 0.7567 0.7598 0.7663 0.7682 0.7717 0.7640 0.7707 0.7705 0.7758
##  [676] 0.7772 0.7756 0.7817 0.7779 0.7849 0.7966 0.7825 0.7862 0.7937
##  [685] 0.7902 0.7972 0.7933 0.7858 0.7964 0.7949 0.7984 0.8011 0.7970
##  [694] 0.8021 0.8076 0.8037 0.8106 0.8091 0.8119 0.8136 0.8124 0.8093
##  [703] 0.8086 0.8218 0.8195 0.8156 0.8202 0.8175 0.8287 0.8196 0.8284
##  [712] 0.8299 0.8276 0.8316 0.8237 0.8336 0.8291 0.8418 0.8362 0.8353
##  [721] 0.8376 0.8431 0.8465 0.8381 0.8416 0.8450 0.8416 0.8426 0.8440
##  [730] 0.8507 0.8440 0.8547 0.8493 0.8567 0.8528 0.8445 0.8588 0.8619
##  [739] 0.8548 0.8633 0.8534 0.8626 0.8626 0.8691 0.8658 0.8619 0.8686
##  [748] 0.8607 0.8693 0.8702 0.8726 0.8731 0.8749 0.8682 0.8763 0.8730
##  [757] 0.8717 0.8712 0.8751 0.8813 0.8830 0.8812 0.8811 0.8749 0.8830
##  [766] 0.8878 0.8838 0.8891 0.8871 0.8890 0.8866 0.8868 0.8898 0.8901
##  [775] 0.8863 0.8902 0.8962 0.8903 0.8947 0.8978 0.8944 0.8937 0.8985
##  [784] 0.9033 0.8936 0.8980 0.8984 0.9069 0.9011 0.8996 0.8956 0.9031
##  [793] 0.9007 0.8993 0.9015 0.9057 0.9060 0.9110 0.9130 0.9038 0.9040
##  [802] 0.9148 0.9115 0.9110 0.9132 0.9139 0.9179 0.9140 0.9089 0.9152
##  [811] 0.9162 0.9185 0.9105 0.9173 0.9160 0.9187 0.9176 0.9189 0.9184
##  [820] 0.9203 0.9191 0.9229 0.9214 0.9153 0.9251 0.9215 0.9259 0.9267
##  [829] 0.9177 0.9222 0.9312 0.9247 0.9366 0.9263 0.9303 0.9284 0.9276
##  [838] 0.9307 0.9288 0.9293 0.9293 0.9359 0.9324 0.9297 0.9346 0.9323
##  [847] 0.9352 0.9317 0.9365 0.9337 0.9354 0.9308 0.9316 0.9351 0.9386
##  [856] 0.9377 0.9344 0.9406 0.9388 0.9386 0.9362 0.9395 0.9396 0.9396
##  [865] 0.9381 0.9383 0.9376 0.9385 0.9422 0.9423 0.9404 0.9430 0.9417
##  [874] 0.9422 0.9380 0.9408 0.9443 0.9423 0.9462 0.9469 0.9467 0.9440
##  [883] 0.9466 0.9462 0.9533 0.9399 0.9487 0.9491 0.9492 0.9480 0.9446
##  [892] 0.9450 0.9462 0.9528 0.9501 0.9479 0.9495 0.9484 0.9501 0.9504
##  [901] 0.9486 0.9540 0.9497 0.9506 0.9538 0.9524 0.9517 0.9547 0.9528
##  [910] 0.9522 0.9551 0.9545 0.9558 0.9550 0.9537 0.9496 0.9515 0.9547
##  [919] 0.9542 0.9546 0.9520 0.9533 0.9557 0.9555 0.9530 0.9516 0.9603
##  [928] 0.9574 0.9584 0.9553 0.9574 0.9586 0.9584 0.9573 0.9540 0.9619
##  [937] 0.9597 0.9583 0.9554 0.9599 0.9586 0.9595 0.9598 0.9620 0.9625
##  [946] 0.9624 0.9638 0.9605 0.9597 0.9616 0.9631 0.9603 0.9644 0.9626
##  [955] 0.9598 0.9601 0.9647 0.9657 0.9643 0.9647 0.9619 0.9639 0.9644
##  [964] 0.9620 0.9635 0.9673 0.9666 0.9671 0.9658 0.9644 0.9648 0.9688
##  [973] 0.9643 0.9638 0.9624 0.9648 0.9658 0.9657 0.9673 0.9637 0.9679
##  [982] 0.9681 0.9678 0.9677 0.9683 0.9684 0.9658 0.9706 0.9703 0.9677
##  [991] 0.9661 0.9655 0.9676 0.9691 0.9698 0.9702 0.9675 0.9672 0.9725
## [1000] 0.9708 0.9704 0.9706 0.9707 0.9703 0.9679 0.9640 0.9719 0.9720
## [1009] 0.9712 0.9698 0.9710 0.9732 0.9711 0.9724 0.9704 0.9704 0.9698
## [1018] 0.9714 0.9717 0.9699 0.9720 0.9709 0.9748 0.9730 0.9718 0.9738
## [1027] 0.9721 0.9722 0.9709 0.9705 0.9683 0.9743 0.9737 0.9710 0.9735
## [1036] 0.9735 0.9758 0.9720 0.9747 0.9756 0.9686 0.9750 0.9740 0.9748
## [1045] 0.9740 0.9746 0.9752 0.9734 0.9760 0.9744 0.9736 0.9752 0.9767
## [1054] 0.9764 0.9734 0.9788 0.9773 0.9754 0.9750 0.9791 0.9791 0.9772
## [1063] 0.9789 0.9760 0.9789 0.9758 0.9777
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{res_iv_uncertain}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##    [1] 0.0203 0.0200 0.0220 0.0211 0.0236 0.0200 0.0208 0.0193 0.0220
##   [10] 0.0222 0.0224 0.0228 0.0194 0.0218 0.0226 0.0221 0.0220 0.0210
##   [19] 0.0234 0.0216 0.0222 0.0242 0.0229 0.0254 0.0216 0.0225 0.0218
##   [28] 0.0224 0.0215 0.0221 0.0227 0.0221 0.0238 0.0233 0.0259 0.0243
##   [37] 0.0227 0.0255 0.0242 0.0216 0.0226 0.0210 0.0263 0.0218 0.0228
##   [46] 0.0232 0.0242 0.0214 0.0247 0.0252 0.0255 0.0240 0.0251 0.0252
##   [55] 0.0238 0.0246 0.0237 0.0265 0.0262 0.0272 0.0244 0.0262 0.0226
##   [64] 0.0245 0.0276 0.0243 0.0225 0.0251 0.0277 0.0297 0.0228 0.0263
##   [73] 0.0256 0.0274 0.0239 0.0259 0.0282 0.0287 0.0270 0.0267 0.0270
##   [82] 0.0265 0.0298 0.0298 0.0275 0.0285 0.0277 0.0278 0.0275 0.0269
##   [91] 0.0307 0.0260 0.0309 0.0273 0.0286 0.0292 0.0291 0.0334 0.0316
##  [100] 0.0297 0.0311 0.0299 0.0305 0.0296 0.0308 0.0303 0.0284 0.0297
##  [109] 0.0285 0.0274 0.0320 0.0287 0.0312 0.0306 0.0325 0.0300 0.0297
##  [118] 0.0306 0.0332 0.0333 0.0323 0.0319 0.0337 0.0306 0.0335 0.0354
##  [127] 0.0306 0.0338 0.0291 0.0314 0.0321 0.0346 0.0338 0.0365 0.0323
##  [136] 0.0327 0.0336 0.0341 0.0351 0.0372 0.0327 0.0319 0.0365 0.0321
##  [145] 0.0360 0.0339 0.0350 0.0364 0.0378 0.0344 0.0350 0.0362 0.0337
##  [154] 0.0364 0.0352 0.0348 0.0350 0.0355 0.0373 0.0383 0.0367 0.0369
##  [163] 0.0397 0.0360 0.0396 0.0432 0.0398 0.0409 0.0412 0.0438 0.0401
##  [172] 0.0407 0.0390 0.0422 0.0397 0.0434 0.0413 0.0418 0.0416 0.0410
##  [181] 0.0454 0.0402 0.0401 0.0429 0.0409 0.0427 0.0425 0.0461 0.0428
##  [190] 0.0479 0.0466 0.0436 0.0489 0.0464 0.0476 0.0469 0.0458 0.0454
##  [199] 0.0484 0.0445 0.0465 0.0466 0.0506 0.0487 0.0502 0.0494 0.0483
##  [208] 0.0483 0.0518 0.0550 0.0493 0.0508 0.0509 0.0535 0.0542 0.0502
##  [217] 0.0495 0.0510 0.0503 0.0486 0.0500 0.0530 0.0521 0.0534 0.0514
##  [226] 0.0545 0.0527 0.0581 0.0545 0.0619 0.0589 0.0595 0.0611 0.0572
##  [235] 0.0631 0.0582 0.0592 0.0607 0.0587 0.0571 0.0606 0.0600 0.0653
##  [244] 0.0645 0.0675 0.0614 0.0620 0.0682 0.0712 0.0648 0.0654 0.0677
##  [253] 0.0719 0.0650 0.0737 0.0692 0.0707 0.0736 0.0731 0.0748 0.0746
##  [262] 0.0789 0.0752 0.0760 0.0775 0.0749 0.0767 0.0753 0.0821 0.0847
##  [271] 0.0832 0.0814 0.0853 0.0878 0.0872 0.0870 0.0836 0.0862 0.0866
##  [280] 0.0869 0.0946 0.0909 0.0922 0.0960 0.0984 0.0901 0.0972 0.0955
##  [289] 0.1010 0.0977 0.0983 0.0986 0.1055 0.1048 0.1016 0.1065 0.1095
##  [298] 0.1046 0.1085 0.1134 0.1145 0.1135 0.1139 0.1188 0.1217 0.1133
##  [307] 0.1179 0.1158 0.1245 0.1229 0.1231 0.1235 0.1245 0.1212 0.1270
##  [316] 0.1278 0.1325 0.1304 0.1370 0.1337 0.1349 0.1355 0.1378 0.1432
##  [325] 0.1420 0.1479 0.1468 0.1422 0.1472 0.1499 0.1557 0.1567 0.1587
##  [334] 0.1564 0.1533 0.1631 0.1646 0.1596 0.1598 0.1548 0.1736 0.1659
##  [343] 0.1752 0.1718 0.1753 0.1759 0.1755 0.1795 0.1797 0.1881 0.1829
##  [352] 0.1843 0.1945 0.1868 0.1933 0.1931 0.1848 0.2012 0.1943 0.2025
##  [361] 0.1993 0.2069 0.2048 0.2036 0.2108 0.2068 0.2143 0.2092 0.2164
##  [370] 0.2184 0.2206 0.2221 0.2288 0.2289 0.2326 0.2314 0.2369 0.2359
##  [379] 0.2324 0.2386 0.2393 0.2404 0.2439 0.2385 0.2511 0.2413 0.2440
##  [388] 0.2543 0.2580 0.2559 0.2518 0.2481 0.2623 0.2649 0.2605 0.2628
##  [397] 0.2636 0.2688 0.2772 0.2788 0.2757 0.2775 0.2756 0.2785 0.2773
##  [406] 0.2870 0.2854 0.2948 0.2991 0.2885 0.3011 0.2991 0.2939 0.2976
##  [415] 0.2907 0.3048 0.3077 0.3064 0.3135 0.3096 0.3175 0.3294 0.3119
##  [424] 0.3192 0.3161 0.3245 0.3258 0.3236 0.3325 0.3279 0.3396 0.3315
##  [433] 0.3313 0.3337 0.3421 0.3384 0.3432 0.3402 0.3376 0.3558 0.3432
##  [442] 0.3524 0.3501 0.3479 0.3537 0.3587 0.3659 0.3627 0.3582 0.3616
##  [451] 0.3725 0.3635 0.3788 0.3722 0.3642 0.3854 0.3723 0.3742 0.3799
##  [460] 0.3859 0.3866 0.3885 0.3952 0.3940 0.3910 0.3908 0.3938 0.4019
##  [469] 0.4070 0.3966 0.3980 0.4045 0.3993 0.4055 0.4053 0.4133 0.4179
##  [478] 0.4192 0.4171 0.4168 0.4247 0.4202 0.4188 0.4326 0.4260 0.4196
##  [487] 0.4277 0.4336 0.4303 0.4257 0.4372 0.4404 0.4395 0.4392 0.4327
##  [496] 0.4418 0.4494 0.4461 0.4390 0.4495 0.4590 0.4546 0.4537 0.4519
##  [505] 0.4460 0.4540 0.4534 0.4698 0.4608 0.4616 0.4593 0.4614 0.4659
##  [514] 0.4682 0.4629 0.4585 0.4783 0.4775 0.4741 0.4782 0.4802 0.4769
##  [523] 0.4790 0.4801 0.4823 0.4851 0.4874 0.4939 0.4923 0.4940 0.4911
##  [532] 0.4889 0.4970 0.4955 0.5035 0.5017 0.5049 0.5121 0.5016 0.5049
##  [541] 0.5041 0.5053 0.5104 0.5097 0.5230 0.5097 0.5204 0.5032 0.5089
##  [550] 0.5190 0.5213 0.5199 0.5272 0.5320 0.5313 0.5231 0.5257 0.5276
##  [559] 0.5318 0.5321 0.5390 0.5372 0.5319 0.5322 0.5459 0.5396 0.5486
##  [568] 0.5418 0.5477 0.5511 0.5489 0.5570 0.5658 0.5614 0.5556 0.5489
##  [577] 0.5569 0.5605 0.5567 0.5614 0.5672 0.5694 0.5661 0.5718 0.5689
##  [586] 0.5719 0.5755 0.5691 0.5735 0.5779 0.5802 0.5864 0.5805 0.5809
##  [595] 0.5763 0.5703 0.5746 0.5821 0.5854 0.5916 0.5892 0.5919 0.5948
##  [604] 0.5991 0.5898 0.6004 0.5983 0.5961 0.6088 0.6063 0.6072 0.6078
##  [613] 0.6058 0.6243 0.6124 0.6114 0.6168 0.6248 0.6158 0.6283 0.6155
##  [622] 0.6222 0.6186 0.6247 0.6250 0.6258 0.6222 0.6306 0.6247 0.6238
##  [631] 0.6276 0.6392 0.6426 0.6351 0.6461 0.6339 0.6489 0.6414 0.6406
##  [640] 0.6401 0.6465 0.6515 0.6560 0.6574 0.6483 0.6600 0.6504 0.6608
##  [649] 0.6518 0.6617 0.6599 0.6639 0.6545 0.6692 0.6686 0.6675 0.6731
##  [658] 0.6696 0.6732 0.6681 0.6755 0.6783 0.6748 0.6822 0.6804 0.6802
##  [667] 0.6843 0.6893 0.6954 0.6899 0.6880 0.6819 0.6996 0.6906 0.6864
##  [676] 0.6991 0.6962 0.7011 0.6999 0.7065 0.7034 0.7037 0.7043 0.7149
##  [685] 0.7074 0.7106 0.7080 0.7146 0.7199 0.7260 0.7231 0.7202 0.7155
##  [694] 0.7211 0.7289 0.7250 0.7178 0.7237 0.7371 0.7387 0.7330 0.7391
##  [703] 0.7387 0.7426 0.7294 0.7346 0.7387 0.7457 0.7472 0.7480 0.7514
##  [712] 0.7501 0.7521 0.7467 0.7479 0.7546 0.7498 0.7546 0.7549 0.7535
##  [721] 0.7552 0.7542 0.7634 0.7556 0.7590 0.7643 0.7677 0.7712 0.7702
##  [730] 0.7744 0.7708 0.7688 0.7768 0.7713 0.7727 0.7798 0.7861 0.7800
##  [739] 0.7825 0.7847 0.7816 0.7846 0.7856 0.7929 0.7954 0.7891 0.7898
##  [748] 0.7843 0.7915 0.7953 0.8005 0.8009 0.7934 0.8031 0.8034 0.8073
##  [757] 0.8010 0.8072 0.8125 0.8056 0.8151 0.8141 0.8083 0.8074 0.8020
##  [766] 0.8125 0.8152 0.8203 0.8119 0.8167 0.8143 0.8127 0.8193 0.8193
##  [775] 0.8188 0.8210 0.8293 0.8235 0.8259 0.8318 0.8239 0.8277 0.8241
##  [784] 0.8374 0.8307 0.8342 0.8323 0.8373 0.8370 0.8360 0.8375 0.8330
##  [793] 0.8388 0.8415 0.8426 0.8421 0.8426 0.8472 0.8337 0.8404 0.8430
##  [802] 0.8473 0.8423 0.8483 0.8408 0.8470 0.8522 0.8498 0.8496 0.8439
##  [811] 0.8536 0.8491 0.8451 0.8551 0.8571 0.8538 0.8518 0.8539 0.8588
##  [820] 0.8545 0.8589 0.8673 0.8547 0.8616 0.8602 0.8636 0.8658 0.8603
##  [829] 0.8707 0.8607 0.8673 0.8674 0.8672 0.8673 0.8656 0.8644 0.8654
##  [838] 0.8678 0.8683 0.8651 0.8743 0.8706 0.8752 0.8745 0.8716 0.8723
##  [847] 0.8762 0.8783 0.8749 0.8763 0.8734 0.8802 0.8745 0.8727 0.8775
##  [856] 0.8779 0.8788 0.8765 0.8786 0.8780 0.8740 0.8793 0.8816 0.8834
##  [865] 0.8815 0.8813 0.8800 0.8875 0.8832 0.8805 0.8880 0.8828 0.8852
##  [874] 0.8851 0.8911 0.8855 0.8886 0.8855 0.8962 0.8932 0.8837 0.8907
##  [883] 0.8907 0.8872 0.8886 0.8920 0.8978 0.8904 0.8922 0.8960 0.8917
##  [892] 0.8955 0.8886 0.8929 0.8923 0.8923 0.8928 0.9007 0.8939 0.8925
##  [901] 0.8950 0.8953 0.8870 0.8895 0.8952 0.8948 0.8958 0.9033 0.8991
##  [910] 0.9003 0.9008 0.9035 0.8995 0.8957 0.9016 0.8998 0.9013 0.9023
##  [919] 0.9000 0.8962 0.8934 0.8994 0.9051 0.9051 0.9016 0.9038 0.9058
##  [928] 0.9060 0.9030 0.8974 0.9083 0.9008 0.9028 0.9042 0.9031 0.9008
##  [937] 0.9063 0.9012 0.9014 0.9014 0.9046 0.9080 0.9093 0.9086 0.9109
##  [946] 0.9033 0.9070 0.9033 0.9042 0.9046 0.9081 0.9080 0.9026 0.9049
##  [955] 0.9064 0.9106 0.9099 0.9087 0.9089 0.9128 0.9109 0.9149 0.9105
##  [964] 0.9071 0.9134 0.9076 0.9127 0.9090 0.9074 0.9090 0.9049 0.9083
##  [973] 0.9094 0.9086 0.9107 0.9043 0.9126 0.9056 0.9114 0.9050 0.9095
##  [982] 0.9147 0.9069 0.9124 0.9130 0.9139 0.9124 0.9100 0.9114 0.9109
##  [991] 0.9114 0.9096 0.9103 0.9110 0.9104 0.9137 0.9121 0.9109 0.9114
## [1000] 0.9102 0.9130 0.9104 0.9137 0.9158 0.9171 0.9108 0.9086 0.9145
## [1009] 0.9174 0.9133 0.9078 0.9137 0.9143 0.9102 0.9157 0.9169 0.9151
## [1018] 0.9154 0.9210 0.9158 0.9171 0.9167 0.9200 0.9166 0.9122 0.9182
## [1027] 0.9162 0.9106 0.9196 0.9195 0.9147 0.9214 0.9211 0.9153 0.9203
## [1036] 0.9168 0.9189 0.9190 0.9158 0.9210 0.9195 0.9240 0.9176 0.9144
## [1045] 0.9218 0.9140 0.9178 0.9170 0.9213 0.9195 0.9177 0.9200 0.9200
## [1054] 0.9204 0.9150 0.9173 0.9172 0.9197 0.9206 0.9148 0.9208 0.9225
## [1063] 0.9242 0.9289 0.9358 0.9499 0.9789
\end{verbatim}

\section{Discussion}\label{discussion}

We note that the variance in the non-DECREASE studies was estimated at
0, indicating that there would supposedly be a fixed-effect of
beta-blockers on perioperative mortality despite differences in type of
beta-blockers administered and dosage/duration of the beta-blockers.
This homogeneity is not excessive, considering that the
\emph{Q}-statistic is sufficiently large to begin with (i.e.,
5).\textsuperscript{18} Nonetheless, the original non-DECREASE trials
show substantive variation in the treatment, for example the usage of
different beta-blockers (i.e., atenolol, propanolol, bisoprolol,
metoprolol) in different (standardized) dosages, which have been argued
to affect effectiveness of the interventions\textsuperscript{19}. As
such, the lack of heterogeneity can be genuine and indicative of no
difference between these different beta-blockers and their
dosage/duration. Nonetheless, it could also be due to uncertainty in the
estimate of heterogeneity \(\tau^2\) due to the small set of studies
included.

However, due to various types of beta-blockers, evidence might be
moderated\textsuperscript{19}.

Additionally, during the process of this research paper, the first
author tried to ascertain a raw dataset of one of the DECREASE trials
via a Freedom of Information request (FOI). The Erasmus MC refused to
share these data

The ESC/ESA guidelines from 2014{[}@{]} were published in October 2014,
whereas the meta-analysis indicating a reversal of the effectiveness of
beta-blockers was published online at the end of July 2013{[}@{]}. The
committee revising the ESC/ESA guidelines stated: ``The respective
writing committees independently performed their literature review and
analysis, and then developed their
recommendations.''\textsuperscript{20} Regardless, the revisions made
``recommend continuation of beta-blocker therapy in the perioperative
period in patients currently receiving this medication'' and do not
recommend the ``initiation of beta-blockers in patients undergoing
low-risk surgery''\textsuperscript{21}. Review pieces note no
substantial clinical change with respect to
beta-blockers.\textsuperscript{22,23}

The revisions to the ESC/ESA guidelines do not discourage use of
beta-blockers, despite the obvious lack of evidence for systematic
evidence of the effectiveness of beta-blockers to reduce perioperative
mortality in trials. If anything, the trials show that insufficient
evidence has been collected. This becomes apparent when the estimated
effect for all non-DECREASE trials is inspected; its \emph{p}-value is
.04, which is insufficient evidence and is actually more likely when
there is no effect in of beta-blockers in reality\textsuperscript{24}

In order , the first author tried to ascertain the data of DECREASE VI,
the only DECREASE study for which a dataset is still
available\textsuperscript{1} via a Freedom of Information (FOI) request.
Within this paper, we inspected data anomalies based on summary results,
which contain less information to be able to determine what isThe board
of Erasmus Medical Centrum (EMC) was unwilling to comply with this
request, stating that it was not a matter of the board. Regardless of
whether it was a board related matter,

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-Coleg5210}{}
1. Cole GD, Francis DP. Perioperative beta blockade: Guidelines do not
reflect the problems with the evidence from the decrease trials.
\emph{BMJ}. 2014;349.
doi:\href{https://doi.org/10.1136/bmj.g5210}{10.1136/bmj.g5210}.

\hypertarget{ref-commissie2011}{}
2. Onderzoekscommissie Wetenschappelijke Integriteit. \emph{Onderzoek
Naar Mogelijke Schending van de Wetenschappelijke Integriteit: Beknopte
Versie}. Erasmus MC; 2011.
\url{https://web.archive.org/web/20151113121125/http://www.erasmusmc.nl/cs-research/bijlagen/integriteit/rapport-poldermans-2011}.

\hypertarget{ref-commissie2012}{}
3. Commissie Vervolgonderzoek 2012. \emph{Rapport Vervolgonderzoek Naar
Mogelijke Schending van de Wetenschappelijke Integriteit}. Erasmus MC;
2012.
\url{https://web.archive.org/web/20151027084205/http://www.erasmusmc.nl/5663/135857/3675250/3706798/erasmusmc.commissie.verv.onderzoek.2012}.

\hypertarget{ref-commissie2013}{}
4. Commissie Vervolgonderzoek Wetenschappelijke Integriteit 2013.
\emph{Rapport}. Erasmus MC; 2014.

\hypertarget{ref-Devereaux313}{}
5. Devereaux PJ, Beattie WS, Choi PT-L, et al. How strong is the
evidence for the use of perioperative beta blockers in non-cardiac
surgery? Systematic review and meta-analysis of randomised controlled
trials. \emph{BMJ}. 2005;331(7512):313-321.
doi:\href{https://doi.org/10.1136/bmj.38503.623646.8F}{10.1136/bmj.38503.623646.8F}.

\hypertarget{ref-Angeli2010}{}
6. Angeli F, Verdecchia P, Karthikeyan G, Mazzotta G, Gentile G, Reboldi
G. \(\beta\)-blockers reduce mortality in patients undergoing high-risk
non-cardiac surgery. \emph{American Journal of Cardiovascular Drugs}.
2010;10(4):247-259.
doi:\href{https://doi.org/10.2165/11539510-000000000-00000}{10.2165/11539510-000000000-00000}.

\hypertarget{ref-bouri2014}{}
7. Bouri S, Shun-Shin MJ, Cole GD, Mayet J, Francis DP. Meta-analysis of
secure randomised controlled trials of -blockade to prevent
perioperative death in non-cardiac surgery. \emph{Heart}.
2014;100(6):456-464.
doi:\href{https://doi.org/10.1136/heartjnl-2013-304262}{10.1136/heartjnl-2013-304262}.

\hypertarget{ref-esc2014}{}
8. ESC/ESA. Guidelines on non-cardiac surgery: Cardiovascular assessment
and management. \emph{European Heart Journal}. 2014;35:2383-2243.

\hypertarget{ref-dunkelgrun2009}{}
9. Dunkelgrun M, Boersma E, Schouten O, et al. Bisoprolol and
fluvastatin for the reduction of perioperative cardiac mortality and
myocardial infarction in intermediate-risk patients undergoing
noncardiovascular surgery: A randomized controlled trial (DECREASE-IV).
\emph{Annals of surgery}. 2009;249(6):921-926.
doi:\href{https://doi.org/10.1097/SLA.0b013e3181a77d00}{10.1097/SLA.0b013e3181a77d00}.

\hypertarget{ref-poldermans1999}{}
10. Poldermans D, Boersma E, Bax JJ, et al. The effect of bisoprolol on
perioperative mortality and myocardial infarction in High-Risk patients
undergoing vascular surgery. \emph{The New England journal of medicine}.
1999;341(24):1789-1794.
doi:\href{https://doi.org/10.1056/NEJM199912093412402}{10.1056/NEJM199912093412402}.

\hypertarget{ref-Buyse1999-jq}{}
11. Buyse M, George SL, Evans S, et al. The role of biostatistics in the
prevention, detection and treatment of fraud in clinical trials.
\emph{Statistics in medicine}. 1999;18(24):3435-3451.
doi:\href{https://doi.org/10.1002/(SICI)1097-0258(19991230)18:24\%3C3435::AID-SIM365\%3E3.0.CO;2-O}{10.1002/(SICI)1097-0258(19991230)18:24\textless{}3435::AID-SIM365\textgreater{}3.0.CO;2-O}.

\hypertarget{ref-Knepper2016-la}{}
12. Knepper D, Lindblad AS, Sharma G, et al. Statistical monitoring in
clinical trials: Best practices for detecting data anomalies suggestive
of fabrication or misconduct. \emph{Therapeutic Innovation \& Regulatory
Science}. 4\textasciitilde{}feb 2016.
doi:\href{https://doi.org/10.1177/2168479016630576}{10.1177/2168479016630576}.

\hypertarget{ref-viechtbauer2010}{}
13. Viechtbauer W. Conducting meta-analyses in R with the metafor
package. \emph{Journal of Statistical Software}. 2010;36:1-48.
\url{http://www.jstatsoft.org/v36/i03/}.

\hypertarget{ref-viechtbauer2005}{}
14. Viechtbauer W. Bias and efficiency of meta-analytic variance
estimators in the Random-Effects model. \emph{Journal of educational and
behavioral statistics: a quarterly publication sponsored by the American
Educational Research Association and the American Statistical
Association}. 2005;30(3):261-293.
doi:\href{https://doi.org/10.3102/10769986030003261}{10.3102/10769986030003261}.

\hypertarget{ref-agresti2002}{}
15. Agresti A. \emph{Categorical Data Analysis}. Hoboken, NJ: John Wiley
\& Sons Inc.; 2002.

\hypertarget{ref-peters2015}{}
16. Peters CFW, Klaassen CAJ, Wiel MA van de. \emph{Evaluating the
Scientific Veracity of Publications by Dr. Jens Forster}. University of
Amsterdam; 2015.

\hypertarget{ref-casella2002}{}
17. Casella G, Berger RL. \emph{Statistical Interference}. Pacific
Grove, CA: Duxbury; 2002.

\hypertarget{ref-ioannidis2006}{}
18. Ioannidis JPA, Trikalinos TA, Zintzaras E. Extreme between-study
homogeneity in meta-analyses could offer useful insights. \emph{Journal
of clinical epidemiology}. 2006;59(10):1023-1032.
doi:\href{https://doi.org/10.1016/j.jclinepi.2006.02.013}{10.1016/j.jclinepi.2006.02.013}.

\hypertarget{ref-klei2015}{}
19. Klei W van. Welke perioperatieve bètablokker heeft de voorkeur?
{[}Which perioperative beta-blocker is preferred?{]}. \emph{Nederlands
Tijdschrift voor Geneeskunde}. 2015;159:A9798.

\hypertarget{ref-Kristensen_2014}{}
20. Kristensen SD, Knuuti J, Saraste A, et al. 2014 ESC/ESA guidelines
on non-cardiac surgery. \emph{European Journal of Anaesthesiology}.
2014;31(10):517-573.
doi:\href{https://doi.org/10.1097/eja.0000000000000150}{10.1097/eja.0000000000000150}.

\hypertarget{ref-guarracino2015}{}
21. Guarracino F PH Baldassarri R. Revised esc/esa guidelines on
non-cardiac surgery: Cardiovascular assessment and management.
implications for preoperative clinical evaluation. \emph{Minerva
Anestesiology}. 2015;2:225-233.

\hypertarget{ref-Port2016}{}
22. Port SC. 2014 esc/esa guidelines on noncardiac surgery:
Cardiovascular assessment and management. \emph{Journal of Nuclear
Cardiology}. 2016:1-3.
doi:\href{https://doi.org/10.1007/s12350-016-0641-x}{10.1007/s12350-016-0641-x}.

\hypertarget{ref-Kristensen2016}{}
23. Kristensen SD. 2014 esc/esa guidelines on noncardiac surgery:
Cardiovascular assessment and management. \emph{Journal of Nuclear
Cardiology}. 2016:1-3.
doi:\href{https://doi.org/10.1007/s12350-016-0642-9}{10.1007/s12350-016-0642-9}.

\hypertarget{ref-van_Aert_2016}{}
24. Aert RCM van, Wicherts JM, Assen MALM van. Conducting meta-analyses
based on p values: Reservations and recommendations for applying
p-uniform and p-curve. \emph{Perspectives on Psychological Science}.
2016;11(5):713-729.
doi:\href{https://doi.org/10.1177/1745691616650874}{10.1177/1745691616650874}.

\end{document}

### 2017-02-01 submission/manuscript.Rmd
Interim commit

@@
     keep_tex: yes
+    fig_caption: yes
   word_document: default
@@
 
+# Old plotting. Saved temporarily. Ignore if forgotten to delete.
 # pdf(sprintf('%sfigures/fig3.pdf', x12), width = 8, height = 6)
@@
      ylab = 'Proportion',
-     main = 'DECREASE I',
+     main = 'DECREASE-I',
      ylim = 0:1,
@@
      ylab = '',
-     main = 'DECREASE IV',
+     main = 'DECREASE-IV',
      ylim = 0:1,
@@
 
-```{r figure 3, fig.cap = 'Test', out.width = '80%', fig.asp = .5, fig.align = 'center', echo = FALSE}
+```{r figure 3, fig.cap = 'Inversion method results used to estimate the number of data points fabricated in the DECREASE-I and DECREASE-IV trials.
+The top row panels indicate $p_F$ (y-axis) for all $X$ out of $N$ fabricated data points (x-axis).
+The bottom row indicates the estimated number of fabricated data points (y-axis) when varying the degree of confidence (x-axis).
+Dotted lines indicate the bounds for a 95 percent CI.', out.width = '80%', fig.asp = .5, fig.align = 'center', echo = FALSE}
 knitr::include_graphics(sprintf("%sfigures/fig3.pdf", x12))
@@
 
-```{r}
-res_i_certain
-res_i_uncertain
-res_iv_certain
-res_iv_uncertain
-```
+For DECREASE-I, the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_i_certain - .025) - 1` - `r which.max(res_i_certain - .975) - 1`] or [`r which.min(res_i_uncertain - .025) - 1` - `r which.max(res_i_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
+The left column of Figure 3 depicts the $p_F$ per *X* fabricated data points (top panel) and the bounds of the confidence interval when the degree of confidence is altered (lower panel).
+Staying clearly between the dotted lines in the top panel, depicting the 95\% CI (top: .975; bottom: .025), it becomes apparent that the degree of uncertainty is too high to make any estimates with a high degree of confidence.
+This due to the small sample size (i.e., $N=`r length(res_i_certain) - 1`$) combined with the availability of just the summary results.
+Only when the degree of confidence is lowered to around 75\% does the interval not span the entire sample size.
+As such, based on the summary results, little can be said about the extent of the data fabrication that occurred in the DECREASE-I trial, affirming the conclusions of the original committee report [@commissie2013].
+
+For DECREASE-IV, the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_iv_certain - .025) - 1` - `r which.max(res_iv_certain - .975) - 1`] or [`r which.min(res_iv_uncertain - .025) - 1` - `r which.max(res_iv_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
+The minor difference between the estimates indicates that there is a high degree of confidence that data fabrication did occur.
+Nonetheless, the range of potentially fabricated data points is still estimated at approximately 1000; this indicates that the summary results are insufficient to provide more than an estimated lowerbound.
+However, this does indicate that it is possible not all data were fabricated (i.e., $N=`r length(res_iv_certain) - 1`$), increasing the importance of data provenance to discern between genuine and falsified data.
+This affirms the conclusion of the scientific integrity committee that the results of the DECREASE-IV trial are "scientifically incorrect" (p.11) [@commissie2012].
+
+The results also highlight that, despite the lack of availability of raw data, summary results from larger samples result in more certainty about the estimated number of fabricated data points using the inversion method.
+This is due to decreased standard errors at the population level of the effects, providing less variation in the simulations and resulting in higher sensitivity.
+However, this also highlights that in order to prevent detection it would be in the fabricators' interest to fabricate small studies (assuming the fabricator wants to remain undetected).
 
@@
 
+Throughout steps 1, 2, and 3 we provided evidence that the , the DECREASE trials are highly unlikely to come from a genuine data pool, and that there is a high degree of confidence that data was fab
+
 <!--Move to discussion?-->
@@
 
-However, due to various types of beta-blockers, evidence might be moderated[@klei2015].
+However, due to various types of beta-blockers, evidence might be moderated [@klei2015].
 

### 2017-02-01 submission/manuscript.tex
Interim commit

@@
 \urlstyle{same}  % don't use monospace font for urls
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\ImportTok}[1]{{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
-\newcommand{\BuiltInTok}[1]{{#1}}
-\newcommand{\ExtensionTok}[1]{{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
-\newcommand{\RegionMarkerTok}[1]{{#1}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
-\newcommand{\NormalTok}[1]{{#1}}
 \usepackage{longtable,booktabs}
@@
   \postdate{\par}
-  \date{26 January, 2017}
+  \date{01 February, 2017}
 
@@
 
-\caption{Test}\label{fig:figure 3}
+\caption{Inversion method results used to estimate the number of data points fabricated in the DECREASE-I and DECREASE-IV trials.
+The top row panels indicate $p_F$ (y-axis) for all $X$ out of $N$ fabricated data points (x-axis).
+The bottom row indicates the estimated number of fabricated data points (y-axis) when varying the degree of confidence (x-axis).
+Dotted lines indicate the bounds for a 95 percent CI.}\label{fig:figure 3}
 \end{figure}
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{res_i_certain}
-\end{Highlighting}
-\end{Shaded}
+For DECREASE-I, the 95\% confidence interval for the estimated number of
+fabricated data points is {[}0 - 112{]} or {[}3 - 112{]} when based on a
+point estimate or a more uncertain distribution estimate, respectively.
+The left column of Figure 3 depicts the \(p_F\) per \emph{X} fabricated
+data points (top panel) and the bounds of the confidence interval when
+the degree of confidence is altered (lower panel). Staying clearly
+between the dotted lines in the top panel, depicting the 95\% CI (top:
+.975; bottom: .025), it becomes apparent that the degree of uncertainty
+is too high to make any estimates with a high degree of confidence. This
+due to the small sample size (i.e., \(N=112\)) combined with the
+availability of just the summary results. Only when the degree of
+confidence is lowered to around 75\% does the interval not span the
+entire sample size. As such, based on the summary results, little can be
+said about the extent of the data fabrication that occurred in the
+DECREASE-I trial, affirming the conclusions of the original committee
+report\textsuperscript{4}.
 
-\begin{verbatim}
-##   [1] 0.1507 0.1624 0.1700 0.1674 0.1792 0.1807 0.1868 0.1867 0.1879 0.1930
-##  [11] 0.2012 0.2182 0.2083 0.2114 0.2257 0.2296 0.2361 0.2476 0.2377 0.2488
-##  [21] 0.2545 0.2631 0.2742 0.2788 0.2823 0.2894 0.3002 0.2996 0.3032 0.3094
-##  [31] 0.3180 0.3243 0.3305 0.3352 0.3380 0.3547 0.3575 0.3649 0.3740 0.3873
-##  [41] 0.3852 0.3957 0.4099 0.4088 0.4113 0.4205 0.4314 0.4378 0.4430 0.4524
-##  [51] 0.4643 0.4655 0.4651 0.4851 0.4983 0.4931 0.5006 0.5106 0.5148 0.5241
-##  [61] 0.5338 0.5428 0.5528 0.5549 0.5558 0.5768 0.5752 0.5797 0.5912 0.6043
-##  [71] 0.6060 0.6177 0.6243 0.6283 0.6346 0.6433 0.6515 0.6586 0.6667 0.6684
-##  [81] 0.6787 0.6895 0.6947 0.6978 0.7021 0.7121 0.7150 0.7264 0.7287 0.7402
-##  [91] 0.7401 0.7522 0.7506 0.7533 0.7591 0.7762 0.7832 0.7782 0.7821 0.7873
-## [101] 0.7894 0.7925 0.8009 0.8095 0.8175 0.8135 0.8211 0.8251 0.8306 0.8363
-## [111] 0.8349 0.8372 0.8453
-\end{verbatim}
+For DECREASE-IV, the 95\% confidence interval for the estimated number
+of fabricated data points is {[}8 - 1059{]} or {[}7 - 1066{]} when based
+on a point estimate or a more uncertain distribution estimate,
+respectively. The minor difference between the estimates indicates that
+there is a high degree of confidence that data fabrication did occur.
+Nonetheless, the range of potentially fabricated data points is still
+estimated at approximately 1000; this indicates that the summary results
+are insufficient to provide more than an estimated lowerbound. However,
+this does indicate that it is possible not all data were fabricated
+(i.e., \(N=1066\)), increasing the importance of data provenance to
+discern between genuine and falsified data. This affirms the conclusion
+of the scientific integrity committee that the results of the
+DECREASE-IV trial are ``scientifically incorrect''
+(p.11)\textsuperscript{3}.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{res_i_uncertain}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##   [1] 0.1224 0.1247 0.1263 0.1213 0.1224 0.1318 0.1288 0.1355 0.1370 0.1419
-##  [11] 0.1502 0.1448 0.1557 0.1582 0.1619 0.1730 0.1801 0.1758 0.1796 0.1869
-##  [21] 0.1938 0.1968 0.2086 0.2106 0.2239 0.2267 0.2350 0.2414 0.2446 0.2602
-##  [31] 0.2619 0.2716 0.2697 0.2896 0.3045 0.3067 0.3302 0.3313 0.3332 0.3543
-##  [41] 0.3552 0.3801 0.3755 0.3816 0.4068 0.4056 0.4106 0.4231 0.4297 0.4384
-##  [51] 0.4573 0.4694 0.4742 0.4911 0.4864 0.5000 0.5125 0.5172 0.5247 0.5337
-##  [61] 0.5451 0.5518 0.5554 0.5640 0.5666 0.5798 0.5924 0.5896 0.6015 0.6100
-##  [71] 0.6154 0.6297 0.6251 0.6376 0.6446 0.6480 0.6616 0.6649 0.6725 0.6786
-##  [81] 0.6887 0.6913 0.6999 0.7077 0.7116 0.7248 0.7153 0.7227 0.7302 0.7379
-##  [91] 0.7432 0.7499 0.7534 0.7530 0.7567 0.7708 0.7712 0.7762 0.7723 0.7988
-## [101] 0.7908 0.7958 0.7998 0.8060 0.8082 0.8051 0.8234 0.8162 0.8240 0.8323
-## [111] 0.8445 0.8647 0.8862
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{res_iv_certain}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##    [1] 0.0223 0.0255 0.0234 0.0240 0.0263 0.0263 0.0226 0.0232 0.0220
-##   [10] 0.0234 0.0251 0.0254 0.0247 0.0234 0.0276 0.0249 0.0268 0.0244
-##   [19] 0.0244 0.0275 0.0264 0.0267 0.0267 0.0250 0.0279 0.0253 0.0254
-##   [28] 0.0268 0.0267 0.0267 0.0267 0.0257 0.0264 0.0279 0.0264 0.0274
-##   [37] 0.0285 0.0295 0.0269 0.0290 0.0301 0.0305 0.0288 0.0281 0.0271
-##   [46] 0.0293 0.0280 0.0271 0.0266 0.0289 0.0283 0.0277 0.0299 0.0283
-##   [55] 0.0296 0.0326 0.0344 0.0291 0.0321 0.0299 0.0287 0.0302 0.0289
-##   [64] 0.0301 0.0297 0.0312 0.0315 0.0291 0.0308 0.0300 0.0298 0.0319
-##   [73] 0.0283 0.0339 0.0311 0.0317 0.0313 0.0322 0.0324 0.0342 0.0302
-##   [82] 0.0334 0.0332 0.0340 0.0334 0.0328 0.0349 0.0334 0.0348 0.0339
-##   [91] 0.0339 0.0326 0.0350 0.0324 0.0353 0.0307 0.0339 0.0386 0.0379
-##  [100] 0.0343 0.0347 0.0356 0.0370 0.0358 0.0343 0.0345 0.0356 0.0373
-##  [109] 0.0399 0.0382 0.0376 0.0372 0.0376 0.0383 0.0374 0.0391 0.0349
-##  [118] 0.0389 0.0377 0.0406 0.0443 0.0394 0.0402 0.0415 0.0406 0.0415
-##  [127] 0.0412 0.0433 0.0430 0.0376 0.0399 0.0444 0.0418 0.0445 0.0464
-##  [136] 0.0450 0.0461 0.0421 0.0447 0.0398 0.0418 0.0440 0.0461 0.0463
-##  [145] 0.0500 0.0447 0.0438 0.0439 0.0492 0.0465 0.0464 0.0466 0.0487
-##  [154] 0.0482 0.0467 0.0530 0.0501 0.0466 0.0510 0.0470 0.0540 0.0517
-##  [163] 0.0542 0.0523 0.0549 0.0488 0.0509 0.0550 0.0550 0.0525 0.0541
-##  [172] 0.0522 0.0579 0.0551 0.0548 0.0533 0.0590 0.0573 0.0557 0.0572
-##  [181] 0.0548 0.0531 0.0598 0.0566 0.0613 0.0595 0.0599 0.0545 0.0627
-##  [190] 0.0581 0.0625 0.0603 0.0643 0.0595 0.0650 0.0621 0.0657 0.0673
-##  [199] 0.0648 0.0666 0.0684 0.0676 0.0654 0.0673 0.0679 0.0702 0.0687
-##  [208] 0.0749 0.0713 0.0731 0.0696 0.0709 0.0702 0.0719 0.0764 0.0758
-##  [217] 0.0733 0.0790 0.0725 0.0773 0.0770 0.0765 0.0778 0.0822 0.0789
-##  [226] 0.0801 0.0814 0.0858 0.0820 0.0821 0.0854 0.0803 0.0824 0.0831
-##  [235] 0.0864 0.0895 0.0845 0.0943 0.0883 0.0963 0.0930 0.0919 0.0871
-##  [244] 0.0921 0.0940 0.0896 0.0967 0.0941 0.1000 0.0969 0.0980 0.0951
-##  [253] 0.0988 0.1051 0.1031 0.1013 0.1059 0.0974 0.1055 0.1064 0.1074
-##  [262] 0.1058 0.1068 0.1064 0.1117 0.1100 0.1069 0.1174 0.1078 0.1120
-##  [271] 0.1107 0.1124 0.1180 0.1165 0.1190 0.1238 0.1187 0.1207 0.1219
-##  [280] 0.1310 0.1247 0.1233 0.1288 0.1228 0.1319 0.1248 0.1337 0.1326
-##  [289] 0.1339 0.1346 0.1329 0.1365 0.1366 0.1443 0.1395 0.1391 0.1377
-##  [298] 0.1433 0.1424 0.1435 0.1466 0.1503 0.1519 0.1494 0.1459 0.1519
-##  [307] 0.1578 0.1559 0.1500 0.1585 0.1550 0.1520 0.1651 0.1632 0.1662
-##  [316] 0.1605 0.1623 0.1672 0.1663 0.1654 0.1730 0.1665 0.1721 0.1663
-##  [325] 0.1789 0.1690 0.1754 0.1792 0.1782 0.1825 0.1827 0.1846 0.1820
-##  [334] 0.1861 0.1884 0.1893 0.1936 0.1931 0.1956 0.1903 0.1942 0.1914
-##  [343] 0.1901 0.2079 0.2054 0.2033 0.1976 0.2029 0.2099 0.2079 0.2084
-##  [352] 0.2212 0.2063 0.2079 0.2232 0.2101 0.2222 0.2166 0.2184 0.2144
-##  [361] 0.2220 0.2219 0.2351 0.2272 0.2271 0.2311 0.2337 0.2291 0.2382
-##  [370] 0.2409 0.2362 0.2391 0.2327 0.2468 0.2372 0.2416 0.2520 0.2460
-##  [379] 0.2548 0.2598 0.2457 0.2585 0.2620 0.2636 0.2608 0.2623 0.2661
-##  [388] 0.2669 0.2655 0.2664 0.2662 0.2679 0.2702 0.2770 0.2752 0.2769
-##  [397] 0.2836 0.2842 0.2828 0.2806 0.2860 0.2825 0.2788 0.2874 0.2987
-##  [406] 0.2879 0.2975 0.2895 0.2920 0.3043 0.3040 0.3044 0.3065 0.3083
-##  [415] 0.3066 0.3182 0.3098 0.3150 0.3167 0.3066 0.3137 0.3200 0.3153
-##  [424] 0.3271 0.3237 0.3273 0.3252 0.3359 0.3381 0.3298 0.3320 0.3384
-##  [433] 0.3328 0.3353 0.3417 0.3377 0.3521 0.3491 0.3432 0.3503 0.3540
-##  [442] 0.3581 0.3569 0.3499 0.3632 0.3561 0.3652 0.3658 0.3703 0.3630
-##  [451] 0.3637 0.3745 0.3743 0.3858 0.3751 0.3728 0.3801 0.3978 0.3775
-##  [460] 0.3776 0.3882 0.3847 0.3893 0.3876 0.3899 0.3889 0.3906 0.3921
-##  [469] 0.4023 0.4096 0.4099 0.4129 0.4029 0.4086 0.4096 0.4102 0.4108
-##  [478] 0.4269 0.4119 0.4128 0.4277 0.4196 0.4194 0.4236 0.4307 0.4253
-##  [487] 0.4279 0.4335 0.4413 0.4325 0.4449 0.4420 0.4371 0.4446 0.4488
-##  [496] 0.4494 0.4533 0.4548 0.4574 0.4723 0.4585 0.4768 0.4649 0.4670
-##  [505] 0.4617 0.4672 0.4727 0.4755 0.4753 0.4681 0.4862 0.4814 0.4835
-##  [514] 0.4873 0.4832 0.4794 0.4749 0.4809 0.4973 0.4886 0.4986 0.4971
-##  [523] 0.5020 0.4994 0.5077 0.5026 0.5016 0.5028 0.5094 0.5111 0.5157
-##  [532] 0.5163 0.5252 0.5159 0.5145 0.5301 0.5237 0.5264 0.5364 0.5265
-##  [541] 0.5337 0.5342 0.5420 0.5380 0.5473 0.5335 0.5496 0.5520 0.5463
-##  [550] 0.5514 0.5471 0.5465 0.5572 0.5573 0.5569 0.5582 0.5618 0.5581
-##  [559] 0.5711 0.5598 0.5705 0.5692 0.5758 0.5841 0.5741 0.5778 0.5785
-##  [568] 0.5856 0.5898 0.5809 0.5956 0.5882 0.5916 0.6016 0.5906 0.5955
-##  [577] 0.6025 0.6019 0.5995 0.5995 0.6165 0.6099 0.6058 0.6030 0.6195
-##  [586] 0.6180 0.6142 0.6274 0.6210 0.6202 0.6308 0.6285 0.6253 0.6296
-##  [595] 0.6348 0.6360 0.6446 0.6328 0.6408 0.6404 0.6390 0.6490 0.6529
-##  [604] 0.6408 0.6524 0.6547 0.6440 0.6573 0.6582 0.6538 0.6579 0.6646
-##  [613] 0.6699 0.6688 0.6707 0.6745 0.6714 0.6767 0.6739 0.6773 0.6863
-##  [622] 0.6772 0.6827 0.6887 0.6903 0.6890 0.6941 0.7070 0.6965 0.6978
-##  [631] 0.7017 0.7021 0.7107 0.6981 0.7082 0.7076 0.7137 0.7035 0.7197
-##  [640] 0.7202 0.7168 0.7055 0.7230 0.7291 0.7289 0.7301 0.7281 0.7277
-##  [649] 0.7348 0.7312 0.7336 0.7395 0.7424 0.7412 0.7415 0.7463 0.7491
-##  [658] 0.7462 0.7457 0.7517 0.7573 0.7553 0.7580 0.7545 0.7525 0.7570
-##  [667] 0.7567 0.7598 0.7663 0.7682 0.7717 0.7640 0.7707 0.7705 0.7758
-##  [676] 0.7772 0.7756 0.7817 0.7779 0.7849 0.7966 0.7825 0.7862 0.7937
-##  [685] 0.7902 0.7972 0.7933 0.7858 0.7964 0.7949 0.7984 0.8011 0.7970
-##  [694] 0.8021 0.8076 0.8037 0.8106 0.8091 0.8119 0.8136 0.8124 0.8093
-##  [703] 0.8086 0.8218 0.8195 0.8156 0.8202 0.8175 0.8287 0.8196 0.8284
-##  [712] 0.8299 0.8276 0.8316 0.8237 0.8336 0.8291 0.8418 0.8362 0.8353
-##  [721] 0.8376 0.8431 0.8465 0.8381 0.8416 0.8450 0.8416 0.8426 0.8440
-##  [730] 0.8507 0.8440 0.8547 0.8493 0.8567 0.8528 0.8445 0.8588 0.8619
-##  [739] 0.8548 0.8633 0.8534 0.8626 0.8626 0.8691 0.8658 0.8619 0.8686
-##  [748] 0.8607 0.8693 0.8702 0.8726 0.8731 0.8749 0.8682 0.8763 0.8730
-##  [757] 0.8717 0.8712 0.8751 0.8813 0.8830 0.8812 0.8811 0.8749 0.8830
-##  [766] 0.8878 0.8838 0.8891 0.8871 0.8890 0.8866 0.8868 0.8898 0.8901
-##  [775] 0.8863 0.8902 0.8962 0.8903 0.8947 0.8978 0.8944 0.8937 0.8985
-##  [784] 0.9033 0.8936 0.8980 0.8984 0.9069 0.9011 0.8996 0.8956 0.9031
-##  [793] 0.9007 0.8993 0.9015 0.9057 0.9060 0.9110 0.9130 0.9038 0.9040
-##  [802] 0.9148 0.9115 0.9110 0.9132 0.9139 0.9179 0.9140 0.9089 0.9152
-##  [811] 0.9162 0.9185 0.9105 0.9173 0.9160 0.9187 0.9176 0.9189 0.9184
-##  [820] 0.9203 0.9191 0.9229 0.9214 0.9153 0.9251 0.9215 0.9259 0.9267
-##  [829] 0.9177 0.9222 0.9312 0.9247 0.9366 0.9263 0.9303 0.9284 0.9276
-##  [838] 0.9307 0.9288 0.9293 0.9293 0.9359 0.9324 0.9297 0.9346 0.9323
-##  [847] 0.9352 0.9317 0.9365 0.9337 0.9354 0.9308 0.9316 0.9351 0.9386
-##  [856] 0.9377 0.9344 0.9406 0.9388 0.9386 0.9362 0.9395 0.9396 0.9396
-##  [865] 0.9381 0.9383 0.9376 0.9385 0.9422 0.9423 0.9404 0.9430 0.9417
-##  [874] 0.9422 0.9380 0.9408 0.9443 0.9423 0.9462 0.9469 0.9467 0.9440
-##  [883] 0.9466 0.9462 0.9533 0.9399 0.9487 0.9491 0.9492 0.9480 0.9446
-##  [892] 0.9450 0.9462 0.9528 0.9501 0.9479 0.9495 0.9484 0.9501 0.9504
-##  [901] 0.9486 0.9540 0.9497 0.9506 0.9538 0.9524 0.9517 0.9547 0.9528
-##  [910] 0.9522 0.9551 0.9545 0.9558 0.9550 0.9537 0.9496 0.9515 0.9547
-##  [919] 0.9542 0.9546 0.9520 0.9533 0.9557 0.9555 0.9530 0.9516 0.9603
-##  [928] 0.9574 0.9584 0.9553 0.9574 0.9586 0.9584 0.9573 0.9540 0.9619
-##  [937] 0.9597 0.9583 0.9554 0.9599 0.9586 0.9595 0.9598 0.9620 0.9625
-##  [946] 0.9624 0.9638 0.9605 0.9597 0.9616 0.9631 0.9603 0.9644 0.9626
-##  [955] 0.9598 0.9601 0.9647 0.9657 0.9643 0.9647 0.9619 0.9639 0.9644
-##  [964] 0.9620 0.9635 0.9673 0.9666 0.9671 0.9658 0.9644 0.9648 0.9688
-##  [973] 0.9643 0.9638 0.9624 0.9648 0.9658 0.9657 0.9673 0.9637 0.9679
-##  [982] 0.9681 0.9678 0.9677 0.9683 0.9684 0.9658 0.9706 0.9703 0.9677
-##  [991] 0.9661 0.9655 0.9676 0.9691 0.9698 0.9702 0.9675 0.9672 0.9725
-## [1000] 0.9708 0.9704 0.9706 0.9707 0.9703 0.9679 0.9640 0.9719 0.9720
-## [1009] 0.9712 0.9698 0.9710 0.9732 0.9711 0.9724 0.9704 0.9704 0.9698
-## [1018] 0.9714 0.9717 0.9699 0.9720 0.9709 0.9748 0.9730 0.9718 0.9738
-## [1027] 0.9721 0.9722 0.9709 0.9705 0.9683 0.9743 0.9737 0.9710 0.9735
-## [1036] 0.9735 0.9758 0.9720 0.9747 0.9756 0.9686 0.9750 0.9740 0.9748
-## [1045] 0.9740 0.9746 0.9752 0.9734 0.9760 0.9744 0.9736 0.9752 0.9767
-## [1054] 0.9764 0.9734 0.9788 0.9773 0.9754 0.9750 0.9791 0.9791 0.9772
-## [1063] 0.9789 0.9760 0.9789 0.9758 0.9777
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{res_iv_uncertain}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##    [1] 0.0203 0.0200 0.0220 0.0211 0.0236 0.0200 0.0208 0.0193 0.0220
-##   [10] 0.0222 0.0224 0.0228 0.0194 0.0218 0.0226 0.0221 0.0220 0.0210
-##   [19] 0.0234 0.0216 0.0222 0.0242 0.0229 0.0254 0.0216 0.0225 0.0218
-##   [28] 0.0224 0.0215 0.0221 0.0227 0.0221 0.0238 0.0233 0.0259 0.0243
-##   [37] 0.0227 0.0255 0.0242 0.0216 0.0226 0.0210 0.0263 0.0218 0.0228
-##   [46] 0.0232 0.0242 0.0214 0.0247 0.0252 0.0255 0.0240 0.0251 0.0252
-##   [55] 0.0238 0.0246 0.0237 0.0265 0.0262 0.0272 0.0244 0.0262 0.0226
-##   [64] 0.0245 0.0276 0.0243 0.0225 0.0251 0.0277 0.0297 0.0228 0.0263
-##   [73] 0.0256 0.0274 0.0239 0.0259 0.0282 0.0287 0.0270 0.0267 0.0270
-##   [82] 0.0265 0.0298 0.0298 0.0275 0.0285 0.0277 0.0278 0.0275 0.0269
-##   [91] 0.0307 0.0260 0.0309 0.0273 0.0286 0.0292 0.0291 0.0334 0.0316
-##  [100] 0.0297 0.0311 0.0299 0.0305 0.0296 0.0308 0.0303 0.0284 0.0297
-##  [109] 0.0285 0.0274 0.0320 0.0287 0.0312 0.0306 0.0325 0.0300 0.0297
-##  [118] 0.0306 0.0332 0.0333 0.0323 0.0319 0.0337 0.0306 0.0335 0.0354
-##  [127] 0.0306 0.0338 0.0291 0.0314 0.0321 0.0346 0.0338 0.0365 0.0323
-##  [136] 0.0327 0.0336 0.0341 0.0351 0.0372 0.0327 0.0319 0.0365 0.0321
-##  [145] 0.0360 0.0339 0.0350 0.0364 0.0378 0.0344 0.0350 0.0362 0.0337
-##  [154] 0.0364 0.0352 0.0348 0.0350 0.0355 0.0373 0.0383 0.0367 0.0369
-##  [163] 0.0397 0.0360 0.0396 0.0432 0.0398 0.0409 0.0412 0.0438 0.0401
-##  [172] 0.0407 0.0390 0.0422 0.0397 0.0434 0.0413 0.0418 0.0416 0.0410
-##  [181] 0.0454 0.0402 0.0401 0.0429 0.0409 0.0427 0.0425 0.0461 0.0428
-##  [190] 0.0479 0.0466 0.0436 0.0489 0.0464 0.0476 0.0469 0.0458 0.0454
-##  [199] 0.0484 0.0445 0.0465 0.0466 0.0506 0.0487 0.0502 0.0494 0.0483
-##  [208] 0.0483 0.0518 0.0550 0.0493 0.0508 0.0509 0.0535 0.0542 0.0502
-##  [217] 0.0495 0.0510 0.0503 0.0486 0.0500 0.0530 0.0521 0.0534 0.0514
-##  [226] 0.0545 0.0527 0.0581 0.0545 0.0619 0.0589 0.0595 0.0611 0.0572
-##  [235] 0.0631 0.0582 0.0592 0.0607 0.0587 0.0571 0.0606 0.0600 0.0653
-##  [244] 0.0645 0.0675 0.0614 0.0620 0.0682 0.0712 0.0648 0.0654 0.0677
-##  [253] 0.0719 0.0650 0.0737 0.0692 0.0707 0.0736 0.0731 0.0748 0.0746
-##  [262] 0.0789 0.0752 0.0760 0.0775 0.0749 0.0767 0.0753 0.0821 0.0847
-##  [271] 0.0832 0.0814 0.0853 0.0878 0.0872 0.0870 0.0836 0.0862 0.0866
-##  [280] 0.0869 0.0946 0.0909 0.0922 0.0960 0.0984 0.0901 0.0972 0.0955
-##  [289] 0.1010 0.0977 0.0983 0.0986 0.1055 0.1048 0.1016 0.1065 0.1095
-##  [298] 0.1046 0.1085 0.1134 0.1145 0.1135 0.1139 0.1188 0.1217 0.1133
-##  [307] 0.1179 0.1158 0.1245 0.1229 0.1231 0.1235 0.1245 0.1212 0.1270
-##  [316] 0.1278 0.1325 0.1304 0.1370 0.1337 0.1349 0.1355 0.1378 0.1432
-##  [325] 0.1420 0.1479 0.1468 0.1422 0.1472 0.1499 0.1557 0.1567 0.1587
-##  [334] 0.1564 0.1533 0.1631 0.1646 0.1596 0.1598 0.1548 0.1736 0.1659
-##  [343] 0.1752 0.1718 0.1753 0.1759 0.1755 0.1795 0.1797 0.1881 0.1829
-##  [352] 0.1843 0.1945 0.1868 0.1933 0.1931 0.1848 0.2012 0.1943 0.2025
-##  [361] 0.1993 0.2069 0.2048 0.2036 0.2108 0.2068 0.2143 0.2092 0.2164
-##  [370] 0.2184 0.2206 0.2221 0.2288 0.2289 0.2326 0.2314 0.2369 0.2359
-##  [379] 0.2324 0.2386 0.2393 0.2404 0.2439 0.2385 0.2511 0.2413 0.2440
-##  [388] 0.2543 0.2580 0.2559 0.2518 0.2481 0.2623 0.2649 0.2605 0.2628
-##  [397] 0.2636 0.2688 0.2772 0.2788 0.2757 0.2775 0.2756 0.2785 0.2773
-##  [406] 0.2870 0.2854 0.2948 0.2991 0.2885 0.3011 0.2991 0.2939 0.2976
-##  [415] 0.2907 0.3048 0.3077 0.3064 0.3135 0.3096 0.3175 0.3294 0.3119
-##  [424] 0.3192 0.3161 0.3245 0.3258 0.3236 0.3325 0.3279 0.3396 0.3315
-##  [433] 0.3313 0.3337 0.3421 0.3384 0.3432 0.3402 0.3376 0.3558 0.3432
-##  [442] 0.3524 0.3501 0.3479 0.3537 0.3587 0.3659 0.3627 0.3582 0.3616
-##  [451] 0.3725 0.3635 0.3788 0.3722 0.3642 0.3854 0.3723 0.3742 0.3799
-##  [460] 0.3859 0.3866 0.3885 0.3952 0.3940 0.3910 0.3908 0.3938 0.4019
-##  [469] 0.4070 0.3966 0.3980 0.4045 0.3993 0.4055 0.4053 0.4133 0.4179
-##  [478] 0.4192 0.4171 0.4168 0.4247 0.4202 0.4188 0.4326 0.4260 0.4196
-##  [487] 0.4277 0.4336 0.4303 0.4257 0.4372 0.4404 0.4395 0.4392 0.4327
-##  [496] 0.4418 0.4494 0.4461 0.4390 0.4495 0.4590 0.4546 0.4537 0.4519
-##  [505] 0.4460 0.4540 0.4534 0.4698 0.4608 0.4616 0.4593 0.4614 0.4659
-##  [514] 0.4682 0.4629 0.4585 0.4783 0.4775 0.4741 0.4782 0.4802 0.4769
-##  [523] 0.4790 0.4801 0.4823 0.4851 0.4874 0.4939 0.4923 0.4940 0.4911
-##  [532] 0.4889 0.4970 0.4955 0.5035 0.5017 0.5049 0.5121 0.5016 0.5049
-##  [541] 0.5041 0.5053 0.5104 0.5097 0.5230 0.5097 0.5204 0.5032 0.5089
-##  [550] 0.5190 0.5213 0.5199 0.5272 0.5320 0.5313 0.5231 0.5257 0.5276
-##  [559] 0.5318 0.5321 0.5390 0.5372 0.5319 0.5322 0.5459 0.5396 0.5486
-##  [568] 0.5418 0.5477 0.5511 0.5489 0.5570 0.5658 0.5614 0.5556 0.5489
-##  [577] 0.5569 0.5605 0.5567 0.5614 0.5672 0.5694 0.5661 0.5718 0.5689
-##  [586] 0.5719 0.5755 0.5691 0.5735 0.5779 0.5802 0.5864 0.5805 0.5809
-##  [595] 0.5763 0.5703 0.5746 0.5821 0.5854 0.5916 0.5892 0.5919 0.5948
-##  [604] 0.5991 0.5898 0.6004 0.5983 0.5961 0.6088 0.6063 0.6072 0.6078
-##  [613] 0.6058 0.6243 0.6124 0.6114 0.6168 0.6248 0.6158 0.6283 0.6155
-##  [622] 0.6222 0.6186 0.6247 0.6250 0.6258 0.6222 0.6306 0.6247 0.6238
-##  [631] 0.6276 0.6392 0.6426 0.6351 0.6461 0.6339 0.6489 0.6414 0.6406
-##  [640] 0.6401 0.6465 0.6515 0.6560 0.6574 0.6483 0.6600 0.6504 0.6608
-##  [649] 0.6518 0.6617 0.6599 0.6639 0.6545 0.6692 0.6686 0.6675 0.6731
-##  [658] 0.6696 0.6732 0.6681 0.6755 0.6783 0.6748 0.6822 0.6804 0.6802
-##  [667] 0.6843 0.6893 0.6954 0.6899 0.6880 0.6819 0.6996 0.6906 0.6864
-##  [676] 0.6991 0.6962 0.7011 0.6999 0.7065 0.7034 0.7037 0.7043 0.7149
-##  [685] 0.7074 0.7106 0.7080 0.7146 0.7199 0.7260 0.7231 0.7202 0.7155
-##  [694] 0.7211 0.7289 0.7250 0.7178 0.7237 0.7371 0.7387 0.7330 0.7391
-##  [703] 0.7387 0.7426 0.7294 0.7346 0.7387 0.7457 0.7472 0.7480 0.7514
-##  [712] 0.7501 0.7521 0.7467 0.7479 0.7546 0.7498 0.7546 0.7549 0.7535
-##  [721] 0.7552 0.7542 0.7634 0.7556 0.7590 0.7643 0.7677 0.7712 0.7702
-##  [730] 0.7744 0.7708 0.7688 0.7768 0.7713 0.7727 0.7798 0.7861 0.7800
-##  [739] 0.7825 0.7847 0.7816 0.7846 0.7856 0.7929 0.7954 0.7891 0.7898
-##  [748] 0.7843 0.7915 0.7953 0.8005 0.8009 0.7934 0.8031 0.8034 0.8073
-##  [757] 0.8010 0.8072 0.8125 0.8056 0.8151 0.8141 0.8083 0.8074 0.8020
-##  [766] 0.8125 0.8152 0.8203 0.8119 0.8167 0.8143 0.8127 0.8193 0.8193
-##  [775] 0.8188 0.8210 0.8293 0.8235 0.8259 0.8318 0.8239 0.8277 0.8241
-##  [784] 0.8374 0.8307 0.8342 0.8323 0.8373 0.8370 0.8360 0.8375 0.8330
-##  [793] 0.8388 0.8415 0.8426 0.8421 0.8426 0.8472 0.8337 0.8404 0.8430
-##  [802] 0.8473 0.8423 0.8483 0.8408 0.8470 0.8522 0.8498 0.8496 0.8439
-##  [811] 0.8536 0.8491 0.8451 0.8551 0.8571 0.8538 0.8518 0.8539 0.8588
-##  [820] 0.8545 0.8589 0.8673 0.8547 0.8616 0.8602 0.8636 0.8658 0.8603
-##  [829] 0.8707 0.8607 0.8673 0.8674 0.8672 0.8673 0.8656 0.8644 0.8654
-##  [838] 0.8678 0.8683 0.8651 0.8743 0.8706 0.8752 0.8745 0.8716 0.8723
-##  [847] 0.8762 0.8783 0.8749 0.8763 0.8734 0.8802 0.8745 0.8727 0.8775
-##  [856] 0.8779 0.8788 0.8765 0.8786 0.8780 0.8740 0.8793 0.8816 0.8834
-##  [865] 0.8815 0.8813 0.8800 0.8875 0.8832 0.8805 0.8880 0.8828 0.8852
-##  [874] 0.8851 0.8911 0.8855 0.8886 0.8855 0.8962 0.8932 0.8837 0.8907
-##  [883] 0.8907 0.8872 0.8886 0.8920 0.8978 0.8904 0.8922 0.8960 0.8917
-##  [892] 0.8955 0.8886 0.8929 0.8923 0.8923 0.8928 0.9007 0.8939 0.8925
-##  [901] 0.8950 0.8953 0.8870 0.8895 0.8952 0.8948 0.8958 0.9033 0.8991
-##  [910] 0.9003 0.9008 0.9035 0.8995 0.8957 0.9016 0.8998 0.9013 0.9023
-##  [919] 0.9000 0.8962 0.8934 0.8994 0.9051 0.9051 0.9016 0.9038 0.9058
-##  [928] 0.9060 0.9030 0.8974 0.9083 0.9008 0.9028 0.9042 0.9031 0.9008
-##  [937] 0.9063 0.9012 0.9014 0.9014 0.9046 0.9080 0.9093 0.9086 0.9109
-##  [946] 0.9033 0.9070 0.9033 0.9042 0.9046 0.9081 0.9080 0.9026 0.9049
-##  [955] 0.9064 0.9106 0.9099 0.9087 0.9089 0.9128 0.9109 0.9149 0.9105
-##  [964] 0.9071 0.9134 0.9076 0.9127 0.9090 0.9074 0.9090 0.9049 0.9083
-##  [973] 0.9094 0.9086 0.9107 0.9043 0.9126 0.9056 0.9114 0.9050 0.9095
-##  [982] 0.9147 0.9069 0.9124 0.9130 0.9139 0.9124 0.9100 0.9114 0.9109
-##  [991] 0.9114 0.9096 0.9103 0.9110 0.9104 0.9137 0.9121 0.9109 0.9114
-## [1000] 0.9102 0.9130 0.9104 0.9137 0.9158 0.9171 0.9108 0.9086 0.9145
-## [1009] 0.9174 0.9133 0.9078 0.9137 0.9143 0.9102 0.9157 0.9169 0.9151
-## [1018] 0.9154 0.9210 0.9158 0.9171 0.9167 0.9200 0.9166 0.9122 0.9182
-## [1027] 0.9162 0.9106 0.9196 0.9195 0.9147 0.9214 0.9211 0.9153 0.9203
-## [1036] 0.9168 0.9189 0.9190 0.9158 0.9210 0.9195 0.9240 0.9176 0.9144
-## [1045] 0.9218 0.9140 0.9178 0.9170 0.9213 0.9195 0.9177 0.9200 0.9200
-## [1054] 0.9204 0.9150 0.9173 0.9172 0.9197 0.9206 0.9148 0.9208 0.9225
-## [1063] 0.9242 0.9289 0.9358 0.9499 0.9789
-\end{verbatim}
+The results also highlight that, despite the lack of availability of raw
+data, summary results from larger samples result in more certainty about
+the estimated number of fabricated data points using the inversion
+method. This is due to decreased standard errors at the population level
+of the effects, providing less variation in the simulations and
+resulting in higher sensitivity. However, this also highlights that in
+order to prevent detection it would be in the fabricators' interest to
+fabricate small studies (assuming the fabricator wants to remain
+undetected).
 
 \section{Discussion}\label{discussion}
+
+Throughout steps 1, 2, and 3 we provided evidence that the , the
+DECREASE trials are highly unlikely to come from a genuine data pool,
+and that there is a high degree of confidence that data was fab
 
@@
 \emph{Rapport}. Erasmus MC; 2014.
+\url{http://web.archive.org/web/20161104135848/http://www.erasmusmc.nl/cs-research/bijlagen/integriteit/eindrapport2014nl}.
 

### 2017-02-10 submission/manuscript.Rmd
Add coding of type of beta-blocker to data, revise intro.

@@
   pdf_document:
+    fig_caption: yes
     keep_tex: yes
-    fig_caption: yes
+  html_document: default
   word_document: default
@@
 3. Why this study. -->
-The effect of beta-blockers on perioperative mortality in non-cardiac surgery for patients with coronary artery disease (CAD) has been subject of discussion [@Coleg5210] due to findings of research  misconduct [@commissie2011;@commissie2012;@commissie2013] that were central to claims of their effectiveness [@Devereaux313;@Angeli2010;@bouri2014].
+The effect of beta-blockers on perioperative mortality in non-cardiac surgery for patients with coronary artery disease (CAD) has been subject of discussion [@Coleg5210] due to findings of research  misconduct [@commissie2011;@commissie2012;@commissie2013].
+Consequently, it appeared that the related trials were central to claims of the effectiveness of beta-blocker in decreasing perioperative mortality [@Devereaux313;@Angeli2010;@bouri2014].
 Perioperative mortality is the deathrate of patients during the period of the surgical procedure, which typically includes admission, anaesthesia, surgery, and recovery.
-This related finding of research misconduct altered meta-analytic results of perioperative mortality in non-cardiac surgery for CAD patients substantively[@bouri2014], where meta-analyses that include these studies conclude that beta-blockers decrease perioperative mortality [@Devereaux313;@Angeli2010;@bouri2014] but a meta-analysis that excludes these studies concludes that beta-blockers increase perioperative mortality[@bouri2014].
+Meta-analyses that include these studies subject to research misconduct conclude that beta-blockers decrease perioperative mortality [@Devereaux313;@Angeli2010;@bouri2014] whereas a meta-analysis that excludes the studies subject to research misconduct concludes that beta-blockers increase perioperative mortality [@bouri2014].
 
-The trials deemed to be the result of research misconduct were the Dutch DECREASE trials[@commissie2011;@commissie2012;@commissie2013].
-The committees that investigated the integrity of the DECREASE studies reported that data fabrication was likely [@commissie2011;@commissie2012;@commissie2013], but that the extent of the data fabrication remained unclear.
-Given that the inclusion or exclusion of these trials in meta-analyses is crucial in determining whether beta-blockers are indeed effective during the perioperative period and the latest guidelines still recommend the usage of beta-blockers [@esc2014], we aim to estimate the extent of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999] in order to further this debate.
+The trials subject to research misconduct were the Dutch DECREASE trials [@commissie2011;@commissie2012;@commissie2013].
+The committees that investigated the integrity of the DECREASE studies reported that data fabrication was likely but that the extent of the data fabrication remained unclear [@commissie2011;@commissie2012;@commissie2013].
+Given that the inclusion or exclusion of these trials in meta-analyses is crucial in determining whether beta-blockers are an effective treatment and considering that the latest guidelines still recommend the usage of beta-blockers [@esc2014], we aim to estimate the extent of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999] in order to further the debate.
 
-The reports on the DECREASE trials primarily focused on the provenance of the raw data and patient files (e.g., informed consent, whether data corresponded to patient files), but neglected the extent to which the DECREASE studies deviated from similar trials that tested the effectiveness of beta-blockers on perioperative mortality.
-Comparing how (dis)similar trials are from comparable studies is not a new idea [@Buyse1999-jq] and has proven effective in detecting data fabrication [@Knepper2016-la].
-This is a different approach from the forensic statistical methods previously applied [@commissie2013], where results within one study are evaluated as opposed to other similar studies.
-This explains the discrepancy between the feasibility of evaluating the DECREASE trials by the statistical expert of the committee report and ours.
+The reports on the integrity of the DECREASE trials primarily focused on the provenance of the raw data and patient files (e.g., informed consent, whether data corresponded to patient files) but did not investigate the extent to which the DECREASE trials deviated from comparable trials.
+The committee reports did not neglect statistical evaluation; a statistical expert evaluated the applicability of forensic statistical methods [@commissie2013] to evaluate results within one trial.
+However, comparing across trials can be used as a forensic statistical method as well.
+This method has previously been used to monitor trial data quality or to test for potential data anomalies [@Buyse1999-jq].
+Moreover, this method has previously proven to be effective in detecting data fabrication [@Knepper2016-la].
+As such, evaluating the DECREASE trials against comparable trials investigating the effectiveness of beta-blockers with respect to perioperative mortality could prove worthwhile.
 
-<!-- Hier evt nog wat vertellen over FOI? -->
-<!-- Finally, we share our unsuccesful experiences in trying to further investigate the issue by requesting the DECREASE-VI raw data and call for more transparency. -->
 To statistically investigate the evidence of data fabrication in the results of the DECREASE studies [@dunkelgrun2009;@poldermans1999], we took three steps.
-First, we replicate the findings from a 2014 meta-analysis [@bouri2014] that contains sufficient information to estimate the deviation of the DECREASE-I and DECREASE-IV studies from the remaining studies.
-Second, we evaluate the veracity of the DECREASE-I and DECREASE-IV studies (i.e., the probability that these study findings occurred assuming no data fabrication occurred).
-Third, we reverse this assumption and assume that data fabrication did occur and estimate how many data points would have to be fabricated to reproduce the results of the DECREASE-I and DECREASE-IV studies if the remaining studies are regarded as estimating the true effect of beta-blockers on perioperative mortality.
+First, we reproduced the findings from a 2014 meta-analysis [@bouri2014] that contained sufficient information to estimate the deviation of the DECREASE-I and DECREASE-IV studies from the remaining studies.
+Second, we evaluated the veracity of the DECREASE-I and DECREASE-IV studies (i.e., the probability that these study findings occurred assuming no data fabrication occurred).
+Third, we reversed this assumption and assumed that data fabrication did occur and estimate how many data points would have to be fabricated to reproduce the results of the DECREASE-I and DECREASE-IV studies if the remaining studies are regarded as estimating the true effect of beta-blockers on perioperative mortality.
 
@@
 This ensured that (1) their results are reproducible and (2) we are using the correct estimates in subsequent steps of our analyses.
-Using figures 2 and 3 from the original paper [@bouri2014], we extracted the raw event data for the 2 (control vs experimental) by 2 (event or no event) design, which we used to recompute the natural logarithm of the risk ratio and its standard error.
+Using figures 2 and 3 from the original paper [@bouri2014], we extracted the raw event data for the 2 (control vs experimental) by 2 (event vs no event) design, which we used to recompute the natural logarithm of the risk ratio and its standard error.
 The extracted event data is available at [osf.io/aykeh](https://osf.io/aykeh) and our analysis plan was preregistered at [osf.io/vnmzc](https://osf.io/vnmzc).
@@
 x12 <- '../'
-# x12 <- ''
+x12 <- ''
 bouri_data <- read.csv(sprintf('%sdata/bouri_data.csv', x12))
@@
 bouri_data$d <- bouri_data$beta_event
+
+bouri_data$beta_blocker <- c("atenolol",
+                             "propanolol",
+                             "metoprolol",
+                             "metoprolol",
+                             "metoprolol",
+                             "atenolol",
+                             "bisoprolol",
+                             "metoprolol",
+                             "metoprolol",
+                             "bisoprolol",
+                             "bisoprolol")
 

### 2017-02-10 submission/manuscript.Rmd
Revise steps

@@
 
-Second, we meta-analyzed all studies combined, including a dummy predictor for the DECREASE and non-DECREASE studies to replicate results presented in Figure 4 of the 2014 meta-analysis [@bouri2014].
+Second, we meta-analyzed all studies combined, including a dummy predictor for the DECREASE and non-DECREASE studies to reproduce results presented in Figure 4 of the 2014 meta-analysis [@bouri2014].
 Surprisingly, our results showed more evidence for unequal subgroups than the original meta-analysis [@bouri2014]; whereas the original test for subgroup differences was $\chi^2(1)=3.91,p=.05$, our analyses indicated that $\chi^2(1)=`r round(mod_mods$QM, 2)`,p=`r round(mod_mods$QMp, 3)`$.
@@
 Different heterogeneity estimators (e.g., `DerSimonian-Laird` instead of `REML`) did not resolve this difference.
-We tried to clarify these discrepancies by e-mailing the original authors, but failed to receive a response.
+We tried to clarify these discrepancies by e-mailing the original authors, but did not receive a response.
 
@@
 
-We estimated the number of data points that would need to be fabricated to arrive at the estimates from the DECREASE studies, given that the non-DECREASE studies represent the true effect.
-This assumes, in contrast to Step 2, that the DECREASE studies in fact contain fabricated data.
-The estimates from Step 3 provide an indication of the extent of the potential data fabrication in the DECREASE studies under this assumption [@commissie2011;@commissie2012;@commissie2013;@bouri2014].
+We estimated the number of data points that would need to be fabricated to arrive at the estimates from the DECREASE studies, given that the non-DECREASE studies represent the true population effect distribution.
+This assumes, in stark contrast to Step 2, that the DECREASE studies in fact contain fabricated data.
+The estimates from Step 3 provide an indication of the extent of potential data fabrication in the DECREASE studies [@commissie2011;@commissie2012;@commissie2013;@bouri2014].
 
@@
 
-In order to estimate the number of fabricated data points, we estimated the population effect distribution of mortality (in log odds) per condition in both the DECREASE- and non-DECREASE studies separately.
+In order to estimate the number of fabricated data points, we estimated the population effect distribution of mortality (in log odds) per condition in both the DECREASE- and non-DECREASE studies.
 For each of these, we ran a meta-analysis applying the same, verified methods used in Step 1.
-These separate analyses were necessary, considering that Step 1 only provided estimates of the overall risk ratio but not the odds conditional on exposure (beta-blocker or control) and study set (DECREASE- versus non-DECREASE).
-This resulted in four meta-analytic mortality estimates with an accompanying variance.
-Throughout the simulations, we use the point estimates to simulate genuine- and fabricated data, but supplement this by using the distribution estimates as sensitivity analyses.
+These separate analyses were necessary, considering that Step 1 only provided estimates of the overall risk ratio but not the odds conditional on exposure (beta-blocker or control) and trial set (DECREASE- versus non-DECREASE).
+This resulted in four meta-analytic mortality estimates with an accompanying effect variance.
+Throughout the simulations, we use the point estimates to simulate genuine- and fabricated data but supplement this by using the more uncertain distribution estimates as sensitivity analyses.
 
-We applied the inversion method to estimate the number of fabricated data points in the DECREASE-I and DECREASE-IV studies [@casella2002].
-With the inversion method, we iteratively hypothesized that *X* out of *N* data points were fabricated (i.e., $0, 1, ..., N$) in the DECREASE-I and DECREASE-IV studies separately.
-For each study, we simulated 10,000 datasets per hypothesis of *X* fabricated data points and *N-X* genuine data points.
+We applied the inversion method to estimate the number of fabricated data points in the DECREASE-I and DECREASE-IV trials [@casella2002].
+With the inversion method, we iteratively hypothesized that *X* out of *N* data points were fabricated (i.e., $X={0, 1, ..., N}$).
+For each combination of *X* and trial, we simulated 10,000 datasets.
+Each simulated dataset contained *X* fabricated data points and *N-X* genuine data points.
 For each simulated dataset (exact simulation procedure in next paragraph), we determined the likelihood of the results with
@@
 Table 2 indicates which cell sizes the various $n_{XX}$ refer to within the (simulated) data.
-We compared the likelihoods to determine whether the simulated data were more likely to arise from the genuine- ($L_{genuine}$) or from the fabricated data ($L_{fabricated}$).
+Subsequently, we compared the likelihoods to determine whether the simulated data were more likely to arise from the genuine trials ($L_{genuine}$) or from the fabricated trials ($L_{fabricated}$).
 Note that using the likelihoods for this is a minor deviation from the preregistration, where we initially planned on using $p$-value comparisons ([osf.io/vnmzc](https://osf.io/vnmzc)).
@@
 
-For each hypothesis of *X* out of *N* fabricated datapoints, we computed the probability that the fabricated data are more likely than the genuine data ($p_F=P(L_{fabricated}>L_{genuine})$).
+For each hypothesis of *X* out of *N* fabricated data points, we computed the probability that the fabricated data are more likely than the genuine data ($p_F=P(L_{fabricated}>L_{genuine})$).
 Based on $p_F$, we computed the confidence interval for $X$ (i.e., $X_{LB};X_{UB}$).
@@
 
-We computed $p_F$ for all possibilities of *X* out of *N* fabricated datapoints in 10,000 randomly generated datsets, which were generated in three steps. For each dataset we:
+We computed $p_F$ for all *X* out of *N* fabricated datapoints in 10,000 randomly generated datsets, which were generated in three steps. For each dataset we:
 
-1. Sampled (without replacement) *X* fictitious participants that would be the result of data fabrication (sampled across all conditions).
+1. Sampled (without replacement and across conditions) *X* fictitious participants that would be the result of data fabrication.
 
@@
 Based on the meta-analytic effect from 2 and the cell sizes from 3, we computed the  likelihoods of $L_{fabricated}$ and $L_{genuine}$ using Equation \ref{eq1}.
-As mentioned before, $p_F$ was computed as the probability that the data were more likely under the estimates resulting from the (allegedly) fabricated data (i.e., the DECREASE trials) than under the estimates resulting from the genuine data (i.e., the non-DECREASE trials; $p_F=P(L_{fabricated}>L_{genuine})$).
+As mentioned before, we computed $p_F$ as the probability that the data were more likely under the estimates resulting from the (allegedly) fabricated data (i.e., the DECREASE trials) than under the estimates resulting from the genuine data (i.e., the non-DECREASE trials; $p_F=P(L_{fabricated}>L_{genuine})$).
 
@@
 The left column of Figure 3 depicts the $p_F$ per *X* fabricated data points (top panel) and the bounds of the confidence interval when the degree of confidence is altered (lower panel).
-Staying clearly between the dotted lines in the top panel, depicting the 95\% CI (top: .975; bottom: .025), it becomes apparent that the degree of uncertainty is too high to make any estimates with a high degree of confidence.
-This due to the small sample size (i.e., $N=`r length(res_i_certain) - 1`$) combined with the availability of just the summary results.
+Staying clearly between the dotted lines in the top panel, depicting the 95\% CI (top: .975; bottom: .025), it becomes apparent that the degree of uncertainty is too high to make any estimates about fabricated data points with a high degree of confidence.
+This is partly due to the small sample size (i.e., $N=`r length(res_i_certain) - 1`$) and the availability of just the summary results.
 Only when the degree of confidence is lowered to around 75\% does the interval not span the entire sample size.
@@
 For DECREASE-IV, the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_iv_certain - .025) - 1` - `r which.max(res_iv_certain - .975) - 1`] or [`r which.min(res_iv_uncertain - .025) - 1` - `r which.max(res_iv_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
-The minor difference between the estimates indicates that there is a high degree of confidence that data fabrication did occur.
+The minor difference between the estimates indicates that there is a high degree of confidence that data fabrication did occur based on the difference of the results alone.
 Nonetheless, the range of potentially fabricated data points is still estimated at approximately 1000; this indicates that the summary results are insufficient to provide more than an estimated lowerbound.
-However, this does indicate that it is possible not all data were fabricated (i.e., $N=`r length(res_iv_certain) - 1`$), increasing the importance of data provenance to discern between genuine and falsified data.
+This does indicate that it is possible not all data were fabricated (i.e., $N=`r length(res_iv_certain) - 1`$), increasing the importance of well-documented data provenance to discern between genuine and falsified data.
 This affirms the conclusion of the scientific integrity committee that the results of the DECREASE-IV trial are "scientifically incorrect" (p.11) [@commissie2012].
@@
 However, this also highlights that in order to prevent detection it would be in the fabricators' interest to fabricate small studies (assuming the fabricator wants to remain undetected).
+This adds another factor to the unreliability of studies with small sample sizes.
 

### 2017-02-15 functions/inversion_method_certain.R
Make inversion method script more dynamic

@@ -67,7 +67,7 @@ for (x in xs){
     
     result_almost_i_certain[x + 1] <- result_almost_i_certain[x + 1] + (l_fab > l_tru )
     
-    cat(sprintf('DECREASE-I: %s fabricated data points, iteration %s\n', x, iter))
+    # cat(sprintf('DECREASE-I: %s fabricated data points, iteration %s\n', x, iter))
   }
 }
 
@@ -141,9 +141,9 @@ for (x in xs){
     
     result_almost_iv_certain[x + 1] <- result_almost_iv_certain[x + 1] + (l_fab > l_tru)
 
-    cat(sprintf('DECREASE-IV: %s fabricated data points, iteration %s\n', x, iter))
+    # cat(sprintf('DECREASE-IV: %s fabricated data points, iteration %s\n', x, iter))
   }
 }
 
-save(result_almost_i_certain, file = 'data/result_almost_i_certain')
-save(result_almost_iv_certain, file = 'data/result_almost_iv_certain')
+save(result_almost_i_certain, file = '../data/result_almost_i_certain')
+save(result_almost_iv_certain, file = '../data/result_almost_iv_certain')

### 2017-02-15 functions/inversion_method_uncertain.R

@@ -71,7 +71,7 @@ for (x in xs){
     
     result_almost_i_uncertain[x + 1] <- result_almost_i_uncertain[x + 1] + (l_fab > l_tru )
     
-    cat(sprintf('DECREASE-I: %s fabricated data points, iteration %s\n', x, iter))
+    # cat(sprintf('DECREASE-I: %s fabricated data points, iteration %s\n', x, iter))
   }
 }
 
@@ -149,9 +149,9 @@ for (x in xs){
     
     result_almost_iv_uncertain[x + 1] <- result_almost_iv_uncertain[x + 1] + (l_fab > l_tru)
     
-    cat(sprintf('DECREASE-IV: %s fabricated data points, iteration %s\n', x, iter))
+    # cat(sprintf('DECREASE-IV: %s fabricated data points, iteration %s\n', x, iter))
   }
 }
 
-save(result_almost_i_uncertain, file = 'data/result_almost_i_uncertain')
-save(result_almost_iv_uncertain, file = 'data/result_almost_iv_uncertain')
+save(result_almost_i_uncertain, file = '../data/result_almost_i_uncertain')
+save(result_almost_iv_uncertain, file = '../data/result_almost_iv_uncertain')

### 2017-02-15 submission/manuscript.Rmd
Re-estimate inversion method with seed, revising through step 2

@@
 output:
+  html_document: default
   pdf_document:
@@
     keep_tex: yes
-  html_document: default
   word_document: default
@@
 
-<!-- Introduction
-What should be in here?
-1. Effectiveness of beta-blockers (decrease probability))
-2. Effects of findings of misconduct
-3. Why this study. -->
 The effect of beta-blockers on perioperative mortality in non-cardiac surgery for patients with coronary artery disease (CAD) has been subject of discussion [@Coleg5210] due to findings of research  misconduct [@commissie2011;@commissie2012;@commissie2013].
-Consequently, it appeared that the related trials were central to claims of the effectiveness of beta-blocker in decreasing perioperative mortality [@Devereaux313;@Angeli2010;@bouri2014].
+Consequently, it appeared that the related trials were central to claims of the effectiveness of beta-blockers in decreasing perioperative mortality [@Devereaux313;@Angeli2010;@bouri2014].
 Perioperative mortality is the deathrate of patients during the period of the surgical procedure, which typically includes admission, anaesthesia, surgery, and recovery.
-Meta-analyses that include these studies subject to research misconduct conclude that beta-blockers decrease perioperative mortality [@Devereaux313;@Angeli2010;@bouri2014] whereas a meta-analysis that excludes the studies subject to research misconduct concludes that beta-blockers increase perioperative mortality [@bouri2014].
+Meta-analyses that include the studies subject to research misconduct conclude that beta-blockers decrease perioperative mortality [@Devereaux313;@Angeli2010;@bouri2014] whereas a meta-analysis that excludes the trials subject to research misconduct concludes that beta-blockers increase perioperative mortality [@bouri2014].
 
-The trials subject to research misconduct were the Dutch DECREASE trials [@commissie2011;@commissie2012;@commissie2013].
-The committees that investigated the integrity of the DECREASE studies reported that data fabrication was likely but that the extent of the data fabrication remained unclear [@commissie2011;@commissie2012;@commissie2013].
-Given that the inclusion or exclusion of these trials in meta-analyses is crucial in determining whether beta-blockers are an effective treatment and considering that the latest guidelines still recommend the usage of beta-blockers [@esc2014], we aim to estimate the extent of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999] in order to further the debate.
+The trials subject to research misconduct were the Dutch DECREASE-I and DECREASE-IV trials [@commissie2011;@commissie2012;@commissie2013].
+The committees that investigated the integrity of the DECREASE trials reported that data fabrication was likely but that the extent of the data fabrication remained unclear [@commissie2011;@commissie2012;@commissie2013].
+As such, the inclusion or exclusion of these trials in meta-analyses is crucial in determining whether beta-blockers are an effective treatment.
+Moreover, the latest guidelines still recommend the usage of beta-blockers in the perioperative period  in certain cases [@esc2014;@Fleishere278], despite evidence either pointing towards detrimental effects or no effects at all.
+Considering these consequences, we aim to estimate the extent of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999] in order to further the debate on using beta-blockers in the perioperative period of CAD patients undergoing non-cardiac surgeries.
 
-The reports on the integrity of the DECREASE trials primarily focused on the provenance of the raw data and patient files (e.g., informed consent, whether data corresponded to patient files) but did not investigate the extent to which the DECREASE trials deviated from comparable trials.
-The committee reports did not neglect statistical evaluation; a statistical expert evaluated the applicability of forensic statistical methods [@commissie2013] to evaluate results within one trial.
-However, comparing across trials can be used as a forensic statistical method as well.
-This method has previously been used to monitor trial data quality or to test for potential data anomalies [@Buyse1999-jq].
+The reports on the integrity of the DECREASE trials primarily focused on the provenance of the raw data and patient files but did not investigate the extent to which the DECREASE trials deviated from comparable trials.
+The provenance is primarily concerned with the origins of the data, verifying things such as (but not limited to) the informed consent and whether data corresponded to patient files.
+The committee reports did not neglect statistical evaluation however: a statistical expert evaluated the applicability of forensic statistical methods [@commissie2013] to evaluate results within one trial.
+Nonetheless, comparing across trials can be used as a forensic statistical method as well; this has previously been used to monitor trial data quality or to test for potential data anomalies [@Buyse1999-jq].
 Moreover, this method has previously proven to be effective in detecting data fabrication [@Knepper2016-la].
-As such, evaluating the DECREASE trials against comparable trials investigating the effectiveness of beta-blockers with respect to perioperative mortality could prove worthwhile.
+Evaluating the DECREASE trials against comparable trials studying the effectiveness of beta-blockers with respect to perioperative mortality could prove informative of the potential extent of the problems in the DECREASE trials.
 
-To statistically investigate the evidence of data fabrication in the results of the DECREASE studies [@dunkelgrun2009;@poldermans1999], we took three steps.
-First, we reproduced the findings from a 2014 meta-analysis [@bouri2014] that contained sufficient information to estimate the deviation of the DECREASE-I and DECREASE-IV studies from the remaining studies.
-Second, we evaluated the veracity of the DECREASE-I and DECREASE-IV studies (i.e., the probability that these study findings occurred assuming no data fabrication occurred).
-Third, we reversed this assumption and assumed that data fabrication did occur and estimate how many data points would have to be fabricated to reproduce the results of the DECREASE-I and DECREASE-IV studies if the remaining studies are regarded as estimating the true effect of beta-blockers on perioperative mortality.
+To statistically investigate the evidence of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999], we took three steps.
+First, we reproduced the findings from a 2014 meta-analysis [@bouri2014] that contained sufficient information to estimate the deviation of the DECREASE trials from comparable trials.
+Second, we evaluated the veracity of the DECREASE trials (i.e., the probability that the trial results occurred assuming no data fabrication took place).
+Third, we reversed this assumption and assumed that data fabrication did occur and estimated how many data points would have to be fabricated to reproduce the results of the DECREASE trials, if the comparable trials are regarded as estimating the true effect of beta-blockers on perioperative mortality in CAD patients undergoing non-cardiac surgery.
 
@@
 x12 <- '../'
-x12 <- ''
+# x12 <- ''
 bouri_data <- read.csv(sprintf('%sdata/bouri_data.csv', x12))
@@
 We used the default weighting procedure in the `metafor` package.
-When there was a zero-count for a cell, .5 was added to each cell of that trial, as is common in meta-analyses on risk- and odds ratios in order to deal with computational artefacts [@agresti2002].
-The 2014 meta-analysis [@bouri2014] did not specify the effect variance estimate used; hence, minor discrepancies between our estimates and the original estimates could be due to differences in the estimation procedure.
+When there was a zero-count for a cell (e.g., zero events), .5 was added to that cell, as is common in meta-analyses on risk- and odds ratios in order to prevent computational artefacts [@agresti2002].
+The 2014 meta-analysis [@bouri2014] did not specify the variance estimate used; hence, minor discrepancies between our estimates and the original estimates could be due to differences in the estimation procedure.
 
@@
 Second, we meta-analyzed all studies combined, including a dummy predictor for the DECREASE and non-DECREASE studies to reproduce results presented in Figure 4 of the 2014 meta-analysis [@bouri2014].
-Surprisingly, our results showed more evidence for unequal subgroups than the original meta-analysis [@bouri2014]; whereas the original test for subgroup differences was $\chi^2(1)=3.91,p=.05$, our analyses indicated that $\chi^2(1)=`r round(mod_mods$QM, 2)`,p=`r round(mod_mods$QMp, 3)`$.
-Additionally, the original analyses showed substantial heterogeneity ($I^2=74.4$\%) where we found an $I^2=`r round(mod_mods$I2, 3)`$\%.
-Different heterogeneity estimators (e.g., `DerSimonian-Laird` instead of `REML`) did not resolve this difference.
+Surprisingly, our results showed stronger evidence against equal subgroups than the original meta-analysis [@bouri2014] (original: $\chi^2(1)=3.91,p=.05$, reproduced $\chi^2(1)=`r round(mod_mods$QM, 2)`,p=`r round(mod_mods$QMp, 3)`$).
+Additionally, the original analyses showed substantial residual heterogeneity ($I^2=74.4$\%) whereas we found no residual heterogeneit ($I^2=$`r round(mod_mods$I2, 3)`\%).
+Different variance estimates (e.g., `DerSimonian-Laird` instead of `REML`) did not resolve this difference.
 We tried to clarify these discrepancies by e-mailing the original authors, but did not receive a response.
+Nonetheless, the broad strokes of the meta-regression confirmed that the DECREASE trials were the determining predictor for the effectiveness of beta-blockers (including DECREASE: $RR=`r round(exp(mod_mods$b[1] + mod_mods$b[2]), 3)`$; excluding DECREASE: $RR=`r round(exp(mod_mods$b[1]), 3)`$).
+<!--- Emailed bouri about this on 2016 03 16 -->
 
-<!--- Emailed bouri about this on 2016 03 16 -->
+```{r, echo = FALSE}
+mod_mods_type <- metafor::rma(yi = ln_rr,
+                         sei = se_ln_rr,
+                         mod =~ as.factor(decrease) + as.factor(beta_blocker),
+                         method = "REML",
+                         data = bouri_data,
+                         weighted = TRUE)
+```
+
+Additionally, and exploratively, we evaluated the predictive effect of the type of beta-blocker used in the trials.
+The DECREASE trials remain predictive of decreased mortality ($RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[2] + mod_mods_type$b[3]), 3)`$), <!-- DECREASE only uses bisoprolol!
+-->
+whereas the non-DECREASE trials provide tentative, but uncertain, evidence that atenolol results in lower mortality ($RR=`r round(exp(mod_mods_type$b[1]), 3)`$).
+Nonetheless, for other beta-blockers in the non-DECREASE trials, there is still tentative and uncertain evidence that beta-blockers increase mortality (bisoprolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[3]), 3)`$; metoprolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[4]), 3)`$; propanolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[5]), 3)`$).
+Table 3 shows the meta-regression results in full.
+
+```{r echo = FALSE}
+table3 <- matrix(NA, nrow = 8, ncol = 3)
+table3[1,1] <- ''
+table3[2,1] <- 'Intercept'
+table3[3,1] <- 'Non-DECREASE'
+table3[4,1] <- 'DECREASE'
+table3[5,1] <- 'Atenolol'
+table3[6,1] <- 'Bisoprolol'
+table3[7,1] <- 'Metoprolol'
+table3[8,1] <- 'Propanolol'
+
+table3[1,2] <- 'Estimate'
+table3[2,2] <- round(mod_mods_type$b[1], 3)
+table3[3,2] <- ''
+table3[4,2] <- round(mod_mods_type$b[2], 3)
+table3[5,2] <- ''
+table3[6,2] <- round(mod_mods_type$b[3], 3)
+table3[7,2] <- round(mod_mods_type$b[4], 3)
+table3[8,2] <- round(mod_mods_type$b[5], 3)
+
+table3[1,3] <- '95% CI'
+table3[2,3] <- sprintf('%s; %s', round(mod_mods_type$ci.lb[1], 3),
+                       round(mod_mods_type$ci.ub[1], 3))
+table3[3,3] <- ''
+table3[4,3] <- sprintf('%s; %s', round(mod_mods_type$ci.lb[2], 3),
+                       round(mod_mods_type$ci.ub[2], 3))
+table3[5,3] <- ''
+table3[6,3] <- sprintf('%s; %s', round(mod_mods_type$ci.lb[3], 3),
+                       round(mod_mods_type$ci.ub[3], 3))
+table3[7,3] <- sprintf('%s; %s', round(mod_mods_type$ci.lb[4], 3),
+                       round(mod_mods_type$ci.ub[4], 3))
+table3[8,3] <- sprintf('%s; %s', round(mod_mods_type$ci.lb[5], 3),
+                       round(mod_mods_type$ci.ub[5], 3))
+
+knitr::kable(table3, caption = 'Meta-regression results for log(RR), including dummy predictors for DECREASE trials (reference: non-DECREASE trials) and type of beta-blockers used in the trial (reference: atenolol).')
+```
 
@@
 
-Based on the non-DECREASE estimates from Step 1, we estimated the probability of obtaining the results in the DECREASE studies.
-To this end, we assumed that the non-DECREASE studies provide a valid representation of the true effect of beta-blockers on perioperative mortality (similar to Bouri et al.
+Based on the non-DECREASE estimates from Step 1, we estimated the probability of obtaining the results in the DECREASE trials.
+To this end, we assumed that the non-DECREASE trials provide a valid representation of the true effect of beta-blockers on perioperative mortality (similar to Bouri et al.
 [@bouri2014]; but see the discussion).
 The estimated probability is also known as the veracity of the data [@peters2015], which indicates the probability of the observed data under a given true effect.
-We assumed that the non-DECREASE studies estimate the true population distribution of effects, not perturbed by publication bias due to statistical (non)significance.
-Publication bias was assumed to not be a problem because a substantial number of nonsignificant effects are included in the dataset (9 of 11 nonsignificant).
+We assumed that the non-DECREASE studies estimated the true population distribution of effects, not perturbed by publication bias due to statistical (non)significance.
+Publication bias was assumed to not be a problem because a substantial number of nonsignificant effects are included in the dataset (9 of 11 nonsignificant results).
 
@@
 
-Based on the estimated mean log RR and its credible interval in the non-DECREASE studies, we computed the probability of log RR in the DECREASE trials.
+Based on the estimated mean log RR and its credible interval in the non-DECREASE studies, we computed the probability of the observed log RR in the DECREASE trials.
 The estimates of the non-DECREASE studies were obtained from Step 1, which include the estimated log RR (i.e., `r round(mod_clean$b, 2)`), and its 95\% credibility interval as provided by the package `metafor` (i.e., `r sprintf('[%s; %s]', round(mod_clean_pred$cr.lb, 3), round(mod_clean_pred$cr.ub, 3))`.
-We assumed a normal distribution of population effects with the estimated effect as the mean of the distribution.
+The meta-analysis model assumes a normal distribution of population effects with the estimated effect as the mean of the distribution.
 The 95\% credibility interval denotes the bounds of the normal distribution that covers 95\% of the density, where the standard deviation is calculated as the distance from the mean to either bound, divided by 1.96.
@@
 Based on the estimated effect distribution from the non-DECREASE trials, we calculated the probability of each DECREASE trial result, or a more extreme result.
-In other words, we computed the *p*-value for the hypothesis that the DECREASE studies occurred based on the information available from the other trials.
+In other words, we computed the *p*-value for the hypothesis that the DECREASE studies occurred, assuming the information available from the other trials is informative of the true population effect.
 
@@
 Figure `r fig` indicates that the DECREASE trials are highly unlikely under the estimated effect distribution based on the non-DECREASE trials.
-More specifically, DECREASE-I (or more extreme) has a probability of $`r prob[1]`$ (less than 1 in a quintillion) and DECREASE-IV $`r prob[2]`$ (3 in a billion), which indicates they are unlikely to come from the same population effect distribution as the non-DECREASE trials.
-Observing two of such extremely unlikely results in this population effect distribution jointly is highly improbable, $`r prob[1]*prob[2]`$.
-This would therefore indicate that the DECREASE trials are severely different from the non-DECREASE studies.
+More specifically, the results from DECREASE-I (or more extreme) have a probability of $`r prob[1]`$ (less than 1 in a quintillion) and the results from DECREASE-IV have a probability of $`r prob[2]`$ (3 in a billion).
+This indicates the DECREASE trial results are unlikely to have come from the same population effect distribution as the non-DECREASE trials.
+Moreover, observing two of such extremely unlikely results jointly, as in the DECREASE trials, is highly improbable, $`r prob[1]*prob[2]`$.
+This would therefore indicate that the DECREASE trials are severely different from the non-DECREASE trials.
 
 Results from Step 1 indicated that no variance (i.e., homogeneity; $\tau^2=0$) of the effects was observed; given the small number of studies included (i.e., `r mod_clean$k`) this estimate is uncertain, however.
-We conducted sensitivity analyses in order to see how dependent results are on the specific heterogeneity estimate.
-The probability of observing both DECREASE trials stays approximately below 1 out of 1000 until the variance estimate is .25, and the probability stays approximately below 1 out of 100 until the variance estimate is .43 (see Figure `r fig <- fig + 1; fig`).
-Further research would be worthwhile to better estimate the variance in the effects of beta-blockers on perioperative mortality and what moderators affect the effect size (e.g., type of beta-blocker).
+We conducted sensitivity analyses in order to see how dependent results are on the heterogeneity estimate.
+The probability of observing the DECREASE trials stays approximately below 1 out of 1000 until the variance estimate is .25; the probability stays approximately below 1 out of 100 until the variance estimate is .43 (see Figure `r fig <- fig + 1; fig`).
 
@@
 
-We estimated the number of data points that would need to be fabricated to arrive at the estimates from the DECREASE studies, given that the non-DECREASE studies represent the true population effect distribution.
-This assumes, in stark contrast to Step 2, that the DECREASE studies in fact contain fabricated data.
+```{r echo = FALSE}
+# This is for easy reference
+# And to easily include in text
+iterations <- 10000
+```
+We estimated the number of data points that would need to be fabricated to arrive at the estimates from the DECREASE trials, given that the non-DECREASE trials represent the true population effect distribution.
+This assumes, in stark contrast to Step 2, that the DECREASE trials in fact contain fabricated data.
 The estimates from Step 3 provide an indication of the extent of potential data fabrication in the DECREASE studies [@commissie2011;@commissie2012;@commissie2013;@bouri2014].
@@
 
-In order to estimate the number of fabricated data points, we estimated the population effect distribution of mortality (in log odds) per condition in both the DECREASE- and non-DECREASE studies.
+In order to estimate the number of fabricated data points, we estimated the population effect distribution of mortality (in log odds) per condition in both the DECREASE- and non-DECREASE trials.
 For each of these, we ran a meta-analysis applying the same, verified methods used in Step 1.
-These separate analyses were necessary, considering that Step 1 only provided estimates of the overall risk ratio but not the odds conditional on exposure (beta-blocker or control) and trial set (DECREASE- versus non-DECREASE).
+These separate analyses were necessary, considering that Step 1 only provided estimates of the overall risk ratio, but not the odds conditional on exposure (beta-blocker or control) and trial set (DECREASE- versus non-DECREASE).
 This resulted in four meta-analytic mortality estimates with an accompanying effect variance.
-Throughout the simulations, we use the point estimates to simulate genuine- and fabricated data but supplement this by using the more uncertain distribution estimates as sensitivity analyses.
+Throughout the simulations, we used the point estimates (i.e., fixed effect) to simulate genuine- and fabricated data, but supplement this by using the more uncertain distribution estimates (i.e., random effects) as sensitivity analyses.
 
-We applied the inversion method to estimate the number of fabricated data points in the DECREASE-I and DECREASE-IV trials [@casella2002].
+We applied the inversion method to estimate the number of fabricated data points in the DECREASE trials [@casella2002].
 With the inversion method, we iteratively hypothesized that *X* out of *N* data points were fabricated (i.e., $X={0, 1, ..., N}$).
-For each combination of *X* and trial, we simulated 10,000 datasets.
+For each combination of *X* and trial, we simulated `r iterations` datasets.
 Each simulated dataset contained *X* fabricated data points and *N-X* genuine data points.
-For each simulated dataset (exact simulation procedure in next paragraph), we determined the likelihood of the results with
+For each simulated dataset (exact simulation procedure in the next paragraph), we determined the likelihood of the results with
 \begin{equation}
@@
 \end{equation}
-where $\pi_{E}$ indicates the mortality rate of the experimental condition as drawn from the meta-analytic effect distribution ($\pi_{C}$ indicates the same for the control condition).
+where $\pi_{E}$ indicates the mortality rate of the experimental condition as drawn from the meta-analytic effect distribution ($\pi_{C}$ indicates the mortality rate for the control condition).
 The likelihood was computed under both the fabricated effect estimates (i.e., $L_{fabricated}$) and the genuine data (i.e., $L_{genuine}$).
-Table 2 indicates which cell sizes the various $n_{XX}$ refer to within the (simulated) data.
-Subsequently, we compared the likelihoods to determine whether the simulated data were more likely to arise from the genuine trials ($L_{genuine}$) or from the fabricated trials ($L_{fabricated}$).
-Note that using the likelihoods for this is a minor deviation from the preregistration, where we initially planned on using $p$-value comparisons ([osf.io/vnmzc](https://osf.io/vnmzc)).
+Table 3 indicates which cell sizes the various $n_{XX}$ refer to within the (simulated) data.
+After computing the likelihoods, we compared them to determine whether the simulated data were more likely to arise from the genuine trials ($L_{genuine}>L_{fabricated}$) or from the fabricated trials ($L_{fabricated}>L_{genuine}$).
+Note that comparing the likelihoods is a minor deviation from the preregistration, where we initially planned on using $p$-value comparisons ([osf.io/vnmzc](https://osf.io/vnmzc)).
 
-```{r echo=FALSE,results='asis'}
+```{r echo=FALSE, results='asis'}
 table2 <- matrix('', ncol = 3, nrow = 3)
@@
 
-We computed $p_F$ for all *X* out of *N* fabricated datapoints in 10,000 randomly generated datsets, which were generated in three steps. For each dataset we:
+We computed $p_F$ for all *X* out of *N* fabricated datapoints in `r iterations` randomly generated datasets, which were generated in three steps. For each dataset we:
 
-1. Sampled (without replacement and across conditions) *X* fictitious participants that would be the result of data fabrication.
+1. Sampled (across conditions, without replacement) *X* fictitious participants that would be the result of data fabrication.
 
@@
 
-Based on the meta-analytic effect from 2 and the cell sizes from 3, we computed the  likelihoods of $L_{fabricated}$ and $L_{genuine}$ using Equation \ref{eq1}.
-As mentioned before, we computed $p_F$ as the probability that the data were more likely under the estimates resulting from the (allegedly) fabricated data (i.e., the DECREASE trials) than under the estimates resulting from the genuine data (i.e., the non-DECREASE trials; $p_F=P(L_{fabricated}>L_{genuine})$).
+Based on the meta-analytic effect from 2 and the cell sizes from 3, we computed the  likelihoods $L_{fabricated}$ and $L_{genuine}$ using Equation \ref{eq1}.
+As mentioned before, we computed $p_F$, which indicated the probability that the data are more likely under the estimates resulting from the (allegedly) fabricated data (i.e., the DECREASE trials) than under the estimates resulting from the genuine data (i.e., the non-DECREASE trials; $p_F=P(L_{fabricated}>L_{genuine})$).
 
@@
 
-iterations <- 10000
-
 # Uncomment if you want to rerun the inversion method
@@
 # t1 <- Sys.time()
+# set.seed(123)
 # source('../functions/inversion_method_uncertain.R')
@@
 # t2 <- Sys.time()
+# set.seed(123)
 # source('../functions/inversion_method_certain.R')
@@
 res_iv_certain_ub[98:101] <- 1067
-
-# Old plotting. Saved temporarily. Ignore if forgotten to delete.
-# pdf(sprintf('%sfigures/fig3.pdf', x12), width = 8, height = 6)
-# 
-# par(mar = c(2.5, 5, 0.5, 0.5),
-#     mfrow = c(2, 2))
-# plot(res_i_uncertain,
-#      xlab = '',
-#      ylab = 'Proportion',
-#      ylim = 0:1,
-#      type = 'l',
-#      lty = 1)
-# lines(res_i_certain, pch = 2, type = 'l', lty = 2)
-# abline(h = c(.025, .975), lty = 3)
-# 
-# legend(x = 0, y = .9, bty = 'n', lty = c(1, 2), legend = c("Distribution estimate", "Point estimate"))
-# 
-# plot(x = NA, y = NA,
-#      xlim = c(0,1), ylim = c(0, length(res_i_certain)),
-#      xlab = "Confidence in %", ylab = "Data points fabricated")
-# 
-# lines(x = seq(0, 1, .01), y = res_i_uncertain_lb)
-# lines(x = seq(0, 1, .01), y = res_i_uncertain_ub)
-# lines(x = seq(0, 1, .01), y = res_i_certain_lb, lty = 2, pch = 2, type = 'l')
-# lines(x = seq(0, 1, .01), y = res_i_certain_ub, lty = 2, pch = 2, type = 'l')
-# abline(v = .95, lty = 3)
-# 
-# # clip(x1 = 0, x2 = length(res_i_uncertain), y1 = 0.025, y2 = .975)
-# # abline(v = which(abs(res_i_uncertain - .975) == min(abs(res_i_uncertain - .975))), lty = 1)
-# # abline(v = which(abs(res_i_certain - .975) == min(abs(res_i_certain - .975))), lty = 2)
-# 
-# par(mar = c(4, 5, 0.5, 0.5))
-# plot(res_iv_uncertain,
-#      xlab = 'Data-points fabricated',
-#      ylab = 'Proportion',
-#      ylim = 0:1,
-#      type = 'l',
-#      lty = 1)
-# lines(res_iv_certain, pch = 2, type = 'l',
-#       lty = 2)
-# abline(h = c(.025, .975), lty = 3)
-# 
-# plot(x = NA, y = NA,
-#      xlim = c(0,1), ylim = c(0, length(res_iv_certain)),
-#      xlab = "Confidence in %", ylab = "Data points fabricated")
-# 
-# lines(x = seq(0, 1, .01), y = res_iv_uncertain_lb)
-# lines(x = seq(0, 1, .01), y = res_iv_uncertain_ub)
-# lines(x = seq(0, 1, .01), y = res_iv_certain_lb, lty = 2, pch = 2, type = 'l')
-# lines(x = seq(0, 1, .01), y = res_iv_certain_ub, lty = 2, pch = 2, type = 'l')
-# abline(v = .95, lty = 3)
-# invisible(dev.off())
 
@@
 The left column of Figure 3 depicts the $p_F$ per *X* fabricated data points (top panel) and the bounds of the confidence interval when the degree of confidence is altered (lower panel).
-Staying clearly between the dotted lines in the top panel, depicting the 95\% CI (top: .975; bottom: .025), it becomes apparent that the degree of uncertainty is too high to make any estimates about fabricated data points with a high degree of confidence.
-This is partly due to the small sample size (i.e., $N=`r length(res_i_certain) - 1`$) and the availability of just the summary results.
+Staying clearly between the dotted lines in the top panel, depicting the 95\% CI (top: .975; bottom: .025), it becomes apparent that the degree of uncertainty is too high to make any reasonable estimates about the number of fabricated data points with sufficient confidence.
+This is partly due to the small sample size of the DECREASE-I trial (i.e., $N=`r length(res_i_certain) - 1`$) and the availability of just the summary results.
 Only when the degree of confidence is lowered to around 75\% does the interval not span the entire sample size.
@@
 For DECREASE-IV, the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_iv_certain - .025) - 1` - `r which.max(res_iv_certain - .975) - 1`] or [`r which.min(res_iv_uncertain - .025) - 1` - `r which.max(res_iv_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
-The minor difference between the estimates indicates that there is a high degree of confidence that data fabrication did occur based on the difference of the results alone.
+The relatively minor difference between the estimates indicates that there is a high degree of confidence that data fabrication did occur based on the difference of the trial results alone.
 Nonetheless, the range of potentially fabricated data points is still estimated at approximately 1000; this indicates that the summary results are insufficient to provide more than an estimated lowerbound.
@@
 
-The results also highlight that, despite the lack of availability of raw data, summary results from larger samples result in more certainty about the estimated number of fabricated data points using the inversion method.
-This is due to decreased standard errors at the population level of the effects, providing less variation in the simulations and resulting in higher sensitivity.
-However, this also highlights that in order to prevent detection it would be in the fabricators' interest to fabricate small studies (assuming the fabricator wants to remain undetected).
-This adds another factor to the unreliability of studies with small sample sizes.
+The results also highlight that, despite the lack of availability of the raw data, summary results from larger samples do allow estimating the number of fabricated data points when similar trials are available.
+Moreover, larger trial samples result in relatively more certainty about the estimated number of fabricated data points using the inversion method, although much residual uncertainty remains.
+This increased certainty is due to decreased standard errors at the population level of the estimated effects, resulting in higher sensitivity of data anomalies.
+Nonetheless, substantially less information is available in just the summary results, and raw data availability would improve results verification and the detection of potential anomalies.
+These results also highlight that in order to prevent detection, it would be in the fabricators' interest to fabricate small studies even if raw data are hidden away (assuming the fabricator wants to remain undetected).
 

### 2017-02-15 submission/manuscript.tex
Re-estimate inversion method with seed, revising through step 2

@@
   \postdate{\par}
-  \date{01 February, 2017}
+  \date{13 februari, 2017}
 
@@
 of discussion\textsuperscript{1} due to findings of research
-misconduct\textsuperscript{2--4} that were central to claims of their
-effectiveness\textsuperscript{5--7}. Perioperative mortality is the
+misconduct\textsuperscript{2--4}. Consequently, it appeared that the
+related trials were central to claims of the effectiveness of
+beta-blocker in decreasing perioperative
+mortality\textsuperscript{5--7}. Perioperative mortality is the
 deathrate of patients during the period of the surgical procedure, which
-typically includes admission, anaesthesia, surgery, and recovery. This
-related finding of research misconduct altered meta-analytic results of
-perioperative mortality in non-cardiac surgery for CAD patients
-substantively\textsuperscript{7}, where meta-analyses that include these
-studies conclude that beta-blockers decrease perioperative
-mortality\textsuperscript{5--7} but a meta-analysis that excludes these
-studies concludes that beta-blockers increase perioperative
-mortality\textsuperscript{7}.
+typically includes admission, anaesthesia, surgery, and recovery.
+Meta-analyses that include these studies subject to research misconduct
+conclude that beta-blockers decrease perioperative
+mortality\textsuperscript{5--7} whereas a meta-analysis that excludes
+the studies subject to research misconduct concludes that beta-blockers
+increase perioperative mortality\textsuperscript{7}.
 
-The trials deemed to be the result of research misconduct were the Dutch
-DECREASE trials\textsuperscript{2--4}. The committees that investigated
-the integrity of the DECREASE studies reported that data fabrication was
-likely\textsuperscript{2--4}, but that the extent of the data
-fabrication remained unclear. Given that the inclusion or exclusion of
+The trials subject to research misconduct were the Dutch DECREASE
+trials\textsuperscript{2--4}. The committees that investigated the
+integrity of the DECREASE studies reported that data fabrication was
+likely but that the extent of the data fabrication remained
+unclear\textsuperscript{2--4}. Given that the inclusion or exclusion of
 these trials in meta-analyses is crucial in determining whether
-beta-blockers are indeed effective during the perioperative period and
-the latest guidelines still recommend the usage of
+beta-blockers are an effective treatment and considering that the latest
+guidelines still recommend the usage of
 beta-blockers\textsuperscript{8}, we aim to estimate the extent of data
 fabrication in the DECREASE studies\textsuperscript{9,10} in order to
-further this debate.
+further the debate.
 
-The reports on the DECREASE trials primarily focused on the provenance
-of the raw data and patient files (e.g., informed consent, whether data
-corresponded to patient files), but neglected the extent to which the
-DECREASE studies deviated from similar trials that tested the
-effectiveness of beta-blockers on perioperative mortality. Comparing how
-(dis)similar trials are from comparable studies is not a new
-idea\textsuperscript{11} and has proven effective in detecting data
-fabrication\textsuperscript{12}. This is a different approach from the
-forensic statistical methods previously applied\textsuperscript{4},
-where results within one study are evaluated as opposed to other similar
-studies. This explains the discrepancy between the feasibility of
-evaluating the DECREASE trials by the statistical expert of the
-committee report and ours.
+The reports on the integrity of the DECREASE trials primarily focused on
+the provenance of the raw data and patient files (e.g., informed
+consent, whether data corresponded to patient files) but did not
+investigate the extent to which the DECREASE trials deviated from
+comparable trials. The committee reports did not neglect statistical
+evaluation; a statistical expert evaluated the applicability of forensic
+statistical methods\textsuperscript{4} to evaluate results within one
+trial. However, comparing across trials can be used as a forensic
+statistical method as well. This method has previously been used to
+monitor trial data quality or to test for potential data
+anomalies\textsuperscript{11}. Moreover, this method has previously
+proven to be effective in detecting data
+fabrication\textsuperscript{12}. As such, evaluating the DECREASE trials
+against comparable trials investigating the effectiveness of
+beta-blockers with respect to perioperative mortality could prove
+worthwhile.
 
@@
 results of the DECREASE studies\textsuperscript{9,10}, we took three
-steps. First, we replicate the findings from a 2014
-meta-analysis\textsuperscript{7} that contains sufficient information to
-estimate the deviation of the DECREASE-I and DECREASE-IV studies from
-the remaining studies. Second, we evaluate the veracity of the
+steps. First, we reproduced the findings from a 2014
+meta-analysis\textsuperscript{7} that contained sufficient information
+to estimate the deviation of the DECREASE-I and DECREASE-IV studies from
+the remaining studies. Second, we evaluated the veracity of the
 DECREASE-I and DECREASE-IV studies (i.e., the probability that these
 study findings occurred assuming no data fabrication occurred). Third,
-we reverse this assumption and assume that data fabrication did occur
+we reversed this assumption and assumed that data fabrication did occur
 and estimate how many data points would have to be fabricated to
@@
 extracted the raw event data for the 2 (control vs experimental) by 2
-(event or no event) design, which we used to recompute the natural
+(event vs no event) design, which we used to recompute the natural
 logarithm of the risk ratio and its standard error. The extracted event
@@
 \href{https://osf.io/vnmzc}{osf.io/vnmzc}.
+
+\begin{verbatim}
+## Warning: package 'latex2exp' was built under R version 3.2.5
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: package 'knitr' was built under R version 3.2.5
+\end{verbatim}
 
@@
 Second, we meta-analyzed all studies combined, including a dummy
-predictor for the DECREASE and non-DECREASE studies to replicate results
+predictor for the DECREASE and non-DECREASE studies to reproduce results
 presented in Figure 4 of the 2014 meta-analysis\textsuperscript{7}.
@@
 this difference. We tried to clarify these discrepancies by e-mailing
-the original authors, but failed to receive a response.
+the original authors, but did not receive a response.
 
@@
 to arrive at the estimates from the DECREASE studies, given that the
-non-DECREASE studies represent the true effect. This assumes, in
-contrast to Step 2, that the DECREASE studies in fact contain fabricated
-data. The estimates from Step 3 provide an indication of the extent of
-the potential data fabrication in the DECREASE studies under this
-assumption\textsuperscript{2--4,7}.
+non-DECREASE studies represent the true population effect distribution.
+This assumes, in stark contrast to Step 2, that the DECREASE studies in
+fact contain fabricated data. The estimates from Step 3 provide an
+indication of the extent of potential data fabrication in the DECREASE
+studies\textsuperscript{2--4,7}.
 
@@
 the population effect distribution of mortality (in log odds) per
-condition in both the DECREASE- and non-DECREASE studies separately. For
-each of these, we ran a meta-analysis applying the same, verified
-methods used in Step 1. These separate analyses were necessary,
-considering that Step 1 only provided estimates of the overall risk
-ratio but not the odds conditional on exposure (beta-blocker or control)
-and study set (DECREASE- versus non-DECREASE). This resulted in four
-meta-analytic mortality estimates with an accompanying variance.
-Throughout the simulations, we use the point estimates to simulate
-genuine- and fabricated data, but supplement this by using the
+condition in both the DECREASE- and non-DECREASE studies. For each of
+these, we ran a meta-analysis applying the same, verified methods used
+in Step 1. These separate analyses were necessary, considering that Step
+1 only provided estimates of the overall risk ratio but not the odds
+conditional on exposure (beta-blocker or control) and trial set
+(DECREASE- versus non-DECREASE). This resulted in four meta-analytic
+mortality estimates with an accompanying effect variance. Throughout the
+simulations, we use the point estimates to simulate genuine- and
+fabricated data but supplement this by using the more uncertain
 distribution estimates as sensitivity analyses.
@@
 data points in the DECREASE-I and DECREASE-IV
-studies\textsuperscript{17}. With the inversion method, we iteratively
+trials\textsuperscript{17}. With the inversion method, we iteratively
 hypothesized that \emph{X} out of \emph{N} data points were fabricated
-(i.e., \(0, 1, ..., N\)) in the DECREASE-I and DECREASE-IV studies
-separately. For each study, we simulated 10,000 datasets per hypothesis
-of \emph{X} fabricated data points and \emph{N-X} genuine data points.
-For each simulated dataset (exact simulation procedure in next
-paragraph), we determined the likelihood of the results with
+(i.e., \(X={0, 1, ..., N}\)). For each combination of \emph{X} and
+trial, we simulated 10,000 datasets. Each simulated dataset contained
+\emph{X} fabricated data points and \emph{N-X} genuine data points. For
+each simulated dataset (exact simulation procedure in next paragraph),
+we determined the likelihood of the results with
 
@@
 Table 2 indicates which cell sizes the various \(n_{XX}\) refer to
-within the (simulated) data. We compared the likelihoods to determine
-whether the simulated data were more likely to arise from the genuine-
-(\(L_{genuine}\)) or from the fabricated data (\(L_{fabricated}\)). Note
-that using the likelihoods for this is a minor deviation from the
-preregistration, where we initially planned on using \(p\)-value
-comparisons (\href{https://osf.io/vnmzc}{osf.io/vnmzc}).
+within the (simulated) data. Subsequently, we compared the likelihoods
+to determine whether the simulated data were more likely to arise from
+the genuine trials (\(L_{genuine}\)) or from the fabricated trials
+(\(L_{fabricated}\)). Note that using the likelihoods for this is a
+minor deviation from the preregistration, where we initially planned on
+using \(p\)-value comparisons
+(\href{https://osf.io/vnmzc}{osf.io/vnmzc}).
 
@@
 
-For each hypothesis of \emph{X} out of \emph{N} fabricated datapoints,
+For each hypothesis of \emph{X} out of \emph{N} fabricated data points,
 we computed the probability that the fabricated data are more likely
@@
 
-We computed \(p_F\) for all possibilities of \emph{X} out of \emph{N}
-fabricated datapoints in 10,000 randomly generated datsets, which were
-generated in three steps. For each dataset we:
+We computed \(p_F\) for all \emph{X} out of \emph{N} fabricated
+datapoints in 10,000 randomly generated datsets, which were generated in
+three steps. For each dataset we:
 
@@
 \item
-  Sampled (without replacement) \emph{X} fictitious participants that
-  would be the result of data fabrication (sampled across all
-  conditions).
+  Sampled (without replacement and across conditions) \emph{X}
+  fictitious participants that would be the result of data fabrication.
 \item
@@
 computed the likelihoods of \(L_{fabricated}\) and \(L_{genuine}\) using
-Equation \ref{eq1}. As mentioned before, \(p_F\) was computed as the
+Equation \ref{eq1}. As mentioned before, we computed \(p_F\) as the
 probability that the data were more likely under the estimates resulting
@@
 .975; bottom: .025), it becomes apparent that the degree of uncertainty
-is too high to make any estimates with a high degree of confidence. This
-due to the small sample size (i.e., \(N=112\)) combined with the
-availability of just the summary results. Only when the degree of
-confidence is lowered to around 75\% does the interval not span the
-entire sample size. As such, based on the summary results, little can be
-said about the extent of the data fabrication that occurred in the
-DECREASE-I trial, affirming the conclusions of the original committee
-report\textsuperscript{4}.
+is too high to make any estimates about fabricated data points with a
+high degree of confidence. This is partly due to the small sample size
+(i.e., \(N=112\)) and the availability of just the summary results. Only
+when the degree of confidence is lowered to around 75\% does the
+interval not span the entire sample size. As such, based on the summary
+results, little can be said about the extent of the data fabrication
+that occurred in the DECREASE-I trial, affirming the conclusions of the
+original committee report\textsuperscript{4}.
 
@@
 respectively. The minor difference between the estimates indicates that
-there is a high degree of confidence that data fabrication did occur.
-Nonetheless, the range of potentially fabricated data points is still
-estimated at approximately 1000; this indicates that the summary results
-are insufficient to provide more than an estimated lowerbound. However,
-this does indicate that it is possible not all data were fabricated
-(i.e., \(N=1066\)), increasing the importance of data provenance to
-discern between genuine and falsified data. This affirms the conclusion
-of the scientific integrity committee that the results of the
-DECREASE-IV trial are ``scientifically incorrect''
-(p.11)\textsuperscript{3}.
+there is a high degree of confidence that data fabrication did occur
+based on the difference of the results alone. Nonetheless, the range of
+potentially fabricated data points is still estimated at approximately
+1000; this indicates that the summary results are insufficient to
+provide more than an estimated lowerbound. This does indicate that it is
+possible not all data were fabricated (i.e., \(N=1066\)), increasing the
+importance of well-documented data provenance to discern between genuine
+and falsified data. This affirms the conclusion of the scientific
+integrity committee that the results of the DECREASE-IV trial are
+``scientifically incorrect'' (p.11)\textsuperscript{3}.
 
@@
 fabricate small studies (assuming the fabricator wants to remain
-undetected).
+undetected). This adds another factor to the unreliability of studies
+with small sample sizes.
 

### 2017-03-02 submission/manuscript.Rmd
Revisions

@@
 output:
-  html_document: default
   pdf_document:
@@
     keep_tex: yes
+  html_document: default
   word_document: default
@@
 
-The effect of beta-blockers on perioperative mortality in non-cardiac surgery for patients with coronary artery disease (CAD) has been subject of discussion [@Coleg5210] due to findings of research  misconduct [@commissie2011;@commissie2012;@commissie2013].
-Consequently, it appeared that the related trials were central to claims of the effectiveness of beta-blockers in decreasing perioperative mortality [@Devereaux313;@Angeli2010;@bouri2014].
-Perioperative mortality is the deathrate of patients during the period of the surgical procedure, which typically includes admission, anaesthesia, surgery, and recovery.
-Meta-analyses that include the studies subject to research misconduct conclude that beta-blockers decrease perioperative mortality [@Devereaux313;@Angeli2010;@bouri2014] whereas a meta-analysis that excludes the trials subject to research misconduct concludes that beta-blockers increase perioperative mortality [@bouri2014].
+The effect of beta-blockers on perioperative mortality in non-cardiac surgery for patients with coronary artery disease (CAD) has been controversial [@Coleg5210] due to findings of research  misconduct in two related clinical trials  [@commissie2011;@commissie2012;@commissie2013].
+Three meta-analyses that included the trials subject to research misconduct concluded that beta-blockers decrease perioperative mortality [@Devereaux313;@Angeli2010;@bouri2014] whereas a meta-analysis that excluded the suspect trials concluded that beta-blockers increase perioperative mortality [@bouri2014].
+In these studies, perioperative mortality was defined as the deathrate of patients during the period of the surgical procedure, which typically includes admission, anaesthesia, surgery, and recovery.
 
+<!-- Should we quote the (translated) main conclusion form the reports? -->
 The trials subject to research misconduct were the Dutch DECREASE-I and DECREASE-IV trials [@commissie2011;@commissie2012;@commissie2013].
 The committees that investigated the integrity of the DECREASE trials reported that data fabrication was likely but that the extent of the data fabrication remained unclear [@commissie2011;@commissie2012;@commissie2013].
-As such, the inclusion or exclusion of these trials in meta-analyses is crucial in determining whether beta-blockers are an effective treatment.
-Moreover, the latest guidelines still recommend the usage of beta-blockers in the perioperative period  in certain cases [@esc2014;@Fleishere278], despite evidence either pointing towards detrimental effects or no effects at all.
-Considering these consequences, we aim to estimate the extent of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999] in order to further the debate on using beta-blockers in the perioperative period of CAD patients undergoing non-cardiac surgeries.
+Moreover, the latest guidelines still recommend the usage of beta-blockers in the perioperative period in certain cases [@esc2014;@Fleishere278].
+Considering these consequences, we aim to estimate the extent of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999] to further the debate on using beta-blockers in the perioperative period of CAD patients undergoing non-cardiac surgery.
 
-The reports on the integrity of the DECREASE trials primarily focused on the provenance of the raw data and patient files but did not investigate the extent to which the DECREASE trials deviated from comparable trials.
+The reports on the integrity of the DECREASE trials primarily focused on the provenance of the raw data but did not investigate the extent to which the DECREASE trials deviated from comparable trials.
 The provenance is primarily concerned with the origins of the data, verifying things such as (but not limited to) the informed consent and whether data corresponded to patient files.
-The committee reports did not neglect statistical evaluation however: a statistical expert evaluated the applicability of forensic statistical methods [@commissie2013] to evaluate results within one trial.
-Nonetheless, comparing across trials can be used as a forensic statistical method as well; this has previously been used to monitor trial data quality or to test for potential data anomalies [@Buyse1999-jq].
+The committee reports did not neglect statistical evaluation however: a statistical expert evaluated the applicability of forensic statistical methods [@commissie2013] to evaluate results of trials separately (i.e., DECREASE-I, DECREASE-IV).
+Nonetheless, comparing across trials is a method that has previously been used to monitor trial data quality or to test for potential data anomalies [@Buyse1999-jq].
 Moreover, this method has previously proven to be effective in detecting data fabrication [@Knepper2016-la].
-Evaluating the DECREASE trials against comparable trials studying the effectiveness of beta-blockers with respect to perioperative mortality could prove informative of the potential extent of the problems in the DECREASE trials.
+Evaluating the DECREASE trials to other published trials studying the effectiveness of beta-blockers with respect to perioperative mortality could prove informative of the potential extent of the fabrication in the DECREASE trials.
+
+<!-- nieuw  revise-->
+However, the effectiveness of perioperative beta-blockade is obfuscated by both trials afflicted by data fabrication and the type of beta-blocker used in the trial.
+Various kinds of beta-blockers could vary in their pharmalogical effectiveness with respect to perioperative mortality, and anomalies in the effectiveness of the DECREASE trials might be due to differences in the beta-blockers used [@] and not purely due to data fabrication (given not all data points can be considered fabricated at this point).
 
 To statistically investigate the evidence of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999], we took three steps.
-First, we reproduced the findings from a 2014 meta-analysis [@bouri2014] that contained sufficient information to estimate the deviation of the DECREASE trials from comparable trials.
-Second, we evaluated the veracity of the DECREASE trials (i.e., the probability that the trial results occurred assuming no data fabrication took place).
-Third, we reversed this assumption and assumed that data fabrication did occur and estimated how many data points would have to be fabricated to reproduce the results of the DECREASE trials, if the comparable trials are regarded as estimating the true effect of beta-blockers on perioperative mortality in CAD patients undergoing non-cardiac surgery.
+First, we reproduced the findings from a 2014 meta-analysis [@bouri2014] that contained sufficient information to estimate the deviation of the DECREASE trials from other published trials on beta-blockers.
+We also include type of beta-blocker to inspect whether this is predictive of the effect of beta-blockers on perioperative mortality.
+Second, we evaluated the the probability that the DECREASE trial results occurred assuming no data fabrication.
+Third, we reversed this assumption and assumed that data fabrication did occur and estimated how many data points would have to be fabricated to reproduce the results of the DECREASE trials, if the other published trials are regarded as estimating the true effect of beta-blockers on perioperative mortality in CAD patients undergoing non-cardiac surgery.
 
@@
 ```{r, echo = FALSE}
+options(scipen = 999)
 suppressPackageStartupMessages(if(!require(metafor)){install.packages('metafor', repos = "http://cran.us.r-project.org")})
@@
 We used the default weighting procedure in the `metafor` package.
-When there was a zero-count for a cell (e.g., zero events), .5 was added to that cell, as is common in meta-analyses on risk- and odds ratios in order to prevent computational artefacts [@agresti2002].
+When there was a zero-count for a cell (e.g., zero events), 0.5 was added to that cell, as is common in meta-analyses on risk- and odds ratios in order to prevent computational artefacts [@agresti2002].
 The 2014 meta-analysis [@bouri2014] did not specify the variance estimate used; hence, minor discrepancies between our estimates and the original estimates could be due to differences in the estimation procedure.
@@
 
-We analyzed the RRs to verify the original estimates (Figure 2 of the 2014 meta-analysis [@bouri2014]) and found that we were able to reproduce the estimates for the different sets of studies.
+We were able to closely reproduce the estimates for the different sets of studies (Figure 2 of the 2014 meta-analysis [@bouri2014]).
 Bouri et al.
@@
 The estimated probability is also known as the veracity of the data [@peters2015], which indicates the probability of the observed data under a given true effect.
-We assumed that the non-DECREASE studies estimated the true population distribution of effects, not perturbed by publication bias due to statistical (non)significance.
+We assumed that the non-DECREASE studies estimated the true effect distribution of perioperative beta-blockade on mortality, not perturbed by publication bias due to statistical (non)significance.
 Publication bias was assumed to not be a problem because a substantial number of nonsignificant effects are included in the dataset (9 of 11 nonsignificant results).
@@
 
-Results from Step 1 indicated that no variance (i.e., homogeneity; $\tau^2=0$) of the effects was observed; given the small number of studies included (i.e., `r mod_clean$k`) this estimate is uncertain, however.
-We conducted sensitivity analyses in order to see how dependent results are on the heterogeneity estimate.
-The probability of observing the DECREASE trials stays approximately below 1 out of 1000 until the variance estimate is .25; the probability stays approximately below 1 out of 100 until the variance estimate is .43 (see Figure `r fig <- fig + 1; fig`).
+Results from Step 1 indicated that no between-study variance (i.e., homogeneity; $\tau^2=0$) of the effects was observed; given the small number of studies included (i.e., `r mod_clean$k`) this estimate is uncertain, however.
+We conducted sensitivity analyses to see how dependent results are on the heterogeneity estimate.
+The probability of observing the DECREASE trials stays approximately below 1 out of 1000 until the variance estimate is 0.25; the probability stays approximately below 1 out of 100 until the variance estimate is 0.43 (see Figure `r fig <- fig + 1; fig`).
+To put these numbers into context, a variance of 0.25 would suggest that results of perioperative beta-blockade vary substantially even if perioperative beta-blockade has no effect whatsoever (RRs between `r round(exp(0 - .25), 3)` and `r round(exp(0 + .25), 3)` in 64\% of the cases) all due to contextual circumstances of the study.
 
@@
 ```
-We estimated the number of data points that would need to be fabricated to arrive at the estimates from the DECREASE trials, given that the non-DECREASE trials represent the true population effect distribution.
-This assumes, in stark contrast to Step 2, that the DECREASE trials in fact contain fabricated data.
+We estimated the number of data points that would need to be fabricated to arrive at the estimates from the DECREASE trials, given that the non-DECREASE trials represent the true effect of perioperative beta-blockade.
+In contrast to Step 2 this assumes that the DECREASE trials might in fact contain fabricated data.
 The estimates from Step 3 provide an indication of the extent of potential data fabrication in the DECREASE studies [@commissie2011;@commissie2012;@commissie2013;@bouri2014].
@@
 
-In order to estimate the number of fabricated data points, we estimated the population effect distribution of mortality (in log odds) per condition in both the DECREASE- and non-DECREASE trials.
-For each of these, we ran a meta-analysis applying the same, verified methods used in Step 1.
-These separate analyses were necessary, considering that Step 1 only provided estimates of the overall risk ratio, but not the odds conditional on exposure (beta-blocker or control) and trial set (DECREASE- versus non-DECREASE).
-This resulted in four meta-analytic mortality estimates with an accompanying effect variance.
-Throughout the simulations, we used the point estimates (i.e., fixed effect) to simulate genuine- and fabricated data, but supplement this by using the more uncertain distribution estimates (i.e., random effects) as sensitivity analyses.
-
-We applied the inversion method to estimate the number of fabricated data points in the DECREASE trials [@casella2002].
-With the inversion method, we iteratively hypothesized that *X* out of *N* data points were fabricated (i.e., $X={0, 1, ..., N}$).
-For each combination of *X* and trial, we simulated `r iterations` datasets.
-Each simulated dataset contained *X* fabricated data points and *N-X* genuine data points.
-For each simulated dataset (exact simulation procedure in the next paragraph), we determined the likelihood of the results with
-\begin{equation}
-\label{eq1}
-L(\theta|\pi_{E},\pi_{C})=\pi_{E}^{n_{11}}(1-\pi_{E})^{n_{12}} \times \pi_{C}^{n_{21}}(1 - \pi_{C})^{n_{22}}
-\end{equation}
-where $\pi_{E}$ indicates the mortality rate of the experimental condition as drawn from the meta-analytic effect distribution ($\pi_{C}$ indicates the mortality rate for the control condition).
-The likelihood was computed under both the fabricated effect estimates (i.e., $L_{fabricated}$) and the genuine data (i.e., $L_{genuine}$).
-Table 3 indicates which cell sizes the various $n_{XX}$ refer to within the (simulated) data.
-After computing the likelihoods, we compared them to determine whether the simulated data were more likely to arise from the genuine trials ($L_{genuine}>L_{fabricated}$) or from the fabricated trials ($L_{fabricated}>L_{genuine}$).
-Note that comparing the likelihoods is a minor deviation from the preregistration, where we initially planned on using $p$-value comparisons ([osf.io/vnmzc](https://osf.io/vnmzc)).
+In order to estimate the number of fabricated data points, we first estimated the effect of perioperative beta-blockade on mortality (in log odds) in each trial arm.
+In total there are four trial arms: one per condition (beta-blocker or control) per trial type (DECREASE- and non-DECREASE trials).
+For each of these trial arms, we ran a meta-analysis applying the same methods used in Step 1.
+For each of the two DECREASE trials separately and the non-DECREASE trials combined, this resulted in four meta-analytic mortality estimates with corresponding effect variances (see Table 3).
+Throughout the simulations, we used the point estimates (i.e., fixed effect) to simulate genuine- and fabricated data, but supplemented this by using the more uncertain distribution estimates (i.e., random effects) as sensitivity analyses.
 
@@
 ```
+
+We applied the inversion method to estimate the number of fabricated data points in the DECREASE trials [@casella2002].
+The inversion method iteratively hypothesizes that *X* out of *N* data points were fabricated (i.e., $X={0, 1, ..., N}$).
+For each combination of *X* and trial, we simulated `r iterations` datasets.
+Each simulated dataset contained *X* fabricated data points and *N-X* genuine data points.
+For each simulated dataset (exact simulation procedure in the next paragraph), we determined the likelihood of the results with
+\begin{equation}
+\label{eq1}
+L(\theta|\pi_{E},\pi_{C})=\pi_{E}^{n_{11}}(1-\pi_{E})^{n_{12}} \times \pi_{C}^{n_{21}}(1 - \pi_{C})^{n_{22}}
+\end{equation}
+where $\pi_{E}$ indicates the mortality rate in the beta-blocker condition as drawn from the meta-analytic effect distribution ($\pi_{C}$ indicates the mortality rate in the control condition).
+The likelihood was computed under both the fabricated effect estimates (i.e., $L_{fabricated}$) and the genuine data (i.e., $L_{genuine}$).
+Table 3 indicates which cell sizes the various $n_{XX}$ refer to within the (simulated) data.
+After computing the likelihoods, we compared them to determine whether the simulated data were more likely to arise from the genuine trials ($L_{genuine}>L_{fabricated}$) or from the fabricated trials ($L_{fabricated}>L_{genuine}$).
+Note that comparing the likelihoods is a minor deviation from the preregistration, where we initially planned on using $p$-value comparisons ([osf.io/vnmzc](https://osf.io/vnmzc)).
 
@@
 
-For DECREASE-I, the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_i_certain - .025) - 1` - `r which.max(res_i_certain - .975) - 1`] or [`r which.min(res_i_uncertain - .025) - 1` - `r which.max(res_i_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
+For DECREASE-I ($N=`r bouri_data$control_n[10] + bouri_data$beta_n[10]`), the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_i_certain - .025) - 1` - `r which.max(res_i_certain - .975) - 1`] or [`r which.min(res_i_uncertain - .025) - 1` - `r which.max(res_i_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
 The left column of Figure 3 depicts the $p_F$ per *X* fabricated data points (top panel) and the bounds of the confidence interval when the degree of confidence is altered (lower panel).
@@
 
-For DECREASE-IV, the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_iv_certain - .025) - 1` - `r which.max(res_iv_certain - .975) - 1`] or [`r which.min(res_iv_uncertain - .025) - 1` - `r which.max(res_iv_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
+For DECREASE-IV ($N=`r bouri_data$control_n[11] + bouri_data$beta_n[11]`), the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_iv_certain - .025) - 1` - `r which.max(res_iv_certain - .975) - 1`] or [`r which.min(res_iv_uncertain - .025) - 1` - `r which.max(res_iv_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
 The relatively minor difference between the estimates indicates that there is a high degree of confidence that data fabrication did occur based on the difference of the trial results alone.
 Nonetheless, the range of potentially fabricated data points is still estimated at approximately 1000; this indicates that the summary results are insufficient to provide more than an estimated lowerbound.
-This does indicate that it is possible not all data were fabricated (i.e., $N=`r length(res_iv_certain) - 1`$), increasing the importance of well-documented data provenance to discern between genuine and falsified data.
+This indicates that it is possible not all data were fabricated (i.e., $N=`r length(res_iv_certain) - 1`$), increasing the importance of well-documented data provenance to discern between genuine and falsified data.
 This affirms the conclusion of the scientific integrity committee that the results of the DECREASE-IV trial are "scientifically incorrect" (p.11) [@commissie2012].
 
-The results also highlight that, despite the lack of availability of the raw data, summary results from larger samples do allow estimating the number of fabricated data points when similar trials are available.
-Moreover, larger trial samples result in relatively more certainty about the estimated number of fabricated data points using the inversion method, although much residual uncertainty remains.
-This increased certainty is due to decreased standard errors at the population level of the estimated effects, resulting in higher sensitivity of data anomalies.
+The results also highlight that, despite the lack of availability of the raw data, summary results from larger samples allow estimation of the number of fabricated data points when similar trials are available.
+Moreover, larger trials result in relatively more certainty (e.g., DECREASE-IV) about the estimated number of fabricated data points using the inversion method compared to smaller trials (e.g., DECREASE-I), although much residual uncertainty remains.
+This increased certainty is due to decreased standard errors at the population level of the estimated effects, resulting in higher sensitivity to data anomalies.
 Nonetheless, substantially less information is available in just the summary results, and raw data availability would improve results verification and the detection of potential anomalies.

### 2017-03-02 submission/manuscript.tex
Revisions

@@
   \postdate{\par}
-  \date{13 februari, 2017}
+  \date{02 March, 2017}
 
@@
 The effect of beta-blockers on perioperative mortality in non-cardiac
-surgery for patients with coronary artery disease (CAD) has been subject
-of discussion\textsuperscript{1} due to findings of research
-misconduct\textsuperscript{2--4}. Consequently, it appeared that the
-related trials were central to claims of the effectiveness of
-beta-blocker in decreasing perioperative
-mortality\textsuperscript{5--7}. Perioperative mortality is the
-deathrate of patients during the period of the surgical procedure, which
-typically includes admission, anaesthesia, surgery, and recovery.
-Meta-analyses that include these studies subject to research misconduct
-conclude that beta-blockers decrease perioperative
-mortality\textsuperscript{5--7} whereas a meta-analysis that excludes
-the studies subject to research misconduct concludes that beta-blockers
-increase perioperative mortality\textsuperscript{7}.
+surgery for patients with coronary artery disease (CAD) has been
+controversial\textsuperscript{1} due to findings of research misconduct
+in two related clinical trials\textsuperscript{2--4}. Three
+meta-analyses that included the trials subject to research misconduct
+concluded that beta-blockers decrease perioperative
+mortality\textsuperscript{5--7} whereas a meta-analysis that excluded
+the suspect trials concluded that beta-blockers increase perioperative
+mortality\textsuperscript{7}. In these studies, perioperative mortality
+was defined as the deathrate of patients during the period of the
+surgical procedure, which typically includes admission, anaesthesia,
+surgery, and recovery.
 
-The trials subject to research misconduct were the Dutch DECREASE
-trials\textsuperscript{2--4}. The committees that investigated the
-integrity of the DECREASE studies reported that data fabrication was
-likely but that the extent of the data fabrication remained
-unclear\textsuperscript{2--4}. Given that the inclusion or exclusion of
-these trials in meta-analyses is crucial in determining whether
-beta-blockers are an effective treatment and considering that the latest
-guidelines still recommend the usage of
-beta-blockers\textsuperscript{8}, we aim to estimate the extent of data
-fabrication in the DECREASE studies\textsuperscript{9,10} in order to
-further the debate.
+The trials subject to research misconduct were the Dutch DECREASE-I and
+DECREASE-IV trials\textsuperscript{2--4}. The committees that
+investigated the integrity of the DECREASE trials reported that data
+fabrication was likely but that the extent of the data fabrication
+remained unclear\textsuperscript{2--4}. Moreover, the latest guidelines
+still recommend the usage of beta-blockers in the perioperative period
+in certain cases\textsuperscript{8,9}. Considering these consequences,
+we aim to estimate the extent of data fabrication in the DECREASE
+studies\textsuperscript{10,11} to further the debate on using
+beta-blockers in the perioperative period of CAD patients undergoing
+non-cardiac surgery.
 
 The reports on the integrity of the DECREASE trials primarily focused on
-the provenance of the raw data and patient files (e.g., informed
-consent, whether data corresponded to patient files) but did not
-investigate the extent to which the DECREASE trials deviated from
-comparable trials. The committee reports did not neglect statistical
-evaluation; a statistical expert evaluated the applicability of forensic
-statistical methods\textsuperscript{4} to evaluate results within one
-trial. However, comparing across trials can be used as a forensic
-statistical method as well. This method has previously been used to
-monitor trial data quality or to test for potential data
-anomalies\textsuperscript{11}. Moreover, this method has previously
-proven to be effective in detecting data
-fabrication\textsuperscript{12}. As such, evaluating the DECREASE trials
-against comparable trials investigating the effectiveness of
-beta-blockers with respect to perioperative mortality could prove
-worthwhile.
+the provenance of the raw data but did not investigate the extent to
+which the DECREASE trials deviated from comparable trials. The
+provenance is primarily concerned with the origins of the data,
+verifying things such as (but not limited to) the informed consent and
+whether data corresponded to patient files. The committee reports did
+not neglect statistical evaluation however: a statistical expert
+evaluated the applicability of forensic statistical
+methods\textsuperscript{4} to evaluate results of trials separately
+(i.e., DECREASE-I, DECREASE-IV). Nonetheless, comparing across trials is
+a method that has previously been used to monitor trial data quality or
+to test for potential data anomalies\textsuperscript{12}. Moreover, this
+method has previously proven to be effective in detecting data
+fabrication\textsuperscript{13}. Evaluating the DECREASE trials to other
+published trials studying the effectiveness of beta-blockers with
+respect to perioperative mortality could prove informative of the
+potential extent of the fabrication in the DECREASE trials.
+
+However, the effectiveness of perioperative beta-blockade is obfuscated
+by both trials afflicted by data fabrication and the type of
+beta-blocker used in the trial. Various kinds of beta-blockers could
+vary in their pharmalogical effectiveness with respect to perioperative
+mortality, and anomalies in the effectiveness of the DECREASE trials
+might be due to differences in the beta-blockers used {[}@{]} and not
+purely due to data fabrication (given not all data points can be
+considered fabricated at this point).
 
 To statistically investigate the evidence of data fabrication in the
-results of the DECREASE studies\textsuperscript{9,10}, we took three
-steps. First, we reproduced the findings from a 2014
-meta-analysis\textsuperscript{7} that contained sufficient information
-to estimate the deviation of the DECREASE-I and DECREASE-IV studies from
-the remaining studies. Second, we evaluated the veracity of the
-DECREASE-I and DECREASE-IV studies (i.e., the probability that these
-study findings occurred assuming no data fabrication occurred). Third,
-we reversed this assumption and assumed that data fabrication did occur
-and estimate how many data points would have to be fabricated to
-reproduce the results of the DECREASE-I and DECREASE-IV studies if the
-remaining studies are regarded as estimating the true effect of
-beta-blockers on perioperative mortality.
+DECREASE studies\textsuperscript{10,11}, we took three steps. First, we
+reproduced the findings from a 2014 meta-analysis\textsuperscript{7}
+that contained sufficient information to estimate the deviation of the
+DECREASE trials from other published trials on beta-blockers. We also
+include type of beta-blocker to inspect whether this is predictive of
+the effect of beta-blockers on perioperative mortality. Second, we
+evaluated the the probability that the DECREASE trial results occurred
+assuming no data fabrication. Third, we reversed this assumption and
+assumed that data fabrication did occur and estimated how many data
+points would have to be fabricated to reproduce the results of the
+DECREASE trials, if the other published trials are regarded as
+estimating the true effect of beta-blockers on perioperative mortality
+in CAD patients undergoing non-cardiac surgery.
 
@@
 
-\begin{verbatim}
-## Warning: package 'latex2exp' was built under R version 3.2.5
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: package 'knitr' was built under R version 3.2.5
-\end{verbatim}
-
 We computed the log risk ratio (i.e., log RR) for each study and pooled
-these using the \texttt{R} package \texttt{metafor}\textsuperscript{13}.
+these using the \texttt{R} package \texttt{metafor}\textsuperscript{14}.
 We estimated a weighted random-effects model using the restricted
-maximum-likelihood estimator (i.e., \texttt{REML})\textsuperscript{14}
+maximum-likelihood estimator (i.e., \texttt{REML})\textsuperscript{15}
 to estimate the variance of effects. We used the default weighting
 procedure in the \texttt{metafor} package. When there was a zero-count
-for a cell, .5 was added to each cell of that trial, as is common in
-meta-analyses on risk- and odds ratios in order to deal with
-computational artefacts\textsuperscript{15}. The 2014
-meta-analysis\textsuperscript{7} did not specify the effect variance
-estimate used; hence, minor discrepancies between our estimates and the
-original estimates could be due to differences in the estimation
-procedure.
+for a cell (e.g., zero events), 0.5 was added to that cell, as is common
+in meta-analyses on risk- and odds ratios in order to prevent
+computational artefacts\textsuperscript{16}. The 2014
+meta-analysis\textsuperscript{7} did not specify the variance estimate
+used; hence, minor discrepancies between our estimates and the original
+estimates could be due to differences in the estimation procedure.
 
@@
 
-We analyzed the RRs to verify the original estimates (Figure 2 of the
-2014 meta-analysis\textsuperscript{7}) and found that we were able to
-reproduce the estimates for the different sets of studies. Bouri et al.
-differentiated between the estimates from the non-DECREASE trials and
-the DECREASE trials. We confirmed the effect size estimates and the
-variance estimates for both the non-DECREASE- and the DECREASE trials,
-save for some minor discrepancies due to the estimation method. Table 1
-depicts the original and reproduced values for both sets of studies.
+We were able to closely reproduce the estimates for the different sets
+of studies (Figure 2 of the 2014 meta-analysis\textsuperscript{7}).
+Bouri et al. differentiated between the estimates from the non-DECREASE
+trials and the DECREASE trials. We confirmed the effect size estimates
+and the variance estimates for both the non-DECREASE- and the DECREASE
+trials, save for some minor discrepancies due to the estimation method.
+Table 1 depicts the original and reproduced values for both sets of
+studies.
 
@@
 presented in Figure 4 of the 2014 meta-analysis\textsuperscript{7}.
-Surprisingly, our results showed more evidence for unequal subgroups
-than the original meta-analysis\textsuperscript{7}; whereas the original
-test for subgroup differences was \(\chi^2(1)=3.91,p=.05\), our analyses
-indicated that \(\chi^2(1)=6.12,p=0.013\). Additionally, the original
-analyses showed substantial heterogeneity (\(I^2=74.4\)\%) where we
-found an \(I^2=0\)\%. Different heterogeneity estimators (e.g.,
+Surprisingly, our results showed stronger evidence against equal
+subgroups than the original meta-analysis\textsuperscript{7} (original:
+\(\chi^2(1)=3.91,p=.05\), reproduced \(\chi^2(1)=6.12,p=0.013\)).
+Additionally, the original analyses showed substantial residual
+heterogeneity (\(I^2=74.4\)\%) whereas we found no residual heterogeneit
+(\$I\^{}2=\$0\%). Different variance estimates (e.g.,
 \texttt{DerSimonian-Laird} instead of \texttt{REML}) did not resolve
 this difference. We tried to clarify these discrepancies by e-mailing
-the original authors, but did not receive a response.
+the original authors, but did not receive a response. Nonetheless, the
+broad strokes of the meta-regression confirmed that the DECREASE trials
+were the determining predictor for the effectiveness of beta-blockers
+(including DECREASE: \(RR=0.496\); excluding DECREASE: \(RR=1.28\)).
+
+Additionally, and exploratively, we evaluated the predictive effect of
+the type of beta-blocker used in the trials. The DECREASE trials remain
+predictive of decreased mortality (\(RR=0.496\)), whereas the
+non-DECREASE trials provide tentative, but uncertain, evidence that
+atenolol results in lower mortality (\(RR=0.751\)). Nonetheless, for
+other beta-blockers in the non-DECREASE trials, there is still tentative
+and uncertain evidence that beta-blockers increase mortality
+(bisoprolol: \(RR=2.973\); metoprolol: \(RR=1.307\); propanolol:
+\(RR=2.041\)). Table 3 shows the meta-regression results in full.
+
+\begin{longtable}[]{@{}lll@{}}
+\caption{Meta-regression results for log(RR), including dummy predictors
+for DECREASE trials (reference: non-DECREASE trials) and type of
+beta-blockers used in the trial (reference: atenolol).}\tabularnewline
+\toprule
+& Estimate & 95\% CI\tabularnewline
+Intercept & -0.287 & -1.317; 0.743\tabularnewline
+Non-DECREASE & &\tabularnewline
+DECREASE & -1.792 & -5.082; 1.498\tabularnewline
+Atenolol & &\tabularnewline
+Bisoprolol & 1.376 & -1.996; 4.749\tabularnewline
+Metoprolol & 0.554 & -0.504; 1.613\tabularnewline
+Propanolol & 1 & -1.642; 3.643\tabularnewline
+\bottomrule
+\end{longtable}
 
@@
 Based on the non-DECREASE estimates from Step 1, we estimated the
-probability of obtaining the results in the DECREASE studies. To this
-end, we assumed that the non-DECREASE studies provide a valid
+probability of obtaining the results in the DECREASE trials. To this
+end, we assumed that the non-DECREASE trials provide a valid
 representation of the true effect of beta-blockers on perioperative
@@
 discussion). The estimated probability is also known as the veracity of
-the data\textsuperscript{16}, which indicates the probability of the
+the data\textsuperscript{17}, which indicates the probability of the
 observed data under a given true effect. We assumed that the
-non-DECREASE studies estimate the true population distribution of
-effects, not perturbed by publication bias due to statistical
-(non)significance. Publication bias was assumed to not be a problem
-because a substantial number of nonsignificant effects are included in
-the dataset (9 of 11 nonsignificant).
+non-DECREASE studies estimated the true effect distribution of
+perioperative beta-blockade on mortality, not perturbed by publication
+bias due to statistical (non)significance. Publication bias was assumed
+to not be a problem because a substantial number of nonsignificant
+effects are included in the dataset (9 of 11 nonsignificant results).
 
@@
 Based on the estimated mean log RR and its credible interval in the
-non-DECREASE studies, we computed the probability of log RR in the
-DECREASE trials. The estimates of the non-DECREASE studies were obtained
-from Step 1, which include the estimated log RR (i.e., 0.25), and its
-95\% credibility interval as provided by the package \texttt{metafor}
-(i.e., {[}0.01; 0.484{]}. We assumed a normal distribution of population
-effects with the estimated effect as the mean of the distribution. The
-95\% credibility interval denotes the bounds of the normal distribution
-that covers 95\% of the density, where the standard deviation is
-calculated as the distance from the mean to either bound, divided by
-1.96. This allows for an approximation of the population effect
-distribution, as depicted in Figure 1.
+non-DECREASE studies, we computed the probability of the observed log RR
+in the DECREASE trials. The estimates of the non-DECREASE studies were
+obtained from Step 1, which include the estimated log RR (i.e., 0.25),
+and its 95\% credibility interval as provided by the package
+\texttt{metafor} (i.e., {[}0.01; 0.484{]}. The meta-analysis model
+assumes a normal distribution of population effects with the estimated
+effect as the mean of the distribution. The 95\% credibility interval
+denotes the bounds of the normal distribution that covers 95\% of the
+density, where the standard deviation is calculated as the distance from
+the mean to either bound, divided by 1.96. This allows for an
+approximation of the population effect distribution, as depicted in
+Figure 1.
 
@@
 extreme result. In other words, we computed the \emph{p}-value for the
-hypothesis that the DECREASE studies occurred based on the information
-available from the other trials.
+hypothesis that the DECREASE studies occurred, assuming the information
+available from the other trials is informative of the true population
+effect.
 
@@
 the estimated effect distribution based on the non-DECREASE trials. More
-specifically, DECREASE-I (or more extreme) has a probability of
-\(2.9610213\times 10^{-53}\) (less than 1 in a quintillion) and
-DECREASE-IV \(3.0974376\times 10^{-9}\) (3 in a billion), which
-indicates they are unlikely to come from the same population effect
-distribution as the non-DECREASE trials. Observing two of such extremely
-unlikely results in this population effect distribution jointly is
-highly improbable, \(9.1715786\times 10^{-62}\). This would therefore
+specifically, the results from DECREASE-I (or more extreme) have a
+probability of \(0\) (less than 1 in a quintillion) and the results from
+DECREASE-IV have a probability of \(0\) (3 in a billion). This indicates
+the DECREASE trial results are unlikely to have come from the same
+population effect distribution as the non-DECREASE trials. Moreover,
+observing two of such extremely unlikely results jointly, as in the
+DECREASE trials, is highly improbable, \(0\). This would therefore
 indicate that the DECREASE trials are severely different from the
-non-DECREASE studies.
+non-DECREASE trials.
 
-Results from Step 1 indicated that no variance (i.e., homogeneity;
-\(\tau^2=0\)) of the effects was observed; given the small number of
-studies included (i.e., 9) this estimate is uncertain, however. We
-conducted sensitivity analyses in order to see how dependent results are
-on the specific heterogeneity estimate. The probability of observing
-both DECREASE trials stays approximately below 1 out of 1000 until the
-variance estimate is .25, and the probability stays approximately below
-1 out of 100 until the variance estimate is .43 (see Figure 2). Further
-research would be worthwhile to better estimate the variance in the
-effects of beta-blockers on perioperative mortality and what moderators
-affect the effect size (e.g., type of beta-blocker).
+Results from Step 1 indicated that no between-study variance (i.e.,
+homogeneity; \(\tau^2=0\)) of the effects was observed; given the small
+number of studies included (i.e., 9) this estimate is uncertain,
+however. We conducted sensitivity analyses to see how dependent results
+are on the heterogeneity estimate. The probability of observing the
+DECREASE trials stays approximately below 1 out of 1000 until the
+variance estimate is 0.25; the probability stays approximately below 1
+out of 100 until the variance estimate is 0.43 (see Figure 2). To put
+these numbers into context, a variance of 0.25 would suggest that
+results of perioperative beta-blockade vary substantially even if
+perioperative beta-blockade has no effect whatsoever (RRs between 0.779
+and 1.284 in 64\% of the cases) all due to contextual circumstances of
+the study.
 
@@
 We estimated the number of data points that would need to be fabricated
-to arrive at the estimates from the DECREASE studies, given that the
-non-DECREASE studies represent the true population effect distribution.
-This assumes, in stark contrast to Step 2, that the DECREASE studies in
-fact contain fabricated data. The estimates from Step 3 provide an
-indication of the extent of potential data fabrication in the DECREASE
-studies\textsuperscript{2--4,7}.
+to arrive at the estimates from the DECREASE trials, given that the
+non-DECREASE trials represent the true effect of perioperative
+beta-blockade. In contrast to Step 2 this assumes that the DECREASE
+trials might in fact contain fabricated data. The estimates from Step 3
+provide an indication of the extent of potential data fabrication in the
+DECREASE studies\textsuperscript{2--4,7}.
 
@@
 
-In order to estimate the number of fabricated data points, we estimated
-the population effect distribution of mortality (in log odds) per
-condition in both the DECREASE- and non-DECREASE studies. For each of
-these, we ran a meta-analysis applying the same, verified methods used
-in Step 1. These separate analyses were necessary, considering that Step
-1 only provided estimates of the overall risk ratio but not the odds
-conditional on exposure (beta-blocker or control) and trial set
-(DECREASE- versus non-DECREASE). This resulted in four meta-analytic
-mortality estimates with an accompanying effect variance. Throughout the
-simulations, we use the point estimates to simulate genuine- and
-fabricated data but supplement this by using the more uncertain
-distribution estimates as sensitivity analyses.
-
-We applied the inversion method to estimate the number of fabricated
-data points in the DECREASE-I and DECREASE-IV
-trials\textsuperscript{17}. With the inversion method, we iteratively
-hypothesized that \emph{X} out of \emph{N} data points were fabricated
-(i.e., \(X={0, 1, ..., N}\)). For each combination of \emph{X} and
-trial, we simulated 10,000 datasets. Each simulated dataset contained
-\emph{X} fabricated data points and \emph{N-X} genuine data points. For
-each simulated dataset (exact simulation procedure in next paragraph),
-we determined the likelihood of the results with
-
-\begin{equation}
-\label{eq1}
-L(\theta|\pi_{E},\pi_{C})=\pi_{E}^{n_{11}}(1-\pi_{E})^{n_{12}} \times \pi_{C}^{n_{21}}(1 - \pi_{C})^{n_{22}}
-\end{equation}
-
-where \(\pi_{E}\) indicates the mortality rate of the experimental
-condition as drawn from the meta-analytic effect distribution
-(\(\pi_{C}\) indicates the same for the control condition). The
-likelihood was computed under both the fabricated effect estimates
-(i.e., \(L_{fabricated}\)) and the genuine data (i.e., \(L_{genuine}\)).
-Table 2 indicates which cell sizes the various \(n_{XX}\) refer to
-within the (simulated) data. Subsequently, we compared the likelihoods
-to determine whether the simulated data were more likely to arise from
-the genuine trials (\(L_{genuine}\)) or from the fabricated trials
-(\(L_{fabricated}\)). Note that using the likelihoods for this is a
-minor deviation from the preregistration, where we initially planned on
-using \(p\)-value comparisons
-(\href{https://osf.io/vnmzc}{osf.io/vnmzc}).
+In order to estimate the number of fabricated data points, we first
+estimated the effect of perioperative beta-blockade on mortality (in log
+odds) in each trial arm. In total there are four trial arms: one per
+condition (beta-blocker or control) per trial type (DECREASE- and
+non-DECREASE trials). For each of these trial arms, we ran a
+meta-analysis applying the same methods used in Step 1. For each of the
+two DECREASE trials separately and the non-DECREASE trials combined,
+this resulted in four meta-analytic mortality estimates with
+corresponding effect variances (see Table 3). Throughout the
+simulations, we used the point estimates (i.e., fixed effect) to
+simulate genuine- and fabricated data, but supplemented this by using
+the more uncertain distribution estimates (i.e., random effects) as
+sensitivity analyses.
 
@@
 
+We applied the inversion method to estimate the number of fabricated
+data points in the DECREASE trials\textsuperscript{18}. The inversion
+method iteratively hypothesizes that \emph{X} out of \emph{N} data
+points were fabricated (i.e., \(X={0, 1, ..., N}\)). For each
+combination of \emph{X} and trial, we simulated 10000 datasets. Each
+simulated dataset contained \emph{X} fabricated data points and
+\emph{N-X} genuine data points. For each simulated dataset (exact
+simulation procedure in the next paragraph), we determined the
+likelihood of the results with
+
+\begin{equation}
+\label{eq1}
+L(\theta|\pi_{E},\pi_{C})=\pi_{E}^{n_{11}}(1-\pi_{E})^{n_{12}} \times \pi_{C}^{n_{21}}(1 - \pi_{C})^{n_{22}}
+\end{equation}
+
+where \(\pi_{E}\) indicates the mortality rate in the beta-blocker
+condition as drawn from the meta-analytic effect distribution
+(\(\pi_{C}\) indicates the mortality rate in the control condition). The
+likelihood was computed under both the fabricated effect estimates
+(i.e., \(L_{fabricated}\)) and the genuine data (i.e., \(L_{genuine}\)).
+Table 3 indicates which cell sizes the various \(n_{XX}\) refer to
+within the (simulated) data. After computing the likelihoods, we
+compared them to determine whether the simulated data were more likely
+to arise from the genuine trials (\(L_{genuine}>L_{fabricated}\)) or
+from the fabricated trials (\(L_{fabricated}>L_{genuine}\)). Note that
+comparing the likelihoods is a minor deviation from the preregistration,
+where we initially planned on using \(p\)-value comparisons
+(\href{https://osf.io/vnmzc}{osf.io/vnmzc}).
+
 For each hypothesis of \emph{X} out of \emph{N} fabricated data points,
@@
 We computed \(p_F\) for all \emph{X} out of \emph{N} fabricated
-datapoints in 10,000 randomly generated datsets, which were generated in
+datapoints in 10000 randomly generated datasets, which were generated in
 three steps. For each dataset we:
@@
 \item
-  Sampled (without replacement and across conditions) \emph{X}
-  fictitious participants that would be the result of data fabrication.
+  Sampled (across conditions, without replacement) \emph{X} fictitious
+  participants that would be the result of data fabrication.
 \item
@@
 Based on the meta-analytic effect from 2 and the cell sizes from 3, we
-computed the likelihoods of \(L_{fabricated}\) and \(L_{genuine}\) using
-Equation \ref{eq1}. As mentioned before, we computed \(p_F\) as the
-probability that the data were more likely under the estimates resulting
-from the (allegedly) fabricated data (i.e., the DECREASE trials) than
-under the estimates resulting from the genuine data (i.e., the
-non-DECREASE trials; \(p_F=P(L_{fabricated}>L_{genuine})\)).
+computed the likelihoods \(L_{fabricated}\) and \(L_{genuine}\) using
+Equation \ref{eq1}. As mentioned before, we computed \(p_F\), which
+indicated the probability that the data are more likely under the
+estimates resulting from the (allegedly) fabricated data (i.e., the
+DECREASE trials) than under the estimates resulting from the genuine
+data (i.e., the non-DECREASE trials;
+\(p_F=P(L_{fabricated}>L_{genuine})\)).
 
@@
 
-For DECREASE-I, the 95\% confidence interval for the estimated number of
-fabricated data points is {[}0 - 112{]} or {[}3 - 112{]} when based on a
-point estimate or a more uncertain distribution estimate, respectively.
-The left column of Figure 3 depicts the \(p_F\) per \emph{X} fabricated
-data points (top panel) and the bounds of the confidence interval when
-the degree of confidence is altered (lower panel). Staying clearly
-between the dotted lines in the top panel, depicting the 95\% CI (top:
-.975; bottom: .025), it becomes apparent that the degree of uncertainty
-is too high to make any estimates about fabricated data points with a
-high degree of confidence. This is partly due to the small sample size
-(i.e., \(N=112\)) and the availability of just the summary results. Only
-when the degree of confidence is lowered to around 75\% does the
-interval not span the entire sample size. As such, based on the summary
-results, little can be said about the extent of the data fabrication
-that occurred in the DECREASE-I trial, affirming the conclusions of the
-original committee report\textsuperscript{4}.
+For DECREASE-I (\$N=112), the 95\% confidence interval for the estimated
+number of fabricated data points is {[}0 - 112{]} or {[}2 - 112{]} when
+based on a point estimate or a more uncertain distribution estimate,
+respectively. The left column of Figure 3 depicts the \(p_F\) per
+\emph{X} fabricated data points (top panel) and the bounds of the
+confidence interval when the degree of confidence is altered (lower
+panel). Staying clearly between the dotted lines in the top panel,
+depicting the 95\% CI (top: .975; bottom: .025), it becomes apparent
+that the degree of uncertainty is too high to make any reasonable
+estimates about the number of fabricated data points with sufficient
+confidence. This is partly due to the small sample size of the
+DECREASE-I trial (i.e., \(N=112\)) and the availability of just the
+summary results. Only when the degree of confidence is lowered to around
+75\% does the interval not span the entire sample size. As such, based
+on the summary results, little can be said about the extent of the data
+fabrication that occurred in the DECREASE-I trial, affirming the
+conclusions of the original committee report\textsuperscript{4}.
 
-For DECREASE-IV, the 95\% confidence interval for the estimated number
-of fabricated data points is {[}8 - 1059{]} or {[}7 - 1066{]} when based
-on a point estimate or a more uncertain distribution estimate,
-respectively. The minor difference between the estimates indicates that
-there is a high degree of confidence that data fabrication did occur
-based on the difference of the results alone. Nonetheless, the range of
-potentially fabricated data points is still estimated at approximately
-1000; this indicates that the summary results are insufficient to
-provide more than an estimated lowerbound. This does indicate that it is
-possible not all data were fabricated (i.e., \(N=1066\)), increasing the
-importance of well-documented data provenance to discern between genuine
-and falsified data. This affirms the conclusion of the scientific
-integrity committee that the results of the DECREASE-IV trial are
-``scientifically incorrect'' (p.11)\textsuperscript{3}.
+For DECREASE-IV (\$N=1066), the 95\% confidence interval for the
+estimated number of fabricated data points is {[}0 - 1053{]} or {[}8 -
+1066{]} when based on a point estimate or a more uncertain distribution
+estimate, respectively. The relatively minor difference between the
+estimates indicates that there is a high degree of confidence that data
+fabrication did occur based on the difference of the trial results
+alone. Nonetheless, the range of potentially fabricated data points is
+still estimated at approximately 1000; this indicates that the summary
+results are insufficient to provide more than an estimated lowerbound.
+This indicates that it is possible not all data were fabricated (i.e.,
+\(N=1066\)), increasing the importance of well-documented data
+provenance to discern between genuine and falsified data. This affirms
+the conclusion of the scientific integrity committee that the results of
+the DECREASE-IV trial are ``scientifically incorrect''
+(p.11)\textsuperscript{3}.
 
-The results also highlight that, despite the lack of availability of raw
-data, summary results from larger samples result in more certainty about
-the estimated number of fabricated data points using the inversion
-method. This is due to decreased standard errors at the population level
-of the effects, providing less variation in the simulations and
-resulting in higher sensitivity. However, this also highlights that in
-order to prevent detection it would be in the fabricators' interest to
-fabricate small studies (assuming the fabricator wants to remain
-undetected). This adds another factor to the unreliability of studies
-with small sample sizes.
+The results also highlight that, despite the lack of availability of the
+raw data, summary results from larger samples allow estimation of the
+number of fabricated data points when similar trials are available.
+Moreover, larger trials result in relatively more certainty (e.g.,
+DECREASE-IV) about the estimated number of fabricated data points using
+the inversion method compared to smaller trials (e.g., DECREASE-I),
+although much residual uncertainty remains. This increased certainty is
+due to decreased standard errors at the population level of the
+estimated effects, resulting in higher sensitivity to data anomalies.
+Nonetheless, substantially less information is available in just the
+summary results, and raw data availability would improve results
+verification and the detection of potential anomalies. These results
+also highlight that in order to prevent detection, it would be in the
+fabricators' interest to fabricate small studies even if raw data are
+hidden away (assuming the fabricator wants to remain undetected).
 
@@
 \emph{Q}-statistic is sufficiently large to begin with (i.e.,
-5).\textsuperscript{18} Nonetheless, the original non-DECREASE trials
+5).\textsuperscript{19} Nonetheless, the original non-DECREASE trials
 show substantive variation in the treatment, for example the usage of
@@
 metoprolol) in different (standardized) dosages, which have been argued
-to affect effectiveness of the interventions\textsuperscript{19}. As
+to affect effectiveness of the interventions\textsuperscript{20}. As
 such, the lack of heterogeneity can be genuine and indicative of no
@@
 However, due to various types of beta-blockers, evidence might be
-moderated\textsuperscript{19}.
+moderated\textsuperscript{20}.
 
@@
 analysis, and then developed their
-recommendations.''\textsuperscript{20} Regardless, the revisions made
+recommendations.''\textsuperscript{21} Regardless, the revisions made
 ``recommend continuation of beta-blocker therapy in the perioperative
@@
 recommend the ``initiation of beta-blockers in patients undergoing
-low-risk surgery''\textsuperscript{21}. Review pieces note no
+low-risk surgery''\textsuperscript{22}. Review pieces note no
 substantial clinical change with respect to
-beta-blockers.\textsuperscript{22,23}
+beta-blockers.\textsuperscript{23,24}
 
@@
 .04, which is insufficient evidence and is actually more likely when
-there is no effect in of beta-blockers in reality\textsuperscript{24}
+there is no effect in of beta-blockers in reality\textsuperscript{25}
 
@@
 
+\hypertarget{ref-Fleishere278}{}
+9. Fleisher LA, Fleischmann KE, Auerbach AD, et al. 2014 acc/aha
+guideline on perioperative cardiovascular evaluation and management of
+patients undergoing noncardiac surgery. \emph{Circulation}.
+2014;130(24):e278-e333.
+doi:\href{https://doi.org/10.1161/CIR.0000000000000106}{10.1161/CIR.0000000000000106}.
+
 \hypertarget{ref-dunkelgrun2009}{}
-9. Dunkelgrun M, Boersma E, Schouten O, et al. Bisoprolol and
+10. Dunkelgrun M, Boersma E, Schouten O, et al. Bisoprolol and
 fluvastatin for the reduction of perioperative cardiac mortality and
@@
 \hypertarget{ref-poldermans1999}{}
-10. Poldermans D, Boersma E, Bax JJ, et al. The effect of bisoprolol on
+11. Poldermans D, Boersma E, Bax JJ, et al. The effect of bisoprolol on
 perioperative mortality and myocardial infarction in High-Risk patients
@@
 \hypertarget{ref-Buyse1999-jq}{}
-11. Buyse M, George SL, Evans S, et al. The role of biostatistics in the
+12. Buyse M, George SL, Evans S, et al. The role of biostatistics in the
 prevention, detection and treatment of fraud in clinical trials.
@@
 \hypertarget{ref-Knepper2016-la}{}
-12. Knepper D, Lindblad AS, Sharma G, et al. Statistical monitoring in
+13. Knepper D, Lindblad AS, Sharma G, et al. Statistical monitoring in
 clinical trials: Best practices for detecting data anomalies suggestive
@@
 \hypertarget{ref-viechtbauer2010}{}
-13. Viechtbauer W. Conducting meta-analyses in R with the metafor
+14. Viechtbauer W. Conducting meta-analyses in R with the metafor
 package. \emph{Journal of Statistical Software}. 2010;36:1-48.
@@
 \hypertarget{ref-viechtbauer2005}{}
-14. Viechtbauer W. Bias and efficiency of meta-analytic variance
+15. Viechtbauer W. Bias and efficiency of meta-analytic variance
 estimators in the Random-Effects model. \emph{Journal of educational and
@@
 \hypertarget{ref-agresti2002}{}
-15. Agresti A. \emph{Categorical Data Analysis}. Hoboken, NJ: John Wiley
+16. Agresti A. \emph{Categorical Data Analysis}. Hoboken, NJ: John Wiley
 \& Sons Inc.; 2002.
@@
 \hypertarget{ref-peters2015}{}
-16. Peters CFW, Klaassen CAJ, Wiel MA van de. \emph{Evaluating the
+17. Peters CFW, Klaassen CAJ, Wiel MA van de. \emph{Evaluating the
 Scientific Veracity of Publications by Dr. Jens Forster}. University of
@@
 \hypertarget{ref-casella2002}{}
-17. Casella G, Berger RL. \emph{Statistical Interference}. Pacific
+18. Casella G, Berger RL. \emph{Statistical Interference}. Pacific
 Grove, CA: Duxbury; 2002.
@@
 \hypertarget{ref-ioannidis2006}{}
-18. Ioannidis JPA, Trikalinos TA, Zintzaras E. Extreme between-study
+19. Ioannidis JPA, Trikalinos TA, Zintzaras E. Extreme between-study
 homogeneity in meta-analyses could offer useful insights. \emph{Journal
@@
 \hypertarget{ref-klei2015}{}
-19. Klei W van. Welke perioperatieve bètablokker heeft de voorkeur?
+20. Klei W van. Welke perioperatieve bètablokker heeft de voorkeur?
 {[}Which perioperative beta-blocker is preferred?{]}. \emph{Nederlands
@@
 \hypertarget{ref-Kristensen_2014}{}
-20. Kristensen SD, Knuuti J, Saraste A, et al. 2014 ESC/ESA guidelines
+21. Kristensen SD, Knuuti J, Saraste A, et al. 2014 ESC/ESA guidelines
 on non-cardiac surgery. \emph{European Journal of Anaesthesiology}.
@@
 \hypertarget{ref-guarracino2015}{}
-21. Guarracino F PH Baldassarri R. Revised esc/esa guidelines on
+22. Guarracino F PH Baldassarri R. Revised esc/esa guidelines on
 non-cardiac surgery: Cardiovascular assessment and management.
@@
 \hypertarget{ref-Port2016}{}
-22. Port SC. 2014 esc/esa guidelines on noncardiac surgery:
+23. Port SC. 2014 esc/esa guidelines on noncardiac surgery:
 Cardiovascular assessment and management. \emph{Journal of Nuclear
@@
 \hypertarget{ref-Kristensen2016}{}
-23. Kristensen SD. 2014 esc/esa guidelines on noncardiac surgery:
+24. Kristensen SD. 2014 esc/esa guidelines on noncardiac surgery:
 Cardiovascular assessment and management. \emph{Journal of Nuclear
@@
 \hypertarget{ref-van_Aert_2016}{}
-24. Aert RCM van, Wicherts JM, Assen MALM van. Conducting meta-analyses
+25. Aert RCM van, Wicherts JM, Assen MALM van. Conducting meta-analyses
 based on p values: Reservations and recommendations for applying

### 2017-03-09 functions/inversion_method_certain.R
Correct error in estimation procedure with certain estimates from MA.

@@ -15,12 +15,8 @@ for (x in xs){
     
     # Sample the proportion of events from population effects
     # min() ensures that probabilities are truncated to be at most 1
-    cont_prop <- min(exp(rnorm(1,
-                               cont_pred,
-                               cont_var)), 1)
-    beta_prop <- min(exp(rnorm(1,
-                               beta_pred,
-                               beta_var)), 1)
+    cont_prop <- exp(cont_pred)
+    beta_prop <- exp(beta_pred)
     
     cont_decrease_prop <- exp(cont_decrease_pred)
     beta_decrease_prop <- exp(beta_decrease_pred)
@@ -89,12 +85,8 @@ for (x in xs){
     
     # Sample the proportion of events from population effects
     # min() ensures that probabilities are truncated to be at most 1
-    cont_prop <- min(exp(rnorm(1,
-                               cont_pred,
-                               cont_var)), 1)
-    beta_prop <- min(exp(rnorm(1,
-                               beta_pred,
-                               beta_var)), 1)
+    cont_prop <- exp(cont_pred)
+    beta_prop <- exp(beta_pred)
     
     cont_decrease_prop <- exp(cont_decrease_pred)
     beta_decrease_prop <- exp(beta_decrease_pred)

### 2017-03-09 submission/manuscript.Rmd
Revisions manuscript, draft discussion.

@@
 However, the effectiveness of perioperative beta-blockade is obfuscated by both trials afflicted by data fabrication and the type of beta-blocker used in the trial.
-Various kinds of beta-blockers could vary in their pharmalogical effectiveness with respect to perioperative mortality, and anomalies in the effectiveness of the DECREASE trials might be due to differences in the beta-blockers used [@] and not purely due to data fabrication (given not all data points can be considered fabricated at this point).
+Various kinds of beta-blockers could vary in their pharmalogical effectiveness with respect to perioperative mortality, and anomalies in the effectiveness of the DECREASE trials might be due to differences in the beta-blockers used [@klei2015] and not purely due to data fabrication (given not all data points can be considered fabricated at this point).
 
@@
 We also include type of beta-blocker to inspect whether this is predictive of the effect of beta-blockers on perioperative mortality.
-Second, we evaluated the the probability that the DECREASE trial results occurred assuming no data fabrication.
+Second, we evaluated the probability that the DECREASE trials occurred assuming no data fabrication.
 Third, we reversed this assumption and assumed that data fabrication did occur and estimated how many data points would have to be fabricated to reproduce the results of the DECREASE trials, if the other published trials are regarded as estimating the true effect of beta-blockers on perioperative mortality in CAD patients undergoing non-cardiac surgery.
@@
 ```{r, echo = FALSE}
-options(scipen = 999)
+options(scipen = 5)
 suppressPackageStartupMessages(if(!require(metafor)){install.packages('metafor', repos = "http://cran.us.r-project.org")})
@@
                          weighted = TRUE)
+
+i2_mods <- mod_mods$I2
 ```
@@
 Second, we meta-analyzed all studies combined, including a dummy predictor for the DECREASE and non-DECREASE studies to reproduce results presented in Figure 4 of the 2014 meta-analysis [@bouri2014].
-Surprisingly, our results showed stronger evidence against equal subgroups than the original meta-analysis [@bouri2014] (original: $\chi^2(1)=3.91,p=.05$, reproduced $\chi^2(1)=`r round(mod_mods$QM, 2)`,p=`r round(mod_mods$QMp, 3)`$).
-Additionally, the original analyses showed substantial residual heterogeneity ($I^2=74.4$\%) whereas we found no residual heterogeneit ($I^2=$`r round(mod_mods$I2, 3)`\%).
+Surprisingly, our results showed stronger evidence against equal subgroups than the original meta-analysis [@bouri2014] (original: $\chi^2(1)=3.91,p=.05$; reproduced: $\chi^2(1)=`r round(mod_mods$QM, 2)`,p=`r round(mod_mods$QMp, 3)`$).
+Additionally, the original analyses showed substantial residual heterogeneity ($I^2=74.4$\%) whereas we found no residual heterogeneity ($I^2=`r round(i2_mods, 3)`$\%).
 Different variance estimates (e.g., `DerSimonian-Laird` instead of `REML`) did not resolve this difference.
@@
 To this end, we assumed that the non-DECREASE trials provide a valid representation of the true effect of beta-blockers on perioperative mortality (similar to Bouri et al.
-[@bouri2014]; but see the discussion).
+[@bouri2014]).
 The estimated probability is also known as the veracity of the data [@peters2015], which indicates the probability of the observed data under a given true effect.
 We assumed that the non-DECREASE studies estimated the true effect distribution of perioperative beta-blockade on mortality, not perturbed by publication bias due to statistical (non)significance.
-Publication bias was assumed to not be a problem because a substantial number of nonsignificant effects are included in the dataset (9 of 11 nonsignificant results).
+Publication bias was assumed to not be a problem because a substantial number of nonsignificant effects are included in the dataset (9 of 11 results are nonsignificant).
 
@@
 Based on the estimated mean log RR and its credible interval in the non-DECREASE studies, we computed the probability of the observed log RR in the DECREASE trials.
-The estimates of the non-DECREASE studies were obtained from Step 1, which include the estimated log RR (i.e., `r round(mod_clean$b, 2)`), and its 95\% credibility interval as provided by the package `metafor` (i.e., `r sprintf('[%s; %s]', round(mod_clean_pred$cr.lb, 3), round(mod_clean_pred$cr.ub, 3))`.
+The estimates of the non-DECREASE studies were obtained from Step 1, which include the estimated log RR (i.e., `r round(mod_clean$b, 2)`), and its 95\% credibility interval as provided by the package `metafor` (i.e., `r sprintf('[%s; %s]', round(mod_clean_pred$cr.lb, 3), round(mod_clean_pred$cr.ub, 3))`).
 The meta-analysis model assumes a normal distribution of population effects with the estimated effect as the mean of the distribution.
@@
 pdf(sprintf('%sfigures/fig1.pdf', x12), width = 8, height = 6)
-par(mar = c(.5,0,0,0))
+par(mar = c(3,0,0,0))
 plot(x = temp,
@@
      bty = 'n',
-     xaxt = 'n',
+     # xaxt = 'n',
      yaxt = 'n',
-     xlab = '',
+     xlab = 'log(RR)',
      ylab = '')
@@
 points(x = bouri_data$ln_rr[bouri_data$decrease == 1][2], y = 0)
-text(x = bouri_data$ln_rr[bouri_data$decrease == 1][2],
-     y = -.1,
+text(x = bouri_data$ln_rr[bouri_data$decrease == 1][2] - .05,
+     y = .15,
      labels = 'DECREASE-IV',
@@
 Based on the estimated effect distribution from the non-DECREASE trials, we calculated the probability of each DECREASE trial result, or a more extreme result.
-In other words, we computed the *p*-value for the hypothesis that the DECREASE studies occurred, assuming the information available from the other trials is informative of the true population effect.
+In other words, we computed the *p*-value for the null hypothesis that the DECREASE trials arise from the same effect distribution as the non-DECREASE trials.
+This assumes that the information available from the other trials is informative of the true population effect.
 
@@
 This indicates the DECREASE trial results are unlikely to have come from the same population effect distribution as the non-DECREASE trials.
-Moreover, observing two of such extremely unlikely results jointly, as in the DECREASE trials, is highly improbable, $`r prob[1]*prob[2]`$.
-This would therefore indicate that the DECREASE trials are severely different from the non-DECREASE trials.
+Moreover, observing two of such extremely unlikely results jointly, as in the DECREASE trials, is nearly impossible, $`r prob[1]*prob[2]`$.
+Hence, this result indicates that the DECREASE trials are severely different from the non-DECREASE trials.
 
-Results from Step 1 indicated that no between-study variance (i.e., homogeneity; $\tau^2=0$) of the effects was observed; given the small number of studies included (i.e., `r mod_clean$k`) this estimate is uncertain, however.
+Results from Step 1 indicated that no between-trial variance (i.e., homogeneity; $\tau^2=0$) of the effects was observed; given the small number of trials included (i.e., `r mod_clean$k`) this estimate is uncertain, however.
 We conducted sensitivity analyses to see how dependent results are on the heterogeneity estimate.
@@
 For each of these trial arms, we ran a meta-analysis applying the same methods used in Step 1.
-For each of the two DECREASE trials separately and the non-DECREASE trials combined, this resulted in four meta-analytic mortality estimates with corresponding effect variances (see Table 3).
+For each of the two DECREASE trials separately and the non-DECREASE trials combined, this resulted in four meta-analytic mortality estimates with corresponding effect variances (see Table 3; one estimate per cell).
 Throughout the simulations, we used the point estimates (i.e., fixed effect) to simulate genuine- and fabricated data, but supplemented this by using the more uncertain distribution estimates (i.e., random effects) as sensitivity analyses.
@@
 
-For DECREASE-I ($N=`r bouri_data$control_n[10] + bouri_data$beta_n[10]`), the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_i_certain - .025) - 1` - `r which.max(res_i_certain - .975) - 1`] or [`r which.min(res_i_uncertain - .025) - 1` - `r which.max(res_i_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
+For DECREASE-I ($N=`r bouri_data$control_n[10] + bouri_data$beta_n[10]`$), the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_i_certain - .025) - 1` - `r which.max(res_i_certain - .975) - 1`] or [`r which.min(res_i_uncertain - .025) - 1` - `r which.max(res_i_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
 The left column of Figure 3 depicts the $p_F$ per *X* fabricated data points (top panel) and the bounds of the confidence interval when the degree of confidence is altered (lower panel).
@@
 
-For DECREASE-IV ($N=`r bouri_data$control_n[11] + bouri_data$beta_n[11]`), the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_iv_certain - .025) - 1` - `r which.max(res_iv_certain - .975) - 1`] or [`r which.min(res_iv_uncertain - .025) - 1` - `r which.max(res_iv_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
+For DECREASE-IV ($N=`r bouri_data$control_n[11] + bouri_data$beta_n[11]`$), the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_iv_certain - .025) - 1` - `r which.max(res_iv_certain - .975) - 1`] or [`r which.min(res_iv_uncertain - .025) - 1` - `r which.max(res_iv_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
 The relatively minor difference between the estimates indicates that there is a high degree of confidence that data fabrication did occur based on the difference of the trial results alone.
@@
 
-The results also highlight that, despite the lack of availability of the raw data, summary results from larger samples allow estimation of the number of fabricated data points when similar trials are available.
-Moreover, larger trials result in relatively more certainty (e.g., DECREASE-IV) about the estimated number of fabricated data points using the inversion method compared to smaller trials (e.g., DECREASE-I), although much residual uncertainty remains.
-This increased certainty is due to decreased standard errors at the population level of the estimated effects, resulting in higher sensitivity to data anomalies.
-Nonetheless, substantially less information is available in just the summary results, and raw data availability would improve results verification and the detection of potential anomalies.
-These results also highlight that in order to prevent detection, it would be in the fabricators' interest to fabricate small studies even if raw data are hidden away (assuming the fabricator wants to remain undetected).
-
 # Discussion
 
-<!---
-GtR
-1. Is natuurlijk wle gevaarlijk als we in gedachten nemen dat DECREASE op een heel
-andere benadering van perioperatieve blockade is gestoeld.
-Ik begrijp dat een dergeliijk bezwaar het fundament van onze exercitie aantast,
-maar we moeten daar goed over nadenken.
-In discussie kunnen we zeggen:
-Decrease approach is niet wezenlijk anders dan non-decrease enonze analyse geldt
-Decrease appraoch is wezenijk anders qua aanpak en effect en daarom kunnen
-verschillen te wijten zijin aan fraude maar ook aan beter effect en die twee
-kunnen wij niet scheiden. 
-Nieuwe decrease aanpak trials zijj nodig (zie ook paper uit NTvG van Marleen)
-2. In Discussion we must debate the statistical aspects and the content aspects where DECREASE have a very different approach to the prevention of periop mortality than non-DECREASE.
-This makes our own work much more difficult to interpret
--->
+The effect of beta-blockade on perioperative mortality was already unclear following the findings of scientific misconduct; these results strongly affirm that the DECREASE trials should be neglected when assessing the effectiveness of beta-blockers.
+Our results indicate that the results from the DECREASE trials are nearly impossible to have arisen from the same effect inspected by the non-DECREASE trials, except when we assume at least some of the data were manipulated.
+As such, the scientific validity of the DECREASE-I and DECREASE-IV trials should by now be clear: they are to be regarded as irrelevant when assessing the effectiveness of beta-blockade on perioperative mortality.
+Nonetheless, the original papers that presented these trial results are not yet retracted [@poldermans1999;@dunkelgrun2009], something we strongly recommend based on the combination of our results and the integrity reports [@commissie2011;@commissie2012;@commissie2013].
 
-Throughout steps 1, 2, and 3 we provided evidence that the , the DECREASE trials are highly unlikely to come from a genuine data pool, and that there is a high degree of confidence that data was fab
+The ESC/ESA and ACC/AHA guidelines [@Kristensen_2014;@Fleisher_2014] on perioperative beta-blockade already excluded the DECREASE trials in their assessment, but also state that other trials by Poldermans are excluded.
+However, upon close inspection of the reference lists, the ACC/AHA guidelines still cites four trials as evidence base [@Boersma_2001;@poldermans1999;@van_Kuijk_2009;@Flu_2010], of which two were already inspected by the scientific integrity committees of Erasmus MC [@Boersma_2001;@poldermans1999].
+In the ACC/AHA guidelines, the following is said about studies conducted by Poldermans:
 
-<!--Move to discussion?-->
-We note that the variance in the non-DECREASE studies was estimated at 0, indicating that there would supposedly be a fixed-effect of beta-blockers on perioperative mortality despite differences in type of beta-blockers administered and dosage/duration of the beta-blockers.
-This homogeneity is not excessive, considering that the *Q*-statistic is sufficiently large to begin with (i.e., `r round(mod_clean$QE, .2)`).[@ioannidis2006] Nonetheless, the original non-DECREASE trials show substantive variation in the treatment, for example the usage of different beta-blockers (i.e., atenolol, propanolol, bisoprolol, metoprolol) in different (standardized) dosages, which have been argued to affect effectiveness of the interventions[@klei2015].
-As such, the lack of heterogeneity can be genuine and indicative of no difference between these different beta-blockers and their dosage/duration.
-Nonetheless, it could also be due to uncertainty in the estimate of heterogeneity $\tau^2$ due to the small set of studies included.
+> _"If nonretracted DECREASE publications and/or other derivative studies by Poldermans are relevant to the topic, they can only be cited in the text with a comment about the finding compared with the current recommendation but should not form the basis of that recommendation or be used as a reference for the recommendation."_ [@Fleisher_2014]
 
-However, due to various types of beta-blockers, evidence might be moderated [@klei2015].
+Nonetheless, references _are_ made without clear comments.
+Given the confirmation of problems in the DECREASE-I and DECREASE-IV trials in our results, it stresses that there is reason to distrust trials by Poldermans.
+We pose that investigations should be initiated into works where Poldermans was clearly involved and which were not cleared by the scientific committees of Erasmus MC in their misconduct investigations.
+Especially those papers cited as evidence in the ACC/AHA guidelines should be investigated, considering they affect patients and their treatment directly.
 
-Additionally, during the process of this research paper, the first author tried to ascertain a raw dataset of one of the DECREASE trials via a Freedom of Information request (FOI).
-The Erasmus MC refused to share these data
+Previously, further investigation of trials by Poldermans was deemed unfeasible due to the lack of raw data; here we indicate methods that do make it feasible.
+Based on just event-count data and comparable trials, we were able to estimate whether part of the data were in fact fabricated and whether the results were within reason of comparable trials.
+The results clearly indicated they were not.
 
-The ESC/ESA guidelines from 2014[@] were published in October 2014, whereas the meta-analysis indicating a reversal of the effectiveness of beta-blockers was published online at the end of July 2013[@].
-The committee revising the ESC/ESA guidelines stated: "The respective writing committees independently performed their literature review and analysis, and then developed their recommendations."[@Kristensen_2014] Regardless, the revisions made "recommend continuation of beta-blocker therapy in the perioperative period in patients currently receiving this medication" and do not recommend the "initiation of beta-blockers in patients undergoing low-risk surgery"[@guarracino2015].
-Review pieces note no substantial clinical change with respect to beta-blockers.[@Port2016;@Kristensen2016]
+The results of our analyses also highlight that, despite the lack of availability of the raw data, summary results from larger samples allow for better estimates of the number of fabricated data points when similar trials are available.
+Moreover, larger trials result in relatively more certainty (e.g., DECREASE-IV) about the estimated number of fabricated data points, when using the inversion method, compared to smaller trials (e.g., DECREASE-I).
+This increased certainty is due to decreased standard errors at the population level of the estimated effects, resulting in higher sensitivity to data anomalies.
+Nonetheless, much residual uncertainty remains and simply fewer  information is available in the summary results.
+Raw data availability would improve the options open to detect potential anomalies (note: raw data are available for DECREASE VI, but upon a freedom of information request Erasmus MC refused to share these data).
+The results also highlight that in order to prevent detection, it would be in the fabricators' interest to fabricate small studies even if raw data are hidden away (assuming the fabricator wants to remain undetected).
 
-The revisions to the ESC/ESA guidelines do not discourage use of beta-blockers, despite the obvious lack of evidence for systematic evidence of the effectiveness of beta-blockers to reduce perioperative mortality in trials.
-If anything, the trials show that insufficient evidence has been collected.
-This becomes apparent when the estimated effect for all non-DECREASE trials is inspected; its *p*-value is .04, which is insufficient evidence and is actually more likely when there is no effect in of beta-blockers in reality[@van_Aert_2016]
+With respect to clinical practice, the results provide some tentative evidence that type of beta-blockade can severely influence perioperative mortality.
+Our reanalysis of the Bouri et al.
+[@bouri2014] data indicates that type of beta-blockade can reverse the effect on perioperative mortality, even after taking into account whether a study belongs to the DECREASE family.
+As such, atenolol seems to tentatively decrease perioperative mortality, whereas the others (metoprolol, propanolol, bisoprolol) increase perioperative mortality.
+However, there seems to be covariation with respect to treatment administration, duration, and dose, which further confounds whether the treatment effect is due to type of beta-blocker or due to one of these other parameters.
+This affirms the statement from the ESC/ESA guidelines that _"high priority needs to be given to new randomized clinical trials to better identify which patients derive benefit from beta-blocker therapy in the perioperative setting, and to determine the optimal method of beta-blockade"_ [@Kristensen_2014].
 
-In order , the first author tried to ascertain the data of DECREASE VI, the only DECREASE study for which a dataset is still available[@Coleg5210] via a Freedom of Information (FOI) request.
-Within this paper, we inspected data anomalies based on summary results, which contain less information to be able to determine what isThe board of Erasmus Medical Centrum (EMC) was unwilling to comply with this request, stating that it was not a matter of the board.
-Regardless of whether it was a board related matter,
+In sum, our research indicates that the DECREASE trials are nearly impossible and contain at least some manipulated data points.
+We recommend retraction of the DECREASE-I and DECREASE-IV trials, and recommend renewed investigations into Poldermans' work --- especially that work still referenced by guidelines on the use of beta-blockers.
+Moreover, it remains unclear whether beta-blockers might be effective given the right treatment.
+We recommend new and more extensively controlled, confirmatory trials to determine whether there is any use in administering beta-blockers in order to decrease perioperative mortality --- at the moment there is insufficient evidence to determine any useful effect of beta-blockers.
 

### 2017-03-09 submission/manuscript.tex
Revisions manuscript, draft discussion.

@@
   \postdate{\par}
-  \date{02 March, 2017}
+  \date{09 March, 2017}
 
@@
 still recommend the usage of beta-blockers in the perioperative period
-in certain cases\textsuperscript{8,9}. Considering these consequences,
-we aim to estimate the extent of data fabrication in the DECREASE
-studies\textsuperscript{10,11} to further the debate on using
+in certain cases\textsuperscript{{\textbf{???}},8}. Considering these
+consequences, we aim to estimate the extent of data fabrication in the
+DECREASE studies\textsuperscript{9,10} to further the debate on using
 beta-blockers in the perioperative period of CAD patients undergoing
@@
 a method that has previously been used to monitor trial data quality or
-to test for potential data anomalies\textsuperscript{12}. Moreover, this
+to test for potential data anomalies\textsuperscript{11}. Moreover, this
 method has previously proven to be effective in detecting data
-fabrication\textsuperscript{13}. Evaluating the DECREASE trials to other
+fabrication\textsuperscript{12}. Evaluating the DECREASE trials to other
 published trials studying the effectiveness of beta-blockers with
@@
 mortality, and anomalies in the effectiveness of the DECREASE trials
-might be due to differences in the beta-blockers used {[}@{]} and not
-purely due to data fabrication (given not all data points can be
-considered fabricated at this point).
+might be due to differences in the beta-blockers
+used\textsuperscript{13} and not purely due to data fabrication (given
+not all data points can be considered fabricated at this point).
 
 To statistically investigate the evidence of data fabrication in the
-DECREASE studies\textsuperscript{10,11}, we took three steps. First, we
+DECREASE studies\textsuperscript{9,10}, we took three steps. First, we
 reproduced the findings from a 2014 meta-analysis\textsuperscript{7}
@@
 the effect of beta-blockers on perioperative mortality. Second, we
-evaluated the the probability that the DECREASE trial results occurred
-assuming no data fabrication. Third, we reversed this assumption and
-assumed that data fabrication did occur and estimated how many data
-points would have to be fabricated to reproduce the results of the
-DECREASE trials, if the other published trials are regarded as
-estimating the true effect of beta-blockers on perioperative mortality
-in CAD patients undergoing non-cardiac surgery.
+evaluated the probability that the DECREASE trials occurred assuming no
+data fabrication. Third, we reversed this assumption and assumed that
+data fabrication did occur and estimated how many data points would have
+to be fabricated to reproduce the results of the DECREASE trials, if the
+other published trials are regarded as estimating the true effect of
+beta-blockers on perioperative mortality in CAD patients undergoing
+non-cardiac surgery.
 
@@
 subgroups than the original meta-analysis\textsuperscript{7} (original:
-\(\chi^2(1)=3.91,p=.05\), reproduced \(\chi^2(1)=6.12,p=0.013\)).
+\(\chi^2(1)=3.91,p=.05\); reproduced: \(\chi^2(1)=6.12,p=0.013\)).
 Additionally, the original analyses showed substantial residual
-heterogeneity (\(I^2=74.4\)\%) whereas we found no residual heterogeneit
-(\$I\^{}2=\$0\%). Different variance estimates (e.g.,
+heterogeneity (\(I^2=74.4\)\%) whereas we found no residual
+heterogeneity (\(I^2=0\)\%). Different variance estimates (e.g.,
 \texttt{DerSimonian-Laird} instead of \texttt{REML}) did not resolve
@@
 representation of the true effect of beta-blockers on perioperative
-mortality (similar to Bouri et al.\textsuperscript{7}; but see the
-discussion). The estimated probability is also known as the veracity of
-the data\textsuperscript{17}, which indicates the probability of the
+mortality (similar to Bouri et al.\textsuperscript{7}). The estimated
+probability is also known as the veracity of the
+data\textsuperscript{17}, which indicates the probability of the
 observed data under a given true effect. We assumed that the
@@
 to not be a problem because a substantial number of nonsignificant
-effects are included in the dataset (9 of 11 nonsignificant results).
+effects are included in the dataset (9 of 11 results are
+nonsignificant).
 
@@
 and its 95\% credibility interval as provided by the package
-\texttt{metafor} (i.e., {[}0.01; 0.484{]}. The meta-analysis model
+\texttt{metafor} (i.e., {[}0.01; 0.484{]}). The meta-analysis model
 assumes a normal distribution of population effects with the estimated
@@
 extreme result. In other words, we computed the \emph{p}-value for the
-hypothesis that the DECREASE studies occurred, assuming the information
-available from the other trials is informative of the true population
-effect.
+null hypothesis that the DECREASE trials arise from the same effect
+distribution as the non-DECREASE trials. This assumes that the
+information available from the other trials is informative of the true
+population effect.
 
@@
 specifically, the results from DECREASE-I (or more extreme) have a
-probability of \(0\) (less than 1 in a quintillion) and the results from
-DECREASE-IV have a probability of \(0\) (3 in a billion). This indicates
-the DECREASE trial results are unlikely to have come from the same
+probability of \(2.9610213\times 10^{-53}\) (less than 1 in a
+quintillion) and the results from DECREASE-IV have a probability of
+\(3.0974376\times 10^{-9}\) (3 in a billion). This indicates the
+DECREASE trial results are unlikely to have come from the same
 population effect distribution as the non-DECREASE trials. Moreover,
 observing two of such extremely unlikely results jointly, as in the
-DECREASE trials, is highly improbable, \(0\). This would therefore
-indicate that the DECREASE trials are severely different from the
-non-DECREASE trials.
+DECREASE trials, is nearly impossible, \(9.1715786\times 10^{-62}\).
+Hence, this result indicates that the DECREASE trials are severely
+different from the non-DECREASE trials.
 
-Results from Step 1 indicated that no between-study variance (i.e.,
+Results from Step 1 indicated that no between-trial variance (i.e.,
 homogeneity; \(\tau^2=0\)) of the effects was observed; given the small
-number of studies included (i.e., 9) this estimate is uncertain,
-however. We conducted sensitivity analyses to see how dependent results
-are on the heterogeneity estimate. The probability of observing the
-DECREASE trials stays approximately below 1 out of 1000 until the
-variance estimate is 0.25; the probability stays approximately below 1
-out of 100 until the variance estimate is 0.43 (see Figure 2). To put
-these numbers into context, a variance of 0.25 would suggest that
-results of perioperative beta-blockade vary substantially even if
-perioperative beta-blockade has no effect whatsoever (RRs between 0.779
-and 1.284 in 64\% of the cases) all due to contextual circumstances of
-the study.
+number of trials included (i.e., 9) this estimate is uncertain, however.
+We conducted sensitivity analyses to see how dependent results are on
+the heterogeneity estimate. The probability of observing the DECREASE
+trials stays approximately below 1 out of 1000 until the variance
+estimate is 0.25; the probability stays approximately below 1 out of 100
+until the variance estimate is 0.43 (see Figure 2). To put these numbers
+into context, a variance of 0.25 would suggest that results of
+perioperative beta-blockade vary substantially even if perioperative
+beta-blockade has no effect whatsoever (RRs between 0.779 and 1.284 in
+64\% of the cases) all due to contextual circumstances of the study.
 
@@
 this resulted in four meta-analytic mortality estimates with
-corresponding effect variances (see Table 3). Throughout the
-simulations, we used the point estimates (i.e., fixed effect) to
-simulate genuine- and fabricated data, but supplemented this by using
-the more uncertain distribution estimates (i.e., random effects) as
-sensitivity analyses.
+corresponding effect variances (see Table 3; one estimate per cell).
+Throughout the simulations, we used the point estimates (i.e., fixed
+effect) to simulate genuine- and fabricated data, but supplemented this
+by using the more uncertain distribution estimates (i.e., random
+effects) as sensitivity analyses.
 
@@
 
-For DECREASE-I (\$N=112), the 95\% confidence interval for the estimated
-number of fabricated data points is {[}0 - 112{]} or {[}2 - 112{]} when
-based on a point estimate or a more uncertain distribution estimate,
-respectively. The left column of Figure 3 depicts the \(p_F\) per
-\emph{X} fabricated data points (top panel) and the bounds of the
+For DECREASE-I (\(N=112\)), the 95\% confidence interval for the
+estimated number of fabricated data points is {[}0 - 112{]} or {[}2 -
+112{]} when based on a point estimate or a more uncertain distribution
+estimate, respectively. The left column of Figure 3 depicts the \(p_F\)
+per \emph{X} fabricated data points (top panel) and the bounds of the
 confidence interval when the degree of confidence is altered (lower
@@
 
-For DECREASE-IV (\$N=1066), the 95\% confidence interval for the
-estimated number of fabricated data points is {[}0 - 1053{]} or {[}8 -
+For DECREASE-IV (\(N=1066\)), the 95\% confidence interval for the
+estimated number of fabricated data points is {[}3 - 1059{]} or {[}8 -
 1066{]} when based on a point estimate or a more uncertain distribution
@@
 
-The results also highlight that, despite the lack of availability of the
-raw data, summary results from larger samples allow estimation of the
-number of fabricated data points when similar trials are available.
-Moreover, larger trials result in relatively more certainty (e.g.,
-DECREASE-IV) about the estimated number of fabricated data points using
-the inversion method compared to smaller trials (e.g., DECREASE-I),
-although much residual uncertainty remains. This increased certainty is
-due to decreased standard errors at the population level of the
-estimated effects, resulting in higher sensitivity to data anomalies.
-Nonetheless, substantially less information is available in just the
-summary results, and raw data availability would improve results
-verification and the detection of potential anomalies. These results
-also highlight that in order to prevent detection, it would be in the
-fabricators' interest to fabricate small studies even if raw data are
-hidden away (assuming the fabricator wants to remain undetected).
-
 \section{Discussion}\label{discussion}
 
-Throughout steps 1, 2, and 3 we provided evidence that the , the
-DECREASE trials are highly unlikely to come from a genuine data pool,
-and that there is a high degree of confidence that data was fab
+The effect of beta-blockade on perioperative mortality was already
+unclear following the findings of scientific misconduct; these results
+strongly affirm that the DECREASE trials should be neglected when
+assessing the effectiveness of beta-blockers. Our results indicate that
+the results from the DECREASE trials are nearly impossible to have
+arisen from the same effect inspected by the non-DECREASE trials, except
+when we assume at least some of the data were manipulated. As such, the
+scientific validity of the DECREASE-I and DECREASE-IV trials should by
+now be clear: they are to be regarded as irrelevant when assessing the
+effectiveness of beta-blockade on perioperative mortality. Nonetheless,
+the original papers that presented these trial results are not yet
+retracted\textsuperscript{9,10}, something we strongly recommend based
+on the combination of our results and the integrity
+reports\textsuperscript{2--4}.
 
-We note that the variance in the non-DECREASE studies was estimated at
-0, indicating that there would supposedly be a fixed-effect of
-beta-blockers on perioperative mortality despite differences in type of
-beta-blockers administered and dosage/duration of the beta-blockers.
-This homogeneity is not excessive, considering that the
-\emph{Q}-statistic is sufficiently large to begin with (i.e.,
-5).\textsuperscript{19} Nonetheless, the original non-DECREASE trials
-show substantive variation in the treatment, for example the usage of
-different beta-blockers (i.e., atenolol, propanolol, bisoprolol,
-metoprolol) in different (standardized) dosages, which have been argued
-to affect effectiveness of the interventions\textsuperscript{20}. As
-such, the lack of heterogeneity can be genuine and indicative of no
-difference between these different beta-blockers and their
-dosage/duration. Nonetheless, it could also be due to uncertainty in the
-estimate of heterogeneity \(\tau^2\) due to the small set of studies
-included.
+The ESC/ESA and ACC/AHA guidelines\textsuperscript{19,20} on
+perioperative beta-blockade already excluded the DECREASE trials in
+their assessment, but also state that other trials by Poldermans are
+excluded. However, upon close inspection of the reference lists, the
+ACC/AHA guidelines still cites four trials as evidence
+base\textsuperscript{10,21--23}, of which two were already inspected by
+the scientific integrity committees of Erasmus
+MC\textsuperscript{10,21}. In the ACC/AHA guidelines, the following is
+said about studies conducted by Poldermans:
 
-However, due to various types of beta-blockers, evidence might be
-moderated\textsuperscript{20}.
+\begin{quote}
+\emph{``If nonretracted DECREASE publications and/or other derivative
+studies by Poldermans are relevant to the topic, they can only be cited
+in the text with a comment about the finding compared with the current
+recommendation but should not form the basis of that recommendation or
+be used as a reference for the recommendation.''}\textsuperscript{20}
+\end{quote}
 
-Additionally, during the process of this research paper, the first
-author tried to ascertain a raw dataset of one of the DECREASE trials
-via a Freedom of Information request (FOI). The Erasmus MC refused to
-share these data
+Nonetheless, references \emph{are} made without clear comments. Given
+the confirmation of problems in the DECREASE-I and DECREASE-IV trials in
+our results, it stresses that there is reason to distrust trials by
+Poldermans. We pose that investigations should be initiated into works
+where Poldermans was clearly involved and which were not cleared by the
+scientific committees of Erasmus MC in their misconduct investigations.
+Especially those papers cited as evidence in the ACC/AHA guidelines
+should be investigated, considering they affect patients and their
+treatment directly.
 
-The ESC/ESA guidelines from 2014{[}@{]} were published in October 2014,
-whereas the meta-analysis indicating a reversal of the effectiveness of
-beta-blockers was published online at the end of July 2013{[}@{]}. The
-committee revising the ESC/ESA guidelines stated: ``The respective
-writing committees independently performed their literature review and
-analysis, and then developed their
-recommendations.''\textsuperscript{21} Regardless, the revisions made
-``recommend continuation of beta-blocker therapy in the perioperative
-period in patients currently receiving this medication'' and do not
-recommend the ``initiation of beta-blockers in patients undergoing
-low-risk surgery''\textsuperscript{22}. Review pieces note no
-substantial clinical change with respect to
-beta-blockers.\textsuperscript{23,24}
+Previously, further investigation of trials by Poldermans was deemed
+unfeasible due to the lack of raw data; here we indicate methods that do
+make it feasible. Based on just event-count data and comparable trials,
+we were able to estimate whether part of the data were in fact
+fabricated and whether the results were within reason of comparable
+trials. The results clearly indicated they were not.
 
-The revisions to the ESC/ESA guidelines do not discourage use of
-beta-blockers, despite the obvious lack of evidence for systematic
-evidence of the effectiveness of beta-blockers to reduce perioperative
-mortality in trials. If anything, the trials show that insufficient
-evidence has been collected. This becomes apparent when the estimated
-effect for all non-DECREASE trials is inspected; its \emph{p}-value is
-.04, which is insufficient evidence and is actually more likely when
-there is no effect in of beta-blockers in reality\textsuperscript{25}
+The results of our analyses also highlight that, despite the lack of
+availability of the raw data, summary results from larger samples allow
+for better estimates of the number of fabricated data points when
+similar trials are available. Moreover, larger trials result in
+relatively more certainty (e.g., DECREASE-IV) about the estimated number
+of fabricated data points, when using the inversion method, compared to
+smaller trials (e.g., DECREASE-I). This increased certainty is due to
+decreased standard errors at the population level of the estimated
+effects, resulting in higher sensitivity to data anomalies. Nonetheless,
+much residual uncertainty remains and simply fewer information is
+available in the summary results. Raw data availability would improve
+the options open to detect potential anomalies (note: raw data are
+available for DECREASE VI, but upon a freedom of information request
+Erasmus MC refused to share these data). The results also highlight that
+in order to prevent detection, it would be in the fabricators' interest
+to fabricate small studies even if raw data are hidden away (assuming
+the fabricator wants to remain undetected).
 
-In order , the first author tried to ascertain the data of DECREASE VI,
-the only DECREASE study for which a dataset is still
-available\textsuperscript{1} via a Freedom of Information (FOI) request.
-Within this paper, we inspected data anomalies based on summary results,
-which contain less information to be able to determine what isThe board
-of Erasmus Medical Centrum (EMC) was unwilling to comply with this
-request, stating that it was not a matter of the board. Regardless of
-whether it was a board related matter,
+With respect to clinical practice, the results provide some tentative
+evidence that type of beta-blockade can severely influence perioperative
+mortality. Our reanalysis of the Bouri et al.\textsuperscript{7} data
+indicates that type of beta-blockade can reverse the effect on
+perioperative mortality, even after taking into account whether a study
+belongs to the DECREASE family. As such, atenolol seems to tentatively
+decrease perioperative mortality, whereas the others (metoprolol,
+propanolol, bisoprolol) increase perioperative mortality. However, there
+seems to be covariation with respect to treatment administration,
+duration, and dose, which further confounds whether the treatment effect
+is due to type of beta-blocker or due to one of these other parameters.
+This affirms the statement from the ESC/ESA guidelines that \emph{``high
+priority needs to be given to new randomized clinical trials to better
+identify which patients derive benefit from beta-blocker therapy in the
+perioperative setting, and to determine the optimal method of
+beta-blockade''}\textsuperscript{19}.
+
+In sum, our research indicates that the DECREASE trials are nearly
+impossible and contain at least some manipulated data points. We
+recommend retraction of the DECREASE-I and DECREASE-IV trials, and
+recommend renewed investigations into Poldermans' work --- especially
+that work still referenced by guidelines on the use of beta-blockers.
+Moreover, it remains unclear whether beta-blockers might be effective
+given the right treatment. We recommend new and more extensively
+controlled, confirmatory trials to determine whether there is any use in
+administering beta-blockers in order to decrease perioperative mortality
+--- at the moment there is insufficient evidence to determine any useful
+effect of beta-blockers.
 
@@
 
-\hypertarget{ref-Fleishere278}{}
-9. Fleisher LA, Fleischmann KE, Auerbach AD, et al. 2014 acc/aha
-guideline on perioperative cardiovascular evaluation and management of
-patients undergoing noncardiac surgery. \emph{Circulation}.
-2014;130(24):e278-e333.
-doi:\href{https://doi.org/10.1161/CIR.0000000000000106}{10.1161/CIR.0000000000000106}.
-
 \hypertarget{ref-dunkelgrun2009}{}
-10. Dunkelgrun M, Boersma E, Schouten O, et al. Bisoprolol and
+9. Dunkelgrun M, Boersma E, Schouten O, et al. Bisoprolol and
 fluvastatin for the reduction of perioperative cardiac mortality and
@@
 \hypertarget{ref-poldermans1999}{}
-11. Poldermans D, Boersma E, Bax JJ, et al. The effect of bisoprolol on
+10. Poldermans D, Boersma E, Bax JJ, et al. The effect of bisoprolol on
 perioperative mortality and myocardial infarction in High-Risk patients
@@
 \hypertarget{ref-Buyse1999-jq}{}
-12. Buyse M, George SL, Evans S, et al. The role of biostatistics in the
+11. Buyse M, George SL, Evans S, et al. The role of biostatistics in the
 prevention, detection and treatment of fraud in clinical trials.
@@
 \hypertarget{ref-Knepper2016-la}{}
-13. Knepper D, Lindblad AS, Sharma G, et al. Statistical monitoring in
+12. Knepper D, Lindblad AS, Sharma G, et al. Statistical monitoring in
 clinical trials: Best practices for detecting data anomalies suggestive
@@
 doi:\href{https://doi.org/10.1177/2168479016630576}{10.1177/2168479016630576}.
+
+\hypertarget{ref-klei2015}{}
+13. Klei W van. Welke perioperatieve bètablokker heeft de voorkeur?
+{[}Which perioperative beta-blocker is preferred?{]}. \emph{Nederlands
+Tijdschrift voor Geneeskunde}. 2015;159:A9798.
 
@@
 
-\hypertarget{ref-ioannidis2006}{}
-19. Ioannidis JPA, Trikalinos TA, Zintzaras E. Extreme between-study
-homogeneity in meta-analyses could offer useful insights. \emph{Journal
-of clinical epidemiology}. 2006;59(10):1023-1032.
-doi:\href{https://doi.org/10.1016/j.jclinepi.2006.02.013}{10.1016/j.jclinepi.2006.02.013}.
-
-\hypertarget{ref-klei2015}{}
-20. Klei W van. Welke perioperatieve bètablokker heeft de voorkeur?
-{[}Which perioperative beta-blocker is preferred?{]}. \emph{Nederlands
-Tijdschrift voor Geneeskunde}. 2015;159:A9798.
-
 \hypertarget{ref-Kristensen_2014}{}
-21. Kristensen SD, Knuuti J, Saraste A, et al. 2014 ESC/ESA guidelines
+19. Kristensen SD, Knuuti J, Saraste A, et al. 2014 ESC/ESA guidelines
 on non-cardiac surgery. \emph{European Journal of Anaesthesiology}.
@@
 
-\hypertarget{ref-guarracino2015}{}
-22. Guarracino F PH Baldassarri R. Revised esc/esa guidelines on
-non-cardiac surgery: Cardiovascular assessment and management.
-implications for preoperative clinical evaluation. \emph{Minerva
-Anestesiology}. 2015;2:225-233.
+\hypertarget{ref-Fleisher_2014}{}
+20. Fleisher LA, Fleischmann KE, Auerbach AD, et al. 2014 ACC/AHA
+guideline on perioperative cardiovascular evaluation and management of
+patients undergoing noncardiac surgery: Executive summary: A report of
+the american college of cardiology/american heart association task force
+on practice guidelines. \emph{Circulation}. 2014;130(24):2215-2245.
+doi:\href{https://doi.org/10.1161/cir.0000000000000105}{10.1161/cir.0000000000000105}.
 
-\hypertarget{ref-Port2016}{}
-23. Port SC. 2014 esc/esa guidelines on noncardiac surgery:
-Cardiovascular assessment and management. \emph{Journal of Nuclear
-Cardiology}. 2016:1-3.
-doi:\href{https://doi.org/10.1007/s12350-016-0641-x}{10.1007/s12350-016-0641-x}.
+\hypertarget{ref-Boersma_2001}{}
+21. Boersma E, Poldermans D, Bax JJ, et al. Predictors of cardiac events
+after major vascular surgery. \emph{JAMA}. 2001;285(14):1865.
+doi:\href{https://doi.org/10.1001/jama.285.14.1865}{10.1001/jama.285.14.1865}.
 
-\hypertarget{ref-Kristensen2016}{}
-24. Kristensen SD. 2014 esc/esa guidelines on noncardiac surgery:
-Cardiovascular assessment and management. \emph{Journal of Nuclear
-Cardiology}. 2016:1-3.
-doi:\href{https://doi.org/10.1007/s12350-016-0642-9}{10.1007/s12350-016-0642-9}.
+\hypertarget{ref-van_Kuijk_2009}{}
+22. Kuijk J-P van, Flu W-J, Schouten O, et al. Timing of noncardiac
+surgery after coronary artery stenting with bare metal or drug-eluting
+stents. \emph{The American Journal of Cardiology}.
+2009;104(9):1229-1234.
+doi:\href{https://doi.org/10.1016/j.amjcard.2009.06.038}{10.1016/j.amjcard.2009.06.038}.
 
-\hypertarget{ref-van_Aert_2016}{}
-25. Aert RCM van, Wicherts JM, Assen MALM van. Conducting meta-analyses
-based on p values: Reservations and recommendations for applying
-p-uniform and p-curve. \emph{Perspectives on Psychological Science}.
-2016;11(5):713-729.
-doi:\href{https://doi.org/10.1177/1745691616650874}{10.1177/1745691616650874}.
+\hypertarget{ref-Flu_2010}{}
+23. Flu W-J, Kuijk J-P van, Chonchol M, et al. Timing of pre-operative
+beta-blocker treatment in vascular surgery patients. \emph{Journal of
+the American College of Cardiology}. 2010;56(23):1922-1929.
+doi:\href{https://doi.org/10.1016/j.jacc.2010.05.056}{10.1016/j.jacc.2010.05.056}.
 

### 2017-05-28 submission/manuscript.Rmd
Residual commit

@@
 ---
-title: "Analyzing DECREASE trials for extent of data fabrication"
+title: "Analyzing DECREASE trials for extent of data fabrication [working title]"
 author: "Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper"
@@
 
-The effect of beta-blockers on perioperative mortality in non-cardiac surgery for patients with coronary artery disease (CAD) has been controversial [@Coleg5210] due to findings of research  misconduct in two related clinical trials  [@commissie2011;@commissie2012;@commissie2013].
+The effect of beta-blockers on perioperative mortality in non-cardiac surgery has been controversial [@Coleg5210] due to findings of research  misconduct in two related clinical trials  [@poldermans1999;@dunkelgrun2009;@commissie2011;@commissie2012;@commissie2013].
 Three meta-analyses that included the trials subject to research misconduct concluded that beta-blockers decrease perioperative mortality [@Devereaux313;@Angeli2010;@bouri2014] whereas a meta-analysis that excluded the suspect trials concluded that beta-blockers increase perioperative mortality [@bouri2014].
@@
 
-<!-- Should we quote the (translated) main conclusion form the reports? -->
 The trials subject to research misconduct were the Dutch DECREASE-I and DECREASE-IV trials [@commissie2011;@commissie2012;@commissie2013].
@@
 Moreover, the latest guidelines still recommend the usage of beta-blockers in the perioperative period in certain cases [@esc2014;@Fleishere278].
-Considering these consequences, we aim to estimate the extent of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999] to further the debate on using beta-blockers in the perioperative period of CAD patients undergoing non-cardiac surgery.
+Considering these consequences, we aim to estimate the extent of data fabrication in the DECREASE studies [@poldermans1999;@dunkelgrun2009;@commissie2011;@commissie2012;@commissie2013] to further the debate on using beta-blockers in the perioperative period of CAD patients undergoing non-cardiac surgery.
 
@@
 Moreover, this method has previously proven to be effective in detecting data fabrication [@Knepper2016-la].
-Evaluating the DECREASE trials to other published trials studying the effectiveness of beta-blockers with respect to perioperative mortality could prove informative of the potential extent of the fabrication in the DECREASE trials.
+Comparing the DECREASE trials to other published trials studying the effectiveness of beta-blockers with respect to perioperative mortality could prove informative of the potential extent of the fabrication in the DECREASE trials.
 
-<!-- nieuw  revise-->
-However, the effectiveness of perioperative beta-blockade is obfuscated by both trials afflicted by data fabrication and the type of beta-blocker used in the trial.
-Various kinds of beta-blockers could vary in their pharmalogical effectiveness with respect to perioperative mortality, and anomalies in the effectiveness of the DECREASE trials might be due to differences in the beta-blockers used [@klei2015] and not purely due to data fabrication (given not all data points can be considered fabricated at this point).
+The effectiveness of perioperative beta-blockade is obfuscated by the afflicted DECREASE trials, potentially interacting with the type of beta-blocker and the way that beta-blocker was administered (i.e., dose and duration of treatment).
+Throughout the clinical trials regarding beta-blockers, patients were administered various beta-blockers (e.g., metroprolol, bisoprolol, atenolol) and in various ways (e.g., intraveneously, orally; half an hour before surgery or multiple days before surgery).
+Factors such as dosage and duration can have an effect on the pharmacological effectiveness with respect to perioperative mortality.
+Moreover, the anomalous results from the DECREASE trials might partly be caused by such differences [@klei2015] and not purely due to data fabrication (given not all data points can be considered fabricated at this point).
+
+The latest guidelines of the European Society of Cardiology and the American Heart Association still recommend the usage of beta-blockers in certain cases [@;@].
+Therefore, we evaluate the usage of beta-blockers with respect to its effectiveness for the the perioperative period in patients undergoing non-cardiac surgery.
+First, we estimated the extent of data fabrication in the DECREASE studies [@poldermans1999;@dunkelgrun2009] Secondly, we evaluated the clinical studies used in the different meta-analysis from a pharmacological point of view.
+
+# Part 1: statistical estimation of data fabrication in DECREASE trials
 
 To statistically investigate the evidence of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999], we took three steps.
-First, we reproduced the findings from a 2014 meta-analysis [@bouri2014] that contained sufficient information to estimate the deviation of the DECREASE trials from other published trials on beta-blockers.
+First, we reproduced the findings from the 2014  meta-analysis by Bouri et al.
+[@bouri2014] that contained sufficient information to estimate the deviation of the DECREASE trials from other published trials on beta-blockers.
 We also include type of beta-blocker to inspect whether this is predictive of the effect of beta-blockers on perioperative mortality.
 Second, we evaluated the probability that the DECREASE trials occurred assuming no data fabrication.
-Third, we reversed this assumption and assumed that data fabrication did occur and estimated how many data points would have to be fabricated to reproduce the results of the DECREASE trials, if the other published trials are regarded as estimating the true effect of beta-blockers on perioperative mortality in CAD patients undergoing non-cardiac surgery.
+Third, we reversed this assumption and assumed that data fabrication did occur and estimated how many data points would have to be fabricated to reproduce the results of the DECREASE trials, if the other published trials are regarded as estimating the true effect of beta-blockers on perioperative mortality in patients undergoing non-cardiac surgery.
 
-# Step 1: reproducing meta-analysis of Bouri et al. (2014)
+## Step 1: reproducing meta-analysis of Bouri et al. (2014)
 
-## Methods
+### Methods
 
@@
 
-## Results
+### Results
 
@@
 
-kable(table1, caption = 'The original- and reproduced meta-analytic results based on the data provided in the 2014 meta-analysis by Bouri et al.')
+knitr::kable(table1, caption = '__Table XX.__ The original- and reproduced meta-analytic results based on the data provided in the 2014 meta-analysis by Bouri et al.')
 ```
@@
 
-# Step 2: evaluating the veracity of DECREASE studies
+## Step 2: evaluating the veracity of DECREASE studies
 
@@
 
-## Method
+### Method
 
@@
 
-# Step 3: estimating the amount of fabricated data
+## Step 3: estimating the amount of fabricated data
 
@@
 
+# Part 2:
+
 # Discussion

### 2017-05-28 submission/manuscript.tex
Residual commit

@@
 \hypersetup{unicode=true,
-            pdftitle={Analyzing DECREASE trials for extent of data fabrication},
+            pdftitle={Analyzing DECREASE trials for extent of data fabrication {[}working title{]}},
             pdfauthor={Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper},
@@
 \setlength{\droptitle}{-2em}
-  \title{Analyzing DECREASE trials for extent of data fabrication}
+  \title{Analyzing DECREASE trials for extent of data fabrication {[}working
+title{]}}
   \pretitle{\vspace{\droptitle}\centering\huge}
@@
   \postdate{\par}
-  \date{09 March, 2017}
+  \date{24 May, 2017}
 
@@
 The effect of beta-blockers on perioperative mortality in non-cardiac
-surgery for patients with coronary artery disease (CAD) has been
-controversial\textsuperscript{1} due to findings of research misconduct
-in two related clinical trials\textsuperscript{2--4}. Three
-meta-analyses that included the trials subject to research misconduct
-concluded that beta-blockers decrease perioperative
-mortality\textsuperscript{5--7} whereas a meta-analysis that excluded
-the suspect trials concluded that beta-blockers increase perioperative
-mortality\textsuperscript{7}. In these studies, perioperative mortality
-was defined as the deathrate of patients during the period of the
-surgical procedure, which typically includes admission, anaesthesia,
-surgery, and recovery.
+surgery has been controversial\textsuperscript{1} due to findings of
+research misconduct in two related clinical
+trials\textsuperscript{2--6}. Three meta-analyses that included the
+trials subject to research misconduct concluded that beta-blockers
+decrease perioperative mortality\textsuperscript{7--9} whereas a
+meta-analysis that excluded the suspect trials concluded that
+beta-blockers increase perioperative mortality\textsuperscript{9}. In
+these studies, perioperative mortality was defined as the deathrate of
+patients during the period of the surgical procedure, which typically
+includes admission, anaesthesia, surgery, and recovery.
 
 The trials subject to research misconduct were the Dutch DECREASE-I and
-DECREASE-IV trials\textsuperscript{2--4}. The committees that
+DECREASE-IV trials\textsuperscript{4--6}. The committees that
 investigated the integrity of the DECREASE trials reported that data
 fabrication was likely but that the extent of the data fabrication
-remained unclear\textsuperscript{2--4}. Moreover, the latest guidelines
+remained unclear\textsuperscript{4--6}. Moreover, the latest guidelines
 still recommend the usage of beta-blockers in the perioperative period
-in certain cases\textsuperscript{{\textbf{???}},8}. Considering these
+in certain cases\textsuperscript{{\textbf{???}},10}. Considering these
 consequences, we aim to estimate the extent of data fabrication in the
-DECREASE studies\textsuperscript{9,10} to further the debate on using
+DECREASE studies\textsuperscript{2--6} to further the debate on using
 beta-blockers in the perioperative period of CAD patients undergoing
@@
 evaluated the applicability of forensic statistical
-methods\textsuperscript{4} to evaluate results of trials separately
+methods\textsuperscript{6} to evaluate results of trials separately
 (i.e., DECREASE-I, DECREASE-IV). Nonetheless, comparing across trials is
@@
 method has previously proven to be effective in detecting data
-fabrication\textsuperscript{12}. Evaluating the DECREASE trials to other
+fabrication\textsuperscript{12}. Comparing the DECREASE trials to other
 published trials studying the effectiveness of beta-blockers with
@@
 
-However, the effectiveness of perioperative beta-blockade is obfuscated
-by both trials afflicted by data fabrication and the type of
-beta-blocker used in the trial. Various kinds of beta-blockers could
-vary in their pharmalogical effectiveness with respect to perioperative
-mortality, and anomalies in the effectiveness of the DECREASE trials
-might be due to differences in the beta-blockers
-used\textsuperscript{13} and not purely due to data fabrication (given
-not all data points can be considered fabricated at this point).
+The effectiveness of perioperative beta-blockade is obfuscated by the
+afflicted DECREASE trials, potentially interacting with the type of
+beta-blocker and the way that beta-blocker was administered (i.e., dose
+and duration of treatment). Throughout the clinical trials regarding
+beta-blockers, patients were administered various beta-blockers (e.g.,
+metroprolol, bisoprolol, atenolol) and in various ways (e.g.,
+intraveneously, orally; half an hour before surgery or multiple days
+before surgery). Factors such as dosage and duration can have an effect
+on the pharmacological effectiveness with respect to perioperative
+mortality. Moreover, the anomalous results from the DECREASE trials
+might partly be caused by such differences\textsuperscript{13} and not
+purely due to data fabrication (given not all data points can be
+considered fabricated at this point).
+
+The latest guidelines of the European Society of Cardiology and the
+American Heart Association still recommend the usage of beta-blockers in
+certain cases {[}@;@{]}. Therefore, we evaluate the usage of
+beta-blockers with respect to its effectiveness for the the
+perioperative period in patients undergoing non-cardiac surgery. First,
+we estimated the extent of data fabrication in the DECREASE
+studies\textsuperscript{2,3} Secondly, we evaluated the clinical studies
+used in the different meta-analysis from a pharmacological point of
+view.
+
+\section{Part 1: statistical estimation of data fabrication in DECREASE
+trials}\label{part-1-statistical-estimation-of-data-fabrication-in-decrease-trials}
 
 To statistically investigate the evidence of data fabrication in the
-DECREASE studies\textsuperscript{9,10}, we took three steps. First, we
-reproduced the findings from a 2014 meta-analysis\textsuperscript{7}
-that contained sufficient information to estimate the deviation of the
-DECREASE trials from other published trials on beta-blockers. We also
-include type of beta-blocker to inspect whether this is predictive of
-the effect of beta-blockers on perioperative mortality. Second, we
-evaluated the probability that the DECREASE trials occurred assuming no
-data fabrication. Third, we reversed this assumption and assumed that
-data fabrication did occur and estimated how many data points would have
-to be fabricated to reproduce the results of the DECREASE trials, if the
-other published trials are regarded as estimating the true effect of
-beta-blockers on perioperative mortality in CAD patients undergoing
-non-cardiac surgery.
+DECREASE studies\textsuperscript{2,3}, we took three steps. First, we
+reproduced the findings from the 2014 meta-analysis by Bouri et
+al.\textsuperscript{9} that contained sufficient information to estimate
+the deviation of the DECREASE trials from other published trials on
+beta-blockers. We also include type of beta-blocker to inspect whether
+this is predictive of the effect of beta-blockers on perioperative
+mortality. Second, we evaluated the probability that the DECREASE trials
+occurred assuming no data fabrication. Third, we reversed this
+assumption and assumed that data fabrication did occur and estimated how
+many data points would have to be fabricated to reproduce the results of
+the DECREASE trials, if the other published trials are regarded as
+estimating the true effect of beta-blockers on perioperative mortality
+in patients undergoing non-cardiac surgery.
 
-\section{Step 1: reproducing meta-analysis of Bouri et al.
+\subsection{Step 1: reproducing meta-analysis of Bouri et al.
 (2014)}\label{step-1-reproducing-meta-analysis-of-bouri-et-al.-2014}
 
-\subsection{Methods}\label{methods}
+\subsubsection{Methods}\label{methods}
 
 To ensure that we used similar analysis procedures as in the 2014
-meta-analysis\textsuperscript{7}, we initially reproduced Bouri et al.'s
+meta-analysis\textsuperscript{9}, we initially reproduced Bouri et al.'s
 estimates. This ensured that (1) their results are reproducible and (2)
 we are using the correct estimates in subsequent steps of our analyses.
-Using figures 2 and 3 from the original paper\textsuperscript{7}, we
+Using figures 2 and 3 from the original paper\textsuperscript{9}, we
 extracted the raw event data for the 2 (control vs experimental) by 2
@@
 \href{https://osf.io/vnmzc}{osf.io/vnmzc}.
+
+\begin{verbatim}
+## Warning in library(package, lib.loc = lib.loc, character.only = TRUE,
+## logical.return = TRUE, : there is no package called 'metafor'
+\end{verbatim}
+
+\begin{verbatim}
+## Installing package into '/home/chjh/R/x86_64-redhat-linux-gnu-library/3.3'
+## (as 'lib' is unspecified)
+\end{verbatim}
 
@@
 computational artefacts\textsuperscript{16}. The 2014
-meta-analysis\textsuperscript{7} did not specify the variance estimate
+meta-analysis\textsuperscript{9} did not specify the variance estimate
 used; hence, minor discrepancies between our estimates and the original
@@
 
-\subsection{Results}\label{results}
+\subsubsection{Results}\label{results}
 
 We were able to closely reproduce the estimates for the different sets
-of studies (Figure 2 of the 2014 meta-analysis\textsuperscript{7}).
+of studies (Figure 2 of the 2014 meta-analysis\textsuperscript{9}).
 Bouri et al. differentiated between the estimates from the non-DECREASE
@@
 \begin{longtable}[]{@{}lllll@{}}
-\caption{The original- and reproduced meta-analytic results based on the
-data provided in the 2014 meta-analysis by Bouri et al.}\tabularnewline
+\caption{\textbf{Table XX.} The original- and reproduced meta-analytic
+results based on the data provided in the 2014 meta-analysis by Bouri et
+al.}\tabularnewline
 \toprule
@@
 predictor for the DECREASE and non-DECREASE studies to reproduce results
-presented in Figure 4 of the 2014 meta-analysis\textsuperscript{7}.
+presented in Figure 4 of the 2014 meta-analysis\textsuperscript{9}.
 Surprisingly, our results showed stronger evidence against equal
-subgroups than the original meta-analysis\textsuperscript{7} (original:
+subgroups than the original meta-analysis\textsuperscript{9} (original:
 \(\chi^2(1)=3.91,p=.05\); reproduced: \(\chi^2(1)=6.12,p=0.013\)).
@@
 
-\section{Step 2: evaluating the veracity of DECREASE
+\subsection{Step 2: evaluating the veracity of DECREASE
 studies}\label{step-2-evaluating-the-veracity-of-decrease-studies}
@@
 representation of the true effect of beta-blockers on perioperative
-mortality (similar to Bouri et al.\textsuperscript{7}). The estimated
+mortality (similar to Bouri et al.\textsuperscript{9}). The estimated
 probability is also known as the veracity of the
@@
 
-\subsection{Method}\label{method}
+\subsubsection{Method}\label{method}
 
@@
 
-\subsection{Results}\label{results-1}
+\subsubsection{Results}\label{results-1}
 
@@
 
-\section{Step 3: estimating the amount of fabricated
+\subsection{Step 3: estimating the amount of fabricated
 data}\label{step-3-estimating-the-amount-of-fabricated-data}
@@
 provide an indication of the extent of potential data fabrication in the
-DECREASE studies\textsuperscript{2--4,7}.
+DECREASE studies\textsuperscript{4--6,9}.
 
-\subsection{Method}\label{method-1}
+\subsubsection{Method}\label{method-1}
 
@@
 
-\subsection{Results}\label{results-2}
+\subsubsection{Results}\label{results-2}
 
@@
 fabrication that occurred in the DECREASE-I trial, affirming the
-conclusions of the original committee report\textsuperscript{4}.
+conclusions of the original committee report\textsuperscript{6}.
 
@@
 the DECREASE-IV trial are ``scientifically incorrect''
-(p.11)\textsuperscript{3}.
+(p.11)\textsuperscript{5}.
+
+\section{Part 2:}\label{part-2}
 
@@
 the original papers that presented these trial results are not yet
-retracted\textsuperscript{9,10}, something we strongly recommend based
-on the combination of our results and the integrity
-reports\textsuperscript{2--4}.
+retracted\textsuperscript{2,3}, something we strongly recommend based on
+the combination of our results and the integrity
+reports\textsuperscript{4--6}.
 
@@
 ACC/AHA guidelines still cites four trials as evidence
-base\textsuperscript{10,21--23}, of which two were already inspected by
-the scientific integrity committees of Erasmus
-MC\textsuperscript{10,21}. In the ACC/AHA guidelines, the following is
-said about studies conducted by Poldermans:
+base\textsuperscript{2,21--23}, of which two were already inspected by
+the scientific integrity committees of Erasmus MC\textsuperscript{2,21}.
+In the ACC/AHA guidelines, the following is said about studies conducted
+by Poldermans:
 
@@
 evidence that type of beta-blockade can severely influence perioperative
-mortality. Our reanalysis of the Bouri et al.\textsuperscript{7} data
+mortality. Our reanalysis of the Bouri et al.\textsuperscript{9} data
 indicates that type of beta-blockade can reverse the effect on
@@
 
+\hypertarget{ref-poldermans1999}{}
+2. Poldermans D, Boersma E, Bax JJ, et al. The effect of bisoprolol on
+perioperative mortality and myocardial infarction in High-Risk patients
+undergoing vascular surgery. \emph{The New England journal of medicine}.
+1999;341(24):1789-1794.
+doi:\href{https://doi.org/10.1056/NEJM199912093412402}{10.1056/NEJM199912093412402}.
+
+\hypertarget{ref-dunkelgrun2009}{}
+3. Dunkelgrun M, Boersma E, Schouten O, et al. Bisoprolol and
+fluvastatin for the reduction of perioperative cardiac mortality and
+myocardial infarction in intermediate-risk patients undergoing
+noncardiovascular surgery: A randomized controlled trial (DECREASE-IV).
+\emph{Annals of surgery}. 2009;249(6):921-926.
+doi:\href{https://doi.org/10.1097/SLA.0b013e3181a77d00}{10.1097/SLA.0b013e3181a77d00}.
+
 \hypertarget{ref-commissie2011}{}
-2. Onderzoekscommissie Wetenschappelijke Integriteit. \emph{Onderzoek
+4. Onderzoekscommissie Wetenschappelijke Integriteit. \emph{Onderzoek
 Naar Mogelijke Schending van de Wetenschappelijke Integriteit: Beknopte
@@
 \hypertarget{ref-commissie2012}{}
-3. Commissie Vervolgonderzoek 2012. \emph{Rapport Vervolgonderzoek Naar
+5. Commissie Vervolgonderzoek 2012. \emph{Rapport Vervolgonderzoek Naar
 Mogelijke Schending van de Wetenschappelijke Integriteit}. Erasmus MC;
@@
 \hypertarget{ref-commissie2013}{}
-4. Commissie Vervolgonderzoek Wetenschappelijke Integriteit 2013.
+6. Commissie Vervolgonderzoek Wetenschappelijke Integriteit 2013.
 \emph{Rapport}. Erasmus MC; 2014.
@@
 \hypertarget{ref-Devereaux313}{}
-5. Devereaux PJ, Beattie WS, Choi PT-L, et al. How strong is the
+7. Devereaux PJ, Beattie WS, Choi PT-L, et al. How strong is the
 evidence for the use of perioperative beta blockers in non-cardiac
@@
 \hypertarget{ref-Angeli2010}{}
-6. Angeli F, Verdecchia P, Karthikeyan G, Mazzotta G, Gentile G, Reboldi
+8. Angeli F, Verdecchia P, Karthikeyan G, Mazzotta G, Gentile G, Reboldi
 G. \(\beta\)-blockers reduce mortality in patients undergoing high-risk
@@
 \hypertarget{ref-bouri2014}{}
-7. Bouri S, Shun-Shin MJ, Cole GD, Mayet J, Francis DP. Meta-analysis of
+9. Bouri S, Shun-Shin MJ, Cole GD, Mayet J, Francis DP. Meta-analysis of
 secure randomised controlled trials of -blockade to prevent
@@
 \hypertarget{ref-esc2014}{}
-8. ESC/ESA. Guidelines on non-cardiac surgery: Cardiovascular assessment
-and management. \emph{European Heart Journal}. 2014;35:2383-2243.
-
-\hypertarget{ref-dunkelgrun2009}{}
-9. Dunkelgrun M, Boersma E, Schouten O, et al. Bisoprolol and
-fluvastatin for the reduction of perioperative cardiac mortality and
-myocardial infarction in intermediate-risk patients undergoing
-noncardiovascular surgery: A randomized controlled trial (DECREASE-IV).
-\emph{Annals of surgery}. 2009;249(6):921-926.
-doi:\href{https://doi.org/10.1097/SLA.0b013e3181a77d00}{10.1097/SLA.0b013e3181a77d00}.
-
-\hypertarget{ref-poldermans1999}{}
-10. Poldermans D, Boersma E, Bax JJ, et al. The effect of bisoprolol on
-perioperative mortality and myocardial infarction in High-Risk patients
-undergoing vascular surgery. \emph{The New England journal of medicine}.
-1999;341(24):1789-1794.
-doi:\href{https://doi.org/10.1056/NEJM199912093412402}{10.1056/NEJM199912093412402}.
+10. ESC/ESA. Guidelines on non-cardiac surgery: Cardiovascular
+assessment and management. \emph{European Heart Journal}.
+2014;35:2383-2243.
 

### 2017-07-13 submission/manuscript.Rmd
Add .5 to all cells in the original trials, instea of just the zero cells (after advise).
This is also the correct reading of the original reference in the preregistration (https://mathdept.iut.ac.ir/sites/mathdept.iut.ac.ir/files/AGRESTI.PDF; p397)
Revision round
Incorporate revisions Jelte, Markus, and Gerben.

@@
 ---
-title: "Analyzing DECREASE trials for extent of data fabrication [working title]"
-author: "Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper"
+title: "Analyzing DECREASE trials to determine evidence for data manipulation"
+author: "Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper, Markus Hollman"
 date: '`r format(Sys.time(), "%d %B, %Y")`'
@@
 Three meta-analyses that included the trials subject to research misconduct concluded that beta-blockers decrease perioperative mortality [@Devereaux313;@Angeli2010;@bouri2014] whereas a meta-analysis that excluded the suspect trials concluded that beta-blockers increase perioperative mortality [@bouri2014].
-In these studies, perioperative mortality was defined as the deathrate of patients during the period of the surgical procedure, which typically includes admission, anaesthesia, surgery, and recovery.
+In these studies, perioperative mortality was defined as the death rate of patients in the perioperative setting, including the period of admission, anaesthesia with surgery, and postoperative recovery.
 
@@
 The committees that investigated the integrity of the DECREASE trials reported that data fabrication was likely but that the extent of the data fabrication remained unclear [@commissie2011;@commissie2012;@commissie2013].
-Moreover, the latest guidelines still recommend the usage of beta-blockers in the perioperative period in certain cases [@esc2014;@Fleishere278].
-Considering these consequences, we aim to estimate the extent of data fabrication in the DECREASE studies [@poldermans1999;@dunkelgrun2009;@commissie2011;@commissie2012;@commissie2013] to further the debate on using beta-blockers in the perioperative period of CAD patients undergoing non-cardiac surgery.
+Moreover, the latest guidelines still recommend the use of beta-blockers in the perioperative period in certain cases [@esc2014;@Fleishere278], where some of these recommendations are based on other work by the PI from the DECREASE trials.
+Considering the potential harm if guidelines are based on potentially manipulated work, we aim to estimate the extent of data fabrication in the DECREASE studies [@poldermans1999;@dunkelgrun2009;@commissie2011;@commissie2012;@commissie2013] to further the debate on using beta-blockers in the perioperative period of patients undergoing non-cardiac surgery.
 
 The reports on the integrity of the DECREASE trials primarily focused on the provenance of the raw data but did not investigate the extent to which the DECREASE trials deviated from comparable trials.
-The provenance is primarily concerned with the origins of the data, verifying things such as (but not limited to) the informed consent and whether data corresponded to patient files.
-The committee reports did not neglect statistical evaluation however: a statistical expert evaluated the applicability of forensic statistical methods [@commissie2013] to evaluate results of trials separately (i.e., DECREASE-I, DECREASE-IV).
-Nonetheless, comparing across trials is a method that has previously been used to monitor trial data quality or to test for potential data anomalies [@Buyse1999-jq].
-Moreover, this method has previously proven to be effective in detecting data fabrication [@Knepper2016-la].
+Provenance is primarily concerned with the origins of the data, verifying things such as (but not limited to) the informed consent and whether data corresponded to patient files.
+The committee reports did not neglect statistical evaluation however: according to the report a statistical expert evaluated the applicability of forensic statistical methods [@commissie2013] to evaluate results of trials separately (i.e., DECREASE-I, DECREASE-IV) although it lacks details as to how this evaluation took place.
+Nonetheless, in this report we compare across trials, which is a method that has previously been used to monitor trial data quality or to test for potential data anomalies [@Buyse1999-jq].
+Moreover, comparing across trials instead of evaluating them separately  has previously proven to be effective in detecting data fabrication [@Knepper2016-la].
 Comparing the DECREASE trials to other published trials studying the effectiveness of beta-blockers with respect to perioperative mortality could prove informative of the potential extent of the fabrication in the DECREASE trials.
@@
 The effectiveness of perioperative beta-blockade is obfuscated by the afflicted DECREASE trials, potentially interacting with the type of beta-blocker and the way that beta-blocker was administered (i.e., dose and duration of treatment).
-Throughout the clinical trials regarding beta-blockers, patients were administered various beta-blockers (e.g., metroprolol, bisoprolol, atenolol) and in various ways (e.g., intraveneously, orally; half an hour before surgery or multiple days before surgery).
+In the randomized trials on beta-blockers, patients were administered various types of beta-blockers (e.g., metroprolol, bisoprolol, atenolol) and in various ways (e.g., intravenously, orally; half an hour before surgery or multiple days before surgery; with or without titration based on heart rate).
 Factors such as dosage and duration can have an effect on the pharmacological effectiveness with respect to perioperative mortality.
@@
 
-The latest guidelines of the European Society of Cardiology and the American Heart Association still recommend the usage of beta-blockers in certain cases [@;@].
-Therefore, we evaluate the usage of beta-blockers with respect to its effectiveness for the the perioperative period in patients undergoing non-cardiac surgery.
-First, we estimated the extent of data fabrication in the DECREASE studies [@poldermans1999;@dunkelgrun2009] Secondly, we evaluated the clinical studies used in the different meta-analysis from a pharmacological point of view.
-
-# Part 1: statistical estimation of data fabrication in DECREASE trials
-
-To statistically investigate the evidence of data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999], we took three steps.
+To statistically investigate the evidence for data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999], we took three steps.
 First, we reproduced the findings from the 2014  meta-analysis by Bouri et al.
 [@bouri2014] that contained sufficient information to estimate the deviation of the DECREASE trials from other published trials on beta-blockers.
-We also include type of beta-blocker to inspect whether this is predictive of the effect of beta-blockers on perioperative mortality.
+We also included type of beta-blocker to inspect whether this is predictive of the effect of beta-blockers on perioperative mortality.
 Second, we evaluated the probability that the DECREASE trials occurred assuming no data fabrication.
@@
 bouri_data$beta_blocker <- c("atenolol",
-                             "propanolol",
+                             "propranolol",
                              "metoprolol",
@@
 
-# Add .5 to studies with zero-counts
-bouri_data[c(4, 7, 9), 7:10] <-  bouri_data[c(4, 7, 9), 7:10] + .5
+# Add .5 to all studies to correct for zero-counts
+bouri_data[, 7:10] <-  bouri_data[, 7:10] + .5
 
@@
 We used the default weighting procedure in the `metafor` package.
-When there was a zero-count for a cell (e.g., zero events), 0.5 was added to that cell, as is common in meta-analyses on risk- and odds ratios in order to prevent computational artefacts [@agresti2002].
+We added 0.5 to each cell, as is common in meta-analyses on risk- and odds ratios in order to prevent computational artefacts [@agresti2002; we made an error in the preregistration stating that only .5 was added to zero-cells].
 The 2014 meta-analysis [@bouri2014] did not specify the variance estimate used; hence, minor discrepancies between our estimates and the original estimates could be due to differences in the estimation procedure.
@@
 differentiated between the estimates from the non-DECREASE trials and the DECREASE trials.
-We confirmed the effect size estimates and the variance estimates for both the non-DECREASE- and the DECREASE trials, save for some minor discrepancies due to the estimation method.
+We confirmed the effect size estimates and the variance estimates for both the non-DECREASE- and the DECREASE trials, except for some minor discrepancies at the second decimal level.
 Table 1 depicts the original and reproduced values for both sets of studies.
@@
 table1[1,5] <- 'Confidence interval'
-table1[2,1] <- 'Non-DECREASE'
-table1[4,1] <- 'DECREASE'
+table1[2,1] <- 'Non-DECREASE (k=9)'
+table1[4,1] <- 'DECREASE (k=2)'
 table1[c(2,4),2] <- 'Original'
@@
 Different variance estimates (e.g., `DerSimonian-Laird` instead of `REML`) did not resolve this difference.
-We tried to clarify these discrepancies by e-mailing the original authors, but did not receive a response.
+We tried to clarify these discrepancies by e-mailing the original authors, with a reminder after two weeks, but did not receive a response.
 Nonetheless, the broad strokes of the meta-regression confirmed that the DECREASE trials were the determining predictor for the effectiveness of beta-blockers (including DECREASE: $RR=`r round(exp(mod_mods$b[1] + mod_mods$b[2]), 3)`$; excluding DECREASE: $RR=`r round(exp(mod_mods$b[1]), 3)`$).
@@
 Additionally, and exploratively, we evaluated the predictive effect of the type of beta-blocker used in the trials.
-The DECREASE trials remain predictive of decreased mortality ($RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[2] + mod_mods_type$b[3]), 3)`$), <!-- DECREASE only uses bisoprolol!
+The DECREASE trials remained predictive of decreased mortality ($RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[2] + mod_mods_type$b[3]), 3)`$), <!-- DECREASE only uses bisoprolol!
 -->
 whereas the non-DECREASE trials provide tentative, but uncertain, evidence that atenolol results in lower mortality ($RR=`r round(exp(mod_mods_type$b[1]), 3)`$).
-Nonetheless, for other beta-blockers in the non-DECREASE trials, there is still tentative and uncertain evidence that beta-blockers increase mortality (bisoprolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[3]), 3)`$; metoprolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[4]), 3)`$; propanolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[5]), 3)`$).
+Nonetheless, for other beta-blockers in the non-DECREASE trials, there is still tentative and uncertain evidence that beta-blockers increase mortality (bisoprolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[3]), 3)`$; metoprolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[4]), 3)`$; propranolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[5]), 3)`$).
 Table 3 shows the meta-regression results in full.
@@
 table3[7,1] <- 'Metoprolol'
-table3[8,1] <- 'Propanolol'
+table3[8,1] <- 'Propranolol'
 
@@
 
-knitr::kable(table3, caption = 'Meta-regression results for log(RR), including dummy predictors for DECREASE trials (reference: non-DECREASE trials) and type of beta-blockers used in the trial (reference: atenolol).')
+knitr::kable(table3, caption = 'Meta-regression results (k=11) for log(RR), including dummy predictors for DECREASE trials (reference: non-DECREASE trials) and type of beta-blockers used in the trial (reference: atenolol).')
 ```
@@
 [@bouri2014]).
-The estimated probability is also known as the veracity of the data [@peters2015], which indicates the probability of the observed data under a given true effect.
+The estimated probability of the observed data under a given true effect is also known as the veracity of the data [@peters2015].
 We assumed that the non-DECREASE studies estimated the true effect distribution of perioperative beta-blockade on mortality, not perturbed by publication bias due to statistical (non)significance.
@@
 Based on the estimated effect distribution from the non-DECREASE trials, we calculated the probability of each DECREASE trial result, or a more extreme result.
-In other words, we computed the *p*-value for the null hypothesis that the DECREASE trials arise from the same effect distribution as the non-DECREASE trials.
+In other words, we computed the $p$-value for the null hypothesis that the DECREASE trials arise from the same effect distribution as the non-DECREASE trials.
 This assumes that the information available from the other trials is informative of the true population effect.
@@
 More specifically, the results from DECREASE-I (or more extreme) have a probability of $`r prob[1]`$ (less than 1 in a quintillion) and the results from DECREASE-IV have a probability of $`r prob[2]`$ (3 in a billion).
-This indicates the DECREASE trial results are unlikely to have come from the same population effect distribution as the non-DECREASE trials.
+This indicates that the DECREASE trial results are unlikely to have come from the same population effect distribution as the non-DECREASE trials.
 Moreover, observing two of such extremely unlikely results jointly, as in the DECREASE trials, is nearly impossible, $`r prob[1]*prob[2]`$.
@@
 
-Results from Step 1 indicated that no between-trial variance (i.e., homogeneity; $\tau^2=0$) of the effects was observed; given the small number of trials included (i.e., `r mod_clean$k`) this estimate is uncertain, however.
-We conducted sensitivity analyses to see how dependent results are on the heterogeneity estimate.
+Results from Step 1 indicated that no between-trial variance (i.e., homogeneity; $\tau^2=0$) of the effects was observed; given the small number of trials included (i.e., `r mod_clean$k`) this estimate is highly uncertain, however.
+The total $N$ across the non-DECREASE trials was `r bouri_data$control_n[bouri_data$decrease == 0] + bouri_data$beta_n[bouri_data$decrease == 0]`.We conducted sensitivity analyses to see how dependent results are on the heterogeneity estimate (not preregistered; [osf.io/vnmzc](https://osf.io/vnmzc)).
 The probability of observing the DECREASE trials stays approximately below 1 out of 1000 until the variance estimate is 0.25; the probability stays approximately below 1 out of 100 until the variance estimate is 0.43 (see Figure `r fig <- fig + 1; fig`).
-To put these numbers into context, a variance of 0.25 would suggest that results of perioperative beta-blockade vary substantially even if perioperative beta-blockade has no effect whatsoever (RRs between `r round(exp(0 - .25), 3)` and `r round(exp(0 + .25), 3)` in 64\% of the cases) all due to contextual circumstances of the study.
+To put these numbers into context, a variance of 0.25 would suggest that results of perioperative beta-blockade vary substantially due to contextual circumstances of the study, even if perioperative beta-blockade has no effect whatsoever (RRs between `r round(exp(0 - .25), 3)` and `r round(exp(0 + .25), 3)` in ~64\% of the cases).
 
@@
 
-In order to estimate the number of fabricated data points, we first estimated the effect of perioperative beta-blockade on mortality (in log odds) in each trial arm.
-In total there are four trial arms: one per condition (beta-blocker or control) per trial type (DECREASE- and non-DECREASE trials).
-For each of these trial arms, we ran a meta-analysis applying the same methods used in Step 1.
+In order to estimate the number of fabricated data points, we first estimated the probability of perioperative mortality (in log odds) in each trial arm separated per trial stratum.
+As such, there are four estimates of perioperative mortality: one per condition (beta-blocker or control) per trial stratum (DECREASE- or non-DECREASE trials).
+For each of these, we ran a meta-analysis applying the same methods used in Step 1.
 For each of the two DECREASE trials separately and the non-DECREASE trials combined, this resulted in four meta-analytic mortality estimates with corresponding effect variances (see Table 3; one estimate per cell).
@@
 For each hypothesis of *X* out of *N* fabricated data points, we computed the probability that the fabricated data are more likely than the genuine data ($p_F=P(L_{fabricated}>L_{genuine})$).
-Based on $p_F$, we computed the confidence interval for $X$ (i.e., $X_{LB};X_{UB}$).
+Based on $p_F$, we computed the confidence interval for the estimated $X$ fabricated data points (i.e., $X_{LB};X_{UB}$).
 For a 95\% confidence interval, the lowerbound is equal to the $p_F$ closest to .025, whereas the upperbound is equal to the $p_F$ closest to .975.
 
-We computed $p_F$ for all *X* out of *N* fabricated datapoints in `r iterations` randomly generated datasets, which were generated in three steps. For each dataset we:
+We computed $p_F$ for all *X* out of *N* fabricated data points in `r iterations` randomly generated datasets, which were generated in three steps. For each dataset we:
 
@@
 Based on the meta-analytic effect from 2 and the cell sizes from 3, we computed the likelihoods $L_{fabricated}$ and $L_{genuine}$ using Equation \ref{eq1}.
-As mentioned before, we computed $p_F$, which indicated the probability that the data are more likely under the estimates resulting from the (allegedly) fabricated data (i.e., the DECREASE trials) than under the estimates resulting from the genuine data (i.e., the non-DECREASE trials; $p_F=P(L_{fabricated}>L_{genuine})$).
+As mentioned before, we computed $p_F$, which indicates the probability that the data are more likely under the estimates resulting from the (allegedly) fabricated data (i.e., the DECREASE trials) than under the estimates resulting from the genuine data (i.e., the non-DECREASE trials; $p_F=P(L_{fabricated}>L_{genuine})$).
 
@@
 This indicates that it is possible not all data were fabricated (i.e., $N=`r length(res_iv_certain) - 1`$), increasing the importance of well-documented data provenance to discern between genuine and falsified data.
-This affirms the conclusion of the scientific integrity committee that the results of the DECREASE-IV trial are "scientifically incorrect" (p.11) [@commissie2012].
-
-# Part 2:
 
@@
 
-The effect of beta-blockade on perioperative mortality was already unclear following the findings of scientific misconduct; these results strongly affirm that the DECREASE trials should be neglected when assessing the effectiveness of beta-blockers.
-Our results indicate that the results from the DECREASE trials are nearly impossible to have arisen from the same effect inspected by the non-DECREASE trials, except when we assume at least some of the data were manipulated.
-As such, the scientific validity of the DECREASE-I and DECREASE-IV trials should by now be clear: they are to be regarded as irrelevant when assessing the effectiveness of beta-blockade on perioperative mortality.
-Nonetheless, the original papers that presented these trial results are not yet retracted [@poldermans1999;@dunkelgrun2009], something we strongly recommend based on the combination of our results and the integrity reports [@commissie2011;@commissie2012;@commissie2013].
+The effect of beta-blockade on perioperative mortality was already unclear based on the findings of research misconduct in institutional investigations [@commissie2011;@commissie2012;@commissie2013]; our results affirm that the empirical results from the DECREASE trials are potentially problematic in light of the other studies supposedly investigating the same effect.
+We indicate that the results from the DECREASE trials are nearly impossible to have arisen from the same effect inspected by the non-DECREASE trials, except when we assume at least some of the data were manipulated and that these trials investigate the same effect.
+As such, we consider the scientific validity of the DECREASE-I and DECREASE-IV trials as problematic when assessing the effectiveness of beta-blockade on perioperative mortality.
+Nonetheless, the original papers that presented these trial results are not yet retracted [@poldermans1999;@dunkelgrun2009], despite the damning integrity reports [@commissie2011;@commissie2012;@commissie2013].
+
+Nonetheless, we note that statistically estimated number of manipulated data points in the DECREASE trials is subject to multiplicity.
+In other words, for any estimated proportion of data points manipulated, there is another estimate that contains fewer (or more) manipulated data points that are more (or less) extremely manipulated and yields exactly the same outcome.
+This is similar to how the mean of two samples can be equivalent (e.g., 0), but that the individual scores giving rise to that mean are extremely different (e.g, +2.5 and -2.5 in the first sample; -100 and +100 in the second sample).
+Regardless, whether any manipulation occurred does not suffer from this multiplicity because there is only one way to have no manipulated data points.
 
 The ESC/ESA and ACC/AHA guidelines [@Kristensen_2014;@Fleisher_2014] on perioperative beta-blockade already excluded the DECREASE trials in their assessment, but also state that other trials by Poldermans are excluded.
-However, upon close inspection of the reference lists, the ACC/AHA guidelines still cites four trials as evidence base [@Boersma_2001;@poldermans1999;@van_Kuijk_2009;@Flu_2010], of which two were already inspected by the scientific integrity committees of Erasmus MC [@Boersma_2001;@poldermans1999].
+However, upon close inspection of the reference lists, the ACC/AHA guidelines still cites four trials as evidence [@Boersma_2001;@poldermans1999;@van_Kuijk_2009;@Flu_2010], of which two were already inspected by the scientific integrity committees of Erasmus MC [@Boersma_2001;@poldermans1999].
 In the ACC/AHA guidelines, the following is said about studies conducted by Poldermans:
@@
 
-Nonetheless, references _are_ made without clear comments.
+Nonetheless, references are made without clear comments.
 Given the confirmation of problems in the DECREASE-I and DECREASE-IV trials in our results, it stresses that there is reason to distrust trials by Poldermans.
-We pose that investigations should be initiated into works where Poldermans was clearly involved and which were not cleared by the scientific committees of Erasmus MC in their misconduct investigations.
-Especially those papers cited as evidence in the ACC/AHA guidelines should be investigated, considering they affect patients and their treatment directly.
+For the integrity of the guidelines and the safety of the patients, we pose that investigations should be initiated into works where Poldermans was clearly involved and which were not cleared by the scientific committees of Erasmus MC in their misconduct investigations.
+Especially those papers cited as evidence in the ACC/AHA guidelines should be investigated, considering that they affect patients and their treatment directly.
 
-Previously, further investigation of trials by Poldermans was deemed unfeasible due to the lack of raw data; here we indicate methods that do make it feasible.
-Based on just event-count data and comparable trials, we were able to estimate whether part of the data were in fact fabricated and whether the results were within reason of comparable trials.
-The results clearly indicated they were not.
+Previously, further investigation of trials by Poldermans was deemed unfeasible due to the lack of raw data; here we indicate methods that do make it feasible, albeit with their own limitations.
+Based on just event-count data and comparable trials, we were able to estimate whether part of the data were in fact fabricated and whether the results were within reason of other trials supposedly investigating the same effect.
+The results clearly indicated the DECREASE trials are not within reason, if we assume the non-DECREASE trials to investigate the same effect.
 
@@
 This increased certainty is due to decreased standard errors at the population level of the estimated effects, resulting in higher sensitivity to data anomalies.
-Nonetheless, much residual uncertainty remains and simply fewer  information is available in the summary results.
-Raw data availability would improve the options open to detect potential anomalies (note: raw data are available for DECREASE VI, but upon a freedom of information request Erasmus MC refused to share these data).
-The results also highlight that in order to prevent detection, it would be in the fabricators' interest to fabricate small studies even if raw data are hidden away (assuming the fabricator wants to remain undetected).
+Nonetheless, alongside multiplicity much residual uncertainty remains and simply less information is available in the summary results.
+Raw data availability would improve the options open to detect potential anomalies (note: raw data are available for DECREASE VI, but upon a freedom of information request by the first author, Erasmus MC refused to share these data; letters available at [osf.io/fswjt](https://osf.io/fswjt/)).
+The results also highlight that in order to prevent detection, it would be in the fabricators' interest to fabricate small and imprecise studies (assuming the fabricator wants to remain undetected), which ultimately detracts from the scientific value of such a study and hopefully the subsequent reward for fabrication due to less impact.
 
@@
 [@bouri2014] data indicates that type of beta-blockade can reverse the effect on perioperative mortality, even after taking into account whether a study belongs to the DECREASE family.
-As such, atenolol seems to tentatively decrease perioperative mortality, whereas the others (metoprolol, propanolol, bisoprolol) increase perioperative mortality.
+As such, atenolol seems to tentatively decrease perioperative mortality, whereas the others (metoprolol, propranolol, bisoprolol) increase perioperative mortality.
 However, there seems to be covariation with respect to treatment administration, duration, and dose, which further confounds whether the treatment effect is due to type of beta-blocker or due to one of these other parameters.
+Given the limited statistical power in this meta-analysis due to the lack of studies to investigate which factors are most essential, we recommend a clinical trial that aims to dissect some of these potential confounds.
 This affirms the statement from the ESC/ESA guidelines that _"high priority needs to be given to new randomized clinical trials to better identify which patients derive benefit from beta-blocker therapy in the perioperative setting, and to determine the optimal method of beta-blockade"_ [@Kristensen_2014].
 
-In sum, our research indicates that the DECREASE trials are nearly impossible and contain at least some manipulated data points.
-We recommend retraction of the DECREASE-I and DECREASE-IV trials, and recommend renewed investigations into Poldermans' work --- especially that work still referenced by guidelines on the use of beta-blockers.
-Moreover, it remains unclear whether beta-blockers might be effective given the right treatment.
-We recommend new and more extensively controlled, confirmatory trials to determine whether there is any use in administering beta-blockers in order to decrease perioperative mortality --- at the moment there is insufficient evidence to determine any useful effect of beta-blockers.
+Moreover, the DECREASE and non-DECREASE trials seem to apply beta-blockade from different conceptual viewpoints.
+The non-DECREASE trials seem to focus purely on the application of beta-blockers in itself, whereas the DECREASE trials use beta-blockade as a proxy to decrease resting heart rate [@poldermans1999;@dunkelgrun2009].
+As such, the DECREASE studies applied beta-blockade at least a week in advance, specifically in order to lower patient's resting heart rate to <70BPM and potentially habituate the patient to the effects of the beta-blockade.
+Other studies apply the beta-blockade just prior to the surgery (maximum: one day prior) and therefore seem to regard the treatment specifically and not the proxy of lowered BPM.
+As such, the differences between the DECREASE and non-DECREASE trials might also in part be a consequence of the different approaches in the various trials.
+Whether these differences matter in treatment decisions is worthy of further research in a clinical trial with high statistical power to find such differences.
+
+In sum, our research indicates that the DECREASE trials are nearly impossible if we assume they investigate exactly the same effect as the non-DECREASE trials and under that assumption, our results provide some evidence that at least some data points were manipulated.
+However, these differences might also be due to different conceptual approaches as to how beta-blockade might prevent mortality in non-cardiac surgery.
+We recommend renewed investigations into Poldermans' work given these findings --- especially those works still referenced by guidelines on the use of beta-blockers without proper notice.
+Moreover, it remains unclear whether beta-blockers might be effective in preventing mortality rates in non-cardiac surgery patients.
+Considering this, we recommend new and more extensively controlled, confirmatory trials to determine whether there is any use in administering beta-blockers in order to decrease perioperative mortality --- at the moment there is insufficient evidence to determine any positive effect of beta-blockers on mortality rates.
 

### 2017-07-13 submission/manuscript.tex
Revision round
Incorporate revisions Jelte, Markus, and Gerben.

@@
 \hypersetup{unicode=true,
-            pdftitle={Analyzing DECREASE trials for extent of data fabrication {[}working title{]}},
+            pdftitle={Analyzing DECREASE trials to determine evidence for data fabrication},
             pdfauthor={Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper},
@@
 \setlength{\droptitle}{-2em}
-  \title{Analyzing DECREASE trials for extent of data fabrication {[}working
-title{]}}
+  \title{Analyzing DECREASE trials to determine evidence for data fabrication}
   \pretitle{\vspace{\droptitle}\centering\huge}
@@
   \postdate{\par}
-  \date{24 May, 2017}
+  \date{06 June, 2017}
 
@@
 the provenance of the raw data but did not investigate the extent to
-which the DECREASE trials deviated from comparable trials. The
-provenance is primarily concerned with the origins of the data,
-verifying things such as (but not limited to) the informed consent and
-whether data corresponded to patient files. The committee reports did
-not neglect statistical evaluation however: a statistical expert
-evaluated the applicability of forensic statistical
-methods\textsuperscript{6} to evaluate results of trials separately
-(i.e., DECREASE-I, DECREASE-IV). Nonetheless, comparing across trials is
-a method that has previously been used to monitor trial data quality or
-to test for potential data anomalies\textsuperscript{11}. Moreover, this
-method has previously proven to be effective in detecting data
+which the DECREASE trials deviated from comparable trials. Provenance is
+primarily concerned with the origins of the data, verifying things such
+as (but not limited to) the informed consent and whether data
+corresponded to patient files. The committee reports did not neglect
+statistical evaluation however: a statistical expert evaluated the
+applicability of forensic statistical methods\textsuperscript{6} to
+evaluate results of trials separately (i.e., DECREASE-I, DECREASE-IV).
+Nonetheless, comparing across trials is a method that has previously
+been used to monitor trial data quality or to test for potential data
+anomalies\textsuperscript{11}. Moreover, this method has previously
+proven to be effective in detecting data
 fabrication\textsuperscript{12}. Comparing the DECREASE trials to other
@@
 
-The latest guidelines of the European Society of Cardiology and the
-American Heart Association still recommend the usage of beta-blockers in
-certain cases {[}@;@{]}. Therefore, we evaluate the usage of
-beta-blockers with respect to its effectiveness for the the
-perioperative period in patients undergoing non-cardiac surgery. First,
-we estimated the extent of data fabrication in the DECREASE
-studies\textsuperscript{2,3} Secondly, we evaluated the clinical studies
-used in the different meta-analysis from a pharmacological point of
-view.
-
-\section{Part 1: statistical estimation of data fabrication in DECREASE
-trials}\label{part-1-statistical-estimation-of-data-fabrication-in-decrease-trials}
-
-To statistically investigate the evidence of data fabrication in the
+To statistically investigate the evidence for data fabrication in the
 DECREASE studies\textsuperscript{2,3}, we took three steps. First, we
@@
 \href{https://osf.io/vnmzc}{osf.io/vnmzc}.
-
-\begin{verbatim}
-## Warning in library(package, lib.loc = lib.loc, character.only = TRUE,
-## logical.return = TRUE, : there is no package called 'metafor'
-\end{verbatim}
-
-\begin{verbatim}
-## Installing package into '/home/chjh/R/x86_64-redhat-linux-gnu-library/3.3'
-## (as 'lib' is unspecified)
-\end{verbatim}
 
@@
 mortality (similar to Bouri et al.\textsuperscript{9}). The estimated
-probability is also known as the veracity of the
-data\textsuperscript{17}, which indicates the probability of the
-observed data under a given true effect. We assumed that the
+probability of the observed data under a given true effect is also known
+as the veracity of the data\textsuperscript{17}. We assumed that the
 non-DECREASE studies estimated the true effect distribution of
@@
 we calculated the probability of each DECREASE trial result, or a more
-extreme result. In other words, we computed the \emph{p}-value for the
-null hypothesis that the DECREASE trials arise from the same effect
+extreme result. In other words, we computed the \(p\)-value for the null
+hypothesis that the DECREASE trials arise from the same effect
 distribution as the non-DECREASE trials. This assumes that the
@@
 We conducted sensitivity analyses to see how dependent results are on
-the heterogeneity estimate. The probability of observing the DECREASE
-trials stays approximately below 1 out of 1000 until the variance
-estimate is 0.25; the probability stays approximately below 1 out of 100
-until the variance estimate is 0.43 (see Figure 2). To put these numbers
-into context, a variance of 0.25 would suggest that results of
-perioperative beta-blockade vary substantially even if perioperative
+the heterogeneity estimate (not preregistered). The probability of
+observing the DECREASE trials stays approximately below 1 out of 1000
+until the variance estimate is 0.25; the probability stays approximately
+below 1 out of 100 until the variance estimate is 0.43 (see Figure 2).
+To put these numbers into context, a variance of 0.25 would suggest that
+results of perioperative beta-blockade vary substantially due to
+contextual circumstances of the study, even if perioperative
 beta-blockade has no effect whatsoever (RRs between 0.779 and 1.284 in
-64\% of the cases) all due to contextual circumstances of the study.
+\textasciitilde{}64\% of the cases).
 
@@
 than the genuine data (\(p_F=P(L_{fabricated}>L_{genuine})\)). Based on
-\(p_F\), we computed the confidence interval for \(X\) (i.e.,
-\(X_{LB};X_{UB}\)). For a 95\% confidence interval, the lowerbound is
-equal to the \(p_F\) closest to .025, whereas the upperbound is equal to
-the \(p_F\) closest to .975.
+\(p_F\), we computed the confidence interval for the estimated \(X\)
+fabricated datapoints (i.e., \(X_{LB};X_{UB}\)). For a 95\% confidence
+interval, the lowerbound is equal to the \(p_F\) closest to .025,
+whereas the upperbound is equal to the \(p_F\) closest to .975.
 
@@
 Equation \ref{eq1}. As mentioned before, we computed \(p_F\), which
-indicated the probability that the data are more likely under the
+indicates the probability that the data are more likely under the
 estimates resulting from the (allegedly) fabricated data (i.e., the
@@
 
-\section{Part 2:}\label{part-2}
-
 \section{Discussion}\label{discussion}
@@
 unclear following the findings of scientific misconduct; these results
-strongly affirm that the DECREASE trials should be neglected when
-assessing the effectiveness of beta-blockers. Our results indicate that
-the results from the DECREASE trials are nearly impossible to have
-arisen from the same effect inspected by the non-DECREASE trials, except
-when we assume at least some of the data were manipulated. As such, the
-scientific validity of the DECREASE-I and DECREASE-IV trials should by
-now be clear: they are to be regarded as irrelevant when assessing the
-effectiveness of beta-blockade on perioperative mortality. Nonetheless,
-the original papers that presented these trial results are not yet
-retracted\textsuperscript{2,3}, something we strongly recommend based on
-the combination of our results and the integrity
-reports\textsuperscript{4--6}.
+strongly affirm that the empirical results from the DECREASE trials
+should be neglected when assessing the effectiveness of beta-blockers.
+Our results indicate that the results from the DECREASE trials are
+nearly impossible to have arisen from the same effect inspected by the
+non-DECREASE trials, except when we assume at least some of the data
+were manipulated. As such, the scientific validity of the DECREASE-I and
+DECREASE-IV trials should by now be clear: they are to be regarded as
+irrelevant when assessing the effectiveness of beta-blockade on
+perioperative mortality. Nonetheless, the original papers that presented
+these trial results are not yet retracted\textsuperscript{2,3}, despite
+the integrity reports\textsuperscript{4--6}.
 
@@
 
-Nonetheless, references \emph{are} made without clear comments. Given
-the confirmation of problems in the DECREASE-I and DECREASE-IV trials in
-our results, it stresses that there is reason to distrust trials by
-Poldermans. We pose that investigations should be initiated into works
+Nonetheless, references are made without clear comments. Given the
+confirmation of problems in the DECREASE-I and DECREASE-IV trials in our
+results, it stresses that there is reason to distrust trials by
+Poldermans. For the integrity of the guidelines and the safety of the
+patients, we pose that investigations should be initiated into works
 where Poldermans was clearly involved and which were not cleared by the
@@
 the options open to detect potential anomalies (note: raw data are
-available for DECREASE VI, but upon a freedom of information request
-Erasmus MC refused to share these data). The results also highlight that
-in order to prevent detection, it would be in the fabricators' interest
-to fabricate small studies even if raw data are hidden away (assuming
-the fabricator wants to remain undetected).
+available for DECREASE VI, but upon a freedom of information request by
+the first author, Erasmus MC refused to share these data). The results
+also highlight that in order to prevent detection, it would be in the
+fabricators' interest to fabricate small and unreliable studies even if
+raw data are hidden away (assuming the fabricator wants to remain
+undetected), which ultimately detracts from the value of such a study
+and hence the reward for fabrication.
 
@@
 
+Moreover, the DECREASE and non-DECREASE trials seem to apply
+beta-blockade from different conceptual viewpoints. The non-DECREASE
+trials seem to focus purely on the application of beta-blockers in
+itself, whereas the DECREASE trials use beta-blockade as a proxy to
+decrease resting heart rate\textsuperscript{2,3}. As such, the DECREASE
+studies applied beta-blockade at least a week in advance, specifically
+in order to lower patient's resting heart rate to \textless{}70BPM and
+potentially habituate the patient to the effects of the beta-blockade.
+Other studies apply the beta-blockade just prior to the surgery
+(maximum: one day prior) and therefore seem to regard the treatment
+specifically and not the proxy of lowered BPM. As such, the differences
+between the DECREASE and non-DECREASE trials might also in part be a
+consequence of the different approaches in the various trials. Whether
+these differences matter in treatment decisions is worthy of further
+research in a clinical trial with high statistical power to find such
+differences.
+
 In sum, our research indicates that the DECREASE trials are nearly
-impossible and contain at least some manipulated data points. We
-recommend retraction of the DECREASE-I and DECREASE-IV trials, and
-recommend renewed investigations into Poldermans' work --- especially
-that work still referenced by guidelines on the use of beta-blockers.
-Moreover, it remains unclear whether beta-blockers might be effective
-given the right treatment. We recommend new and more extensively
-controlled, confirmatory trials to determine whether there is any use in
-administering beta-blockers in order to decrease perioperative mortality
---- at the moment there is insufficient evidence to determine any useful
-effect of beta-blockers.
+impossible if we assume they investigate exactly the same effect as the
+non-DECREASE trials and under that assumption, our results provide some
+evidence that at least some data points were manipulated. However, these
+differences might also be due to different conceptual approaches to how
+beta-blockade might prevent mortality in non-cardiac surgery. We
+recommend renewed investigations into Poldermans' work given these
+findings --- especially those works still referenced by guidelines on
+the use of beta-blockers without proper notice. Moreover, it remains
+unclear whether beta-blockers might be effective in preventing mortality
+rates in non-cardiac surgery patients. Considering this, we recommend
+new and more extensively controlled, confirmatory trials to determine
+whether there is any use in administering beta-blockers in order to
+decrease perioperative mortality --- at the moment there is insufficient
+evidence to determine any positive effect of beta-blockers on mortality
+rates.
 

### 2017-07-14 submission/manuscript.Rmd
Revision round 1200

@@
 ---
-title: "Analyzing DECREASE trials to determine evidence for data manipulation"
+title: "Analyzing DECREASE trials to estimate evidence of data manipulation"
 author: "Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper, Markus Hollman"
@@
 
-The effect of beta-blockers on perioperative mortality in non-cardiac surgery has been controversial [@Coleg5210] due to findings of research  misconduct in two related clinical trials  [@poldermans1999;@dunkelgrun2009;@commissie2011;@commissie2012;@commissie2013].
-Three meta-analyses that included the trials subject to research misconduct concluded that beta-blockers decrease perioperative mortality [@Devereaux313;@Angeli2010;@bouri2014] whereas a meta-analysis that excluded the suspect trials concluded that beta-blockers increase perioperative mortality [@bouri2014].
+```{r, echo = FALSE}
+tabnr <- 1
+fignr <- 1
+```
+
+The effect of beta-blockers on perioperative mortality in non-cardiac surgery has been controversial [@Coleg5210] due to concerns regarding the  scientific integrity in two related clinical trials [@poldermans1999;@dunkelgrun2009;@commissie2011;@commissie2012;@commissie2013].
+Three meta-analyses that included the trials subject to concerns concluded that beta-blockers decrease perioperative mortality [@Devereaux313;@Angeli2010;@bouri2014] whereas a meta-analysis that excluded the suspect trials concluded that beta-blockers increase perioperative mortality [@bouri2014].
 In these studies, perioperative mortality was defined as the death rate of patients in the perioperative setting, including the period of admission, anaesthesia with surgery, and postoperative recovery.
 
-The trials subject to research misconduct were the Dutch DECREASE-I and DECREASE-IV trials [@commissie2011;@commissie2012;@commissie2013].
-The committees that investigated the integrity of the DECREASE trials reported that data fabrication was likely but that the extent of the data fabrication remained unclear [@commissie2011;@commissie2012;@commissie2013].
-Moreover, the latest guidelines still recommend the use of beta-blockers in the perioperative period in certain cases [@esc2014;@Fleishere278], where some of these recommendations are based on other work by the PI from the DECREASE trials.
-Considering the potential harm if guidelines are based on potentially manipulated work, we aim to estimate the extent of data fabrication in the DECREASE studies [@poldermans1999;@dunkelgrun2009;@commissie2011;@commissie2012;@commissie2013] to further the debate on using beta-blockers in the perioperative period of patients undergoing non-cardiac surgery.
+The trials subject to concerns regarding scientific integrity were the Dutch DECREASE-I and DECREASE-IV trials [@commissie2011;@commissie2012;@commissie2013].
+The committees that investigated the integrity of the DECREASE trials reported that data manipulation was likely but that the extent of the data manipulation remained unclear [@commissie2011;@commissie2012;@commissie2013].
+Moreover, the latest guidelines still recommend the usage of beta-blockers in the perioperative period in certain cases [@esc2014;@Fleisher_2014], where some of these guidelines are based on other work by the PI of the DECREASE trials.
+Considering the potential harmful consequences of guidelines based on work by someone who has been repeatedly investigated for breaching scientific integrity, we aim to estimate the extent of data manipulation in the DECREASE studies [@poldermans1999;@dunkelgrun2009;@commissie2011;@commissie2012;@commissie2013] to further stimulate the debate on using beta-blockers in the perioperative period patients undergoing non-cardiac surgery.
 
@@
 The committee reports did not neglect statistical evaluation however: according to the report a statistical expert evaluated the applicability of forensic statistical methods [@commissie2013] to evaluate results of trials separately (i.e., DECREASE-I, DECREASE-IV) although it lacks details as to how this evaluation took place.
-Nonetheless, in this report we compare across trials, which is a method that has previously been used to monitor trial data quality or to test for potential data anomalies [@Buyse1999-jq].
-Moreover, comparing across trials instead of evaluating them separately  has previously proven to be effective in detecting data fabrication [@Knepper2016-la].
-Comparing the DECREASE trials to other published trials studying the effectiveness of beta-blockers with respect to perioperative mortality could prove informative of the potential extent of the fabrication in the DECREASE trials.
+The expert concluded that the previously applied methods by him were not applicable for use in this case.
+Nonetheless, in this report we compare across trials, which is a method that has previously been used to monitor trial data quality or to test for potential data anomalies [@Buyse1999-jq;@10.1111/anae.13938].
+Moreover, comparing across trials instead of evaluating them separately  has previously proven to be effective in detecting data manipulation [@Knepper2016-la].
+Comparing the DECREASE trials to other published trials studying the effectiveness of beta-blockers with respect to perioperative mortality could prove informative of the potential extent of the manipulation in the DECREASE trials.
 
-The effectiveness of perioperative beta-blockade is obfuscated by the afflicted DECREASE trials, potentially interacting with the type of beta-blocker and the way that beta-blocker was administered (i.e., dose and duration of treatment).
+The effectiveness of perioperative beta-blockade is obfuscated by the afflicted DECREASE trials, potentially interacting with the type of beta-blocker and the way that beta-blockers were administered (i.e., dose and duration of treatment).
 In the randomized trials on beta-blockers, patients were administered various types of beta-blockers (e.g., metroprolol, bisoprolol, atenolol) and in various ways (e.g., intravenously, orally; half an hour before surgery or multiple days before surgery; with or without titration based on heart rate).
 Factors such as dosage and duration can have an effect on the pharmacological effectiveness with respect to perioperative mortality.
-Moreover, the anomalous results from the DECREASE trials might partly be caused by such differences [@klei2015] and not purely due to data fabrication (given not all data points can be considered fabricated at this point).
+Moreover, the highly discrepant results from the DECREASE trials [@bouri2014] might partly be caused by such differences [@klei2015] and not purely due to data manipulation (given that not all data points can be considered manipulated at this point).
 
-To statistically investigate the evidence for data fabrication in the DECREASE studies [@dunkelgrun2009;@poldermans1999], we took three steps.
+To statistically investigate the evidence for data manipulation in the DECREASE studies [@dunkelgrun2009;@poldermans1999], we took three steps.
 First, we reproduced the findings from the 2014 meta-analysis by Bouri et al.
-[@bouri2014] that contained sufficient information to estimate the deviation of the DECREASE trials from other published trials on beta-blockers.
+[@bouri2014], which contained sufficient information to estimate the deviation of the DECREASE trials from other published trials on beta-blockers.
 We also included type of beta-blocker to inspect whether this is predictive of the effect of beta-blockers on perioperative mortality.
-Second, we evaluated the probability that the DECREASE trials occurred assuming no data fabrication.
-Third, we reversed this assumption and assumed that data fabrication did occur and estimated how many data points would have to be fabricated to reproduce the results of the DECREASE trials, if the other published trials are regarded as estimating the true effect of beta-blockers on perioperative mortality in patients undergoing non-cardiac surgery.
+Second, we evaluated the probability that the DECREASE trials occurred assuming no data manipulation.
+Third, we reversed this assumption and assumed that data manipulation did occur and estimated how many data points would have to be manipulated in order to reproduce the results of the DECREASE trials, if the other published trials are regarded as estimating the true effect of beta-blockers on perioperative mortality in patients undergoing non-cardiac surgery.
+Considering the committees investigating the scientific integrity of the DECREASE trials was unable to assess this, we consider it worthwhile to investigate this further.
 
@@
 We used the default weighting procedure in the `metafor` package.
-We added 0.5 to each cell, as is common in meta-analyses on risk- and odds ratios in order to prevent computational artefacts [@agresti2002; we made an error in the preregistration stating that only .5 was added to zero-cells].
+We added 0.5 to each cell count, as is common in meta-analyses on risk- and odds ratios in order to prevent computational artefacts [@agresti2002; p397].
 The 2014 meta-analysis [@bouri2014] did not specify the variance estimate used; hence, minor discrepancies between our estimates and the original estimates could be due to differences in the estimation procedure.
@@
 Bouri et al.
-differentiated between the estimates from the non-DECREASE trials and the DECREASE trials.
+differentiated between the estimates from the non-DECREASE trials ($k=`r sum(bouri_data$decrease == 0)`$) and the DECREASE trials ($k=`r sum(bouri_data$decrease == 1)`$).
 We confirmed the effect size estimates and the variance estimates for both the non-DECREASE- and the DECREASE trials, except for some minor discrepancies at the second decimal level.
-Table 1 depicts the original and reproduced values for both sets of studies.
+Table `r tabnr` depicts the original and reproduced values for both sets of studies.
 
@@
 
-knitr::kable(table1, caption = '__Table XX.__ The original- and reproduced meta-analytic results based on the data provided in the 2014 meta-analysis by Bouri et al.')
+knitr::kable(table1, caption = 'The original- and reproduced meta-analytic results based on the data provided in the 2014 meta-analysis by Bouri et al.')
+
+tabnr <- tabnr + 1
 ```
@@
 Second, we meta-analyzed all studies combined, including a dummy predictor for the DECREASE and non-DECREASE studies to reproduce results presented in Figure 4 of the 2014 meta-analysis [@bouri2014].
-Surprisingly, our results showed stronger evidence against equal subgroups than the original meta-analysis [@bouri2014] (original: $\chi^2(1)=3.91,p=.05$; reproduced: $\chi^2(1)=`r round(mod_mods$QM, 2)`,p=`r round(mod_mods$QMp, 3)`$).
+Surprisingly, our results showed a bit more evidence against equal subgroups than the original meta-analysis [@bouri2014] (original: $\chi^2(1)=3.91,p=.05$; reproduced: $\chi^2(1)=`r round(mod_mods$QM, 2)`,p=`r round(mod_mods$QMp, 3)`$).
 Additionally, the original analyses showed substantial residual heterogeneity ($I^2=74.4$\%) whereas we found no residual heterogeneity ($I^2=`r round(i2_mods, 3)`$\%).
 Different variance estimates (e.g., `DerSimonian-Laird` instead of `REML`) did not resolve this difference.
-We tried to clarify these discrepancies by e-mailing the original authors, with a reminder after two weeks, but did not receive a response.
+We tried to clarify these discrepancies by e-mailing the original authors (including a reminder after several weeks) but did not receive a response.
 Nonetheless, the broad strokes of the meta-regression confirmed that the DECREASE trials were the determining predictor for the effectiveness of beta-blockers (including DECREASE: $RR=`r round(exp(mod_mods$b[1] + mod_mods$b[2]), 3)`$; excluding DECREASE: $RR=`r round(exp(mod_mods$b[1]), 3)`$).
@@
 Nonetheless, for other beta-blockers in the non-DECREASE trials, there is still tentative and uncertain evidence that beta-blockers increase mortality (bisoprolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[3]), 3)`$; metoprolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[4]), 3)`$; propranolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[5]), 3)`$).
-Table 3 shows the meta-regression results in full.
+Table `r tabnr` shows the meta-regression results in full.
 
@@
 table3[2,2] <- round(mod_mods_type$b[1], 3)
-table3[3,2] <- ''
+table3[3,2] <- '-'
 table3[4,2] <- round(mod_mods_type$b[2], 3)
-table3[5,2] <- ''
+table3[5,2] <- '-'
 table3[6,2] <- round(mod_mods_type$b[3], 3)
@@
                        round(mod_mods_type$ci.ub[1], 3))
-table3[3,3] <- ''
+table3[3,3] <- '-'
 table3[4,3] <- sprintf('%s; %s', round(mod_mods_type$ci.lb[2], 3),
                        round(mod_mods_type$ci.ub[2], 3))
-table3[5,3] <- ''
+table3[5,3] <- '-'
 table3[6,3] <- sprintf('%s; %s', round(mod_mods_type$ci.lb[3], 3),
@@
 knitr::kable(table3, caption = 'Meta-regression results (k=11) for log(RR), including dummy predictors for DECREASE trials (reference: non-DECREASE trials) and type of beta-blockers used in the trial (reference: atenolol).')
+tabnr <- tabnr + 1
 ```
@@
 The 95\% credibility interval denotes the bounds of the normal distribution that covers 95\% of the density, where the standard deviation is calculated as the distance from the mean to either bound, divided by 1.96.
-This allows for an approximation of the population effect distribution, as depicted in Figure `r fig <- 1; fig`.
+This allows for an approximation of the population effect distribution, as depicted in Figure `r fignr`.
 
@@
 pdf(sprintf('%sfigures/fig1.pdf', x12), width = 8, height = 6)
-par(mar = c(3,0,0,0))
+par(mar = c(5,0,0,0))
 plot(x = temp,
@@
 knitr::include_graphics(sprintf("%sfigures/fig1.pdf", x12))
+fignr <- fignr + 1
 ```
@@
               (mod_clean_pred$pred - mod_clean_pred$cr.lb)/1.96) * 2
+
+zval <- (bouri_data$ln_rr[bouri_data$decrease == 1] - mod_clean_pred$pred) / ((mod_clean_pred$pred - mod_clean_pred$cr.lb)/1.96)
 ```
 
-Figure `r fig` indicates that the DECREASE trials are highly unlikely under the estimated effect distribution based on the non-DECREASE trials.
-More specifically, the results from DECREASE-I (or more extreme) have a probability of $`r prob[1]`$ (less than 1 in a quintillion) and the results from DECREASE-IV have a probability of $`r prob[2]`$ (3 in a billion).
+Figure `r fignr` indicates that the DECREASE trials are highly unlikely under the estimated effect distribution based on the non-DECREASE trials.
+More specifically, the results from DECREASE-I (or more extreme) have a probability of $`r prob[1]`$ (approximately 7 in a [quattuordecillion](https://en.wikipedia.org/wiki/Names_of_large_numbers); $z=`r round(zval[1], 3)`$) and the results from DECREASE-IV have a probability of $`r prob[2]`$ (approximately 7 in a billion; $z=`r round(zval[2], 3)`$).
 This indicates that the DECREASE trial results are unlikely to have come from the same population effect distribution as the non-DECREASE trials.
@@
 Results from Step 1 indicated that no between-trial variance (i.e., homogeneity; $\tau^2=0$) of the effects was observed; given the small number of trials included (i.e., `r mod_clean$k`) this estimate is highly uncertain, however.
-The total $N$ across the non-DECREASE trials was `r bouri_data$control_n[bouri_data$decrease == 0] + bouri_data$beta_n[bouri_data$decrease == 0]`.We conducted sensitivity analyses to see how dependent results are on the heterogeneity estimate (not preregistered; [osf.io/vnmzc](https://osf.io/vnmzc)).
-The probability of observing the DECREASE trials stays approximately below 1 out of 1000 until the variance estimate is 0.25; the probability stays approximately below 1 out of 100 until the variance estimate is 0.43 (see Figure `r fig <- fig + 1; fig`).
+The total $N$ across the non-DECREASE trials was `r sum(bouri_data$control_n[bouri_data$decrease == 0] + bouri_data$beta_n[bouri_data$decrease == 0])`.We conducted sensitivity analyses to see how dependent results are on the heterogeneity estimate (not preregistered; [osf.io/vnmzc](https://osf.io/vnmzc)).
+The probability of observing the DECREASE trials stays approximately below 1 out of 1000 until the variance estimate is 0.25; the probability stays approximately below 1 out of 100 until the variance estimate is 0.43 (see Figure `r fignr`).
 To put these numbers into context, a variance of 0.25 would suggest that results of perioperative beta-blockade vary substantially due to contextual circumstances of the study, even if perioperative beta-blockade has no effect whatsoever (RRs between `r round(exp(0 - .25), 3)` and `r round(exp(0 + .25), 3)` in ~64\% of the cases).
@@
 pdf(sprintf('%sfigures/fig2.pdf', x12), width = 8, height = 6)
-tau2 <- seq(0, 2, .01)
+tau2 <- seq(0, .5, .01)
 
@@
 text(x = tau2[c(26, 44)],
-     y = prob_sens_t[c(26, 44)] + .02,
+     y = prob_sens_t[c(26, 44)] + .003,
      labels = c(TeX('$\\tau^2=.25$'), TeX('$\\tau^2=.43$')),
@@
 knitr::include_graphics(sprintf("%sfigures/fig2.pdf", x12))
+fignr <- fignr + 1
 ```
 
-## Step 3: estimating the amount of fabricated data
+## Step 3: estimating the amount of manipulated data
 
@@
 ```
-We estimated the number of data points that would need to be fabricated to arrive at the estimates from the DECREASE trials, given that the non-DECREASE trials represent the true effect of perioperative beta-blockade.
-In contrast to Step 2 this assumes that the DECREASE trials might in fact contain fabricated data.
-The estimates from Step 3 provide an indication of the extent of potential data fabrication in the DECREASE studies [@commissie2011;@commissie2012;@commissie2013;@bouri2014].
+We estimated the number of data points that would need to be manipulated to arrive at the estimates from the DECREASE trials, given that the non-DECREASE trials represent the true effect of perioperative beta-blockade.
+In contrast to Step 2, which assumes no data manipulation occurred, Step 3 assumes that the DECREASE trials might in fact contain manipulated data.
+The estimates from Step 3 provide an indication of the extent of potential data manipulation in the DECREASE studies [@commissie2011;@commissie2012;@commissie2013;@bouri2014].
 
@@
 
-In order to estimate the number of fabricated data points, we first estimated the probability of perioperative mortality (in log odds) in each trial arm separated per trial stratum.
-As such, there are four estimates of perioperative mortality: one per condition (beta-blocker or control) per trial stratum (DECREASE- or non-DECREASE trials).
-For each of these, we ran a meta-analysis applying the same methods used in Step 1.
-For each of the two DECREASE trials separately and the non-DECREASE trials combined, this resulted in four meta-analytic mortality estimates with corresponding effect variances (see Table 3; one estimate per cell).
-Throughout the simulations, we used the point estimates (i.e., fixed effect) to simulate genuine- and fabricated data, but supplemented this by using the more uncertain distribution estimates (i.e., random effects) as sensitivity analyses.
+In order to estimate the number of manipulated data points, we first estimated the probability of perioperative mortality (in log odds) in each trial arm for each trail stratum.
+As such, we estimate mortality odds four times: once per condition (beta-blocker or control) per trial type (DECREASE- and non-DECREASE trials).
+For all combinations of condition and trial type, we ran a meta-analysis applying similar methods used in Step 1, resulting in four meta-analytic absolute mortality estimates with corresponding effect variances (see Table 3; one estimate per cell).
+Throughout the simulations, we used the point estimates (i.e., fixed effect) to simulate genuine- and manipulated data, but supplemented this by using the more uncertain distribution estimates (i.e., random effects) as sensitivity analyses.
 
@@
 
-kable(table2, caption = 'Outcome possibilities within a simulated 2 (beta-blocker v control) by 2 (dead v alive) clinical trial.')
+knitr::kable(table2, caption = 'Outcome possibilities within a simulated 2 (beta-blocker v control) by 2 (dead v alive) clinical trial.')
+
+tabnr <- tabnr + 1
 ```
 
-We applied the inversion method to estimate the number of fabricated data points in the DECREASE trials [@casella2002].
-The inversion method iteratively hypothesizes that *X* out of *N* data points were fabricated (i.e., $X={0, 1, ..., N}$).
+We applied the inversion method to estimate the number of manipulated data points in the DECREASE trials [@casella2002].
+The inversion method iteratively hypothesizes that *X* out of *N* data points were manipulated (i.e., $X={0, 1, ..., N}$).
 For each combination of *X* and trial, we simulated `r iterations` datasets.
-Each simulated dataset contained *X* fabricated data points and *N-X* genuine data points.
+Each simulated dataset contained *X* manipulated data points and *N-X* genuine data points.
 For each simulated dataset (exact simulation procedure in the next paragraph), we determined the likelihood of the results with
@@
 where $\pi_{E}$ indicates the mortality rate in the beta-blocker condition as drawn from the meta-analytic effect distribution ($\pi_{C}$ indicates the mortality rate in the control condition).
-The likelihood was computed under both the fabricated effect estimates (i.e., $L_{fabricated}$) and the genuine data (i.e., $L_{genuine}$).
+We estimated those parameters using the meta-analytic procedure described in the previous paragraph, resulting in the estimates depicted in Table `r tabnr`.
+The likelihood was computed under both the manipulated effect estimates (i.e., $L_{manipulated}$) and the genuine data (i.e., $L_{genuine}$).
 Table 3 indicates which cell sizes the various $n_{XX}$ refer to within the (simulated) data.
-After computing the likelihoods, we compared them to determine whether the simulated data were more likely to arise from the genuine trials ($L_{genuine}>L_{fabricated}$) or from the fabricated trials ($L_{fabricated}>L_{genuine}$).
+After computing the likelihoods, we compared them to determine whether the simulated data were more likely to arise from the genuine trials ($L_{genuine}>L_{manipulated}$) or from the manipulated trials ($L_{manipulated}>L_{genuine}$).
 Note that comparing the likelihoods is a minor deviation from the preregistration, where we initially planned on using $p$-value comparisons ([osf.io/vnmzc](https://osf.io/vnmzc)).
-
-For each hypothesis of *X* out of *N* fabricated data points, we computed the probability that the fabricated data are more likely than the genuine data ($p_F=P(L_{fabricated}>L_{genuine})$).
-Based on $p_F$, we computed the confidence interval for the estimated $X$ fabricated data points (i.e., $X_{LB};X_{UB}$).
-For a 95\% confidence interval, the lowerbound is equal to the $p_F$ closest to .025, whereas the upperbound is equal to the $p_F$ closest to .975.
-
-We computed $p_F$ for all *X* out of *N* fabricated data points in `r iterations` randomly generated datasets, which were generated in three steps. For each dataset we:
-
-1. Sampled (across conditions, without replacement) *X* fictitious participants that would be the result of data fabrication.
-
-2.
-Determined the population mortality rate for each condition (i.e., for each cell as in Table 3).
-The meta-analytic point estimate was used or a population effect was randomly drawn from the meta-analytic effect distribution.
-
-3. Simulated the number of deaths for the different conditions using a binomial distribution based on the mortality rate as determined in 2, resulting in the cell counts as in Table 3.
-
-Based on the meta-analytic effect from 2 and the cell sizes from 3, we computed the likelihoods $L_{fabricated}$ and $L_{genuine}$ using Equation \ref{eq1}.
-As mentioned before, we computed $p_F$, which indicates the probability that the data are more likely under the estimates resulting from the (allegedly) fabricated data (i.e., the DECREASE trials) than under the estimates resulting from the genuine data (i.e., the non-DECREASE trials; $p_F=P(L_{fabricated}>L_{genuine})$).
-
-### Results
 
@@
 
+table22 <- matrix('', ncol = 4, nrow = 5)
+table22[1,1] <- ''
+table22[1,2] <- ''
+table22[1,3] <- 'log(odds)'
+table22[1,4] <- "$\\tau$"
+
+table22[2,1] <- 'DECREASE (k=2)'
+table22[2,2] <- 'Beta-blocker'
+table22[2,3] <- round(beta_decrease_pred, 3)
+table22[2,4] <- round(beta_decrease_var, 3)
+
+table22[3,1] <- ''
+table22[3,2] <- 'Control'
+table22[3,3] <- round(cont_decrease_pred, 3)
+table22[3,4] <- round(cont_decrease_var, 3)
+
+table22[4,1] <- 'Non-DECREASE (k=9)'
+table22[4,2] <- 'Beta-blocker'
+table22[4,3] <- round(beta_pred, 3)
+table22[4,4] <- round(beta_var, 3)
+
+table22[5,1] <- ''
+table22[5,2] <- 'Control'
+table22[5,3] <- round(cont_pred, 3)
+table22[5,4] <- round(cont_var, 3)
+
+knitr::kable(table22, caption = "Mortality estimates per condition, per trial stratum. We used these parameters to estimate the number of manipulated data points with the inversion method.")
+
+tabnr <- tabnr + 1
+```
+
+For each hypothesis of *X* out of *N* manipulated data points, we computed the probability that the manipulated data are more likely than the genuine data ($p_M=P(L_{manipulated}>L_{genuine})$).
+Based on $p_M$, we computed the confidence interval for the estimated $X$ manipulated data points (i.e., $X_{LB};X_{UB}$).
+For a 95\% confidence interval, the lower bound is equal to the $p_M$ closest to .025, whereas the upperbound is equal to the $p_M$ closest to .975.
+
+We computed $p_M$ for all *X* out of *N* manipulated data points in `r iterations` randomly generated datasets, which were generated in three steps. For each dataset we:
+
+1. Sampled (across conditions, without replacement) *X* fictitious participants that would be the result of data manipulation.
+
+2.
+Determined the population mortality rate for each condition (i.e., for each cell as in Table 3).
+The meta-analytic point estimate was used or a population effect was randomly drawn from the meta-analytic effect distribution.
+
+3. Simulated the number of deaths for the different conditions using a binomial distribution based on the mortality rate as determined in 2, resulting in the cell counts as in Table 3.
+
+Based on the meta-analytic effect from 2 and the cell sizes from 3, we computed the likelihoods $L_{manipulated}$ and $L_{genuine}$ using Equation \ref{eq1}.
+As mentioned before, we computed $p_M$, which indicates the probability that the data are more likely under the estimates resulting from the (allegedly) manipulated data (i.e., the DECREASE trials) than under the estimates resulting from the genuine data (i.e., the non-DECREASE trials; $p_M=P(L_{manipulated}>L_{genuine})$).
+
+### Results
+
+```{r, echo = FALSE}
 # Uncomment if you want to rerun the inversion method
@@
 
-# Manually correct some errors in the bound seeking from the loop above
-# errors arise from Monte Carlo error:
-# .022 .021 .023 <-- selects .021, but should select 0
-# considering that running >10000 takes ages and does not guarantee
-# it will fix the problem, I do it like this
-res_i_uncertain_lb <- ifelse(res_i_uncertain_lb == 4, 0, res_i_uncertain_lb)
-res_i_certain_lb <- ifelse(res_i_certain_lb == 1, 0, res_i_certain_lb)
-res_iv_uncertain_lb <- ifelse(res_iv_uncertain_lb == 8, 0, res_iv_uncertain_lb)
-res_iv_certain_lb[98:101] <- 0
-res_iv_certain_ub[98:101] <- 1067
-
 pdf(sprintf('%sfigures/fig3.pdf', x12), width = 8, height = 6)
@@
 plot(res_i_uncertain,
-     xlab = 'Data points fabricated',
+     xlab = 'Number of data points manipulated',
      ylab = 'Proportion',
@@
 plot(res_iv_uncertain,
-     xlab = 'Data-points fabricated',
+     xlab = 'Number of data points manipulated',
      ylab = '',
@@
      xlim = c(0,1), ylim = c(0, length(res_i_certain)),
-     xlab = "Confidence in %", ylab = "Data points fabricated")
+     xlab = "Confidence in %", ylab = "Number of data points manipulated")
 
@@
 
-```{r figure 3, fig.cap = 'Inversion method results used to estimate the number of data points fabricated in the DECREASE-I and DECREASE-IV trials.
-The top row panels indicate $p_F$ (y-axis) for all $X$ out of $N$ fabricated data points (x-axis).
-The bottom row indicates the estimated number of fabricated data points (y-axis) when varying the degree of confidence (x-axis).
+For DECREASE-I ($N=`r bouri_data$control_n[10] + bouri_data$beta_n[10]`$), the 95\% confidence interval for the estimated number of manipulated data points is [`r which.min(res_i_certain - .025) - 1` - `r which.max(res_i_certain - .975) - 1`] or [`r which.min(res_i_uncertain - .025) - 1` - `r which.max(res_i_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
+The left column of Figure `r fignr` depicts the $p_M$ per *X* manipulated data points (top panel) and the bounds of the confidence interval when the degree of confidence is altered (lower panel).
+Staying clearly between the dotted lines in the top panel, depicting the 95\% CI (top: .975; bottom: .025), it becomes apparent that the degree of uncertainty is too high to make any reasonable estimates about the number of manipulated data points with sufficient confidence.
+This is partly due to the small sample size of the DECREASE-I trial (i.e., $N=`r length(res_i_certain) - 1`$) and the availability of just the summary results.
+Only when the degree of confidence is lowered to around 75\% does the interval not span the entire sample size.
+As such, based on the summary results, little can be said about the extent of the data manipulation that occurred in the DECREASE-I trial, affirming the conclusions of the original committee report [@commissie2013].
+
+```{r figure 3, fig.cap = 'Inversion method results used to estimate the number of data points manipulated in the DECREASE-I and DECREASE-IV trials.
+The top row panels indicate $p_M$ (y-axis) for all $X$ out of $N$ manipulated data points (x-axis).
+The bottom row indicates the estimated interval of manipulated data points (y-axis) when varying the degree of confidence (x-axis).
 Dotted lines indicate the bounds for a 95 percent CI.', out.width = '80%', fig.asp = .5, fig.align = 'center', echo = FALSE}
 knitr::include_graphics(sprintf("%sfigures/fig3.pdf", x12))
+fignr <- fignr + 1
 ```
 
-For DECREASE-I ($N=`r bouri_data$control_n[10] + bouri_data$beta_n[10]`$), the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_i_certain - .025) - 1` - `r which.max(res_i_certain - .975) - 1`] or [`r which.min(res_i_uncertain - .025) - 1` - `r which.max(res_i_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
-The left column of Figure 3 depicts the $p_F$ per *X* fabricated data points (top panel) and the bounds of the confidence interval when the degree of confidence is altered (lower panel).
-Staying clearly between the dotted lines in the top panel, depicting the 95\% CI (top: .975; bottom: .025), it becomes apparent that the degree of uncertainty is too high to make any reasonable estimates about the number of fabricated data points with sufficient confidence.
-This is partly due to the small sample size of the DECREASE-I trial (i.e., $N=`r length(res_i_certain) - 1`$) and the availability of just the summary results.
-Only when the degree of confidence is lowered to around 75\% does the interval not span the entire sample size.
-As such, based on the summary results, little can be said about the extent of the data fabrication that occurred in the DECREASE-I trial, affirming the conclusions of the original committee report [@commissie2013].
-
-For DECREASE-IV ($N=`r bouri_data$control_n[11] + bouri_data$beta_n[11]`$), the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_iv_certain - .025) - 1` - `r which.max(res_iv_certain - .975) - 1`] or [`r which.min(res_iv_uncertain - .025) - 1` - `r which.max(res_iv_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
-The relatively minor difference between the estimates indicates that there is a high degree of confidence that data fabrication did occur based on the difference of the trial results alone.
-Nonetheless, the range of potentially fabricated data points is still estimated at approximately 1000; this indicates that the summary results are insufficient to provide more than an estimated lowerbound.
-This indicates that it is possible not all data were fabricated (i.e., $N=`r length(res_iv_certain) - 1`$), increasing the importance of well-documented data provenance to discern between genuine and falsified data.
+For DECREASE-IV ($N=`r bouri_data$control_n[11] + bouri_data$beta_n[11]`$), the 95\% confidence interval for the estimated number of manipulated data points is [`r which.min(res_iv_certain - .025) - 1` - `r which.max(res_iv_certain - .975) - 1`] or [`r which.min(res_iv_uncertain - .025) - 1` - `r which.max(res_iv_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
+The relatively minor difference between the estimates indicates that there is a high degree of confidence that data manipulation did occur based on the difference of the trial results alone.
+Nonetheless, the range of potentially manipulated data points is still estimated at approximately 1000; this indicates that the summary results are insufficient to provide more than an estimated lower bound.
+This indicates that it is possible not all data were manipulated (i.e., $N=`r length(res_iv_certain) - 1`$), increasing the importance of well-documented data provenance to discern between genuine and falsified data.
 
@@
 
-The effect of beta-blockade on perioperative mortality was already unclear based on the findings of research misconduct in institutional investigations [@commissie2011;@commissie2012;@commissie2013]; our results affirm that the empirical results from the DECREASE trials are potentially problematic in light of the other studies supposedly investigating the same effect.
-We indicate that the results from the DECREASE trials are nearly impossible to have arisen from the same effect inspected by the non-DECREASE trials, except when we assume at least some of the data were manipulated and that these trials investigate the same effect.
-As such, we consider the scientific validity of the DECREASE-I and DECREASE-IV trials as problematic when assessing the effectiveness of beta-blockade on perioperative mortality.
-Nonetheless, the original papers that presented these trial results are not yet retracted [@poldermans1999;@dunkelgrun2009], despite the damning integrity reports [@commissie2011;@commissie2012;@commissie2013].
+The effect of beta-blockade on perioperative mortality was already unclear based on the investigations regarding scientific integrity; our results strongly affirm that the empirical evidence from the DECREASE trials is highly discrepant from other trials supposedly studying the same effect (i.e., the effectiveness of beta-blockers in decreasing perioperative mortality).
+Our results indicate that the results from the DECREASE trials are nearly impossible to have arisen from the same effect inspected by the non-DECREASE trials, except when we assume at least some of the data were manipulated.
+As such, the scientific validity of the DECREASE-I and DECREASE-IV trials should be regarded as highly problematic and untrustworthy when assessing the effectiveness of beta-blockade on perioperative mortality if they truly investigate the same effects as the non-DECREASE trials, as is often assumed [@bouri2014].
+Nonetheless, the original papers that presented these trial results are not yet retracted [@poldermans1999;@dunkelgrun2009], despite the integrity reports [@commissie2011;@commissie2012;@commissie2013].
 
-Nonetheless, we note that statistically estimated number of manipulated data points in the DECREASE trials is subject to multiplicity.
-In other words, for any estimated proportion of data points manipulated, there is another estimate that contains fewer (or more) manipulated data points that are more (or less) extremely manipulated and yields exactly the same outcome.
-This is similar to how the mean of two samples can be equivalent (e.g., 0), but that the individual scores giving rise to that mean are extremely different (e.g, +2.5 and -2.5 in the first sample; -100 and +100 in the second sample).
-Regardless, whether any manipulation occurred does not suffer from this multiplicity because there is only one way to have no manipulated data points.
+<!-- Ja onderscheid tussen fabricatie en manipulatie is het discussieren waard.
+Voor de werenschappelijke waarde vd decrease trials maakt het weinig uit, maar het is wel interessant voor de vraag hoe die rare data van poldermans totstand gekomen zijn.
+-->
+
+Our approach to estimating the number of manipulated data points has one major limitation that we would like to highlight: multiplicity.
+For each estimated proportion of manipulated data points, there is another smaller (or larger) proportion with more (or less) extremely manipulated data points.
+This problem is similar to how various samples can give rise to the same mean, but contain vastly different individual scores within them (e.g., -2.5 and +2.5 versus -100 and +100; both give the mean zero).
+Nonetheless, this limitation does still allow us to estimate whether any data manipulation occurred because there is no multiplicity in not manipulating data.
 
 The ESC/ESA and ACC/AHA guidelines [@Kristensen_2014;@Fleisher_2014] on perioperative beta-blockade already excluded the DECREASE trials in their assessment, but also state that other trials by Poldermans are excluded.
-However, upon close inspection of the reference lists, the ACC/AHA guidelines still cites four trials as evidence [@Boersma_2001;@poldermans1999;@van_Kuijk_2009;@Flu_2010], of which two were already inspected by the scientific integrity committees of Erasmus MC [@Boersma_2001;@poldermans1999].
+However, upon close inspection of the reference lists, the ACC/AHA guidelines still cites four trials including Poldermans as author as evidence for the guidelines [@Boersma_2001;@poldermans1999;@van_Kuijk_2009;@Flu_2010], of which two were already inspected by the scientific integrity committees of Erasmus MC [@Boersma_2001;@poldermans1999].
 In the ACC/AHA guidelines, the following is said about studies conducted by Poldermans:
@@
 Given the confirmation of problems in the DECREASE-I and DECREASE-IV trials in our results, it stresses that there is reason to distrust trials by Poldermans.
-For the integrity of the guidelines and the safety of the patients, we pose that investigations should be initiated into works where Poldermans was clearly involved and which were not cleared by the scientific committees of Erasmus MC in their misconduct investigations.
+For the integrity of the guidelines and the safety of the patients, we pose that investigations should be initiated into works where Poldermans was involved and which were not cleared by the scientific committees of Erasmus MC in their misconduct investigations.
 Especially those papers cited as evidence in the ACC/AHA guidelines should be investigated, considering that they affect patients and their treatment directly.
 
-Previously, further investigation of trials by Poldermans was deemed unfeasible due to the lack of raw data; here we indicate methods that do make it feasible, albeit with their own limitations.
-Based on just event-count data and comparable trials, we were able to estimate whether part of the data were in fact fabricated and whether the results were within reason of other trials supposedly investigating the same effect.
-The results clearly indicated the DECREASE trials are not within reason, if we assume the non-DECREASE trials to investigate the same effect.
+Previously, further investigation of trials by Poldermans was deemed unfeasible due to the lack of raw data; here we indicate methods that do make it feasible.
+Based on just event-count data and trials that supposedly investigate the same effect, we were able to estimate whether part of the data were in fact manipulated and whether the results were within reason of trials investigating the same effect.
+The results clearly indicated they were not such reason.
 
-The results of our analyses also highlight that, despite the lack of availability of the raw data, summary results from larger samples allow for better estimates of the number of fabricated data points when similar trials are available.
-Moreover, larger trials result in relatively more certainty (e.g., DECREASE-IV) about the estimated number of fabricated data points, when using the inversion method, compared to smaller trials (e.g., DECREASE-I).
-This increased certainty is due to decreased standard errors at the population level of the estimated effects, resulting in higher sensitivity to data anomalies.
-Nonetheless, alongside multiplicity much residual uncertainty remains and simply less information is available in the summary results.
-Raw data availability would improve the options open to detect potential anomalies (note: raw data are available for DECREASE VI, but upon a freedom of information request by the first author, Erasmus MC refused to share these data; letters available at [osf.io/fswjt](https://osf.io/fswjt/)).
-The results also highlight that in order to prevent detection, it would be in the fabricators' interest to fabricate small and imprecise studies (assuming the fabricator wants to remain undetected), which ultimately detracts from the scientific value of such a study and hopefully the subsequent reward for fabrication due to less impact.
+The results of our analyses also highlight that, despite the lack of availability of the raw data, summary results from larger samples allow for better estimates of the number of manipulated data points when similar trials are available.
+Moreover, larger trials result in relatively more certainty (e.g., DECREASE-IV) about the estimated number of manipulated data points, when using the inversion method, compared to smaller trials (e.g., DECREASE-I).
+This increased certainty is due to decreased standard errors of the estimated effects, resulting in higher sensitivity to data anomalies.
+Nonetheless, much residual uncertainty remains and simply less information is available in summary results when compared to raw data.
+As such, raw data availability would improve the options open to detect potential anomalies (note: raw data are available for DECREASE VI, but upon a freedom of information request by the first author, Erasmus MC refused to share these data; see [osf.io/fswjt/](https://osf.io/fswjt/)).
+The results also highlight that in order to prevent detection, it would be in the manipulator's interest to fabricate small and imprecise studies (assuming the manipulator wants to remain undetected), which ultimately detracts from the scientific value of such a study and hence the individual reward for manipulation through reduced impact (hopefully).
 
@@
 However, there seems to be covariation with respect to treatment administration, duration, and dose, which further confounds whether the treatment effect is due to type of beta-blocker or due to one of these other parameters.
-Given the limited statistical power in this meta-analysis due to the lack of studies to investigate which factors are most essential, we recommend a clinical trial that aims to dissect some of these potential confounds.
+There are too few studies ($k=11$) too properly discern the various treatments from each other, requiring a new randomized trial with high statistical power to determine moderating factors (if any).
 This affirms the statement from the ESC/ESA guidelines that _"high priority needs to be given to new randomized clinical trials to better identify which patients derive benefit from beta-blocker therapy in the perioperative setting, and to determine the optimal method of beta-blockade"_ [@Kristensen_2014].
 
-Moreover, the DECREASE and non-DECREASE trials seem to apply beta-blockade from different conceptual viewpoints.
+Moreover, the DECREASE and non-DECREASE trials seem to apply beta-blockade from different conceptual viewpoints that could confound the effectiveness of beta-blockade.
 The non-DECREASE trials seem to focus purely on the application of beta-blockers in itself, whereas the DECREASE trials use beta-blockade as a proxy to decrease resting heart rate [@poldermans1999;@dunkelgrun2009].
@@
 
+# Session info
+
+```{r, echo = FALSE}
+sessionInfo()
+```
+
 # References

### 2017-07-14 submission/manuscript.tex
Revision round 1200

@@
 \hypersetup{unicode=true,
-            pdftitle={Analyzing DECREASE trials to determine evidence for data fabrication},
-            pdfauthor={Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper},
+            pdftitle={Analyzing DECREASE trials to estimate evidence of data manipulation},
+            pdfauthor={Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper, Markus Hollman},
             pdfborder={0 0 0},
@@
 \setlength{\droptitle}{-2em}
-  \title{Analyzing DECREASE trials to determine evidence for data fabrication}
+  \title{Analyzing DECREASE trials to estimate evidence of data manipulation}
   \pretitle{\vspace{\droptitle}\centering\huge}
   \posttitle{\par}
-  \author{Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper}
+  \author{Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper, Markus Hollman}
   \preauthor{\centering\large\emph}
@@
   \postdate{\par}
-  \date{06 June, 2017}
+  \date{14 July, 2017}
 
@@
 The effect of beta-blockers on perioperative mortality in non-cardiac
-surgery has been controversial\textsuperscript{1} due to findings of
-research misconduct in two related clinical
+surgery has been controversial\textsuperscript{1} due to concerns
+regarding the scientific integrity in two related clinical
 trials\textsuperscript{2--6}. Three meta-analyses that included the
-trials subject to research misconduct concluded that beta-blockers
-decrease perioperative mortality\textsuperscript{7--9} whereas a
-meta-analysis that excluded the suspect trials concluded that
-beta-blockers increase perioperative mortality\textsuperscript{9}. In
-these studies, perioperative mortality was defined as the deathrate of
-patients during the period of the surgical procedure, which typically
-includes admission, anaesthesia, surgery, and recovery.
+trials subject to concerns concluded that beta-blockers decrease
+perioperative mortality\textsuperscript{7--9} whereas a meta-analysis
+that excluded the suspect trials concluded that beta-blockers increase
+perioperative mortality\textsuperscript{9}. In these studies,
+perioperative mortality was defined as the death rate of patients in the
+perioperative setting, including the period of admission, anaesthesia
+with surgery, and postoperative recovery.
 
-The trials subject to research misconduct were the Dutch DECREASE-I and
-DECREASE-IV trials\textsuperscript{4--6}. The committees that
-investigated the integrity of the DECREASE trials reported that data
-fabrication was likely but that the extent of the data fabrication
-remained unclear\textsuperscript{4--6}. Moreover, the latest guidelines
-still recommend the usage of beta-blockers in the perioperative period
-in certain cases\textsuperscript{{\textbf{???}},10}. Considering these
-consequences, we aim to estimate the extent of data fabrication in the
-DECREASE studies\textsuperscript{2--6} to further the debate on using
-beta-blockers in the perioperative period of CAD patients undergoing
-non-cardiac surgery.
+The trials subject to concerns regarding scientific integrity were the
+Dutch DECREASE-I and DECREASE-IV trials\textsuperscript{4--6}. The
+committees that investigated the integrity of the DECREASE trials
+reported that data manipulation was likely but that the extent of the
+data manipulation remained unclear\textsuperscript{4--6}. Moreover, the
+latest guidelines still recommend the usage of beta-blockers in the
+perioperative period in certain cases\textsuperscript{10,11}, where some
+of these guidelines are based on other work by the PI of the DECREASE
+trials. Considering the potential harmful consequences of guidelines
+based on work by someone who has been repeatedly investigated for
+breaching scientific integrity, we aim to estimate the extent of data
+manipulation in the DECREASE studies\textsuperscript{2--6} to further
+stimulate the debate on using beta-blockers in the perioperative period
+patients undergoing non-cardiac surgery.
 
@@
 corresponded to patient files. The committee reports did not neglect
-statistical evaluation however: a statistical expert evaluated the
-applicability of forensic statistical methods\textsuperscript{6} to
-evaluate results of trials separately (i.e., DECREASE-I, DECREASE-IV).
-Nonetheless, comparing across trials is a method that has previously
-been used to monitor trial data quality or to test for potential data
-anomalies\textsuperscript{11}. Moreover, this method has previously
+statistical evaluation however: according to the report a statistical
+expert evaluated the applicability of forensic statistical
+methods\textsuperscript{6} to evaluate results of trials separately
+(i.e., DECREASE-I, DECREASE-IV) although it lacks details as to how this
+evaluation took place. The expert concluded that the previously applied
+methods by him were not applicable for use in this case. Nonetheless, in
+this report we compare across trials, which is a method that has
+previously been used to monitor trial data quality or to test for
+potential data anomalies\textsuperscript{12,13}. Moreover, comparing
+across trials instead of evaluating them separately has previously
 proven to be effective in detecting data
-fabrication\textsuperscript{12}. Comparing the DECREASE trials to other
+manipulation\textsuperscript{14}. Comparing the DECREASE trials to other
 published trials studying the effectiveness of beta-blockers with
 respect to perioperative mortality could prove informative of the
-potential extent of the fabrication in the DECREASE trials.
+potential extent of the manipulation in the DECREASE trials.
 
@@
 afflicted DECREASE trials, potentially interacting with the type of
-beta-blocker and the way that beta-blocker was administered (i.e., dose
-and duration of treatment). Throughout the clinical trials regarding
-beta-blockers, patients were administered various beta-blockers (e.g.,
-metroprolol, bisoprolol, atenolol) and in various ways (e.g.,
-intraveneously, orally; half an hour before surgery or multiple days
-before surgery). Factors such as dosage and duration can have an effect
-on the pharmacological effectiveness with respect to perioperative
-mortality. Moreover, the anomalous results from the DECREASE trials
-might partly be caused by such differences\textsuperscript{13} and not
-purely due to data fabrication (given not all data points can be
-considered fabricated at this point).
+beta-blocker and the way that beta-blockers were administered (i.e.,
+dose and duration of treatment). In the randomized trials on
+beta-blockers, patients were administered various types of beta-blockers
+(e.g., metroprolol, bisoprolol, atenolol) and in various ways (e.g.,
+intravenously, orally; half an hour before surgery or multiple days
+before surgery; with or without titration based on heart rate). Factors
+such as dosage and duration can have an effect on the pharmacological
+effectiveness with respect to perioperative mortality. Moreover, the
+highly discrepant results from the DECREASE trials\textsuperscript{9}
+might partly be caused by such differences\textsuperscript{15} and not
+purely due to data manipulation (given that not all data points can be
+considered manipulated at this point).
 
-To statistically investigate the evidence for data fabrication in the
+To statistically investigate the evidence for data manipulation in the
 DECREASE studies\textsuperscript{2,3}, we took three steps. First, we
 reproduced the findings from the 2014 meta-analysis by Bouri et
-al.\textsuperscript{9} that contained sufficient information to estimate
-the deviation of the DECREASE trials from other published trials on
-beta-blockers. We also include type of beta-blocker to inspect whether
-this is predictive of the effect of beta-blockers on perioperative
-mortality. Second, we evaluated the probability that the DECREASE trials
-occurred assuming no data fabrication. Third, we reversed this
-assumption and assumed that data fabrication did occur and estimated how
-many data points would have to be fabricated to reproduce the results of
-the DECREASE trials, if the other published trials are regarded as
-estimating the true effect of beta-blockers on perioperative mortality
-in patients undergoing non-cardiac surgery.
+al.\textsuperscript{9}, which contained sufficient information to
+estimate the deviation of the DECREASE trials from other published
+trials on beta-blockers. We also included type of beta-blocker to
+inspect whether this is predictive of the effect of beta-blockers on
+perioperative mortality. Second, we evaluated the probability that the
+DECREASE trials occurred assuming no data manipulation. Third, we
+reversed this assumption and assumed that data manipulation did occur
+and estimated how many data points would have to be manipulated in order
+to reproduce the results of the DECREASE trials, if the other published
+trials are regarded as estimating the true effect of beta-blockers on
+perioperative mortality in patients undergoing non-cardiac surgery.
+Considering the committees investigating the scientific integrity of the
+DECREASE trials was unable to assess this, we consider it worthwhile to
+investigate this further.
 
@@
 We computed the log risk ratio (i.e., log RR) for each study and pooled
-these using the \texttt{R} package \texttt{metafor}\textsuperscript{14}.
+these using the \texttt{R} package \texttt{metafor}\textsuperscript{16}.
 We estimated a weighted random-effects model using the restricted
-maximum-likelihood estimator (i.e., \texttt{REML})\textsuperscript{15}
+maximum-likelihood estimator (i.e., \texttt{REML})\textsuperscript{17}
 to estimate the variance of effects. We used the default weighting
-procedure in the \texttt{metafor} package. When there was a zero-count
-for a cell (e.g., zero events), 0.5 was added to that cell, as is common
-in meta-analyses on risk- and odds ratios in order to prevent
-computational artefacts\textsuperscript{16}. The 2014
+procedure in the \texttt{metafor} package. We added 0.5 to each cell
+count, as is common in meta-analyses on risk- and odds ratios in order
+to prevent computational artefacts {[}18; p397{]}. The 2014
 meta-analysis\textsuperscript{9} did not specify the variance estimate
@@
 Bouri et al. differentiated between the estimates from the non-DECREASE
-trials and the DECREASE trials. We confirmed the effect size estimates
-and the variance estimates for both the non-DECREASE- and the DECREASE
-trials, save for some minor discrepancies due to the estimation method.
-Table 1 depicts the original and reproduced values for both sets of
-studies.
+trials (\(k=9\)) and the DECREASE trials (\(k=2\)). We confirmed the
+effect size estimates and the variance estimates for both the
+non-DECREASE- and the DECREASE trials, except for some minor
+discrepancies at the second decimal level. Table 1 depicts the original
+and reproduced values for both sets of studies.
 
 \begin{longtable}[]{@{}lllll@{}}
-\caption{\textbf{Table XX.} The original- and reproduced meta-analytic
-results based on the data provided in the 2014 meta-analysis by Bouri et
-al.}\tabularnewline
+\caption{The original- and reproduced meta-analytic results based on the
+data provided in the 2014 meta-analysis by Bouri et al.}\tabularnewline
 \toprule
 & & Risk ratio & \(\tau^2\) & Confidence interval\tabularnewline
-Non-DECREASE & Original & 1.27 & 0 & {[}1.01; 1.60{]}\tabularnewline
-& Reproduced & 1.28 & 0 & {[}1.01; 1.62{]}\tabularnewline
-DECREASE & Original & 0.42 & 0.29 & {[}0.15; 1.23{]}\tabularnewline
-& Reproduced & 0.44 & 0.24 & {[}0.15; 1.23{]}\tabularnewline
+Non-DECREASE (k=9) & Original & 1.27 & 0 & {[}1.01;
+1.60{]}\tabularnewline
+& Reproduced & 1.27 & 0 & {[}1.01; 1.61{]}\tabularnewline
+DECREASE (k=2) & Original & 0.42 & 0.29 & {[}0.15;
+1.23{]}\tabularnewline
+& Reproduced & 0.47 & 0.14 & {[}0.19; 1.15{]}\tabularnewline
 \bottomrule
@@
 presented in Figure 4 of the 2014 meta-analysis\textsuperscript{9}.
-Surprisingly, our results showed stronger evidence against equal
+Surprisingly, our results showed a bit more evidence against equal
 subgroups than the original meta-analysis\textsuperscript{9} (original:
@@
 this difference. We tried to clarify these discrepancies by e-mailing
-the original authors, but did not receive a response. Nonetheless, the
-broad strokes of the meta-regression confirmed that the DECREASE trials
-were the determining predictor for the effectiveness of beta-blockers
-(including DECREASE: \(RR=0.496\); excluding DECREASE: \(RR=1.28\)).
+the original authors (including a reminder after several weeks) but did
+not receive a response. Nonetheless, the broad strokes of the
+meta-regression confirmed that the DECREASE trials were the determining
+predictor for the effectiveness of beta-blockers (including DECREASE:
+\(RR=0.509\); excluding DECREASE: \(RR=1.275\)).
 
 Additionally, and exploratively, we evaluated the predictive effect of
-the type of beta-blocker used in the trials. The DECREASE trials remain
-predictive of decreased mortality (\(RR=0.496\)), whereas the
+the type of beta-blocker used in the trials. The DECREASE trials
+remained predictive of decreased mortality (\(RR=0.509\)), whereas the
 non-DECREASE trials provide tentative, but uncertain, evidence that
-atenolol results in lower mortality (\(RR=0.751\)). Nonetheless, for
+atenolol results in lower mortality (\(RR=0.777\)). Nonetheless, for
 other beta-blockers in the non-DECREASE trials, there is still tentative
 and uncertain evidence that beta-blockers increase mortality
-(bisoprolol: \(RR=2.973\); metoprolol: \(RR=1.307\); propanolol:
-\(RR=2.041\)). Table 3 shows the meta-regression results in full.
+(bisoprolol: \(RR=2.973\); metoprolol: \(RR=1.303\); propranolol:
+\(RR=1.7\)). Table 2 shows the meta-regression results in full.
 
 \begin{longtable}[]{@{}lll@{}}
-\caption{Meta-regression results for log(RR), including dummy predictors
-for DECREASE trials (reference: non-DECREASE trials) and type of
-beta-blockers used in the trial (reference: atenolol).}\tabularnewline
+\caption{Meta-regression results (k=11) for log(RR), including dummy
+predictors for DECREASE trials (reference: non-DECREASE trials) and type
+of beta-blockers used in the trial (reference:
+atenolol).}\tabularnewline
 \toprule
 & Estimate & 95\% CI\tabularnewline
-Intercept & -0.287 & -1.317; 0.743\tabularnewline
-Non-DECREASE & &\tabularnewline
-DECREASE & -1.792 & -5.082; 1.498\tabularnewline
-Atenolol & &\tabularnewline
-Bisoprolol & 1.376 & -1.996; 4.749\tabularnewline
-Metoprolol & 0.554 & -0.504; 1.613\tabularnewline
-Propanolol & 1 & -1.642; 3.643\tabularnewline
+Intercept & -0.252 & -1.228; 0.724\tabularnewline
+Non-DECREASE & - & -\tabularnewline
+DECREASE & -1.765 & -5.05; 1.519\tabularnewline
+Atenolol & - & -\tabularnewline
+Bisoprolol & 1.342 & -2.015; 4.698\tabularnewline
+Metoprolol & 0.517 & -0.49; 1.523\tabularnewline
+Propranolol & 0.783 & -1.5; 3.065\tabularnewline
 \bottomrule
@@
 probability of the observed data under a given true effect is also known
-as the veracity of the data\textsuperscript{17}. We assumed that the
+as the veracity of the data\textsuperscript{19}. We assumed that the
 non-DECREASE studies estimated the true effect distribution of
@@
 in the DECREASE trials. The estimates of the non-DECREASE studies were
-obtained from Step 1, which include the estimated log RR (i.e., 0.25),
+obtained from Step 1, which include the estimated log RR (i.e., 0.24),
 and its 95\% credibility interval as provided by the package
-\texttt{metafor} (i.e., {[}0.01; 0.484{]}). The meta-analysis model
+\texttt{metafor} (i.e., {[}0.008; 0.477{]}). The meta-analysis model
 assumes a normal distribution of population effects with the estimated
@@
 
-Figure 1 indicates that the DECREASE trials are highly unlikely under
+Figure 2 indicates that the DECREASE trials are highly unlikely under
 the estimated effect distribution based on the non-DECREASE trials. More
 specifically, the results from DECREASE-I (or more extreme) have a
-probability of \(2.9610213\times 10^{-53}\) (less than 1 in a
-quintillion) and the results from DECREASE-IV have a probability of
-\(3.0974376\times 10^{-9}\) (3 in a billion). This indicates the
-DECREASE trial results are unlikely to have come from the same
-population effect distribution as the non-DECREASE trials. Moreover,
-observing two of such extremely unlikely results jointly, as in the
-DECREASE trials, is nearly impossible, \(9.1715786\times 10^{-62}\).
-Hence, this result indicates that the DECREASE trials are severely
-different from the non-DECREASE trials.
+probability of \(6.9621238\times 10^{-45}\) (approximately 7 in a
+\href{https://en.wikipedia.org/wiki/Names_of_large_numbers}{quattuordecillion};
+\(z=-14.057\)) and the results from DECREASE-IV have a probability of
+\(6.556057\times 10^{-9}\) (approximately 7 in a billion; \(z=-5.802\)).
+This indicates that the DECREASE trial results are unlikely to have come
+from the same population effect distribution as the non-DECREASE trials.
+Moreover, observing two of such extremely unlikely results jointly, as
+in the DECREASE trials, is nearly impossible,
+\(4.564408\times 10^{-53}\). Hence, this result indicates that the
+DECREASE trials are severely different from the non-DECREASE trials.
 
@@
 homogeneity; \(\tau^2=0\)) of the effects was observed; given the small
-number of trials included (i.e., 9) this estimate is uncertain, however.
-We conducted sensitivity analyses to see how dependent results are on
-the heterogeneity estimate (not preregistered). The probability of
-observing the DECREASE trials stays approximately below 1 out of 1000
-until the variance estimate is 0.25; the probability stays approximately
-below 1 out of 100 until the variance estimate is 0.43 (see Figure 2).
-To put these numbers into context, a variance of 0.25 would suggest that
+number of trials included (i.e., 9) this estimate is highly uncertain,
+however. The total \(N\) across the non-DECREASE trials was 10529.We
+conducted sensitivity analyses to see how dependent results are on the
+heterogeneity estimate (not preregistered;
+\href{https://osf.io/vnmzc}{osf.io/vnmzc}). The probability of observing
+the DECREASE trials stays approximately below 1 out of 1000 until the
+variance estimate is 0.25; the probability stays approximately below 1
+out of 100 until the variance estimate is 0.43 (see Figure 2). To put
+these numbers into context, a variance of 0.25 would suggest that
 results of perioperative beta-blockade vary substantially due to
@@
 
-\subsection{Step 3: estimating the amount of fabricated
-data}\label{step-3-estimating-the-amount-of-fabricated-data}
+\subsection{Step 3: estimating the amount of manipulated
+data}\label{step-3-estimating-the-amount-of-manipulated-data}
 
-We estimated the number of data points that would need to be fabricated
+We estimated the number of data points that would need to be manipulated
 to arrive at the estimates from the DECREASE trials, given that the
 non-DECREASE trials represent the true effect of perioperative
-beta-blockade. In contrast to Step 2 this assumes that the DECREASE
-trials might in fact contain fabricated data. The estimates from Step 3
-provide an indication of the extent of potential data fabrication in the
-DECREASE studies\textsuperscript{4--6,9}.
+beta-blockade. In contrast to Step 2, which assumes no data manipulation
+occurred, Step 3 assumes that the DECREASE trials might in fact contain
+manipulated data. The estimates from Step 3 provide an indication of the
+extent of potential data manipulation in the DECREASE
+studies\textsuperscript{4--6,9}.
 
@@
 
-In order to estimate the number of fabricated data points, we first
-estimated the effect of perioperative beta-blockade on mortality (in log
-odds) in each trial arm. In total there are four trial arms: one per
-condition (beta-blocker or control) per trial type (DECREASE- and
-non-DECREASE trials). For each of these trial arms, we ran a
-meta-analysis applying the same methods used in Step 1. For each of the
-two DECREASE trials separately and the non-DECREASE trials combined,
-this resulted in four meta-analytic mortality estimates with
-corresponding effect variances (see Table 3; one estimate per cell).
-Throughout the simulations, we used the point estimates (i.e., fixed
-effect) to simulate genuine- and fabricated data, but supplemented this
-by using the more uncertain distribution estimates (i.e., random
-effects) as sensitivity analyses.
+In order to estimate the number of manipulated data points, we first
+estimated the probability of perioperative mortality (in log odds) in
+each trial arm for each trail stratum. As such, we estimate mortality
+odds four times: once per condition (beta-blocker or control) per trial
+type (DECREASE- and non-DECREASE trials). For all combinations of
+condition and trial type, we ran a meta-analysis applying similar
+methods used in Step 1, resulting in four meta-analytic absolute
+mortality estimates with corresponding effect variances (see Table 3;
+one estimate per cell). Throughout the simulations, we used the point
+estimates (i.e., fixed effect) to simulate genuine- and manipulated
+data, but supplemented this by using the more uncertain distribution
+estimates (i.e., random effects) as sensitivity analyses.
 
@@
 
-We applied the inversion method to estimate the number of fabricated
-data points in the DECREASE trials\textsuperscript{18}. The inversion
+We applied the inversion method to estimate the number of manipulated
+data points in the DECREASE trials\textsuperscript{20}. The inversion
 method iteratively hypothesizes that \emph{X} out of \emph{N} data
-points were fabricated (i.e., \(X={0, 1, ..., N}\)). For each
+points were manipulated (i.e., \(X={0, 1, ..., N}\)). For each
 combination of \emph{X} and trial, we simulated 10000 datasets. Each
-simulated dataset contained \emph{X} fabricated data points and
+simulated dataset contained \emph{X} manipulated data points and
 \emph{N-X} genuine data points. For each simulated dataset (exact
@@
 condition as drawn from the meta-analytic effect distribution
-(\(\pi_{C}\) indicates the mortality rate in the control condition). The
-likelihood was computed under both the fabricated effect estimates
-(i.e., \(L_{fabricated}\)) and the genuine data (i.e., \(L_{genuine}\)).
-Table 3 indicates which cell sizes the various \(n_{XX}\) refer to
-within the (simulated) data. After computing the likelihoods, we
-compared them to determine whether the simulated data were more likely
-to arise from the genuine trials (\(L_{genuine}>L_{fabricated}\)) or
-from the fabricated trials (\(L_{fabricated}>L_{genuine}\)). Note that
-comparing the likelihoods is a minor deviation from the preregistration,
-where we initially planned on using \(p\)-value comparisons
+(\(\pi_{C}\) indicates the mortality rate in the control condition). We
+estimated those parameters using the meta-analytic procedure described
+in the previous paragraph, resulting in the estimates depicted in Table
+4. The likelihood was computed under both the manipulated effect
+estimates (i.e., \(L_{manipulated}\)) and the genuine data (i.e.,
+\(L_{genuine}\)). Table 3 indicates which cell sizes the various
+\(n_{XX}\) refer to within the (simulated) data. After computing the
+likelihoods, we compared them to determine whether the simulated data
+were more likely to arise from the genuine trials
+(\(L_{genuine}>L_{manipulated}\)) or from the manipulated trials
+(\(L_{manipulated}>L_{genuine}\)). Note that comparing the likelihoods
+is a minor deviation from the preregistration, where we initially
+planned on using \(p\)-value comparisons
 (\href{https://osf.io/vnmzc}{osf.io/vnmzc}).
 
-For each hypothesis of \emph{X} out of \emph{N} fabricated data points,
-we computed the probability that the fabricated data are more likely
-than the genuine data (\(p_F=P(L_{fabricated}>L_{genuine})\)). Based on
-\(p_F\), we computed the confidence interval for the estimated \(X\)
-fabricated datapoints (i.e., \(X_{LB};X_{UB}\)). For a 95\% confidence
-interval, the lowerbound is equal to the \(p_F\) closest to .025,
-whereas the upperbound is equal to the \(p_F\) closest to .975.
+\begin{longtable}[]{@{}llll@{}}
+\caption{Mortality estimates per condition, per trial stratum. We used
+these parameters to estimate the number of manipulated data points with
+the inversion method.}\tabularnewline
+\toprule
+& & log(odds) & \(\tau\)\tabularnewline
+DECREASE (k=2) & Beta-blocker & -3.629 & 0.588\tabularnewline
+& Control & -2.498 & 1.644\tabularnewline
+Non-DECREASE (k=9) & Beta-blocker & -3.034 & 0.712\tabularnewline
+& Control & -3.208 & 1.001\tabularnewline
+\bottomrule
+\end{longtable}
 
-We computed \(p_F\) for all \emph{X} out of \emph{N} fabricated
-datapoints in 10000 randomly generated datasets, which were generated in
+For each hypothesis of \emph{X} out of \emph{N} manipulated data points,
+we computed the probability that the manipulated data are more likely
+than the genuine data (\(p_M=P(L_{manipulated}>L_{genuine})\)). Based on
+\(p_M\), we computed the confidence interval for the estimated \(X\)
+manipulated data points (i.e., \(X_{LB};X_{UB}\)). For a 95\% confidence
+interval, the lower bound is equal to the \(p_M\) closest to .025,
+whereas the upperbound is equal to the \(p_M\) closest to .975.
+
+We computed \(p_M\) for all \emph{X} out of \emph{N} manipulated data
+points in 10000 randomly generated datasets, which were generated in
 three steps. For each dataset we:
@@
   Sampled (across conditions, without replacement) \emph{X} fictitious
-  participants that would be the result of data fabrication.
+  participants that would be the result of data manipulation.
 \item
@@
 Based on the meta-analytic effect from 2 and the cell sizes from 3, we
-computed the likelihoods \(L_{fabricated}\) and \(L_{genuine}\) using
-Equation \ref{eq1}. As mentioned before, we computed \(p_F\), which
+computed the likelihoods \(L_{manipulated}\) and \(L_{genuine}\) using
+Equation \ref{eq1}. As mentioned before, we computed \(p_M\), which
 indicates the probability that the data are more likely under the
-estimates resulting from the (allegedly) fabricated data (i.e., the
+estimates resulting from the (allegedly) manipulated data (i.e., the
 DECREASE trials) than under the estimates resulting from the genuine
 data (i.e., the non-DECREASE trials;
-\(p_F=P(L_{fabricated}>L_{genuine})\)).
+\(p_M=P(L_{manipulated}>L_{genuine})\)).
 
 \subsubsection{Results}\label{results-2}
+
+For DECREASE-I (\(N=112\)), the 95\% confidence interval for the
+estimated number of manipulated data points is {[}0 - 112{]} or {[}0 -
+112{]} when based on a point estimate or a more uncertain distribution
+estimate, respectively. The left column of Figure 3 depicts the \(p_M\)
+per \emph{X} manipulated data points (top panel) and the bounds of the
+confidence interval when the degree of confidence is altered (lower
+panel). Staying clearly between the dotted lines in the top panel,
+depicting the 95\% CI (top: .975; bottom: .025), it becomes apparent
+that the degree of uncertainty is too high to make any reasonable
+estimates about the number of manipulated data points with sufficient
+confidence. This is partly due to the small sample size of the
+DECREASE-I trial (i.e., \(N=112\)) and the availability of just the
+summary results. Only when the degree of confidence is lowered to around
+75\% does the interval not span the entire sample size. As such, based
+on the summary results, little can be said about the extent of the data
+manipulation that occurred in the DECREASE-I trial, affirming the
+conclusions of the original committee report\textsuperscript{6}.
 
@@
 
-\caption{Inversion method results used to estimate the number of data points fabricated in the DECREASE-I and DECREASE-IV trials.
-The top row panels indicate $p_F$ (y-axis) for all $X$ out of $N$ fabricated data points (x-axis).
-The bottom row indicates the estimated number of fabricated data points (y-axis) when varying the degree of confidence (x-axis).
+\caption{Inversion method results used to estimate the number of data points manipulated in the DECREASE-I and DECREASE-IV trials.
+The top row panels indicate $p_M$ (y-axis) for all $X$ out of $N$ manipulated data points (x-axis).
+The bottom row indicates the estimated interval of manipulated data points (y-axis) when varying the degree of confidence (x-axis).
 Dotted lines indicate the bounds for a 95 percent CI.}\label{fig:figure 3}
@@
 
-For DECREASE-I (\(N=112\)), the 95\% confidence interval for the
-estimated number of fabricated data points is {[}0 - 112{]} or {[}2 -
-112{]} when based on a point estimate or a more uncertain distribution
-estimate, respectively. The left column of Figure 3 depicts the \(p_F\)
-per \emph{X} fabricated data points (top panel) and the bounds of the
-confidence interval when the degree of confidence is altered (lower
-panel). Staying clearly between the dotted lines in the top panel,
-depicting the 95\% CI (top: .975; bottom: .025), it becomes apparent
-that the degree of uncertainty is too high to make any reasonable
-estimates about the number of fabricated data points with sufficient
-confidence. This is partly due to the small sample size of the
-DECREASE-I trial (i.e., \(N=112\)) and the availability of just the
-summary results. Only when the degree of confidence is lowered to around
-75\% does the interval not span the entire sample size. As such, based
-on the summary results, little can be said about the extent of the data
-fabrication that occurred in the DECREASE-I trial, affirming the
-conclusions of the original committee report\textsuperscript{6}.
-
 For DECREASE-IV (\(N=1066\)), the 95\% confidence interval for the
-estimated number of fabricated data points is {[}3 - 1059{]} or {[}8 -
+estimated number of manipulated data points is {[}3 - 1066{]} or {[}10 -
 1066{]} when based on a point estimate or a more uncertain distribution
@@
 estimates indicates that there is a high degree of confidence that data
-fabrication did occur based on the difference of the trial results
-alone. Nonetheless, the range of potentially fabricated data points is
+manipulation did occur based on the difference of the trial results
+alone. Nonetheless, the range of potentially manipulated data points is
 still estimated at approximately 1000; this indicates that the summary
-results are insufficient to provide more than an estimated lowerbound.
-This indicates that it is possible not all data were fabricated (i.e.,
+results are insufficient to provide more than an estimated lower bound.
+This indicates that it is possible not all data were manipulated (i.e.,
 \(N=1066\)), increasing the importance of well-documented data
-provenance to discern between genuine and falsified data. This affirms
-the conclusion of the scientific integrity committee that the results of
-the DECREASE-IV trial are ``scientifically incorrect''
-(p.11)\textsuperscript{5}.
+provenance to discern between genuine and falsified data.
 
@@
 The effect of beta-blockade on perioperative mortality was already
-unclear following the findings of scientific misconduct; these results
-strongly affirm that the empirical results from the DECREASE trials
-should be neglected when assessing the effectiveness of beta-blockers.
-Our results indicate that the results from the DECREASE trials are
-nearly impossible to have arisen from the same effect inspected by the
-non-DECREASE trials, except when we assume at least some of the data
-were manipulated. As such, the scientific validity of the DECREASE-I and
-DECREASE-IV trials should by now be clear: they are to be regarded as
-irrelevant when assessing the effectiveness of beta-blockade on
-perioperative mortality. Nonetheless, the original papers that presented
-these trial results are not yet retracted\textsuperscript{2,3}, despite
-the integrity reports\textsuperscript{4--6}.
+unclear based on the investigations regarding scientific integrity; our
+results strongly affirm that the empirical evidence from the DECREASE
+trials is highly discrepant from other trials supposedly studying the
+same effect (i.e., the effectiveness of beta-blockers in decreasing
+perioperative mortality). Our results indicate that the results from the
+DECREASE trials are nearly impossible to have arisen from the same
+effect inspected by the non-DECREASE trials, except when we assume at
+least some of the data were manipulated. As such, the scientific
+validity of the DECREASE-I and DECREASE-IV trials should be regarded as
+highly problematic and untrustworthy when assessing the effectiveness of
+beta-blockade on perioperative mortality if they truly investigate the
+same effects as the non-DECREASE trials, as is often
+assumed\textsuperscript{9}. Nonetheless, the original papers that
+presented these trial results are not yet
+retracted\textsuperscript{2,3}, despite the integrity
+reports\textsuperscript{4--6}.
 
-The ESC/ESA and ACC/AHA guidelines\textsuperscript{19,20} on
+Our approach to estimating the number of manipulated data points has one
+major limitation that we would like to highlight: multiplicity. For each
+estimated proportion of manipulated data points, there is another
+smaller (or larger) proportion with more (or less) extremely manipulated
+data points. This problem is similar to how various samples can give
+rise to the same mean, but contain vastly different individual scores
+within them (e.g., -2.5 and +2.5 versus -100 and +100; both give the
+mean zero). Nonetheless, this limitation does still allow us to estimate
+whether any data manipulation occurred because there is no multiplicity
+in not manipulating data.
+
+The ESC/ESA and ACC/AHA guidelines\textsuperscript{11,21} on
 perioperative beta-blockade already excluded the DECREASE trials in
@@
 excluded. However, upon close inspection of the reference lists, the
-ACC/AHA guidelines still cites four trials as evidence
-base\textsuperscript{2,21--23}, of which two were already inspected by
-the scientific integrity committees of Erasmus MC\textsuperscript{2,21}.
-In the ACC/AHA guidelines, the following is said about studies conducted
-by Poldermans:
+ACC/AHA guidelines still cites four trials including Poldermans as
+author as evidence for the guidelines\textsuperscript{2,22--24}, of
+which two were already inspected by the scientific integrity committees
+of Erasmus MC\textsuperscript{2,22}. In the ACC/AHA guidelines, the
+following is said about studies conducted by Poldermans:
 
@@
 recommendation but should not form the basis of that recommendation or
-be used as a reference for the recommendation.''}\textsuperscript{20}
+be used as a reference for the recommendation.''}\textsuperscript{11}
 \end{quote}
@@
 patients, we pose that investigations should be initiated into works
-where Poldermans was clearly involved and which were not cleared by the
+where Poldermans was involved and which were not cleared by the
 scientific committees of Erasmus MC in their misconduct investigations.
 Especially those papers cited as evidence in the ACC/AHA guidelines
-should be investigated, considering they affect patients and their
+should be investigated, considering that they affect patients and their
 treatment directly.
@@
 unfeasible due to the lack of raw data; here we indicate methods that do
-make it feasible. Based on just event-count data and comparable trials,
-we were able to estimate whether part of the data were in fact
-fabricated and whether the results were within reason of comparable
-trials. The results clearly indicated they were not.
+make it feasible. Based on just event-count data and trials that
+supposedly investigate the same effect, we were able to estimate whether
+part of the data were in fact manipulated and whether the results were
+within reason of trials investigating the same effect. The results
+clearly indicated they were not such reason.
 
@@
 availability of the raw data, summary results from larger samples allow
-for better estimates of the number of fabricated data points when
+for better estimates of the number of manipulated data points when
 similar trials are available. Moreover, larger trials result in
 relatively more certainty (e.g., DECREASE-IV) about the estimated number
-of fabricated data points, when using the inversion method, compared to
+of manipulated data points, when using the inversion method, compared to
 smaller trials (e.g., DECREASE-I). This increased certainty is due to
-decreased standard errors at the population level of the estimated
-effects, resulting in higher sensitivity to data anomalies. Nonetheless,
-much residual uncertainty remains and simply fewer information is
-available in the summary results. Raw data availability would improve
-the options open to detect potential anomalies (note: raw data are
-available for DECREASE VI, but upon a freedom of information request by
-the first author, Erasmus MC refused to share these data). The results
-also highlight that in order to prevent detection, it would be in the
-fabricators' interest to fabricate small and unreliable studies even if
-raw data are hidden away (assuming the fabricator wants to remain
-undetected), which ultimately detracts from the value of such a study
-and hence the reward for fabrication.
+decreased standard errors of the estimated effects, resulting in higher
+sensitivity to data anomalies. Nonetheless, much residual uncertainty
+remains and simply less information is available in summary results when
+compared to raw data. As such, raw data availability would improve the
+options open to detect potential anomalies (note: raw data are available
+for DECREASE VI, but upon a freedom of information request by the first
+author, Erasmus MC refused to share these data; see
+\href{https://osf.io/fswjt/}{osf.io/fswjt/}). The results also highlight
+that in order to prevent detection, it would be in the manipulator's
+interest to fabricate small and imprecise studies (assuming the
+manipulator wants to remain undetected), which ultimately detracts from
+the scientific value of such a study and hence the individual reward for
+manipulation through reduced impact (hopefully).
 
@@
 decrease perioperative mortality, whereas the others (metoprolol,
-propanolol, bisoprolol) increase perioperative mortality. However, there
-seems to be covariation with respect to treatment administration,
+propranolol, bisoprolol) increase perioperative mortality. However,
+there seems to be covariation with respect to treatment administration,
 duration, and dose, which further confounds whether the treatment effect
 is due to type of beta-blocker or due to one of these other parameters.
-This affirms the statement from the ESC/ESA guidelines that \emph{``high
-priority needs to be given to new randomized clinical trials to better
-identify which patients derive benefit from beta-blocker therapy in the
+There are too few studies (\(k=11\)) too properly discern the various
+treatments from each other, requiring a new randomized trial with high
+statistical power to determine moderating factors (if any). This affirms
+the statement from the ESC/ESA guidelines that \emph{``high priority
+needs to be given to new randomized clinical trials to better identify
+which patients derive benefit from beta-blocker therapy in the
 perioperative setting, and to determine the optimal method of
-beta-blockade''}\textsuperscript{19}.
+beta-blockade''}\textsuperscript{21}.
 
 Moreover, the DECREASE and non-DECREASE trials seem to apply
-beta-blockade from different conceptual viewpoints. The non-DECREASE
-trials seem to focus purely on the application of beta-blockers in
-itself, whereas the DECREASE trials use beta-blockade as a proxy to
-decrease resting heart rate\textsuperscript{2,3}. As such, the DECREASE
-studies applied beta-blockade at least a week in advance, specifically
-in order to lower patient's resting heart rate to \textless{}70BPM and
-potentially habituate the patient to the effects of the beta-blockade.
-Other studies apply the beta-blockade just prior to the surgery
-(maximum: one day prior) and therefore seem to regard the treatment
-specifically and not the proxy of lowered BPM. As such, the differences
-between the DECREASE and non-DECREASE trials might also in part be a
-consequence of the different approaches in the various trials. Whether
-these differences matter in treatment decisions is worthy of further
-research in a clinical trial with high statistical power to find such
-differences.
+beta-blockade from different conceptual viewpoints that could confound
+the effectiveness of beta-blockade. The non-DECREASE trials seem to
+focus purely on the application of beta-blockers in itself, whereas the
+DECREASE trials use beta-blockade as a proxy to decrease resting heart
+rate\textsuperscript{2,3}. As such, the DECREASE studies applied
+beta-blockade at least a week in advance, specifically in order to lower
+patient's resting heart rate to \textless{}70BPM and potentially
+habituate the patient to the effects of the beta-blockade. Other studies
+apply the beta-blockade just prior to the surgery (maximum: one day
+prior) and therefore seem to regard the treatment specifically and not
+the proxy of lowered BPM. As such, the differences between the DECREASE
+and non-DECREASE trials might also in part be a consequence of the
+different approaches in the various trials. Whether these differences
+matter in treatment decisions is worthy of further research in a
+clinical trial with high statistical power to find such differences.
 
@@
 evidence that at least some data points were manipulated. However, these
-differences might also be due to different conceptual approaches to how
-beta-blockade might prevent mortality in non-cardiac surgery. We
+differences might also be due to different conceptual approaches as to
+how beta-blockade might prevent mortality in non-cardiac surgery. We
 recommend renewed investigations into Poldermans' work given these
@@
 rates.
+
+\section{Session info}\label{session-info}
+
+\begin{verbatim}
+## R version 3.4.0 (2017-04-21)
+## Platform: x86_64-redhat-linux-gnu (64-bit)
+## Running under: Fedora 25 (Workstation Edition)
+## 
+## Matrix products: default
+## BLAS/LAPACK: /usr/lib64/R/lib/libRblas.so
+## 
+## locale:
+##  [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C              
+##  [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8    
+##  [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8   
+##  [7] LC_PAPER=en_US.UTF-8       LC_NAME=C                 
+##  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
+## [11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C       
+## 
+## attached base packages:
+## [1] stats     graphics  grDevices utils     datasets  methods   base     
+## 
+## other attached packages:
+## [1] knitr_1.16      latex2exp_0.4.0 metafor_2.0-0   Matrix_1.2-9   
+## 
+## loaded via a namespace (and not attached):
+##  [1] Rcpp_0.12.11    lattice_0.20-35 digest_0.6.12   rprojroot_1.2  
+##  [5] grid_3.4.0      nlme_3.1-131    backports_1.1.0 magrittr_1.5   
+##  [9] evaluate_0.10.1 highr_0.6       stringi_1.1.5   rmarkdown_1.6  
+## [13] tools_3.4.0     stringr_1.2.0   yaml_2.1.14     compiler_3.4.0 
+## [17] htmltools_0.3.6
+\end{verbatim}
 
@@
 
+\hypertarget{ref-Fleisher_2014}{}
+11. Fleisher LA, Fleischmann KE, Auerbach AD, et al. 2014 ACC/AHA
+guideline on perioperative cardiovascular evaluation and management of
+patients undergoing noncardiac surgery: Executive summary: A report of
+the american college of cardiology/american heart association task force
+on practice guidelines. \emph{Circulation}. 2014;130(24):2215-2245.
+doi:\href{https://doi.org/10.1161/cir.0000000000000105}{10.1161/cir.0000000000000105}.
+
 \hypertarget{ref-Buyse1999-jq}{}
-11. Buyse M, George SL, Evans S, et al. The role of biostatistics in the
+12. Buyse M, George SL, Evans S, et al. The role of biostatistics in the
 prevention, detection and treatment of fraud in clinical trials.
@@
 
+\hypertarget{ref-10.1111ux2fanae.13938}{}
+13. Carlisle JB. Data fabrication and other reasons for non-random
+sampling in 5087 randomised, controlled trials in anaesthetic and
+general medical journals. \emph{Anaesthesia}. June 2017.
+doi:\href{https://doi.org/10.1111/anae.13938}{10.1111/anae.13938}.
+
 \hypertarget{ref-Knepper2016-la}{}
-12. Knepper D, Lindblad AS, Sharma G, et al. Statistical monitoring in
+14. Knepper D, Lindblad AS, Sharma G, et al. Statistical monitoring in
 clinical trials: Best practices for detecting data anomalies suggestive
@@
 \hypertarget{ref-klei2015}{}
-13. Klei W van. Welke perioperatieve bètablokker heeft de voorkeur?
+15. Klei W van. Welke perioperatieve bètablokker heeft de voorkeur?
 {[}Which perioperative beta-blocker is preferred?{]}. \emph{Nederlands
@@
 \hypertarget{ref-viechtbauer2010}{}
-14. Viechtbauer W. Conducting meta-analyses in R with the metafor
+16. Viechtbauer W. Conducting meta-analyses in R with the metafor
 package. \emph{Journal of Statistical Software}. 2010;36:1-48.
@@
 \hypertarget{ref-viechtbauer2005}{}
-15. Viechtbauer W. Bias and efficiency of meta-analytic variance
+17. Viechtbauer W. Bias and efficiency of meta-analytic variance
 estimators in the Random-Effects model. \emph{Journal of educational and
@@
 \hypertarget{ref-agresti2002}{}
-16. Agresti A. \emph{Categorical Data Analysis}. Hoboken, NJ: John Wiley
+18. Agresti A. \emph{Categorical Data Analysis}. Hoboken, NJ: John Wiley
 \& Sons Inc.; 2002.
@@
 \hypertarget{ref-peters2015}{}
-17. Peters CFW, Klaassen CAJ, Wiel MA van de. \emph{Evaluating the
+19. Peters CFW, Klaassen CAJ, Wiel MA van de. \emph{Evaluating the
 Scientific Veracity of Publications by Dr. Jens Forster}. University of
@@
 \hypertarget{ref-casella2002}{}
-18. Casella G, Berger RL. \emph{Statistical Interference}. Pacific
+20. Casella G, Berger RL. \emph{Statistical Interference}. Pacific
 Grove, CA: Duxbury; 2002.
@@
 \hypertarget{ref-Kristensen_2014}{}
-19. Kristensen SD, Knuuti J, Saraste A, et al. 2014 ESC/ESA guidelines
+21. Kristensen SD, Knuuti J, Saraste A, et al. 2014 ESC/ESA guidelines
 on non-cardiac surgery. \emph{European Journal of Anaesthesiology}.
@@
 
-\hypertarget{ref-Fleisher_2014}{}
-20. Fleisher LA, Fleischmann KE, Auerbach AD, et al. 2014 ACC/AHA
-guideline on perioperative cardiovascular evaluation and management of
-patients undergoing noncardiac surgery: Executive summary: A report of
-the american college of cardiology/american heart association task force
-on practice guidelines. \emph{Circulation}. 2014;130(24):2215-2245.
-doi:\href{https://doi.org/10.1161/cir.0000000000000105}{10.1161/cir.0000000000000105}.
-
 \hypertarget{ref-Boersma_2001}{}
-21. Boersma E, Poldermans D, Bax JJ, et al. Predictors of cardiac events
+22. Boersma E, Poldermans D, Bax JJ, et al. Predictors of cardiac events
 after major vascular surgery. \emph{JAMA}. 2001;285(14):1865.
@@
 \hypertarget{ref-van_Kuijk_2009}{}
-22. Kuijk J-P van, Flu W-J, Schouten O, et al. Timing of noncardiac
+23. Kuijk J-P van, Flu W-J, Schouten O, et al. Timing of noncardiac
 surgery after coronary artery stenting with bare metal or drug-eluting
@@
 \hypertarget{ref-Flu_2010}{}
-23. Flu W-J, Kuijk J-P van, Chonchol M, et al. Timing of pre-operative
+24. Flu W-J, Kuijk J-P van, Chonchol M, et al. Timing of pre-operative
 beta-blocker treatment in vascular surgery patients. \emph{Journal of

### 2017-07-27 submission/manuscript.Rmd
Revision round gazilion

@@
 title: "Analyzing DECREASE trials to estimate evidence of data manipulation"
-author: "Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper, Markus Hollman"
+author: "[Chris H.
+J.
+Hartgerink](https://orcid.org/0000-0003-1050-6809), E.
+M.
+Kemper, [Markus W.
+Hollman](https://orcid.org/0000-0001-8248-0244), [Gerben ter Riet](https://orcid.org/0000-0002-2231-7637)"
 date: '`r format(Sys.time(), "%d %B, %Y")`'
@@
 
+# Abstract
+
+The effect of beta-blockers on perioperative mortality in non-cardiac surgery has been controversial due to concerns regarding the scientific integrity of the DECREASE-I and DECREASE-IV trials.
+In this report, we investigate these DECREASE trials and model their veracity (i.e., the probability that these effects or more extreme occurred naturally) and estimate how many data points might have been manipulated in the DECREASE trials.
+Our research indicates that the DECREASE trials are nearly impossible if we assume they investigate the same effect as the non-DECREASE trials and under that assumption, our results provide evidence that at least some data points were manipulated.
+However, these differences might also be due to different conceptual approaches as to how beta-blockade might prevent mortality in non-cardiac surgery.
+Considering this, we recommend new and more extensively controlled, confirmatory trials to determine whether there is any use in administering beta-blockers in order to decrease perioperative mortality.
+
+# Author note
+
+All manuscript materials are available at https://github.com/chartgerink/2015poldermans and are preserved at Zenodo [@].
+
 ```{r, echo = FALSE}
@@
 ```
+
+# Introduction
 
@@
 Moreover, the latest guidelines still recommend the usage of beta-blockers in the perioperative period in certain cases [@esc2014;@Fleisher_2014], where some of these guidelines are based on other work by the PI of the DECREASE trials.
-Considering the potential harmful consequences of guidelines based on work by someone who has been repeatedly investigated for breaching scientific integrity, we aim to estimate the extent of data manipulation in the DECREASE studies [@poldermans1999;@dunkelgrun2009;@commissie2011;@commissie2012;@commissie2013] to further stimulate the debate on using beta-blockers in the perioperative period patients undergoing non-cardiac surgery.
+Considering the potential harmful consequences of guidelines based on work by someone who has been repeatedly investigated for breaching scientific integrity, we aim to estimate the extent of data manipulation in the DECREASE studies [@poldermans1999;@dunkelgrun2009;@commissie2011;@commissie2012;@commissie2013] to further stimulate the debate on using beta-blockers in the perioperative period for patients undergoing non-cardiac surgery.
 
@@
 Provenance is primarily concerned with the origins of the data, verifying things such as (but not limited to) the informed consent and whether data corresponded to patient files.
-The committee reports did not neglect statistical evaluation however: according to the report a statistical expert evaluated the applicability of forensic statistical methods [@commissie2013] to evaluate results of trials separately (i.e., DECREASE-I, DECREASE-IV) although it lacks details as to how this evaluation took place.
+The committee reports did not neglect statistical evaluation however: according to the report a statistical expert evaluated the applicability of forensic statistical methods [@commissie2013] to evaluate results of trials separately (i.e., DECREASE-I, DECREASE-IV) although the report lacks details as to how this evaluation took place.
 The expert concluded that the previously applied methods by him were not applicable for use in this case.
@@
 We also included type of beta-blocker to inspect whether this is predictive of the effect of beta-blockers on perioperative mortality.
-Second, we evaluated the probability that the DECREASE trials occurred assuming no data manipulation.
-Third, we reversed this assumption and assumed that data manipulation did occur and estimated how many data points would have to be manipulated in order to reproduce the results of the DECREASE trials, if the other published trials are regarded as estimating the true effect of beta-blockers on perioperative mortality in patients undergoing non-cardiac surgery.
+Second, we evaluated the probability that the DECREASE trials (or more extreme effects) arose from the same effect distribution as the non-DECREASE trials, which are assumed to be the true effect of beta-blockers on perioperative mortality in patients undergoing non-cardiac surgery.
+Third, we estimated how many data points would have to be manipulated in order to reproduce the results of the DECREASE trials if the initial, non-manipulated results arose from the effect estimates as obtained from the non-DECREASE trials.
 Considering the committees investigating the scientific integrity of the DECREASE trials was unable to assess this, we consider it worthwhile to investigate this further.
@@
 
-We computed the log risk ratio (i.e., log RR) for each study and pooled these using the `R` package `metafor` [@viechtbauer2010].
+We computed the log risk ratio (i.e., log RR) for each study and pooled these using `v2.0.0` of the `R` package `metafor` [@viechtbauer2010].
 We estimated a weighted random-effects model using the restricted maximum-likelihood estimator (i.e., `REML`) [@viechtbauer2005] to estimate the variance of effects.
 We used the default weighting procedure in the `metafor` package.
-We added 0.5 to each cell count, as is common in meta-analyses on risk- and odds ratios in order to prevent computational artefacts [@agresti2002; p397].
-The 2014 meta-analysis [@bouri2014] did not specify the variance estimate used; hence, minor discrepancies between our estimates and the original estimates could be due to differences in the estimation procedure.
+We added 0.5 to each cell count, as is common in meta-analyses on risk- and odds ratios in order to prevent computational artefacts [@agresti2002].
+<!--p397-->
+The 2014 meta-analysis [@bouri2014] did not specify the variance estimate used, hence, (minor) discrepancies between our estimates and the original estimates could be due to differences in the estimation procedure.
 
@@
 differentiated between the estimates from the non-DECREASE trials ($k=`r sum(bouri_data$decrease == 0)`$) and the DECREASE trials ($k=`r sum(bouri_data$decrease == 1)`$).
-We confirmed the effect size estimates and the variance estimates for both the non-DECREASE- and the DECREASE trials, except for some minor discrepancies at the second decimal level.
+We confirmed the effect size estimates and the variance estimates for both the non-DECREASE- and the DECREASE trials, except for some discrepancies at the second decimal level for the estimated effect sizes and a somewhat larger difference between the variance estimates of the DECREASE studies.
 Table `r tabnr` depicts the original and reproduced values for both sets of studies.
@@
 table1[1,3] <- 'Risk ratio'
-table1[1,4] <- '$\\tau^2$'
-table1[1,5] <- 'Confidence interval'
+table1[1,5] <- '$\\tau^2$'
+table1[1,4] <- 'RR Confidence interval'
 table1[2,1] <- 'Non-DECREASE (k=9)'
@@
 table1[2,3] <- 1.27
-table1[2,4] <- 0.00
-table1[2,5] <- '[1.01; 1.60]'
+table1[2,5] <- 0.00
+table1[2,4] <- '[1.01; 1.60]'
 
 table1[3,3] <- round(exp(mod_clean$b), 2)
-table1[3,4] <- round(mod_clean$tau2, 2)
-table1[3,5] <- sprintf('[%s; %s]', round(exp(mod_clean$ci.lb), 2), round(exp(mod_clean$ci.ub), 2))
+table1[3,5] <- round(mod_clean$tau2, 2)
+table1[3,4] <- sprintf('[%s; %s]', round(exp(mod_clean$ci.lb), 2), round(exp(mod_clean$ci.ub), 2))
 
 table1[4,3] <- 0.42
-table1[4,4] <- 0.29
-table1[4,5] <- '[0.15; 1.23]'
+table1[4,5] <- 0.29
+table1[4,4] <- '[0.15; 1.23]'
 
 table1[5,3] <- round(exp(mod_dirty$b), 2)
-table1[5,4] <- round(mod_dirty$tau2, 2)
-table1[5,5] <- sprintf('[%s; %s]', round(exp(mod_dirty$ci.lb), 2), round(exp(mod_dirty$ci.ub), 2))
+table1[5,5] <- round(mod_dirty$tau2, 2)
+table1[5,4] <- sprintf('[%s; %s]', round(exp(mod_dirty$ci.lb), 2), round(exp(mod_dirty$ci.ub), 2))
 
@@
 Second, we meta-analyzed all studies combined, including a dummy predictor for the DECREASE and non-DECREASE studies to reproduce results presented in Figure 4 of the 2014 meta-analysis [@bouri2014].
-Surprisingly, our results showed a bit more evidence against equal subgroups than the original meta-analysis [@bouri2014] (original: $\chi^2(1)=3.91,p=.05$; reproduced: $\chi^2(1)=`r round(mod_mods$QM, 2)`,p=`r round(mod_mods$QMp, 3)`$).
+Our results showed a bit more evidence against equal subgroups than the original meta-analysis [@bouri2014] (original: $\chi^2(1)=3.91,p=.05$; reproduced: $\chi^2(1)=`r round(mod_mods$QM, 2)`,p=`r round(mod_mods$QMp, 3)`$).
 Additionally, the original analyses showed substantial residual heterogeneity ($I^2=74.4$\%) whereas we found no residual heterogeneity ($I^2=`r round(i2_mods, 3)`$\%).
@@
 Additionally, and exploratively, we evaluated the predictive effect of the type of beta-blocker used in the trials.
-The DECREASE trials remained predictive of decreased mortality ($RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[2] + mod_mods_type$b[3]), 3)`$), <!-- DECREASE only uses bisoprolol!
+Descriptively, the DECREASE trials remained predictive of decreased mortality ($RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[2] + mod_mods_type$b[3]), 3)`$), <!-- DECREASE only uses bisoprolol!
 -->
-whereas the non-DECREASE trials provide tentative, but uncertain, evidence that atenolol results in lower mortality ($RR=`r round(exp(mod_mods_type$b[1]), 3)`$).
-Nonetheless, for other beta-blockers in the non-DECREASE trials, there is still tentative and uncertain evidence that beta-blockers increase mortality (bisoprolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[3]), 3)`$; metoprolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[4]), 3)`$; propranolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[5]), 3)`$).
+whereas the non-DECREASE trials provide tentative evidence that atenolol results in lower mortality ($RR=`r round(exp(mod_mods_type$b[1]), 3)`$).
+Nonetheless, for other beta-blockers in the non-DECREASE trials, there is descriptive evidence that beta-blockers could increase mortality (bisoprolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[3]), 3)`$; metoprolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[4]), 3)`$; propranolol: $RR=`r round(exp(mod_mods_type$b[1] + mod_mods_type$b[5]), 3)`$).
 Table `r tabnr` shows the meta-regression results in full.
+We do note that the DECREASE studies only use bisoprolol and any estimates for other beta-blockers are extrapolations.
 
@@
 
-Based on the non-DECREASE estimates from Step 1, we estimated the probability of obtaining the results in the DECREASE trials.
-To this end, we assumed that the non-DECREASE trials provide a valid representation of the true effect of beta-blockers on perioperative mortality (similar to Bouri et al.
-[@bouri2014]).
-The estimated probability of the observed data under a given true effect is also known as the veracity of the data [@peters2015].
+Based on the effect estimates for the non-DECREASE trials from Step 1, we estimated the probability that the observed effects from the DECREASE studies (or more extreme) occurred naturally.
 We assumed that the non-DECREASE studies estimated the true effect distribution of perioperative beta-blockade on mortality, not perturbed by publication bias due to statistical (non)significance.
 Publication bias was assumed to not be a problem because a substantial number of nonsignificant effects are included in the dataset (9 of 11 results are nonsignificant).
+Based on this effect distribution, we estimated the veracity of the DECREASE trials separately, which is the estimated probability of the observed data under a given true effect [@peters2015].
 
@@
 mod_dirty_pred <- predict(mod_dirty)
-```
 
-Based on the estimated mean log RR and its credible interval in the non-DECREASE studies, we computed the probability of the observed log RR in the DECREASE trials.
-The estimates of the non-DECREASE studies were obtained from Step 1, which include the estimated log RR (i.e., `r round(mod_clean$b, 2)`), and its 95\% credibility interval as provided by the package `metafor` (i.e., `r sprintf('[%s; %s]', round(mod_clean_pred$cr.lb, 3), round(mod_clean_pred$cr.ub, 3))`).
-The meta-analysis model assumes a normal distribution of population effects with the estimated effect as the mean of the distribution.
-The 95\% credibility interval denotes the bounds of the normal distribution that covers 95\% of the density, where the standard deviation is calculated as the distance from the mean to either bound, divided by 1.96.
-This allows for an approximation of the population effect distribution, as depicted in Figure `r fignr`.
+var1 <- mod_clean_pred$se^2 / sum(bouri_data$decrease == 0)
+var_decrI <- (bouri_data$se_ln_rr[bouri_data$decrease == 1][1]^2 / sum(bouri_data$decrease == 0))
+var_decrIV <- (bouri_data$se_ln_rr[bouri_data$decrease == 1][2]^2 / sum(bouri_data$decrease == 0))
+decrI <- bouri_data$ln_rr[bouri_data$decrease == 1][1]
+decrIV <- bouri_data$ln_rr[bouri_data$decrease == 1][2]
+mu_ma <- mod_clean_pred$pred
 
-```{r, echo = FALSE}
-temp <- seq(-3, 3, .001)
+df_I <- 1 + 9 - 2
+t_I <- (decrI - mu_ma) / sqrt(var1 + var_decrI)
+p_I <- pt(abs(t_I), df = df_I, lower.tail = FALSE) * 2
 
-pdf(sprintf('%sfigures/fig1.pdf', x12), width = 8, height = 6)
-par(mar = c(5,0,0,0))
-plot(x = temp,
-     dnorm(x = temp,
-           mod_clean$b,
-           (mod_clean_pred$pred - mod_clean_pred$cr.lb)/1.96),
-     type = 'l',
-     bty = 'n',
-     # xaxt = 'n',
-     yaxt = 'n',
-     xlab = 'log(RR)',
-     ylab = '')
-
-points(x = bouri_data$ln_rr[bouri_data$decrease == 1][1], y = 0)
-text(x = bouri_data$ln_rr[bouri_data$decrease == 1][1],
-     y = .15,
-     labels = 'DECREASE-I',
-     cex = .8)
-
-points(x = bouri_data$ln_rr[bouri_data$decrease == 1][2], y = 0)
-text(x = bouri_data$ln_rr[bouri_data$decrease == 1][2] - .05,
-     y = .15,
-     labels = 'DECREASE-IV',
-     cex = .8)
-
-invisible(dev.off())
-```
-
-```{r figure 1, fig.cap = 'Density plot of the estimated true effect distribution based on the non-DECREASE studies only, with the position of the DECREASE studies highlighted.', out.width = '80%', fig.asp = .5, fig.align = 'center', echo = FALSE}
-knitr::include_graphics(sprintf("%sfigures/fig1.pdf", x12))
-fignr <- fignr + 1
+df_IV <- 1 + 9 - 2
+t_IV <- (decrIV - mu_ma) / sqrt(var1 + var_decrIV)
+p_IV <- pt(abs(t_IV), df = df_IV, lower.tail = FALSE) * 2
 ```
@@
 Based on the estimated effect distribution from the non-DECREASE trials, we calculated the probability of each DECREASE trial result, or a more extreme result.
-In other words, we computed the $p$-value for the null hypothesis that the DECREASE trials arise from the same effect distribution as the non-DECREASE trials.
-This assumes that the information available from the other trials is informative of the true population effect.
+In other words, we computed the two-tailed $p$-value for the null hypothesis that the DECREASE trials arose from the same effect distribution as the non-DECREASE trials ($H_0:\mu_{\bar{x}_1-\bar{x}_2}=0$).
+To this end, we applied a Welch t-test [@doi:10.1093/biomet/34.1-2.28].
+As means we used the observed log RR for the DECREASE trials (i.e., DECREASE-I: `r round(decrI, 3)`; DECREASE-IV: `r round(decrIV, 3)`) and the meta-analyzed log RR for the non-DECREASE trials (i.e., `r round(mu_ma, 3)`).
+As standard deviations we used the standard error for the DECREASE trials (i.e., DECREASE-I: `r round(var_decrI, 3)`; DECREASE-IV: `r round(var_decrIV, 3)`) and the standard error of the estimated log RR for the non-DECREASE trials (i.e., `r round(var1, 3)`).
+We initially preregistered that the DECREASE trials would be regarded as fixed in the computation of the veracity, which was erroneous because these also have their own standard error; hence, we applied the Welch test to take into account the uncertainty in the estimates of both the DECREASE and non-DECREASE trials.
 
@@
 Figure `r fignr` indicates that the DECREASE trials are highly unlikely under the estimated effect distribution based on the non-DECREASE trials.
-More specifically, the results from DECREASE-I (or more extreme) have a probability of $`r prob[1]`$ (approximately 7 in a [quattuordecillion](https://en.wikipedia.org/wiki/Names_of_large_numbers); $z=`r round(zval[1], 3)`$) and the results from DECREASE-IV have a probability of $`r prob[2]`$ (approximately 7 in a billion; $z=`r round(zval[2], 3)`$).
+More specifically, the results from DECREASE-I (or more extreme) have a probability of 1 in 10 000 ($t(`r df_I`)=`r round(t_I, 3)`,p=`r p_I`$) and the results from DECREASE-IV (or more extreme) have a probability of 1 in 1000 ($t(`r df_IV`)=`r round(t_IV, 3)`,p=`r p_IV`$).
 This indicates that the DECREASE trial results are unlikely to have come from the same population effect distribution as the non-DECREASE trials.
-Moreover, observing two of such extremely unlikely results jointly, as in the DECREASE trials, is nearly impossible, $`r prob[1]*prob[2]`$.
+Moreover, observing two of such extremely unlikely results jointly, as in the DECREASE trials, would occur in only 2 out of 10 million sets of two trials (i.e., $`r p_I*p_IV`$) according to this model.
 Hence, this result indicates that the DECREASE trials are severely different from the non-DECREASE trials.
@@
 Results from Step 1 indicated that no between-trial variance (i.e., homogeneity; $\tau^2=0$) of the effects was observed; given the small number of trials included (i.e., `r mod_clean$k`) this estimate is highly uncertain, however.
-The total $N$ across the non-DECREASE trials was `r sum(bouri_data$control_n[bouri_data$decrease == 0] + bouri_data$beta_n[bouri_data$decrease == 0])`.We conducted sensitivity analyses to see how dependent results are on the heterogeneity estimate (not preregistered; [osf.io/vnmzc](https://osf.io/vnmzc)).
-The probability of observing the DECREASE trials stays approximately below 1 out of 1000 until the variance estimate is 0.25; the probability stays approximately below 1 out of 100 until the variance estimate is 0.43 (see Figure `r fignr`).
+The total $N$ across the non-DECREASE trials was `r sum(bouri_data$control_n[bouri_data$decrease == 0] + bouri_data$beta_n[bouri_data$decrease == 0])`.
+We conducted sensitivity analyses to see how dependent results are on the heterogeneity estimate (not preregistered; [osf.io/vnmzc](https://osf.io/vnmzc)).
+Fixing the variance estimate $\tau^2$ to .5, indicates that the probability of observing the DECREASE trials jointly is approximately 1.2 out of 100 000 (see Figure `r fignr`).
 To put these numbers into context, a variance of 0.25 would suggest that results of perioperative beta-blockade vary substantially due to contextual circumstances of the study, even if perioperative beta-blockade has no effect whatsoever (RRs between `r round(exp(0 - .25), 3)` and `r round(exp(0 + .25), 3)` in ~64\% of the cases).
@@
   mod_clean_pred_sens <- predict(mod_clean_sens)
-  prob_sens <- pnorm(q = bouri_data$ln_rr[bouri_data$decrease == 1],
-                     mod_clean_pred_sens$pred,
-                     (mod_clean_pred_sens$pred - mod_clean_pred_sens$cr.lb)/1.96) * 2
-  prob_sens_t[i] <- prob_sens[1] * prob_sens[2]
+
+  var1 <- mod_clean_pred_sens$se^2 / sum(bouri_data$decrease == 0)
+  mu_ma <- mod_clean_pred_sens$pred
+
+  df_I <- 1 + 9 - 2
+  t_I <- (decrI - mu_ma) / sqrt(var1 + var_decrI)
+  p_I <- pt(abs(t_I), df = df_I, lower.tail = FALSE) * 2
+
+  df_IV <- 1 + 9 - 2
+  t_IV <- (decrIV - mu_ma) / sqrt(var1 + var_decrIV)
+  p_IV <- pt(abs(t_IV), df = df_IV, lower.tail = FALSE) * 2
+  prob_sens_t[i] <- p_I* p_IV
 
@@
 
-par(mar = c(4, 4, 0, 0))
+par(mar = c(4, 4, 1, 0))
 plot(tau2, prob_sens_t, type = 'l', xlab = TeX('$\\tau^2$'),
      ylab = 'p-value', bty = 'n')
-points(x = tau2[c(26, 44)], y = prob_sens_t[c(26, 44)])
-text(x = tau2[c(26, 44)],
-     y = prob_sens_t[c(26, 44)] + .003,
-     labels = c(TeX('$\\tau^2=.25$'), TeX('$\\tau^2=.43$')),
-     cex = .8)
+# points(x = tau2[c(26, 44)], y = prob_sens_t[c(26, 44)])
+# text(x = tau2[c(26, 44)],
+#      y = prob_sens_t[c(26, 44)] + .003,
+#      labels = c(TeX('$\\tau^2=.25$'), TeX('$\\tau^2=.43$')),
+#      cex = .8)
 invisible(dev.off())
@@
 We estimated the number of data points that would need to be manipulated to arrive at the estimates from the DECREASE trials, given that the non-DECREASE trials represent the true effect of perioperative beta-blockade.
-In contrast to Step 2, which assumes no data manipulation occurred, Step 3 assumes that the DECREASE trials might in fact contain manipulated data.
+In contrast to Step 2, which assumes no data manipulation occurred and that the DECREASE trials occurred naturally from the same effect distribution as the non-DECREASE trials, Step 3 assumes that the DECREASE trials might in fact contain manipulated data.
 The estimates from Step 3 provide an indication of the extent of potential data manipulation in the DECREASE studies [@commissie2011;@commissie2012;@commissie2013;@bouri2014].
@@
 In order to estimate the number of manipulated data points, we first estimated the probability of perioperative mortality (in log odds) in each trial arm for each trail stratum.
-As such, we estimate mortality odds four times: once per condition (beta-blocker or control) per trial type (DECREASE- and non-DECREASE trials).
-For all combinations of condition and trial type, we ran a meta-analysis applying similar methods used in Step 1, resulting in four meta-analytic absolute mortality estimates with corresponding effect variances (see Table 3; one estimate per cell).
-Throughout the simulations, we used the point estimates (i.e., fixed effect) to simulate genuine- and manipulated data, but supplemented this by using the more uncertain distribution estimates (i.e., random effects) as sensitivity analyses.
-
-```{r echo=FALSE, results='asis'}
-table2 <- matrix('', ncol = 3, nrow = 3)
-table2[1,1] <- ''
-table2[1,2] <- 'Dead'
-table2[1,3] <- 'Alive'
-
-table2[2,1] <- 'Beta-blockers'
-table2[2,2] <- '$n_{11}$'
-table2[2,3] <- '$n_{12}$'
-
-table2[3,1] <- 'Control'
-table2[3,2] <- '$n_{21}$'
-table2[3,3] <- '$n_{22}$'
-
-knitr::kable(table2, caption = 'Outcome possibilities within a simulated 2 (beta-blocker v control) by 2 (dead v alive) clinical trial.')
-
-tabnr <- tabnr + 1
-```
+As such, we estimate mortality odds four times: once per condition (beta-blocker or control) per trial stratum (DECREASE- and non-DECREASE trials).
+For all four combinations of condition and trial type, we ran a meta-analysis applying similar methods used in Step 1, resulting in four meta-analytic absolute mortality estimates with corresponding effect variances.
+Throughout the simulations, we used the point estimates (i.e., fixed effect) to simulate genuine- and manipulated data, but supplemented this by using distribution estimates (i.e., random effects) as sensitivity analyses.
 
 We applied the inversion method to estimate the number of manipulated data points in the DECREASE trials [@casella2002].
-The inversion method iteratively hypothesizes that *X* out of *N* data points were manipulated (i.e., $X={0, 1, ..., N}$).
+We assumed that if data are manipulated, each data point is manipulated in the same way and to the same extent.
+The inversion method iteratively hypothesizes that *X* out of *N* data points were manipulated (i.e., $X={0, 1, ..., N}$), assuming they were manipulated in the same way.
 For each combination of *X* and trial, we simulated `r iterations` datasets.
@@
 The likelihood was computed under both the manipulated effect estimates (i.e., $L_{manipulated}$) and the genuine data (i.e., $L_{genuine}$).
-Table 3 indicates which cell sizes the various $n_{XX}$ refer to within the (simulated) data.
+Table 4 indicates which cell sizes the various $n_{XX}$ refer to within the (simulated) data.
 After computing the likelihoods, we compared them to determine whether the simulated data were more likely to arise from the genuine trials ($L_{genuine}>L_{manipulated}$) or from the manipulated trials ($L_{manipulated}>L_{genuine}$).
@@
 
+```
+
+```{r echo=FALSE, results='asis'}
+table2 <- matrix('', ncol = 3, nrow = 3)
+table2[1,1] <- ''
+table2[1,2] <- 'Dead'
+tabnr <- tabnr + 1
+table2[1,3] <- 'Alive'
+
+table2[2,1] <- 'Beta-blockers'
+table2[2,2] <- '$n_{11}$'
+table2[2,3] <- '$n_{12}$'
+
+table2[3,1] <- 'Control'
+table2[3,2] <- '$n_{21}$'
+table2[3,3] <- '$n_{22}$'
+
+knitr::kable(table2, caption = 'Outcome possibilities within a simulated 2 (beta-blocker v control) by 2 (dead v alive) clinical trial.')
+
 tabnr <- tabnr + 1
@@
 2.
-Determined the population mortality rate for each condition (i.e., for each cell as in Table 3).
+Determined the population mortality rate for each condition (i.e., for each cell based on the estimates from Table 3).
 The meta-analytic point estimate was used or a population effect was randomly drawn from the meta-analytic effect distribution.
@@
 
-The ESC/ESA and ACC/AHA guidelines [@Kristensen_2014;@Fleisher_2014] on perioperative beta-blockade already excluded the DECREASE trials in their assessment, but also state that other trials by Poldermans are excluded.
+The ESC/ESA and ACC/AHA guidelines [@Kristensen_2014;@Fleisher_2014] on perioperative beta-blockade already excluded the DECREASE trials in their assessment, but also explicitly states that other trials by Poldermans are excluded.
 However, upon close inspection of the reference lists, the ACC/AHA guidelines still cites four trials including Poldermans as author as evidence for the guidelines [@Boersma_2001;@poldermans1999;@van_Kuijk_2009;@Flu_2010], of which two were already inspected by the scientific integrity committees of Erasmus MC [@Boersma_2001;@poldermans1999].

### 2017-07-27 submission/manuscript.tex
Revision round gazilion

@@
             pdftitle={Analyzing DECREASE trials to estimate evidence of data manipulation},
-            pdfauthor={Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper, Markus Hollman},
+            pdfauthor={Chris H. J. Hartgerink, E. M. Kemper, Markus W. Hollman, Gerben ter Riet},
             pdfborder={0 0 0},
@@
   \posttitle{\par}
-  \author{Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper, Markus Hollman}
+  \author{\href{https://orcid.org/}{Chris H. J. Hartgerink}, E. M. Kemper,
+\href{https://orcid.org/0000-0001-8248-0244}{Markus W. Hollman},
+\href{https://orcid.org/0000-0002-2231-7637}{Gerben ter Riet}}
   \preauthor{\centering\large\emph}
@@
   \postdate{\par}
-  \date{14 July, 2017}
+  \date{27 July, 2017}
 
@@
 \maketitle
+
+\section{Abstract}\label{abstract}
+
+The effect of beta-blockers on perioperative mortality in non-cardiac
+surgery has been controversial due to concerns regarding the scientific
+integrity of the DECREASE-I and DECREASE-IV trials. In this report, we
+investigate these DECREASE trials and model their veracity (i.e., the
+probability that these effects or more extreme occurred naturally) and
+estimate how many data points might have been manipulated in the
+DECREASE trials. Our research indicates that the DECREASE trials are
+nearly impossible if we assume they investigate the same effect as the
+non-DECREASE trials and under that assumption, our results provide
+evidence that at least some data points were manipulated. However, these
+differences might also be due to different conceptual approaches as to
+how beta-blockade might prevent mortality in non-cardiac surgery.
+Considering this, we recommend new and more extensively controlled,
+confirmatory trials to determine whether there is any use in
+administering beta-blockers in order to decrease perioperative
+mortality.
+
+\section{Author note}\label{author-note}
+
+All manuscript materials are available at
+\url{https://github.com/chartgerink/2015poldermans} and are preserved at
+Zenodo {[}@{]}.
+
+\section{Introduction}\label{introduction}
 
@@
 stimulate the debate on using beta-blockers in the perioperative period
-patients undergoing non-cardiac surgery.
+for patients undergoing non-cardiac surgery.
 
@@
 methods\textsuperscript{6} to evaluate results of trials separately
-(i.e., DECREASE-I, DECREASE-IV) although it lacks details as to how this
-evaluation took place. The expert concluded that the previously applied
-methods by him were not applicable for use in this case. Nonetheless, in
-this report we compare across trials, which is a method that has
-previously been used to monitor trial data quality or to test for
-potential data anomalies\textsuperscript{12,13}. Moreover, comparing
+(i.e., DECREASE-I, DECREASE-IV) although the report lacks details as to
+how this evaluation took place. The expert concluded that the previously
+applied methods by him were not applicable for use in this case.
+Nonetheless, in this report we compare across trials, which is a method
+that has previously been used to monitor trial data quality or to test
+for potential data anomalies\textsuperscript{12,13}. Moreover, comparing
 across trials instead of evaluating them separately has previously
@@
 perioperative mortality. Second, we evaluated the probability that the
-DECREASE trials occurred assuming no data manipulation. Third, we
-reversed this assumption and assumed that data manipulation did occur
-and estimated how many data points would have to be manipulated in order
-to reproduce the results of the DECREASE trials, if the other published
-trials are regarded as estimating the true effect of beta-blockers on
-perioperative mortality in patients undergoing non-cardiac surgery.
-Considering the committees investigating the scientific integrity of the
-DECREASE trials was unable to assess this, we consider it worthwhile to
+DECREASE trials (or more extreme effects) arose from the same effect
+distribution as the non-DECREASE trials, which are assumed to be the
+true effect of beta-blockers on perioperative mortality in patients
+undergoing non-cardiac surgery. Third, we estimated how many data points
+would have to be manipulated in order to reproduce the results of the
+DECREASE trials if the initial, non-manipulated results arose from the
+effect estimates as obtained from the non-DECREASE trials. Considering
+the committees investigating the scientific integrity of the DECREASE
+trials was unable to assess this, we consider it worthwhile to
 investigate this further.
@@
 We computed the log risk ratio (i.e., log RR) for each study and pooled
-these using the \texttt{R} package \texttt{metafor}\textsuperscript{16}.
-We estimated a weighted random-effects model using the restricted
-maximum-likelihood estimator (i.e., \texttt{REML})\textsuperscript{17}
-to estimate the variance of effects. We used the default weighting
-procedure in the \texttt{metafor} package. We added 0.5 to each cell
-count, as is common in meta-analyses on risk- and odds ratios in order
-to prevent computational artefacts {[}18; p397{]}. The 2014
-meta-analysis\textsuperscript{9} did not specify the variance estimate
-used; hence, minor discrepancies between our estimates and the original
-estimates could be due to differences in the estimation procedure.
+these using \texttt{v2.0.0} of the \texttt{R} package
+\texttt{metafor}\textsuperscript{16}. We estimated a weighted
+random-effects model using the restricted maximum-likelihood estimator
+(i.e., \texttt{REML})\textsuperscript{17} to estimate the variance of
+effects. We used the default weighting procedure in the \texttt{metafor}
+package. We added 0.5 to each cell count, as is common in meta-analyses
+on risk- and odds ratios in order to prevent computational
+artefacts\textsuperscript{18}. The 2014 meta-analysis\textsuperscript{9}
+did not specify the variance estimate used, hence, (minor) discrepancies
+between our estimates and the original estimates could be due to
+differences in the estimation procedure.
 
@@
 effect size estimates and the variance estimates for both the
-non-DECREASE- and the DECREASE trials, except for some minor
-discrepancies at the second decimal level. Table 1 depicts the original
-and reproduced values for both sets of studies.
+non-DECREASE- and the DECREASE trials, except for some discrepancies at
+the second decimal level for the estimated effect sizes and a somewhat
+larger difference between the variance estimates of the DECREASE
+studies. Table 1 depicts the original and reproduced values for both
+sets of studies.
 
@@
 \toprule
-& & Risk ratio & \(\tau^2\) & Confidence interval\tabularnewline
-Non-DECREASE (k=9) & Original & 1.27 & 0 & {[}1.01;
-1.60{]}\tabularnewline
-& Reproduced & 1.27 & 0 & {[}1.01; 1.61{]}\tabularnewline
-DECREASE (k=2) & Original & 0.42 & 0.29 & {[}0.15;
-1.23{]}\tabularnewline
-& Reproduced & 0.47 & 0.14 & {[}0.19; 1.15{]}\tabularnewline
+\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+Risk ratio\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+RR Confidence interval\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+\(\tau^2\)\strut
+\end{minipage}\tabularnewline
+Non-DECREASE (k=9) & Original & 1.27 & {[}1.01; 1.60{]} &
+0\tabularnewline
+\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+Reproduced\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+1.27\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+{[}1.01; 1.61{]}\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+0\strut
+\end{minipage}\tabularnewline
+DECREASE (k=2) & Original & 0.42 & {[}0.15; 1.23{]} &
+0.29\tabularnewline
+\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+Reproduced\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+0.47\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+{[}0.19; 1.15{]}\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+0.14\strut
+\end{minipage}\tabularnewline
 \bottomrule
@@
 predictor for the DECREASE and non-DECREASE studies to reproduce results
-presented in Figure 4 of the 2014 meta-analysis\textsuperscript{9}.
-Surprisingly, our results showed a bit more evidence against equal
-subgroups than the original meta-analysis\textsuperscript{9} (original:
+presented in Figure 4 of the 2014 meta-analysis\textsuperscript{9}. Our
+results showed a bit more evidence against equal subgroups than the
+original meta-analysis\textsuperscript{9} (original:
 \(\chi^2(1)=3.91,p=.05\); reproduced: \(\chi^2(1)=6.12,p=0.013\)).
@@
 Additionally, and exploratively, we evaluated the predictive effect of
-the type of beta-blocker used in the trials. The DECREASE trials
-remained predictive of decreased mortality (\(RR=0.509\)), whereas the
-non-DECREASE trials provide tentative, but uncertain, evidence that
-atenolol results in lower mortality (\(RR=0.777\)). Nonetheless, for
-other beta-blockers in the non-DECREASE trials, there is still tentative
-and uncertain evidence that beta-blockers increase mortality
-(bisoprolol: \(RR=2.973\); metoprolol: \(RR=1.303\); propranolol:
-\(RR=1.7\)). Table 2 shows the meta-regression results in full.
+the type of beta-blocker used in the trials. Descriptively, the DECREASE
+trials remained predictive of decreased mortality (\(RR=0.509\)),
+whereas the non-DECREASE trials provide tentative evidence that atenolol
+results in lower mortality (\(RR=0.777\)). Nonetheless, for other
+beta-blockers in the non-DECREASE trials, there is descriptive evidence
+that beta-blockers could increase mortality (bisoprolol: \(RR=2.973\);
+metoprolol: \(RR=1.303\); propranolol: \(RR=1.7\)). Table 2 shows the
+meta-regression results in full. We do note that the DECREASE studies
+only use bisoprolol and any estimates for other beta-blockers are
+extrapolations.
 
@@
 \toprule
-& Estimate & 95\% CI\tabularnewline
+\begin{minipage}[t]{0.32\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright\strut
+Estimate\strut
+\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright\strut
+95\% CI\strut
+\end{minipage}\tabularnewline
 Intercept & -0.252 & -1.228; 0.724\tabularnewline
@@
 
-Based on the non-DECREASE estimates from Step 1, we estimated the
-probability of obtaining the results in the DECREASE trials. To this
-end, we assumed that the non-DECREASE trials provide a valid
-representation of the true effect of beta-blockers on perioperative
-mortality (similar to Bouri et al.\textsuperscript{9}). The estimated
-probability of the observed data under a given true effect is also known
-as the veracity of the data\textsuperscript{19}. We assumed that the
+Based on the effect estimates for the non-DECREASE trials from Step 1,
+we estimated the probability that the observed effects from the DECREASE
+studies (or more extreme) occurred naturally. We assumed that the
 non-DECREASE studies estimated the true effect distribution of
@@
 effects are included in the dataset (9 of 11 results are
-nonsignificant).
+nonsignificant). Based on this effect distribution, we estimated the
+veracity of the DECREASE trials separately, which is the estimated
+probability of the observed data under a given true
+effect\textsuperscript{19}.
 
@@
 
-Based on the estimated mean log RR and its credible interval in the
-non-DECREASE studies, we computed the probability of the observed log RR
-in the DECREASE trials. The estimates of the non-DECREASE studies were
-obtained from Step 1, which include the estimated log RR (i.e., 0.24),
-and its 95\% credibility interval as provided by the package
-\texttt{metafor} (i.e., {[}0.008; 0.477{]}). The meta-analysis model
-assumes a normal distribution of population effects with the estimated
-effect as the mean of the distribution. The 95\% credibility interval
-denotes the bounds of the normal distribution that covers 95\% of the
-density, where the standard deviation is calculated as the distance from
-the mean to either bound, divided by 1.96. This allows for an
-approximation of the population effect distribution, as depicted in
-Figure 1.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.8\linewidth]{../figures/fig1}
-
-}
-
-\caption{Density plot of the estimated true effect distribution based on the non-DECREASE studies only, with the position of the DECREASE studies highlighted.}\label{fig:figure 1}
-\end{figure}
-
 Based on the estimated effect distribution from the non-DECREASE trials,
 we calculated the probability of each DECREASE trial result, or a more
-extreme result. In other words, we computed the \(p\)-value for the null
-hypothesis that the DECREASE trials arise from the same effect
-distribution as the non-DECREASE trials. This assumes that the
-information available from the other trials is informative of the true
-population effect.
+extreme result. In other words, we computed the two-tailed \(p\)-value
+for the null hypothesis that the DECREASE trials arose from the same
+effect distribution as the non-DECREASE trials
+(\(H_0:\mu_{\bar{x}_1-\bar{x}_2}=0\)). To this end, we applied a Welch
+t-test\textsuperscript{20}. As means we used the observed log RR for the
+DECREASE trials (i.e., DECREASE-I: -1.44; DECREASE-IV: -0.452) and the
+meta-analyzed log RR for the non-DECREASE trials (i.e., 0.243). As
+standard deviations we used the standard error for the DECREASE trials
+(i.e., DECREASE-I: 0.061; DECREASE-IV: 0.018) and the standard error of
+the estimated log RR for the non-DECREASE trials (i.e., 0.002). We
+initially preregistered that the DECREASE trials would be regarded as
+fixed in the computation of the veracity, which was erroneous because
+these also have their own standard error; hence, we applied the Welch
+test to take into account the uncertainty in the estimates of both the
+DECREASE and non-DECREASE trials.
 
@@
 
-Figure 2 indicates that the DECREASE trials are highly unlikely under
+Figure 1 indicates that the DECREASE trials are highly unlikely under
 the estimated effect distribution based on the non-DECREASE trials. More
 specifically, the results from DECREASE-I (or more extreme) have a
-probability of \(6.9621238\times 10^{-45}\) (approximately 7 in a
-\href{https://en.wikipedia.org/wiki/Names_of_large_numbers}{quattuordecillion};
-\(z=-14.057\)) and the results from DECREASE-IV have a probability of
-\(6.556057\times 10^{-9}\) (approximately 7 in a billion; \(z=-5.802\)).
-This indicates that the DECREASE trial results are unlikely to have come
-from the same population effect distribution as the non-DECREASE trials.
-Moreover, observing two of such extremely unlikely results jointly, as
-in the DECREASE trials, is nearly impossible,
-\(4.564408\times 10^{-53}\). Hence, this result indicates that the
-DECREASE trials are severely different from the non-DECREASE trials.
+probability of 1 in 10 000 (\(t(8)=-6.75,p=0.000145\)) and the results
+from DECREASE-IV (or more extreme) have a probability of 1 in 1000
+(\(t(8)=-4.996,p=0.0010587\)). This indicates that the DECREASE trial
+results are unlikely to have come from the same population effect
+distribution as the non-DECREASE trials. Moreover, observing two of such
+extremely unlikely results jointly, as in the DECREASE trials, would
+occur in only 2 out of 10 million sets of two trials (i.e.,
+\(0.0000002\)) according to this model. Hence, this result indicates
+that the DECREASE trials are severely different from the non-DECREASE
+trials.
 
@@
 number of trials included (i.e., 9) this estimate is highly uncertain,
-however. The total \(N\) across the non-DECREASE trials was 10529.We
+however. The total \(N\) across the non-DECREASE trials was 10529. We
 conducted sensitivity analyses to see how dependent results are on the
 heterogeneity estimate (not preregistered;
-\href{https://osf.io/vnmzc}{osf.io/vnmzc}). The probability of observing
-the DECREASE trials stays approximately below 1 out of 1000 until the
-variance estimate is 0.25; the probability stays approximately below 1
-out of 100 until the variance estimate is 0.43 (see Figure 2). To put
-these numbers into context, a variance of 0.25 would suggest that
-results of perioperative beta-blockade vary substantially due to
+\href{https://osf.io/vnmzc}{osf.io/vnmzc}). Fixing the variance estimate
+\(\tau^2\) to .5, indicates that the probability of observing the
+DECREASE trials jointly is approximately 1.2 out of 100 000 (see Figure
+1). To put these numbers into context, a variance of 0.25 would suggest
+that results of perioperative beta-blockade vary substantially due to
 contextual circumstances of the study, even if perioperative
@@
 beta-blockade. In contrast to Step 2, which assumes no data manipulation
-occurred, Step 3 assumes that the DECREASE trials might in fact contain
-manipulated data. The estimates from Step 3 provide an indication of the
-extent of potential data manipulation in the DECREASE
-studies\textsuperscript{4--6,9}.
+occurred and that the DECREASE trials occurred naturally from the same
+effect distribution as the non-DECREASE trials, Step 3 assumes that the
+DECREASE trials might in fact contain manipulated data. The estimates
+from Step 3 provide an indication of the extent of potential data
+manipulation in the DECREASE studies\textsuperscript{4--6,9}.
 
@@
 odds four times: once per condition (beta-blocker or control) per trial
-type (DECREASE- and non-DECREASE trials). For all combinations of
-condition and trial type, we ran a meta-analysis applying similar
+stratum (DECREASE- and non-DECREASE trials). For all four combinations
+of condition and trial type, we ran a meta-analysis applying similar
 methods used in Step 1, resulting in four meta-analytic absolute
-mortality estimates with corresponding effect variances (see Table 3;
-one estimate per cell). Throughout the simulations, we used the point
-estimates (i.e., fixed effect) to simulate genuine- and manipulated
-data, but supplemented this by using the more uncertain distribution
-estimates (i.e., random effects) as sensitivity analyses.
-
-\begin{longtable}[]{@{}lll@{}}
-\caption{Outcome possibilities within a simulated 2 (beta-blocker v
-control) by 2 (dead v alive) clinical trial.}\tabularnewline
-\toprule
-& Dead & Alive\tabularnewline
-Beta-blockers & \(n_{11}\) & \(n_{12}\)\tabularnewline
-Control & \(n_{21}\) & \(n_{22}\)\tabularnewline
-\bottomrule
-\end{longtable}
+mortality estimates with corresponding effect variances. Throughout the
+simulations, we used the point estimates (i.e., fixed effect) to
+simulate genuine- and manipulated data, but supplemented this by using
+distribution estimates (i.e., random effects) as sensitivity analyses.
 
 We applied the inversion method to estimate the number of manipulated
-data points in the DECREASE trials\textsuperscript{20}. The inversion
-method iteratively hypothesizes that \emph{X} out of \emph{N} data
-points were manipulated (i.e., \(X={0, 1, ..., N}\)). For each
-combination of \emph{X} and trial, we simulated 10000 datasets. Each
-simulated dataset contained \emph{X} manipulated data points and
+data points in the DECREASE trials\textsuperscript{21}. We assumed that
+if data are manipulated, each data point is manipulated in the same way
+and to the same extent. The inversion method iteratively hypothesizes
+that \emph{X} out of \emph{N} data points were manipulated (i.e.,
+\(X={0, 1, ..., N}\)), assuming they were manipulated in the same way.
+For each combination of \emph{X} and trial, we simulated 10000 datasets.
+Each simulated dataset contained \emph{X} manipulated data points and
 \emph{N-X} genuine data points. For each simulated dataset (exact
@@
 in the previous paragraph, resulting in the estimates depicted in Table
-4. The likelihood was computed under both the manipulated effect
+3. The likelihood was computed under both the manipulated effect
 estimates (i.e., \(L_{manipulated}\)) and the genuine data (i.e.,
-\(L_{genuine}\)). Table 3 indicates which cell sizes the various
+\(L_{genuine}\)). Table 4 indicates which cell sizes the various
 \(n_{XX}\) refer to within the (simulated) data. After computing the
@@
 \toprule
-& & log(odds) & \(\tau\)\tabularnewline
+\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+log(odds)\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+\(\tau\)\strut
+\end{minipage}\tabularnewline
 DECREASE (k=2) & Beta-blocker & -3.629 & 0.588\tabularnewline
-& Control & -2.498 & 1.644\tabularnewline
+\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+Control\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+-2.498\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+1.644\strut
+\end{minipage}\tabularnewline
 Non-DECREASE (k=9) & Beta-blocker & -3.034 & 0.712\tabularnewline
-& Control & -3.208 & 1.001\tabularnewline
+\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+Control\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+-3.208\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+1.001\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
+
+\begin{longtable}[]{@{}lll@{}}
+\caption{Outcome possibilities within a simulated 2 (beta-blocker v
+control) by 2 (dead v alive) clinical trial.}\tabularnewline
+\toprule
+\begin{minipage}[t]{0.32\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright\strut
+Dead\strut
+\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright\strut
+Alive\strut
+\end{minipage}\tabularnewline
+Beta-blockers & \(n_{11}\) & \(n_{12}\)\tabularnewline
+Control & \(n_{21}\) & \(n_{22}\)\tabularnewline
 \bottomrule
@@
   Determined the population mortality rate for each condition (i.e., for
-  each cell as in Table 3). The meta-analytic point estimate was used or
-  a population effect was randomly drawn from the meta-analytic effect
-  distribution.
+  each cell based on the estimates from Table 3). The meta-analytic
+  point estimate was used or a population effect was randomly drawn from
+  the meta-analytic effect distribution.
 \item
@@
 112{]} when based on a point estimate or a more uncertain distribution
-estimate, respectively. The left column of Figure 3 depicts the \(p_M\)
+estimate, respectively. The left column of Figure 2 depicts the \(p_M\)
 per \emph{X} manipulated data points (top panel) and the bounds of the
@@
 
-The ESC/ESA and ACC/AHA guidelines\textsuperscript{11,21} on
+The ESC/ESA and ACC/AHA guidelines\textsuperscript{11,22} on
 perioperative beta-blockade already excluded the DECREASE trials in
-their assessment, but also state that other trials by Poldermans are
-excluded. However, upon close inspection of the reference lists, the
-ACC/AHA guidelines still cites four trials including Poldermans as
-author as evidence for the guidelines\textsuperscript{2,22--24}, of
-which two were already inspected by the scientific integrity committees
-of Erasmus MC\textsuperscript{2,22}. In the ACC/AHA guidelines, the
-following is said about studies conducted by Poldermans:
+their assessment, but also explicitly states that other trials by
+Poldermans are excluded. However, upon close inspection of the reference
+lists, the ACC/AHA guidelines still cites four trials including
+Poldermans as author as evidence for the
+guidelines\textsuperscript{2,23--25}, of which two were already
+inspected by the scientific integrity committees of Erasmus
+MC\textsuperscript{2,23}. In the ACC/AHA guidelines, the following is
+said about studies conducted by Poldermans:
 
@@
 perioperative setting, and to determine the optimal method of
-beta-blockade''}\textsuperscript{21}.
+beta-blockade''}\textsuperscript{22}.
 
@@
 ## Platform: x86_64-redhat-linux-gnu (64-bit)
-## Running under: Fedora 25 (Workstation Edition)
+## Running under: Fedora 26 (Workstation Edition)
 ## 
@@
 
+\hypertarget{ref-doi:10.1093ux2fbiometux2f34.1-2.28}{}
+20. Welch BL. The generalization of ``student's'' problem when several
+different population variances are involved. \emph{Biometrika}.
+1947;34(1-2):28-35.
+doi:\href{https://doi.org/10.1093/biomet/34.1-2.28}{10.1093/biomet/34.1-2.28}.
+
 \hypertarget{ref-casella2002}{}
-20. Casella G, Berger RL. \emph{Statistical Interference}. Pacific
+21. Casella G, Berger RL. \emph{Statistical Interference}. Pacific
 Grove, CA: Duxbury; 2002.
@@
 \hypertarget{ref-Kristensen_2014}{}
-21. Kristensen SD, Knuuti J, Saraste A, et al. 2014 ESC/ESA guidelines
+22. Kristensen SD, Knuuti J, Saraste A, et al. 2014 ESC/ESA guidelines
 on non-cardiac surgery. \emph{European Journal of Anaesthesiology}.
@@
 \hypertarget{ref-Boersma_2001}{}
-22. Boersma E, Poldermans D, Bax JJ, et al. Predictors of cardiac events
+23. Boersma E, Poldermans D, Bax JJ, et al. Predictors of cardiac events
 after major vascular surgery. \emph{JAMA}. 2001;285(14):1865.
@@
 \hypertarget{ref-van_Kuijk_2009}{}
-23. Kuijk J-P van, Flu W-J, Schouten O, et al. Timing of noncardiac
+24. Kuijk J-P van, Flu W-J, Schouten O, et al. Timing of noncardiac
 surgery after coronary artery stenting with bare metal or drug-eluting
@@
 \hypertarget{ref-Flu_2010}{}
-24. Flu W-J, Kuijk J-P van, Chonchol M, et al. Timing of pre-operative
+25. Flu W-J, Kuijk J-P van, Chonchol M, et al. Timing of pre-operative
 beta-blocker treatment in vascular surgery patients. \emph{Journal of

### 2017-08-22 submission/manuscript.Rmd
Preprint stage

@@
 
-All manuscript materials are available at https://github.com/chartgerink/2015poldermans and are preserved at Zenodo [@].
+All manuscript materials are available at https://github.com/chartgerink/2015poldermans and are preserved at Zenodo [@chris_hartgerink_2017_845355].
+CHJH was funded by the Office of Research Integrity during part of this project (ORI; ORIIR160019).
 
@@
 Third, we estimated how many data points would have to be manipulated in order to reproduce the results of the DECREASE trials if the initial, non-manipulated results arose from the effect estimates as obtained from the non-DECREASE trials.
-Considering the committees investigating the scientific integrity of the DECREASE trials was unable to assess this, we consider it worthwhile to investigate this further.
+Considering the committees investigating the scientific integrity of the DECREASE trials were unable to assess this, we consider it worthwhile to investigate this further.
 
@@
 Publication bias was assumed to not be a problem because a substantial number of nonsignificant effects are included in the dataset (9 of 11 results are nonsignificant).
-Based on this effect distribution, we estimated the veracity of the DECREASE trials separately, which is the estimated probability of the observed data under a given true effect [@peters2015].
+Based on this effect distribution, we estimated the veracity of the DECREASE trials separately, which is the estimated probability of the observed data (or more extreme) under a given true effect [@peters2015].
 
@@
 
-Figure `r fignr` indicates that the DECREASE trials are highly unlikely under the estimated effect distribution based on the non-DECREASE trials.
+Results indicate that the DECREASE trials are highly unlikely under the estimated effect distribution from the non-DECREASE trials.
 More specifically, the results from DECREASE-I (or more extreme) have a probability of 1 in 10 000 ($t(`r df_I`)=`r round(t_I, 3)`,p=`r p_I`$) and the results from DECREASE-IV (or more extreme) have a probability of 1 in 1000 ($t(`r df_IV`)=`r round(t_IV, 3)`,p=`r p_IV`$).
@@
 Nonetheless, the range of potentially manipulated data points is still estimated at approximately 1000; this indicates that the summary results are insufficient to provide more than an estimated lower bound.
-This indicates that it is possible not all data were manipulated (i.e., $N=`r length(res_iv_certain) - 1`$), increasing the importance of well-documented data provenance to discern between genuine and falsified data.
+This indicates that it is possible not all data were manipulated (i.e., $N=`r length(res_iv_certain) - 1`$) but at least some were, increasing the importance of well-documented data provenance to discern between genuine and falsified data.
 

### 2017-08-22 submission/manuscript.tex
Preprint stage

@@
   \posttitle{\par}
-  \author{\href{https://orcid.org/}{Chris H. J. Hartgerink}, E. M. Kemper,
-\href{https://orcid.org/0000-0001-8248-0244}{Markus W. Hollman},
-\href{https://orcid.org/0000-0002-2231-7637}{Gerben ter Riet}}
+  \author{\href{https://orcid.org/0000-0003-1050-6809}{Chris H. J. Hartgerink}, E.
+M. Kemper, \href{https://orcid.org/0000-0001-8248-0244}{Markus W.
+Hollman}, \href{https://orcid.org/0000-0002-2231-7637}{Gerben ter Riet}}
   \preauthor{\centering\large\emph}
@@
   \postdate{\par}
-  \date{27 July, 2017}
+  \date{18 August, 2017}
 
@@
 \url{https://github.com/chartgerink/2015poldermans} and are preserved at
-Zenodo {[}@{]}.
+Zenodo\textsuperscript{1}. CHJH was funded by the Office of Research
+Integrity during part of this project (ORI; ORIIR160019).
 
@@
 The effect of beta-blockers on perioperative mortality in non-cardiac
-surgery has been controversial\textsuperscript{1} due to concerns
+surgery has been controversial\textsuperscript{2} due to concerns
 regarding the scientific integrity in two related clinical
-trials\textsuperscript{2--6}. Three meta-analyses that included the
+trials\textsuperscript{3--7}. Three meta-analyses that included the
 trials subject to concerns concluded that beta-blockers decrease
-perioperative mortality\textsuperscript{7--9} whereas a meta-analysis
+perioperative mortality\textsuperscript{8--10} whereas a meta-analysis
 that excluded the suspect trials concluded that beta-blockers increase
-perioperative mortality\textsuperscript{9}. In these studies,
+perioperative mortality\textsuperscript{10}. In these studies,
 perioperative mortality was defined as the death rate of patients in the
@@
 The trials subject to concerns regarding scientific integrity were the
-Dutch DECREASE-I and DECREASE-IV trials\textsuperscript{4--6}. The
+Dutch DECREASE-I and DECREASE-IV trials\textsuperscript{5--7}. The
 committees that investigated the integrity of the DECREASE trials
 reported that data manipulation was likely but that the extent of the
-data manipulation remained unclear\textsuperscript{4--6}. Moreover, the
+data manipulation remained unclear\textsuperscript{5--7}. Moreover, the
 latest guidelines still recommend the usage of beta-blockers in the
-perioperative period in certain cases\textsuperscript{10,11}, where some
+perioperative period in certain cases\textsuperscript{11,12}, where some
 of these guidelines are based on other work by the PI of the DECREASE
@@
 breaching scientific integrity, we aim to estimate the extent of data
-manipulation in the DECREASE studies\textsuperscript{2--6} to further
+manipulation in the DECREASE studies\textsuperscript{3--7} to further
 stimulate the debate on using beta-blockers in the perioperative period
@@
 expert evaluated the applicability of forensic statistical
-methods\textsuperscript{6} to evaluate results of trials separately
+methods\textsuperscript{7} to evaluate results of trials separately
 (i.e., DECREASE-I, DECREASE-IV) although the report lacks details as to
@@
 that has previously been used to monitor trial data quality or to test
-for potential data anomalies\textsuperscript{12,13}. Moreover, comparing
+for potential data anomalies\textsuperscript{13,14}. Moreover, comparing
 across trials instead of evaluating them separately has previously
 proven to be effective in detecting data
-manipulation\textsuperscript{14}. Comparing the DECREASE trials to other
+manipulation\textsuperscript{15}. Comparing the DECREASE trials to other
 published trials studying the effectiveness of beta-blockers with
@@
 effectiveness with respect to perioperative mortality. Moreover, the
-highly discrepant results from the DECREASE trials\textsuperscript{9}
-might partly be caused by such differences\textsuperscript{15} and not
+highly discrepant results from the DECREASE trials\textsuperscript{10}
+might partly be caused by such differences\textsuperscript{16} and not
 purely due to data manipulation (given that not all data points can be
@@
 To statistically investigate the evidence for data manipulation in the
-DECREASE studies\textsuperscript{2,3}, we took three steps. First, we
+DECREASE studies\textsuperscript{3,4}, we took three steps. First, we
 reproduced the findings from the 2014 meta-analysis by Bouri et
-al.\textsuperscript{9}, which contained sufficient information to
+al.\textsuperscript{10}, which contained sufficient information to
 estimate the deviation of the DECREASE trials from other published
@@
 the committees investigating the scientific integrity of the DECREASE
-trials was unable to assess this, we consider it worthwhile to
+trials were unable to assess this, we consider it worthwhile to
 investigate this further.
@@
 To ensure that we used similar analysis procedures as in the 2014
-meta-analysis\textsuperscript{9}, we initially reproduced Bouri et al.'s
-estimates. This ensured that (1) their results are reproducible and (2)
-we are using the correct estimates in subsequent steps of our analyses.
-Using figures 2 and 3 from the original paper\textsuperscript{9}, we
-extracted the raw event data for the 2 (control vs experimental) by 2
-(event vs no event) design, which we used to recompute the natural
-logarithm of the risk ratio and its standard error. The extracted event
-data is available at \href{https://osf.io/aykeh}{osf.io/aykeh} and our
-analysis plan was preregistered at
-\href{https://osf.io/vnmzc}{osf.io/vnmzc}.
+meta-analysis\textsuperscript{10}, we initially reproduced Bouri et
+al.'s estimates. This ensured that (1) their results are reproducible
+and (2) we are using the correct estimates in subsequent steps of our
+analyses. Using figures 2 and 3 from the original
+paper\textsuperscript{10}, we extracted the raw event data for the 2
+(control vs experimental) by 2 (event vs no event) design, which we used
+to recompute the natural logarithm of the risk ratio and its standard
+error. The extracted event data is available at
+\href{https://osf.io/aykeh}{osf.io/aykeh} and our analysis plan was
+preregistered at \href{https://osf.io/vnmzc}{osf.io/vnmzc}.
 
@@
 these using \texttt{v2.0.0} of the \texttt{R} package
-\texttt{metafor}\textsuperscript{16}. We estimated a weighted
+\texttt{metafor}\textsuperscript{17}. We estimated a weighted
 random-effects model using the restricted maximum-likelihood estimator
-(i.e., \texttt{REML})\textsuperscript{17} to estimate the variance of
+(i.e., \texttt{REML})\textsuperscript{18} to estimate the variance of
 effects. We used the default weighting procedure in the \texttt{metafor}
@@
 on risk- and odds ratios in order to prevent computational
-artefacts\textsuperscript{18}. The 2014 meta-analysis\textsuperscript{9}
-did not specify the variance estimate used, hence, (minor) discrepancies
-between our estimates and the original estimates could be due to
-differences in the estimation procedure.
+artefacts\textsuperscript{19}. The 2014
+meta-analysis\textsuperscript{10} did not specify the variance estimate
+used, hence, (minor) discrepancies between our estimates and the
+original estimates could be due to differences in the estimation
+procedure.
 
@@
 We were able to closely reproduce the estimates for the different sets
-of studies (Figure 2 of the 2014 meta-analysis\textsuperscript{9}).
+of studies (Figure 2 of the 2014 meta-analysis\textsuperscript{10}).
 Bouri et al. differentiated between the estimates from the non-DECREASE
@@
 predictor for the DECREASE and non-DECREASE studies to reproduce results
-presented in Figure 4 of the 2014 meta-analysis\textsuperscript{9}. Our
+presented in Figure 4 of the 2014 meta-analysis\textsuperscript{10}. Our
 results showed a bit more evidence against equal subgroups than the
-original meta-analysis\textsuperscript{9} (original:
+original meta-analysis\textsuperscript{10} (original:
 \(\chi^2(1)=3.91,p=.05\); reproduced: \(\chi^2(1)=6.12,p=0.013\)).
@@
 veracity of the DECREASE trials separately, which is the estimated
-probability of the observed data under a given true
-effect\textsuperscript{19}.
+probability of the observed data (or more extreme) under a given true
+effect\textsuperscript{20}.
 
@@
 (\(H_0:\mu_{\bar{x}_1-\bar{x}_2}=0\)). To this end, we applied a Welch
-t-test\textsuperscript{20}. As means we used the observed log RR for the
+t-test\textsuperscript{21}. As means we used the observed log RR for the
 DECREASE trials (i.e., DECREASE-I: -1.44; DECREASE-IV: -0.452) and the
@@
 
-Figure 1 indicates that the DECREASE trials are highly unlikely under
-the estimated effect distribution based on the non-DECREASE trials. More
+Results indicate that the DECREASE trials are highly unlikely under the
+estimated effect distribution from the non-DECREASE trials. More
 specifically, the results from DECREASE-I (or more extreme) have a
@@
 from Step 3 provide an indication of the extent of potential data
-manipulation in the DECREASE studies\textsuperscript{4--6,9}.
+manipulation in the DECREASE studies\textsuperscript{5--7,10}.
 
@@
 We applied the inversion method to estimate the number of manipulated
-data points in the DECREASE trials\textsuperscript{21}. We assumed that
+data points in the DECREASE trials\textsuperscript{22}. We assumed that
 if data are manipulated, each data point is manipulated in the same way
@@
 manipulation that occurred in the DECREASE-I trial, affirming the
-conclusions of the original committee report\textsuperscript{6}.
+conclusions of the original committee report\textsuperscript{7}.
 
@@
 This indicates that it is possible not all data were manipulated (i.e.,
-\(N=1066\)), increasing the importance of well-documented data
-provenance to discern between genuine and falsified data.
+\(N=1066\)) but at least some were, increasing the importance of
+well-documented data provenance to discern between genuine and falsified
+data.
 
@@
 same effects as the non-DECREASE trials, as is often
-assumed\textsuperscript{9}. Nonetheless, the original papers that
+assumed\textsuperscript{10}. Nonetheless, the original papers that
 presented these trial results are not yet
-retracted\textsuperscript{2,3}, despite the integrity
-reports\textsuperscript{4--6}.
+retracted\textsuperscript{3,4}, despite the integrity
+reports\textsuperscript{5--7}.
 
@@
 
-The ESC/ESA and ACC/AHA guidelines\textsuperscript{11,22} on
+The ESC/ESA and ACC/AHA guidelines\textsuperscript{12,23} on
 perioperative beta-blockade already excluded the DECREASE trials in
@@
 Poldermans as author as evidence for the
-guidelines\textsuperscript{2,23--25}, of which two were already
+guidelines\textsuperscript{3,24--26}, of which two were already
 inspected by the scientific integrity committees of Erasmus
-MC\textsuperscript{2,23}. In the ACC/AHA guidelines, the following is
+MC\textsuperscript{3,24}. In the ACC/AHA guidelines, the following is
 said about studies conducted by Poldermans:
@@
 recommendation but should not form the basis of that recommendation or
-be used as a reference for the recommendation.''}\textsuperscript{11}
+be used as a reference for the recommendation.''}\textsuperscript{12}
 \end{quote}
@@
 evidence that type of beta-blockade can severely influence perioperative
-mortality. Our reanalysis of the Bouri et al.\textsuperscript{9} data
+mortality. Our reanalysis of the Bouri et al.\textsuperscript{10} data
 indicates that type of beta-blockade can reverse the effect on
@@
 perioperative setting, and to determine the optimal method of
-beta-blockade''}\textsuperscript{22}.
+beta-blockade''}\textsuperscript{23}.
 
@@
 DECREASE trials use beta-blockade as a proxy to decrease resting heart
-rate\textsuperscript{2,3}. As such, the DECREASE studies applied
+rate\textsuperscript{3,4}. As such, the DECREASE studies applied
 beta-blockade at least a week in advance, specifically in order to lower
@@
 \begin{verbatim}
-## R version 3.4.0 (2017-04-21)
+## R version 3.4.1 (2017-06-30)
 ## Platform: x86_64-redhat-linux-gnu (64-bit)
@@
 ## other attached packages:
-## [1] knitr_1.16      latex2exp_0.4.0 metafor_2.0-0   Matrix_1.2-9   
+## [1] knitr_1.16      latex2exp_0.4.0 metafor_2.0-0   Matrix_1.2-10  
 ## 
 ## loaded via a namespace (and not attached):
-##  [1] Rcpp_0.12.11    lattice_0.20-35 digest_0.6.12   rprojroot_1.2  
-##  [5] grid_3.4.0      nlme_3.1-131    backports_1.1.0 magrittr_1.5   
+##  [1] Rcpp_0.12.12    lattice_0.20-35 digest_0.6.12   rprojroot_1.2  
+##  [5] grid_3.4.1      nlme_3.1-131    backports_1.1.0 magrittr_1.5   
 ##  [9] evaluate_0.10.1 highr_0.6       stringi_1.1.5   rmarkdown_1.6  
-## [13] tools_3.4.0     stringr_1.2.0   yaml_2.1.14     compiler_3.4.0 
+## [13] tools_3.4.1     stringr_1.2.0   yaml_2.1.14     compiler_3.4.1 
 ## [17] htmltools_0.3.6
@@
 \hypertarget{refs}{}
+\hypertarget{ref-chris_hartgerink_2017_845355}{}
+1. Hartgerink C. Chartgerink/2015poldermans: Submission to eLife. August
+2017.
+doi:\href{https://doi.org/10.5281/zenodo.845355}{10.5281/zenodo.845355}.
+
 \hypertarget{ref-Coleg5210}{}
-1. Cole GD, Francis DP. Perioperative beta blockade: Guidelines do not
+2. Cole GD, Francis DP. Perioperative beta blockade: Guidelines do not
 reflect the problems with the evidence from the decrease trials.
@@
 \hypertarget{ref-poldermans1999}{}
-2. Poldermans D, Boersma E, Bax JJ, et al. The effect of bisoprolol on
+3. Poldermans D, Boersma E, Bax JJ, et al. The effect of bisoprolol on
 perioperative mortality and myocardial infarction in High-Risk patients
@@
 \hypertarget{ref-dunkelgrun2009}{}
-3. Dunkelgrun M, Boersma E, Schouten O, et al. Bisoprolol and
+4. Dunkelgrun M, Boersma E, Schouten O, et al. Bisoprolol and
 fluvastatin for the reduction of perioperative cardiac mortality and
@@
 \hypertarget{ref-commissie2011}{}
-4. Onderzoekscommissie Wetenschappelijke Integriteit. \emph{Onderzoek
+5. Onderzoekscommissie Wetenschappelijke Integriteit. \emph{Onderzoek
 Naar Mogelijke Schending van de Wetenschappelijke Integriteit: Beknopte
@@
 \hypertarget{ref-commissie2012}{}
-5. Commissie Vervolgonderzoek 2012. \emph{Rapport Vervolgonderzoek Naar
+6. Commissie Vervolgonderzoek 2012. \emph{Rapport Vervolgonderzoek Naar
 Mogelijke Schending van de Wetenschappelijke Integriteit}. Erasmus MC;
@@
 \hypertarget{ref-commissie2013}{}
-6. Commissie Vervolgonderzoek Wetenschappelijke Integriteit 2013.
+7. Commissie Vervolgonderzoek Wetenschappelijke Integriteit 2013.
 \emph{Rapport}. Erasmus MC; 2014.
@@
 \hypertarget{ref-Devereaux313}{}
-7. Devereaux PJ, Beattie WS, Choi PT-L, et al. How strong is the
+8. Devereaux PJ, Beattie WS, Choi PT-L, et al. How strong is the
 evidence for the use of perioperative beta blockers in non-cardiac
@@
 \hypertarget{ref-Angeli2010}{}
-8. Angeli F, Verdecchia P, Karthikeyan G, Mazzotta G, Gentile G, Reboldi
+9. Angeli F, Verdecchia P, Karthikeyan G, Mazzotta G, Gentile G, Reboldi
 G. \(\beta\)-blockers reduce mortality in patients undergoing high-risk
@@
 \hypertarget{ref-bouri2014}{}
-9. Bouri S, Shun-Shin MJ, Cole GD, Mayet J, Francis DP. Meta-analysis of
-secure randomised controlled trials of -blockade to prevent
+10. Bouri S, Shun-Shin MJ, Cole GD, Mayet J, Francis DP. Meta-analysis
+of secure randomised controlled trials of -blockade to prevent
 perioperative death in non-cardiac surgery. \emph{Heart}.
@@
 \hypertarget{ref-esc2014}{}
-10. ESC/ESA. Guidelines on non-cardiac surgery: Cardiovascular
+11. ESC/ESA. Guidelines on non-cardiac surgery: Cardiovascular
 assessment and management. \emph{European Heart Journal}.
@@
 \hypertarget{ref-Fleisher_2014}{}
-11. Fleisher LA, Fleischmann KE, Auerbach AD, et al. 2014 ACC/AHA
+12. Fleisher LA, Fleischmann KE, Auerbach AD, et al. 2014 ACC/AHA
 guideline on perioperative cardiovascular evaluation and management of
@@
 \hypertarget{ref-Buyse1999-jq}{}
-12. Buyse M, George SL, Evans S, et al. The role of biostatistics in the
+13. Buyse M, George SL, Evans S, et al. The role of biostatistics in the
 prevention, detection and treatment of fraud in clinical trials.
@@
 \hypertarget{ref-10.1111ux2fanae.13938}{}
-13. Carlisle JB. Data fabrication and other reasons for non-random
+14. Carlisle JB. Data fabrication and other reasons for non-random
 sampling in 5087 randomised, controlled trials in anaesthetic and
@@
 \hypertarget{ref-Knepper2016-la}{}
-14. Knepper D, Lindblad AS, Sharma G, et al. Statistical monitoring in
+15. Knepper D, Lindblad AS, Sharma G, et al. Statistical monitoring in
 clinical trials: Best practices for detecting data anomalies suggestive
@@
 \hypertarget{ref-klei2015}{}
-15. Klei W van. Welke perioperatieve bètablokker heeft de voorkeur?
+16. Klei W van. Welke perioperatieve bètablokker heeft de voorkeur?
 {[}Which perioperative beta-blocker is preferred?{]}. \emph{Nederlands
@@
 \hypertarget{ref-viechtbauer2010}{}
-16. Viechtbauer W. Conducting meta-analyses in R with the metafor
+17. Viechtbauer W. Conducting meta-analyses in R with the metafor
 package. \emph{Journal of Statistical Software}. 2010;36:1-48.
@@
 \hypertarget{ref-viechtbauer2005}{}
-17. Viechtbauer W. Bias and efficiency of meta-analytic variance
+18. Viechtbauer W. Bias and efficiency of meta-analytic variance
 estimators in the Random-Effects model. \emph{Journal of educational and
@@
 \hypertarget{ref-agresti2002}{}
-18. Agresti A. \emph{Categorical Data Analysis}. Hoboken, NJ: John Wiley
+19. Agresti A. \emph{Categorical Data Analysis}. Hoboken, NJ: John Wiley
 \& Sons Inc.; 2002.
@@
 \hypertarget{ref-peters2015}{}
-19. Peters CFW, Klaassen CAJ, Wiel MA van de. \emph{Evaluating the
+20. Peters CFW, Klaassen CAJ, Wiel MA van de. \emph{Evaluating the
 Scientific Veracity of Publications by Dr. Jens Forster}. University of
@@
 \hypertarget{ref-doi:10.1093ux2fbiometux2f34.1-2.28}{}
-20. Welch BL. The generalization of ``student's'' problem when several
+21. Welch BL. The generalization of ``student's'' problem when several
 different population variances are involved. \emph{Biometrika}.
@@
 \hypertarget{ref-casella2002}{}
-21. Casella G, Berger RL. \emph{Statistical Interference}. Pacific
+22. Casella G, Berger RL. \emph{Statistical Interference}. Pacific
 Grove, CA: Duxbury; 2002.
@@
 \hypertarget{ref-Kristensen_2014}{}
-22. Kristensen SD, Knuuti J, Saraste A, et al. 2014 ESC/ESA guidelines
+23. Kristensen SD, Knuuti J, Saraste A, et al. 2014 ESC/ESA guidelines
 on non-cardiac surgery. \emph{European Journal of Anaesthesiology}.
@@
 \hypertarget{ref-Boersma_2001}{}
-23. Boersma E, Poldermans D, Bax JJ, et al. Predictors of cardiac events
+24. Boersma E, Poldermans D, Bax JJ, et al. Predictors of cardiac events
 after major vascular surgery. \emph{JAMA}. 2001;285(14):1865.
@@
 \hypertarget{ref-van_Kuijk_2009}{}
-24. Kuijk J-P van, Flu W-J, Schouten O, et al. Timing of noncardiac
+25. Kuijk J-P van, Flu W-J, Schouten O, et al. Timing of noncardiac
 surgery after coronary artery stenting with bare metal or drug-eluting
@@
 \hypertarget{ref-Flu_2010}{}
-25. Flu W-J, Kuijk J-P van, Chonchol M, et al. Timing of pre-operative
+26. Flu W-J, Kuijk J-P van, Chonchol M, et al. Timing of pre-operative
 beta-blocker treatment in vascular surgery patients. \emph{Journal of

### 2017-10-18 submission/manuscript.Rmd
Minor changes and data for interactive figure

@@
 title: "Analyzing DECREASE trials to estimate evidence of data manipulation"
-author: "[Chris H.
-J.
-Hartgerink](https://orcid.org/0000-0003-1050-6809), E.
-M.
-Kemper, [Markus W.
-Hollman](https://orcid.org/0000-0001-8248-0244), [Gerben ter Riet](https://orcid.org/0000-0002-2231-7637)"
+author: '[Chris H. J. Hartgerink](https://orcid.org/0000-0003-1050-6809), E. M. Kemper,
+  [Markus W. Hollman](https://orcid.org/0000-0001-8248-0244), [Gerben ter Riet](https://orcid.org/0000-0002-2231-7637)'
 date: '`r format(Sys.time(), "%d %B, %Y")`'
@@
 
-All manuscript materials are available at https://github.com/chartgerink/2015poldermans and are preserved at Zenodo [@chris_hartgerink_2017_845355].
+All manuscript materials are available at https://github.com/chartgerink/2015poldermans and are preserved at Zenodo ([doi.org/10.5281/zenodo.845354](https://doi.org/10.5281/zenodo.845354)).
 CHJH was funded by the Office of Research Integrity during part of this project (ORI; ORIIR160019).
@@
 ```{r, echo = FALSE}
-pdf(sprintf('%sfigures/fig2.pdf', x12), width = 8, height = 6)
+pdf(sprintf('%sfigures/fig1.pdf', x12), width = 8, height = 6)
 tau2 <- seq(0, .5, .01)
@@
 ```{r figure 2, fig.cap = 'Sensitivity analyses for the p-value that indicates the probability of observing the results from the DECREASE studies, or more extreme results, based on the estimated true effect (non-DECREASE trials) and the accompanying variance estimate.', out.width = '80%', fig.asp = .5, fig.align = 'center', echo = FALSE}
-knitr::include_graphics(sprintf("%sfigures/fig2.pdf", x12))
+knitr::include_graphics(sprintf("%sfigures/fig1.pdf", x12))
 fignr <- fignr + 1
@@
 For each simulated dataset (exact simulation procedure in the next paragraph), we determined the likelihood of the results with
-\begin{equation}
-\label{eq1}
+$$
 L(\theta|\pi_{E},\pi_{C})=\pi_{E}^{n_{11}}(1-\pi_{E})^{n_{12}} \times \pi_{C}^{n_{21}}(1 - \pi_{C})^{n_{22}}
-\end{equation}
+$$
 where $\pi_{E}$ indicates the mortality rate in the beta-blocker condition as drawn from the meta-analytic effect distribution ($\pi_{C}$ indicates the mortality rate in the control condition).
@@
 
-Based on the meta-analytic effect from 2 and the cell sizes from 3, we computed the likelihoods $L_{manipulated}$ and $L_{genuine}$ using Equation \ref{eq1}.
+Based on the meta-analytic effect from 2 and the cell sizes from 3, we computed the likelihoods $L_{manipulated}$ and $L_{genuine}$ using Equation 1.
 As mentioned before, we computed $p_M$, which indicates the probability that the data are more likely under the estimates resulting from the (allegedly) manipulated data (i.e., the DECREASE trials) than under the estimates resulting from the genuine data (i.e., the non-DECREASE trials; $p_M=P(L_{manipulated}>L_{genuine})$).
@@
 
-  res_i_uncertain_lb[j] <- which(abs(res_i_uncertain - lb) == min(abs(res_i_uncertain - lb)))[1]
-  res_i_uncertain_ub[j] <- which(abs(res_i_uncertain - ub) == min(abs(res_i_uncertain - ub)))[1]
+  # MINUS ONE because R does not have zero indexing
+  # AAAA
+  res_i_uncertain_lb[j] <- which(abs(res_i_uncertain - lb) == min(abs(res_i_uncertain - lb)))[1] - 1
+  res_i_uncertain_ub[j] <- which(abs(res_i_uncertain - ub) == min(abs(res_i_uncertain - ub)))[1] - 1
 
-  res_i_certain_lb[j] <- which(abs(res_i_certain - lb) == min(abs(res_i_certain - lb)))[1]
-  res_i_certain_ub[j] <- which(abs(res_i_certain - ub) == min(abs(res_i_certain - ub)))[1]
+  res_i_certain_lb[j] <- which(abs(res_i_certain - lb) == min(abs(res_i_certain - lb)))[1] - 1
+  res_i_certain_ub[j] <- which(abs(res_i_certain - ub) == min(abs(res_i_certain - ub)))[1] - 1
 
-  res_iv_uncertain_lb[j] <- which(abs(res_iv_uncertain - lb) == min(abs(res_iv_uncertain - lb)))[1]
-  res_iv_uncertain_ub[j] <- which(abs(res_iv_uncertain - ub) == min(abs(res_iv_uncertain - ub)))[1]
+  res_iv_uncertain_lb[j] <- which(abs(res_iv_uncertain - lb) == min(abs(res_iv_uncertain - lb)))[1] - 1
+  res_iv_uncertain_ub[j] <- which(abs(res_iv_uncertain - ub) == min(abs(res_iv_uncertain - ub)))[1] - 1
 
-  res_iv_certain_lb[j] <- which(abs(res_iv_certain - lb) == min(abs(res_iv_certain - lb)))[1]
-  res_iv_certain_ub[j] <- which(abs(res_iv_certain - ub) == min(abs(res_iv_certain - ub)))[1]
+  res_iv_certain_lb[j] <- which(abs(res_iv_certain - lb) == min(abs(res_iv_certain - lb)))[1] - 1
+  res_iv_certain_ub[j] <- which(abs(res_iv_certain - ub) == min(abs(res_iv_certain - ub)))[1] - 1
 
@@
 
-pdf(sprintf('%sfigures/fig3.pdf', x12), width = 8, height = 6)
+N_I <- sum(bouri_data[grepl(bouri_data$study, pattern = '_I-'),c(3,5)])
+N_IV <- sum(bouri_data[grepl(bouri_data$study, pattern = '_IV-'),c(3,5)])
+
+pdf(sprintf('%sfigures/fig2.pdf', x12), width = 8, height = 6)
 
@@
 plot(x = NA, y = NA,
-     xlim = c(0,1), ylim = c(0, length(res_i_certain)),
-     xlab = "Confidence in %", ylab = "Number of data points manipulated")
+     xlim = c(0,1), ylim = c(0, 100),
+     xlab = "Confidence in %", ylab = "% of data points manipulated")
 
-lines(x = seq(0, 1, .01), y = res_i_uncertain_lb)
-lines(x = seq(0, 1, .01), y = res_i_uncertain_ub)
-lines(x = seq(0, 1, .01), y = res_i_certain_lb, lty = 2, pch = 2, type = 'l')
-lines(x = seq(0, 1, .01), y = res_i_certain_ub, lty = 2, pch = 2, type = 'l')
+lines(x = seq(0, 1, .01), y = (res_i_uncertain_lb / N_I) * 100)
+lines(x = seq(0, 1, .01), y = (res_i_uncertain_ub / N_I) * 100)
+lines(x = seq(0, 1, .01), y = (res_i_certain_lb / N_I) * 100, lty = 2, pch = 2, type = 'l')
+lines(x = seq(0, 1, .01), y = (res_i_certain_ub / N_I) * 100, lty = 2, pch = 2, type = 'l')
 abline(v = .95, lty = 3)
@@
 plot(x = NA, y = NA,
-     xlim = c(0,1), ylim = c(0, length(res_iv_certain)),
+     xlim = c(0,1), ylim = c(0, 100),
      xlab = "Confidence in %", ylab = "",
@@
 
-lines(x = seq(0, 1, .01), y = res_iv_uncertain_lb)
-lines(x = seq(0, 1, .01), y = res_iv_uncertain_ub)
-lines(x = seq(0, 1, .01), y = res_iv_certain_lb, lty = 2, pch = 2, type = 'l')
-lines(x = seq(0, 1, .01), y = res_iv_certain_ub, lty = 2, pch = 2, type = 'l')
+lines(x = seq(0, 1, .01), y = (res_iv_uncertain_lb / N_IV) * 100)
+lines(x = seq(0, 1, .01), y = (res_iv_uncertain_ub / N_IV) * 100)
+lines(x = seq(0, 1, .01), y = (res_iv_certain_lb / N_IV) * 100, lty = 2, pch = 2, type = 'l')
+lines(x = seq(0, 1, .01), y = (res_iv_certain_ub / N_IV) * 100, lty = 2, pch = 2, type = 'l')
 abline(v = .95, lty = 3)
@@
 Dotted lines indicate the bounds for a 95 percent CI.', out.width = '80%', fig.asp = .5, fig.align = 'center', echo = FALSE}
-knitr::include_graphics(sprintf("%sfigures/fig3.pdf", x12))
+knitr::include_graphics(sprintf("%sfigures/fig2.pdf", x12))
 fignr <- fignr + 1

### 2017-10-18 submission/manuscript.tex
Minor changes and data for interactive figure

--- 2017-07-27 submission/manuscript.tex
@@
   \postdate{\par}
-  \date{27 July, 2017}
+  \date{18 October, 2017}
 
@@
 \url{https://github.com/chartgerink/2015poldermans} and are preserved at
-Zenodo {[}@{]}.
+Zenodo
+(\href{https://doi.org/10.5281/zenodo.845354}{doi.org/10.5281/zenodo.845354}).
+CHJH was funded by the Office of Research Integrity during part of this
+project (ORI; ORIIR160019).
 
@@
 veracity of the DECREASE trials separately, which is the estimated
-probability of the observed data under a given true
+probability of the observed data (or more extreme) under a given true
 effect\textsuperscript{19}.
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{../figures/fig2}
+{\centering \includegraphics[width=0.8\linewidth]{../figures/fig1}
 
@@
 simulation procedure in the next paragraph), we determined the
-likelihood of the results with
-
-\begin{equation}
-\label{eq1}
+likelihood of the results with \[
 L(\theta|\pi_{E},\pi_{C})=\pi_{E}^{n_{11}}(1-\pi_{E})^{n_{12}} \times \pi_{C}^{n_{21}}(1 - \pi_{C})^{n_{22}}
-\end{equation}
-
-where \(\pi_{E}\) indicates the mortality rate in the beta-blocker
+\] where \(\pi_{E}\) indicates the mortality rate in the beta-blocker
 condition as drawn from the meta-analytic effect distribution
@@
 computed the likelihoods \(L_{manipulated}\) and \(L_{genuine}\) using
-Equation \ref{eq1}. As mentioned before, we computed \(p_M\), which
-indicates the probability that the data are more likely under the
-estimates resulting from the (allegedly) manipulated data (i.e., the
-DECREASE trials) than under the estimates resulting from the genuine
-data (i.e., the non-DECREASE trials;
-\(p_M=P(L_{manipulated}>L_{genuine})\)).
+Equation 1. As mentioned before, we computed \(p_M\), which indicates
+the probability that the data are more likely under the estimates
+resulting from the (allegedly) manipulated data (i.e., the DECREASE
+trials) than under the estimates resulting from the genuine data (i.e.,
+the non-DECREASE trials; \(p_M=P(L_{manipulated}>L_{genuine})\)).
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{../figures/fig3}
+{\centering \includegraphics[width=0.8\linewidth]{../figures/fig2}
 
@@
 ## other attached packages:
-## [1] knitr_1.16      latex2exp_0.4.0 metafor_2.0-0   Matrix_1.2-9   
+## [1] knitr_1.17      latex2exp_0.4.0 metafor_2.0-0   Matrix_1.2-10  
 ## 
 ## loaded via a namespace (and not attached):
-##  [1] Rcpp_0.12.11    lattice_0.20-35 digest_0.6.12   rprojroot_1.2  
-##  [5] grid_3.4.0      nlme_3.1-131    backports_1.1.0 magrittr_1.5   
+##  [1] Rcpp_0.12.13    lattice_0.20-35 digest_0.6.12   rprojroot_1.2  
+##  [5] grid_3.4.1      nlme_3.1-131    backports_1.1.1 magrittr_1.5   
 ##  [9] evaluate_0.10.1 highr_0.6       stringi_1.1.5   rmarkdown_1.6  
-## [13] tools_3.4.0     stringr_1.2.0   yaml_2.1.14     compiler_3.4.0 
+## [13] tools_3.4.1     stringr_1.2.0   yaml_2.1.14     compiler_3.4.1 
 ## [17] htmltools_0.3.6

### 2017-10-18 functions/plotly_data.R

@@ -1,0 +1,46 @@
+# Assumes the scripts from Rmarkdown file are already run
+
+top.left <- data.frame(x.dashed = 0:(length(res_i_certain) - 1),
+                       y.dashed = res_i_certain,
+                       x.solid = 0:(length(res_i_uncertain) - 1),
+                       y.solid = res_i_uncertain)
+
+plot(x = top.left$x.dashed, y = top.left$y.dashed, ylim = c(0,1))
+plot(x = top.left$x.solid, y = top.left$y.solid, ylim = c(0,1))
+
+top.right <- data.frame(x.dashed = 0:(length(res_iv_certain) - 1),
+                        y.dashed = res_iv_certain,
+                        x.solid = 0:(length(res_iv_uncertain) - 1),
+                        y.solid = res_iv_uncertain)
+plot(x = top.right$x.dashed, y = top.right$y.dashed, ylim = c(0,1))
+plot(x = top.right$x.solid, y = top.right$y.solid, ylim = c(0,1))
+
+bottom.left <- data.frame(x.dashed = seq(0, 1, .01),
+                          y.dashed.lb = (res_i_certain_lb / N_I) * 100,
+                          y.dashed.ub = (res_i_certain_ub / N_I) * 100,
+                          x.solid = seq(0, 1, .01),
+                          y.solid.lb = (res_i_uncertain_lb / N_I) * 100,
+                          y.solid.ub = (res_i_uncertain_ub / N_I) * 100) 
+
+plot(x = bottom.left$x.dashed, y = bottom.left$y.dashed.lb, ylim = c(0, 100))
+lines(x = bottom.left$x.dashed, y = bottom.left$y.dashed.ub, ylim = c(0, 100))
+plot(x = bottom.left$x.solid, y = bottom.left$y.solid.lb, ylim = c(0, 100))
+lines(x = bottom.left$x.solid, y = bottom.left$y.solid.ub, ylim = c(0, 100))
+
+
+bottom.right <- data.frame(x.dashed = seq(0, 1, .01),
+                           y.dashed.lb = (res_iv_certain_lb / N_IV) * 100,
+                           y.dashed.ub = (res_iv_certain_ub / N_IV) * 100,
+                           x.solid = seq(0, 1, .01),
+                           y.solid.lb = (res_iv_uncertain_lb / N_IV) * 100,
[…]

### 2018-05-01 submission/review-reply.md
Draft review response

This is a point-by-point response to [reviewer one of v1](https://f1000research.com/articles/6-1995/v1). This is *work in progress* and is drafted solely by me (still needs approval of other authors)

> The paper by Hartgerink et al. sought to determine whether or not there is evidence for data manipulation in the DECREASE-I and Decrease-IV trials. First, the authors reproduced the results of a previous meta-analysis from Bouri et al., using raw event data to calculate the ln(relative risk) and its standard error. Next, they modeled the probability that the results from the DECREASE trials could have occurred from the effect distribution of the non-DECREASE trials. Finally, by creating simulated datasets that were subsequently manipulated they determined the probability that the likelihood of the manipulated data (DECREASE meta-analytic effect distribution) is larger than the likelihood of the genuine data (non-DECREASE meta-analytic effect distribution).

> Even though the motivating case for the paper is the questioned reliability of the data within the DECREASE trials that tries to answer a very concrete clinical question (should we use perioperatively use beta blockers to reduce mortality?), the paper ultimately tries to answer a question that is much broader: can we use forensics statistics even when individual data records are not available. In our opinion, the authors succeeded in part to answer that underlying question.

Thank you :)

> General Comments:

> Some circular reasoning in  step 2, In step 2 the authors provide clear evidence that the results of the DECREASE trials are quite unlikely, however, they employ some circular reasoning in their argumentation. Because we know the DECREASE values are extreme, when a p-value is calculated for the DECREASE effect is calculated based off of the non-DECREASE effect distribution it will of course be very small. The results of this approach are methodically sound but suffer from circular reasoning and do not seem to add much to the goal of determining whether data manipulation occurred. Although this step might be useful to help ease the reader into the following steps, the limited added value of these analyses should be recognized.

We recognize the circularity and adapt the text to indicate that is a quantification of the suspicion that the data points are extreme. This quantification allows us to better assess *how* extreme these data points are.

> Robustness of approach In Step 3, the combination of adding simulated datasets with a likelihood calculation is an interesting approach to try to determine data manipulation when only summary statistics are available. In the second paragraph of the methods, when describing the inversion method, the authors say “We assumed that if data are manipulated, each data point is manipulated in the same way and to the same extent.” We have a feeling that this is a very strong assumption that is likely not to hold true and might drive the results. An exploration of the robustness, in text or analyses, could help strengthen the argument for the approach the authors apply. Additionally, since the DECREASE-IV trials have approximately 10x the sample size as DECREASE-I we expected a reduction in the confidence interval, yet the confidence interval still spans almost the entire range. Further deliberations on this phenomenon would help the reader understand the robustness of the applied technique.

We agree this is a strong assumption and recognize it as a limitation in the second paragraph of our discussion section. Because we are building these models solely on summary statistics, we have to compensate the lack of information with assumptions. For that same lack of information, it is difficult to assess the robustness simply because of the wide range of potential ways of fabricating data. We inspected the integrity reports for information on how data were fabricated but found insufficient indications to help in our modeling.

With respect to the confidence intervals, this could very well be a floor effect where the information difference is so minimal the difference also does not really show in the confidence interval range. We updated a sentence in the discussion to clarify this: "Nonetheless, much residual uncertainty remains and simply less information is available in summary results when compared to raw data, as indicated by the large range of the confidence intervals in Step 3"

>  Interpretation of results. In general, the results show that the data simulation does not give a lot of precision when it comes to pinpointing the number of manipulated datapoints. In some sense, and as also described by the authors, this is also not needed as the lower boundary of the confidence interval of manipulated datapoints should always include 0.  In fact, the authors claim that “there is a high degree of confidence that data manipulation did occur based on the difference of the trial results alone”. We do follow that reasoning, but do not fully understand why the authors adhere so much to the difference between the results of the two trials. There does not seem to be a material difference in the 95% confidence intervals between DECREASE-I and DECREASE-IV since in both cases the confidence interval spans essentially the entire range of possible outcomes. Nonetheless, the authors seem to put a lot more emphasis on the analyses of DECREASE IV with its higher sample size, in part because the results show that the lower boundary of the confidence interval does not span the 0 but ranges from 3-1066 or 10-1066, depending on the method used. Given the unknown robustness of this approach, we wonder whether the implied relevance of the small difference between the results of the DECREASE I and DECREASE IV is justified.

We understand the reviewer's hesitation and hope our adjustment to the previous comment helps contextualize the findings a bit better. We note that the dichotomous question of some form of manipulation or no manipulation is more generic than estimating the exact number. Hence, there is more confidence in making that dichotomous statement than making more specific statements.

>  Conclusion. In the discussion section, we take issue with the sentence beginning “Our results indicate that the results from the DECREASE trials ... at least some of the data were manipulated”. The word manipulated implies an intention to manipulate data, yet there are numerous ways for data to become incorrectly classified without overt and intended manipulation. Although the authors show it is likely that something altered the data, we believe that strong claims about the provenance of that change based on the analyses provided in this paper should be avoided. However, we agree with the authors that DECREASE I and DECREASE IV should be regarded as "highly problematic and untrustworthy when assessing the effect of beta blockade in perioperative mortality", but that is only when we place the results from this paper in the context of the reports on scientific integrity of the DECREASE trials.

We explicitly refrained from using the word fabricated here because it implies intent in a legal way. Upon reading the reviewer's comment, we considered alternatives such as "altered", "misprocessed", but given the conclusions of the integrity reports that preceded this study, we also find it a misrepresentation to underplay the results of those reports.

> Next to these general comments, we also have some specific suggestions or questions, mainly to help increase the readability in a potential second version of the manuscript :

> It would be helpful to have a forest plot of Table 1, to better visualize the results and compare DECREASE vs. non-DECREASE.

We did generate this forest plot at some point, and saw exactly the same one as in Bouri et al. (2014). We explicitly did not include a forest plot for that reason.

> In Table 2, the authors could report the relative risk (RR) instead of the log(RR) to improve the table’s readability and help the reader understand the magnitude of the different effects depicted. We would also be interested whether some more general information on this topic, especially regarding the non-DECREASE trials to put the DECREASE trials and their data into perspective. (e.g. how many non decrease trials were registered but not published; or whether there is any other evidence or suggestion for non-publication of relevant data that could alter some of the conclusions in this paper.)

We could report those numbers, but we decided to stick to the log(RR) because that is closer to the output and maintains the symmetry of the confidence intervals (asymmetric CIs might confuse the reader too). With respect to investigating how many trials had been registered, we agree it is an interesting research question on its own and consider a superficial investigation in the current paper to contextualize the findings not proportional to the complexity of the issue (e.g., a review of trial registration would be necessary given the first DECREASE trial took place in the late 90s). Hence, we do not take up this suggestion here.

>  A minor point, re-label the subscripts of the Likelihood function as DECREASE and non-DECREASE to avoid the loaded terminology of manipulated and genuine which seem to presuppose the conclusion.

Considering this investigation was based on the integrity reports that preceded it, we consider it reasonable to use these terms without presupposing the conclusion. We also explicitly state that these labels operate under *the assumption* that the DECREASE trials were fabricated (in step 3; in step 2 that assumption is explicitly reversed).

> We suggest rewording the second paragraph under the results section, and specifically address the sentence that begins “The relatively minor difference…”. This sentence first references “estimates”, but it is unclear what these estimates are referring to.

Adjusted for clarity to "The relatively minor difference between the point- and distribution estimates indicates "

> In Figure 2 please address why, in the bottom lower panel, the confidence interval for the point estimate is larger than the distribution estimate. Because the distribution estimate has more error in it, we were expecting that the confidence interval of the distribution estimate would be larger than that of the point estimate.

The reviewers are correct in finding this odd. We also found this odd and checked the code multiple times to find any mistakes. We invite anyone to take a look and correct us if we're wrong, but we are unable to find the mistake if there is one. Notably, it is the same code that gives the results for DECREASE-I that do follow the notion that the point estimates are more precise. We added a sentence to make this explicit to the reader: "We note that, for reasons unknown to us, the confidence interval of the point estimates are more uncertain than those of the distribution estimates, whereas similar code for DECREASE-I provides the reverse (the code is available for review at https://osf.io/4zs29/)."

### 2018-05-01 submission/manuscript.Rmd
Draft review response

@@
 Based on this effect distribution, we estimated the veracity of the DECREASE trials separately, which is the estimated probability of the observed data (or more extreme) under a given true effect [@peters2015].
+We note that this is primarily aimed to quantify the suspected difference between the DECREASE and non-DECREASE trials.
 
@@
 For DECREASE-IV ($N=`r bouri_data$control_n[11] + bouri_data$beta_n[11]`$), the 95\% confidence interval for the estimated number of manipulated data points is [`r which.min(res_iv_certain - .025) - 1` - `r which.max(res_iv_certain - .975) - 1`] or [`r which.min(res_iv_uncertain - .025) - 1` - `r which.max(res_iv_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively.
-The relatively minor difference between the estimates indicates that there is a high degree of confidence that data manipulation did occur based on the difference of the trial results alone.
+The relatively minor difference between the point- and distribution estimates indicates that there is a high degree of confidence that data manipulation did occur based on the difference of the trial results alone.
 Nonetheless, the range of potentially manipulated data points is still estimated at approximately 1000; this indicates that the summary results are insufficient to provide more than an estimated lower bound.
 This indicates that it is possible not all data were manipulated (i.e., $N=`r length(res_iv_certain) - 1`$) but at least some were, increasing the importance of well-documented data provenance to discern between genuine and falsified data.
+We note that, for reasons unknown to us, the confidence interval of the point estimates are more uncertain than those of the distribution estimates, whereas similar code for DECREASE-I provides the reverse (the code is available for review at https://osf.io/4zs29/).
 
@@
 
-The results of our analyses also highlight that, despite the lack of availability of the raw data, summary results from larger samples allow for better estimates of the number of manipulated data points when similar trials are available.
-Moreover, larger trials result in relatively more certainty (e.g., DECREASE-IV) about the estimated number of manipulated data points, when using the inversion method, compared to smaller trials (e.g., DECREASE-I).
+The results of our analyses also highlight that, despite the lack of availability of the raw data, summary results from larger samples allow for marginally better estimates of the number of manipulated data points when similar trials are available.
+Larger trials result in relatively more certainty (e.g., DECREASE-IV) about the estimated number of manipulated data points, when using the inversion method, compared to smaller trials (e.g., DECREASE-I).
 This increased certainty is due to decreased standard errors of the estimated effects, resulting in higher sensitivity to data anomalies.
-Nonetheless, much residual uncertainty remains and simply less information is available in summary results when compared to raw data.
+Nonetheless, much residual uncertainty remains and simply less information is available in summary results when compared to raw data, as indicated by the large range of the confidence intervals in Step 3.
 As such, raw data availability would improve the options open to detect potential anomalies (note: raw data are available for DECREASE VI, but upon a freedom of information request by the first author, Erasmus MC refused to share these data; see [osf.io/fswjt/](https://osf.io/fswjt/)).

### 2018-05-23 submission/review-reply.md
Update pre-revisions

@@
 
-We recognize the circularity and adapt the text to indicate that is a quantification of the suspicion that the data points are extreme.
-This quantification allows us to better assess *how* extreme these data points are.
+We recognize the circularity and adapt the text to indicate that this is a quantification of the suspicion that the data points are extreme.
+This quantification shows *how* extreme these data points are.
 
@@
 
-We could report those numbers, but we decided to stick to the log(RR) because that is closer to the output and maintains the symmetry of the confidence intervals (asymmetric CIs might confuse the reader too).
+We decided to stick to the log(RR) because otherwise readers would have to backtransform when trying to calculate the log(RR) for a specific set of conditions using the results.
 With respect to investigating how many trials had been registered, we agree it is an interesting research question on its own and consider a superficial investigation in the current paper to contextualize the findings not proportional to the complexity of the issue (e.g., a review of trial registration would be necessary given the first DECREASE trial took place in the late 90s).
@@
 
-The reviewers are correct in finding this odd.
 We also found this odd and checked the code multiple times to find any mistakes.
@@
 Notably, it is the same code that gives the results for DECREASE-I that do follow the notion that the point estimates are more precise.
-We added a sentence to make this explicit to the reader: "We note that, for reasons unknown to us, the confidence interval of the point estimates are more uncertain than those of the distribution estimates, whereas similar code for DECREASE-I provides the reverse (the code is available for review at https://osf.io/4zs29/)."
+We added a sentence to make this explicit to the reader: "We note that, for reasons unknown to us, the confidence interval of the point estimates are wider than those of the distribution estimates, whereas similar code for DECREASE-I yield the reverse (the code is available for review at https://osf.io/4zs29/)."

### 2018-05-23 submission/manuscript.Rmd
Update pre-revisions

@@
 This indicates that it is possible not all data were manipulated (i.e., $N=`r length(res_iv_certain) - 1`$) but at least some were, increasing the importance of well-documented data provenance to discern between genuine and falsified data.
-We note that, for reasons unknown to us, the confidence interval of the point estimates are more uncertain than those of the distribution estimates, whereas similar code for DECREASE-I provides the reverse (the code is available for review at https://osf.io/4zs29/).
+We note that, for reasons unknown to us, the confidence interval of the point estimates are wider than those of the distribution estimates, whereas similar code for DECREASE-I yield the reverse (the code is available for review at https://osf.io/4zs29/).
 
@@
 This increased certainty is due to decreased standard errors of the estimated effects, resulting in higher sensitivity to data anomalies.
-Nonetheless, much residual uncertainty remains and simply less information is available in summary results when compared to raw data, as indicated by the large range of the confidence intervals in Step 3.
+Nonetheless, much residual uncertainty remains and simply less information is available in summary results than in raw data, as indicated by the width of the confidence intervals in Step 3.
 As such, raw data availability would improve the options open to detect potential anomalies (note: raw data are available for DECREASE VI, but upon a freedom of information request by the first author, Erasmus MC refused to share these data; see [osf.io/fswjt/](https://osf.io/fswjt/)).
